### 2023-07-18 Articles/ERK/erk.tex
clear code, article template

% !TEX encoding = UTF-8 Unicode
\documentclass[a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage{erk}
\usepackage{times}
\usepackage{graphicx}
\usepackage[top=22.5mm, bottom=22.5mm, left=22.5mm, right=22.5mm]{geometry}

\usepackage[slovene,english]{babel}

% local definitions
\def\footnotemark{}%  to avoid footnote on cover page

\begin{document}
%make title
\title{Navodila in osnutek prispevka za konferenco ERK 2014}

\author{Prvi Avtor$^{1}$, Drugi Avtor$^{2}$} % use ^1, ^2 for author(s) from different institutions

\affiliation{$^{1}$Ime in naslov organizacije prvega avtorja\\ 
$^{2}$Ime in naslov organizacije drugega avtorja, če je drugačna od prvega }

\email{E-pošta: avtor@naslov.com}

\maketitle

%\thispagestyle{empty}

\begin{abstract}{Instructions and Template for ERK 2014 Conference}
For the papers written in Slovene language an English title and abstract is required. The abstract should be no more than 200 words.

We accept only electronic submissions in pdf (without password protection) or Microsoft Word format. Submission form can be found on ERK web page \cite{ERK}.
\end{abstract}

\selectlanguage{slovene}

\section{Uvod}

Prispevki za Elektrotehniško in računalniško konferenco (ERK) naj bodo napisani na papirju formata A4 v dveh stolpcih z vsemi zunanjimi robovi velikosti 2,25cm. Redni prispevki naj ne obsegajo več kot 4 strani skupaj s slikami, povzetki študentskih referatov pa 2 strani. Oddani prispevek mora biti brez opomb in številk strani.

\section{Pisanje prispevka}

Pri pisanju uporabljajte sloge, ki so definirani v datoteki erk.sty in ne spreminjajte velikosti ali sloga pisave. Ta dokument lahko vzamete kot vzorec za oblikovanje prispevka v orodju \LaTeX.
Tabela \ref{tab1} povzema osnovne sloge in velikosti pisave.

\begin{table}[h]
\caption{Osnovni slogi pri oblikovanju prispevka.} \label{tab1}
\smallskip
\begin{center}
\begin{tabular}{ | r | c | c | }
\hline  
  \textbf{Text Style} & \textbf{Font Size} & \textbf{Attributes}\\ 
\hline  
  Paper title & 17pt & bold\\
  Author names & 12pt & bold\\
  Affiliation & 10pt & italic\\
  Heading & 12pt & bold \\
  Subheading & 10pt & bold\\
  Regular text & 10pt &\\
  Captions \& ref. & 9pt &\\
\hline  
\end{tabular}
\end{center}
\end{table}

\subsection{Enačbe in slike}

Enačbe in formule se pišejo s poševnimi črkami in so zaporedno oštevilčene. V besedilu naj bodo reference na enačbe zapisane v oklepajih (\ref{eq1}).

\begin{equation}
 \int^{r_{2}}_{0}F(r,\varphi)\; dr\; d\varphi= [\sigma r_{2} / (2\mu_{0})]
    \label{eq1}
\end{equation}

Prispevki za ERK se ne tiskajo v barvah, zato pripravite črno-bele slike. Slike naj bodo postavljene v okvirje širine enega ali dveh stolpcev in označene s številko in podnapisom. V besedilu morajo biti reference na slike (slika \ref{slika}).

\begin{figure}[!htb]
    \begin{center}
        \includegraphics[scale=1]{field1.eps}
        \caption{Podajte kratko razlago slike.} \label{slika}
    \end{center}
\end{figure}

\section{Oddaja prispevka}

Prispevek oddajte v elektronski obliki preko formularja na spletni strani \cite{ERK}. Sprejemamo prispevke v formatu pdf (brez nastavljenih zaščitnih gesel) ali Microsoft Word. Pred oddajo prispevka prosimo preverite dodatne informacije za avtorje, ki so na spletnih straneh.

Formular za elektronsko oddajo prispevkov bo aktiviran v juniju, rok za oddajo je \textbf{13. julij 2014}. Avtorje bomo obvestili o izboru konec avgusta.

\small
\begin{thebibliography}{1}

\bibitem{ERK} ERK, http://www.ieee.si/erk/index.html 
\bibitem{Zbornik} B. Zajc, A. Trost: Zbornik triindvajsete mednarodne Elektrotehniške in računalniške konference ERK 2014, 22. - 24. September 2014, Portorož, Slovenija

\end{thebibliography}

\end{document}

### 2023-07-18 Code/directional_kernel_2d.m

@@ -1,40 +1,0 @@
-function [kernel] = directional_kernel_2d(direction, kernel_length, kernel_width, kernel_sigma, kernel_type)
-
-    % x - direction
-    % ------------------------------------
-
-    kernel = zeros(kernel_size, kernel_size);
-    center = ceil(kernel_size/2); % mid of the kernel
-
-
-    if kernel_type == 'box'
-
-        % prepare directional weights
-        weights = [1:1:center center+1 -flip(1:1:center)] / (center+1); 
-
-        % extend weights to kernel width
-        kernel = ones(kernel_width,1) * weights
-
-
-
-
-
-
-    elseif kernel_type == 'gaussian'
-
-        % The standard deviation is usually set to be (Size/2)/3
-        sigma = floor(kernel_width/2)/3; 
-
-        [X,Y] = meshgrid(-(center-1):(center-1));
-        kernel = exp(-(X.^2) / (2* sigma^2));
-        kernel = kernel / sum(kernel(:)) % Normalize the kernel
-
-    end
-
-
-
-
-
-
-
-end

### 2023-07-18 Articles/ERK/erk.tex
article structure

@@
 %make title
-\title{Navodila in osnutek prispevka za konferenco ERK 2014}
+\title{Utilizing Convolution for Obstacle Avoidance during Robotic Manipulator Movement}
 
-\author{Prvi Avtor$^{1}$, Drugi Avtor$^{2}$} % use ^1, ^2 for author(s) from different institutions
+\author{Jakob Baumgartner$^{1}$, Gregor Klančar$^{2}$} % use ^1, ^2 for author(s) from different institutions
 
-\affiliation{$^{1}$Ime in naslov organizacije prvega avtorja\\ 
-$^{2}$Ime in naslov organizacije drugega avtorja, če je drugačna od prvega }
+\affiliation{$^{1}$Fakulteta za Elektrotehniko, Tržaška cesta 25, 1000 Ljubljana\\}
 
-\email{E-pošta: avtor@naslov.com}
+\email{E-pošta: jakob.baumgartner@fe.uni-lj.si}
 
@@
 
-\begin{abstract}{Instructions and Template for ERK 2014 Conference}
+\begin{abstract}{Abstract}
 For the papers written in Slovene language an English title and abstract is required. The abstract should be no more than 200 words.
@@
 
-\section{Uvod}
+\section{Introduction }
 
@@
 Oddani prispevek mora biti brez opomb in številk strani.
+
+\section{Background and Related Work}
+
+\section{Methodology }
+
+\subsection{Obstacle Grid}
+
+\subsection{Convolution Techniques}
+
+\subsection{Robot Segments Control}
+
+\section{Experiment and Results}
 

### 2023-07-18 Code/directional_kernel_3d.m
3D z - directional kernel

@@ -35,7 +35,7 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, kernel_sig
         kernel = exp(-(1:center).^2/(2*sigma^2)); 
         
         % generate symmetric kernel
-        kernel = [flip(kernel) 1 kernel] / sum([flip(kernel) 1 kernel]);
+        kernel = [flip(kernel) 0 -kernel] / center;
         
         % prepare gaussian weights for width
         center_width = floor(kernel_width/2);
@@ -50,6 +50,8 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, kernel_sig
         height_extender = exp(-(1:center_height).^2/(2*sigma^2));
         height_extender = [flip(height_extender) 1 height_extender] / sum([flip(height_extender) 1 height_extender]);
 
+        % create each layer
+        kernel3D = repmat(kernel2D, 1, 1, length(height_extender)) .* reshape(height_extender, 1, 1, []);
 
     end
     
@@ -71,14 +73,16 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, kernel_sig
 
     % z kernel
     elseif strcmp(direction, 'z')
-    % kernel width and height should be equal (imagine a column pushing
-    % down on robot)
+        % kernel width and height should be equal (imagine a column pushing
+        % down on robot)
 
-                % TODO !!!
-        for i = 1:1:kernel_length % TODO
-            kernel3D(i,:,:) = kernel2D % TODO !!!
-        end
-    
+        % reshape arrays for bsxfun operation
+        reshaped_kernel2D = reshape(kernel2D, [1 size(kernel2D)]);
+        reshaped_height_extender = reshape(height_extender, [length(height_extender) 1 1]);
+        
+        % perform element-wise multiplication using bsxfun
+        kernel3D = bsxfun(@times, reshaped_kernel2D, reshaped_height_extender);
+
 
     end
 

### 2023-07-18 Code/directional_kernel_3d.m
Update directional_kernel_3d.m

@@ -8,18 +8,18 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, kernel_sig
    if strcmp('linear',kernel_type)
 
         % prepare directional weights
-        kernel = [1:1:center 0 -flip(1:1:center)] / (center+1); 
+        kernel = [1:1:center 0 -flip(1:1:center)] / (center); 
 
         % extend weights to kernel width
         center_width = floor(kernel_width/2);
-        width_extender = [1:1:center_width center_width+1 flip(1:1:center_width)]/(center_width+1);
+        width_extender = [1:1:center_width center_width flip(1:1:center_width)]/(center_width);
 
         % generate 2D kernel
         kernel2D = width_extender' * kernel;
 
         % extend weight to kernel height
         center_height = floor(kernel_height/2);
-        height_extender = [1:1:center_height center_height+1 flip(1:1:center_height)]/(center_height+1);
+        height_extender = [1:1:center_height center_height flip(1:1:center_height)]/(center_height);
         
         % create each layer
         kernel3D = repmat(kernel2D, 1, 1, length(height_extender)) .* reshape(height_extender, 1, 1, []);
@@ -33,14 +33,16 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, kernel_sig
         
         % generate gaussian weights
         kernel = exp(-(1:center).^2/(2*sigma^2)); 
+
+        kernel_max = kernel(1);
         
         % generate symmetric kernel
-        kernel = [flip(kernel) 0 -kernel] / center;
+        kernel = [flip(kernel) 0 -kernel]/kernel_max;
         
         % prepare gaussian weights for width
         center_width = floor(kernel_width/2);
         width_extender = exp(-(1:center_width).^2/(2*sigma^2)); 
-        width_extender = [flip(width_extender) 1 width_extender] / sum([flip(width_extender) 1 width_extender]);
+        width_extender = [flip(width_extender) 1 width_extender];
         
         % generate 2D kernel
         kernel2D = width_extender' * kernel;
@@ -48,7 +50,7 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, kernel_sig
         % prepare gaussian weights for height
         center_height = floor(kernel_height/2);
         height_extender = exp(-(1:center_height).^2/(2*sigma^2));
-        height_extender = [flip(height_extender) 1 height_extender] / sum([flip(height_extender) 1 height_extender]);
[…]

### 2023-07-18 Code/directional_kernel_3d.m
Update directional_kernel_3d.m

@@ -1,7 +1,7 @@
 function [kernel3D] = directional_kernel_3d(direction, kernel_length, kernel_sigma, kernel_width, kernel_width_sigma, kernel_height, kernel_height_sigma, kernel_type)
 
-    % create kernel
-    % =========================================================================
+   % FORM KERNELS 
+   % ---------------------------------------------------------------------
 
     center = floor(kernel_length/2); % mid of the kernel
 
@@ -26,7 +26,6 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, kernel_sig
 
 
     elseif strcmp('gaussian',kernel_type)
-
 
         % define the standard deviation (sigma)
         sigma = 1;
@@ -55,7 +54,10 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, kernel_sig
         % create each layer
         kernel3D = repmat(kernel2D, 1, 1, length(height_extender)) .* reshape(height_extender, 1, 1, []);
 
-    end
+   end
+
+   % STACK KERNELS IN PICKED DIRECTION
+   % ---------------------------------------------------------------------
     
     % x kernel
     if strcmp(direction, 'x')
@@ -87,10 +89,6 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, kernel_sig
 
 
     end
-
-
-
-
 
     end
 

### 2023-07-18 Code/ObstacleGrid_RTConvolution_OnePoint.m
fixes directional kernel

@@ -1,0 +1,237 @@
+function [joints_positions, EE_positions, goal_distances, q_velocities, ee_velocities, values_APF] = ObstacleGrid_RTConvolution_OnePoint(grid, grid_field, goal_point, Tbase, q, varargin)
+
+% Parse the optional arguments
+p = inputParser;
+addOptional(p, 'mid_joints', true); % Default value is false
+addOptional(p, 'avoid_task', true); % Default value is true
+
+parse(p, varargin{:});
+
+mid_joints = p.Results.mid_joints;
+avoid_task = p.Results.avoid_task;
+
+% -----------------------------------------------------------
+
+% function parameters
+goal_dist = 0.02; % distance which satisfies ending of optimization
+damping_factor = 0.1; % damping factor to avoid inverse Jacobain matrix singularities
+Tstep = 0.01; % time step
+Nmax = 1000; % max number of iterations
+space_resolution = grid.resolution; % resolution of the obstacles grid
+
+% joints range
+q_range = [2.8973 -2.8973;
+           1.7628 -1.7628;
+           2.8973 -2.8973;
+           -0.0698 -3.0718;
+           2.8973 -2.8973;
+           3.7525 -0.0175;
+           2.8973 -2.8973];
+
+% number of points per segment for obstacle avoidance task
+points_per_segment = 1*[1 1 1 1 1 1 1];
+
+% weights for different tasks
+wp = 5; % primary task
+wm = 1; % mid-joints task
+wa = 0.1; % obstacle avoidance task
+
+% -----------------------------------------------------------
+
+% get a list of transforms for every APF detection point on the robot
+transformations_list = getPartialTransformationsListPanda('points_per_segment', points_per_segment);
+
+% get current robot pose
+[robot_transforms] = GeometricPandaMATLAB(q, Tbase);
+
+% set current ee point to start point
+ee_point = robot_transforms(1:3,4,8);
+
+% calculate distance from goal
+current_dist = norm(ee_point'- goal_point(1:3));
+
+% add starting positions to saved array
+EE_positions = [ee_point];
+joints_positions = [q];
+goal_distances = [current_dist];
+q_velocities = [];
+values_APF = [];
+ee_velocities = [];
+
+% set itarations count
+Niter = 0;
+
+% trajectory calculation loop
+while current_dist > goal_dist && Niter < Nmax  
+
+    % OPTION JACOBIAN
+    [~,~,J]=kinmodel_panda(q); % calculate jacobian
+
[…]

### 2023-07-18 Code/directional_kernel_3d.m

@@ -1,4 +1,4 @@
-function [kernel3D] = directional_kernel_3d(direction, kernel_length, kernel_sigma, kernel_width, kernel_width_sigma, kernel_height, kernel_height_sigma, kernel_type)
+function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, kernel_width, sigma_width, kernel_height, sigma_height, kernel_type)
 
    % FORM KERNELS 
    % ---------------------------------------------------------------------
@@ -8,7 +8,7 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, kernel_sig
    if strcmp('linear',kernel_type)
 
         % prepare directional weights
-        kernel = [1:1:center 0 -flip(1:1:center)] / (center); 
+        kernel = [1:1:center center -flip(1:1:center)] / (center); 
 
         % extend weights to kernel width
         center_width = floor(kernel_width/2);
@@ -26,21 +26,18 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, kernel_sig
 
 
     elseif strcmp('gaussian',kernel_type)
-
-        % define the standard deviation (sigma)
-        sigma = 1;
-        
+      
         % generate gaussian weights
         kernel = exp(-(1:center).^2/(2*sigma^2)); 
 
         kernel_max = kernel(1);
         
         % generate symmetric kernel
-        kernel = [flip(kernel) 0 -kernel]/kernel_max;
+        kernel = [flip(kernel) 1 -kernel]/kernel_max;
         
         % prepare gaussian weights for width
         center_width = floor(kernel_width/2);
-        width_extender = exp(-(1:center_width).^2/(2*sigma^2)); 
+        width_extender = exp(-(1:center_width).^2/(2*sigma_width^2)); 
         width_extender = [flip(width_extender) 1 width_extender];
         
         % generate 2D kernel
@@ -48,7 +45,7 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, kernel_sig
         
         % prepare gaussian weights for height
         center_height = floor(kernel_height/2);
-        height_extender = exp(-(1:center_height).^2/(2*sigma^2));
+        height_extender = exp(-(1:center_height).^2/(2*sigma_height^2));
         height_extender = [flip(height_extender) 1 height_extender];
 
         % create each layer
[…]

### 2023-07-18 Code/directional_kernel_3d.m
kernel tunning

@@ -8,18 +8,18 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
    if strcmp('linear',kernel_type)
 
         % prepare directional weights
-        kernel = [1:1:center center -flip(1:1:center)] / (center); 
+        kernel = [1:1:center 0 -flip(1:1:center)] / (center); 
 
         % extend weights to kernel width
         center_width = floor(kernel_width/2);
-        width_extender = [1:1:center_width center_width flip(1:1:center_width)]/(center_width);
+        width_extender = [1:1:center_width 0 flip(1:1:center_width)]/(center_width);
 
         % generate 2D kernel
         kernel2D = width_extender' * kernel;
 
         % extend weight to kernel height
         center_height = floor(kernel_height/2);
-        height_extender = [1:1:center_height center_height flip(1:1:center_height)]/(center_height);
+        height_extender = [1:1:center_height 0 flip(1:1:center_height)]/(center_height);
         
         % create each layer
         kernel3D = repmat(kernel2D, 1, 1, length(height_extender)) .* reshape(height_extender, 1, 1, []);

### 2023-07-18 Code/directional_kernel_3d.m
kernel fixes

@@ -12,14 +12,14 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
 
         % extend weights to kernel width
         center_width = floor(kernel_width/2);
-        width_extender = [1:1:center_width 0 flip(1:1:center_width)]/(center_width);
+        width_extender = [1:1:center_width center_width flip(1:1:center_width)]/(center_width);
 
         % generate 2D kernel
         kernel2D = width_extender' * kernel;
 
         % extend weight to kernel height
         center_height = floor(kernel_height/2);
-        height_extender = [1:1:center_height 0 flip(1:1:center_height)]/(center_height);
+        height_extender = [1:1:center_height center_height flip(1:1:center_height)]/(center_height);
         
         % create each layer
         kernel3D = repmat(kernel2D, 1, 1, length(height_extender)) .* reshape(height_extender, 1, 1, []);
@@ -33,12 +33,12 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         kernel_max = kernel(1);
         
         % generate symmetric kernel
-        kernel = [flip(kernel) 1 -kernel]/kernel_max;
+        kernel = [flip(kernel) 0 -kernel]/kernel_max;
         
         % prepare gaussian weights for width
         center_width = floor(kernel_width/2);
         width_extender = exp(-(1:center_width).^2/(2*sigma_width^2)); 
-        width_extender = [flip(width_extender) 1 width_extender];
+        width_extender = [flip(width_extender) exp(-(center_width).^2/(2*sigma_width^2)) width_extender];
         
         % generate 2D kernel
         kernel2D = width_extender' * kernel;
@@ -46,7 +46,7 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         % prepare gaussian weights for height
         center_height = floor(kernel_height/2);
         height_extender = exp(-(1:center_height).^2/(2*sigma_height^2));
-        height_extender = [flip(height_extender) 1 height_extender];
+        height_extender = [flip(height_extender) exp(-(center_height).^2/(2*sigma_height^2)) height_extender];
 
         % create each layer
         kernel3D = repmat(kernel2D, 1, 1, length(height_extender)) .* reshape(height_extender, 1, 1, []);

### 2023-07-18 Code/ObstacleGrid_RTConvolution_OnePoint.m
changed classic to APF for primary task, to get trajectory like movement

@@ -14,8 +14,8 @@ avoid_task = p.Results.avoid_task;
 
 % function parameters
 goal_dist = 0.02; % distance which satisfies ending of optimization
-damping_factor = 0.1; % damping factor to avoid inverse Jacobain matrix singularities
-Tstep = 0.01; % time step
+damping_factor = 0.01; % damping factor to avoid inverse Jacobain matrix singularities
+Tstep = 0.1; % time step
 Nmax = 1000; % max number of iterations
 space_resolution = grid.resolution; % resolution of the obstacles grid
 
@@ -32,8 +32,8 @@ q_range = [2.8973 -2.8973;
 points_per_segment = 1*[1 1 1 1 1 1 1];
 
 % weights for different tasks
-wp = 5; % primary task
-wm = 1; % mid-joints task
+wp = 1; % primary task
+wm = 5; % mid-joints task
 wa = 0.1; % obstacle avoidance task
 
 % -----------------------------------------------------------
@@ -89,11 +89,27 @@ while current_dist > goal_dist && Niter < Nmax
     % calculate pseudo inverse
     pinv_J = J'*(J*J'+ damping_factor^2 * eye(6))^-1; % damping to avoid singularities
 
-  % calculate ee velocities
-    ee_vel = goal_point(1:3)' - ee_point;
+    % --------------------------------------------------
+    % OPTION KINEMATICS CLASSIC END EFFECTOR
+
+%     % calculate ee velocities
+%     ee_vel = goal_point(1:3)' - ee_point;
+% 
+%     % calculate joint velocities using inverse kinematics
+%     q_vel = pinv_J * (wp * [ee_vel ; 0 ; 0 ; 0]);
+
+    % --------------------------------------------------
+
+    % EE kinematics using APF  %%%% TODODODODODODO TODO !!!! (FIX THIS,
+    % REPULSIVE FIELD NOT WORKING REALLY)
+
+    % calculate ee velocities by interpolating APF
+    [dx,dy,dz] = interpolate_derivative(ee_point, grid_field, space_resolution);
+    ee_vel = -[dx ; dy ; dz ; 0 ; 0 ; 0]; 
 
     % calculate joint velocities using inverse kinematics
-    q_vel = pinv_J * (wp * [ee_vel ; 0 ; 0 ; 0]);
+    q_vel = pinv_J * (wp * ee_vel);   
+
 
     % calculate Null Space 
[…]

### 2023-07-18 Code/showMovementPanda.m

@@ -29,12 +29,14 @@ function [f] = showMovementPanda(grid, grid_repulsive, control_points, joints_po
         'Value', 1,...
         'Callback', @checkboxCallbackObstacles); 
 
-    % Add checkbox for Repulsive field
-    cr = uicontrol('Style', 'checkbox',...
-        'String', 'Repulsive field',...
-        'Position', [20 60 100 20],...
-        'Value', 0,...
-        'Callback', @checkboxCallbackRepulsive); 
+    if size(grid_repulsive,1) > 1
+        % Add checkbox for Repulsive field
+        cr = uicontrol('Style', 'checkbox',...
+            'String', 'Repulsive field',...
+            'Position', [20 60 100 20],...
+            'Value', 0,...
+            'Callback', @checkboxCallbackRepulsive); 
+    end
 
     % On/off APF vectors
     show_arrows = 1;
@@ -43,7 +45,7 @@ function [f] = showMovementPanda(grid, grid_repulsive, control_points, joints_po
     HRepulsive = [];
     
     % display grid
-    HObstacles = grid.showGridVol3D(grid.grid,'floor',true,'height',true);
+    HObstacles = grid.showGridVol3D(grid.grid,'floor',true,'height',false);
     hold on
     axis equal
 
@@ -187,7 +189,7 @@ function [f] = showMovementPanda(grid, grid_repulsive, control_points, joints_po
         % Check if the checkbox is checked
         if checkboxValue == 1
             % display grid
-            HObstacles = grid.showGridVol3D(grid.grid,'floor',false,'height',true);
+            HObstacles = grid.showGridVol3D(grid.grid,'floor',false,'height',false);
         else
             for i = 1:1:length(HObstacles.handles)
                 delete(HObstacles.handles(i))

### 2023-07-19 Code/ObstacleGrid_RTConvolution_OnePoint.m
WORKING JOINT 4 AVOIDANCE

@@ -33,7 +33,7 @@ points_per_segment = 1*[1 1 1 1 1 1 1];
 
 % weights for different tasks
 wp = 1; % primary task
-wm = 5; % mid-joints task
+wm = 4; % mid-joints task
 wa = 0.1; % obstacle avoidance task
 
 % -----------------------------------------------------------
@@ -59,12 +59,15 @@ values_APF = [];
 ee_velocities = [];
 
 % set directional kernels
-kernel_size = 31;
-sigma = 21;
+sigma = 41;
 
-kernel_x = directional_kernel_3d('x', kernel_size, sigma, kernel_size, sigma, kernel_size, sigma, 'gaussian')/1000;
-kernel_y = directional_kernel_3d('y', kernel_size, sigma, kernel_size, sigma, kernel_size, sigma, 'gaussian')/1000;
-kernel_z = directional_kernel_3d('z', kernel_size, sigma, kernel_size, sigma, kernel_size, sigma, 'gaussian')/1000;
+kernel_length = 61;
+kernel_width = 9;
+kernel_height = kernel_width;
+
+kernel_x = directional_kernel_3d('x', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, 'gaussian')/1000;
+kernel_y = directional_kernel_3d('y', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, 'gaussian')/1000;
+kernel_z = directional_kernel_3d('z', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, 'gaussian')/1000;
 
 % set itarations count
 Niter = 0;
@@ -174,19 +177,19 @@ while current_dist > goal_dist && Niter < Nmax
         % --------------------
 
         % cut out local part of the grid around joint 4
-        cutout = grid.grid(((1:kernel_size)+xyz(2)-ceil(kernel_size/2)), ((1:kernel_size)+xyz(1)-ceil(kernel_size/2)), ((1:kernel_size)+xyz(3)-ceil(kernel_size/2)));
+        cutout = grid.grid(((1:kernel_length)+xyz(2)-ceil(kernel_length/2)), ((1:kernel_width)+xyz(1)-ceil(kernel_width/2)), ((1:kernel_height)+xyz(3)-ceil(kernel_height/2)));
 
         % calculate x - distance transform
-        C = cutout .* kernel_x; % Perform element-wise multiplication
-        dx = sum(C(:)); % Sum all values 
+%         C = cutout .* kernel_x; % Perform element-wise multiplication
[…]

### 2023-07-19 Code/ObstacleGrid_RTConvolution_OnePoint.m
joint limits

@@ -32,9 +32,9 @@ q_range = [2.8973 -2.8973;
 points_per_segment = 1*[1 1 1 1 1 1 1];
 
 % weights for different tasks
-wp = 1; % primary task
-wm = 4; % mid-joints task
-wa = 0.1; % obstacle avoidance task
+wp = 5; % primary task
+wm = 5; % mid-joints task
+wa = 0.5; % obstacle avoidance task
 
 % -----------------------------------------------------------
 
@@ -220,12 +220,24 @@ while current_dist > goal_dist && Niter < Nmax
             % (TODO: calculate weighting coefficients)
     
         % calculate pseudo inverse
-        % pinv_J0 = (J0*N)'*((J0*N)*(J0*N)'+ damping_factor^2 * eye(6))^-1; % damping to avoid singularities
-        pinv_J0 = J0'*(J0*J0'+ damping_factor^2 * eye(6))^-1;
+%         pinv_J0 = (J0*N)'*((J0*N)*(J0*N)'+ damping_factor^2 * eye(6))^-1; % damping to avoid singularities
+%         pinv_J0 = J0'*(J0*J0'+ damping_factor^2 * eye(6))^-1;
     
         % calculate avoidance joints velocities
-        % q_vel = q_vel + pinv_J0*(avoid_vel - J0*pinv_J*ee_vel);
-        q_vel = q_vel + pinv_J0*(avoid_vel);
+%         q_vel = q_vel + pinv_J0*(avoid_vel - J0*pinv_J*ee_vel);
+%         q_vel = q_vel + pinv_J0*(avoid_vel);
+
+    % APPROXIMATE SOLUTION
+    % -----------------------
+    
+    % calculate pseudo inverse
+    pinv_J0 = J0'*(J0*J0'+ damping_factor^2 * eye(6))^-1;
+
+
+    % calculate avoidance joints velocities
+    q_vel = q_vel + N * pinv_J0*(avoid_vel);
+
+    % -----------------------
 
 
     end
@@ -233,6 +245,9 @@ while current_dist > goal_dist && Niter < Nmax
 
     % calculate new joint positions
     q = q + q_vel * Tstep;
+
+    % limit positions within joint limits
+    [q, ~] = checkPositionLimits(q);
 
     % calculate new EE position
     [robot_transforms] = GeometricPandaMATLAB(q, Tbase);

### 2023-07-19 Code/checkPositionLimits.m

@@ -1,0 +1,19 @@
+function [jointsWithinLimits, isWithinLimits] = checkPositionLimits(position)
+
+    combinedLimits = [ 
+        -2.7437 2.7437 ; 
+        -1.7837 1.7837 ;
+        -2.9007 2.9007 ;
+        -3.0421 -0.1518 ;
+        -2.8065 2.8065 ;
+        0.5445 4.5169 ;
+        -3.0159 3.0159
+    ];
+
+    isWithinLimits = position >= combinedLimits(:, 1) & position <= combinedLimits(:, 2);
+    
+    jointsWithinLimits = (position <= combinedLimits(:, 1)) .* combinedLimits(:, 1) ...
+                       + (position >= combinedLimits(:, 2)) .* combinedLimits(:, 2) ...
+                       + (position > combinedLimits(:, 1)) .* (position < combinedLimits(:, 2)) .* position;
+
+end

### 2023-07-19 Articles/ERK/erk.tex
ERK article structure comments
ERK article: introduction, background and related work

@@
 \usepackage{graphicx}
+\usepackage{xcolor}
 \usepackage[top=22.5mm, bottom=22.5mm, left=22.5mm, right=22.5mm]{geometry}
@@
 \usepackage[slovene,english]{babel}
+
+\newcommand\todocomment[1]{\textcolor{red}{||TODO:\\ #1\\||}}
 
@@
 
-Prispevki za Elektrotehniško in računalniško konferenco (ERK) naj bodo napisani na papirju formata A4 v dveh stolpcih z vsemi zunanjimi robovi velikosti 2,25cm.
-Redni prispevki naj ne obsegajo več kot 4 strani skupaj s slikami, povzetki študentskih referatov pa 2 strani.
-Oddani prispevek mora biti brez opomb in številk strani.
+Robotic manipulators, specifically those with a high degree of freedom (DOF), possess a crucial characteristic known as kinematic redundancy.
+Kinematic redundancy refers to having more joints (i.e., degrees of freedom - DOF) than required to perform a task, thus allowing the robot to adopt multiple configurations to accomplish the same end-effector position and orientation.
+This redundancy provides a considerable advantage, making it possible for manipulators to optimize movement according to different criteria, such as energy efficiency, obstacle avoidance, or joint limit avoidance.
+
+In this research, we specifically consider a 7-DOF Panda Emika manipulator.
+When controlling both the position and orientation of the end-effector in a 3D space, which requires six degrees of freedom (three for position and three for orientation), a 7-DOF manipulator such as the Panda Emika has one redundant degree.
+However, when controlling only the position, which requires three degrees of freedom, the manipulator possesses four redundant degrees.
+This redundancy grants the manipulator an essential characteristic: an infinite number of possible joint configurations to reach a specific end-effector (EE) point in space.
+
+When dealing with redundant manipulators, task prioritization is crucial.
+In our approach, we distinguish between primary and secondary tasks.
+The primary task of the manipulator is usually defined as reaching a desired EE position and orientation or as following a trajectory made out of such points.
+In this work we instead use APF (Artificial Potential Field), to guide the EE towards the goal position, while avoiding the obstacles in the path of the EE.
+Operating in a real-world environment often requires considering secondary tasks, such as real-time obstacle avoidance, to ensure safe and uninterrupted operation of the manipulator.
+
+This paper proposes a novel method for real-time obstacle avoidance in the Panda Emika 7-DOF manipulator.
+We use an obstacle grid combined with convolution techniques to assign avoidance directions to each manipulator segment, ensuring efficient and safe maneuvering within complex environments.
+
+\todocomment{Dodaj opis poglavij.}
 
 \section{Background and Related Work}
+
+The execution of the secondary task of obstacle avoidance necessitates an understanding of the spatial relationship between the manipulator and surrounding obstacles, usually quantified as a distance or interpreted as a repulsive force.
+By conceptualizing obstacles as sources of repulsive forces, the manipulator can be guided away from potential collisions, akin to how two magnets of the same polarity repel each other.
+
+Two commonly employed methods for obstacle avoidance include potential field methods and artificial neural networks.
+In potential field methods, the robot and obstacles are treated as charged entities.
+The robot is attracted to the target (goal) while being repelled by obstacles, resulting in a potential field.
+The robot navigates this field, moving along the path of steepest descent.
+While this method is intuitive and relatively easy to implement, it can sometimes result in local minima problems where the robot gets stuck in a position that is not the target but cannot find a path due to the repelling obstacles.
+
+Artificial neural networks (ANNs) are another prevalent approach, offering potential solutions to the local minima problem.
+By learning to map sensor readings to appropriate actions, ANNs can be trained to successfully navigate complex environments.
+However, they require extensive training and can be computically intensive.
+
+Our proposed approach seeks to combine the strengths of these methods, utilizing the obstacle grid and convolution techniques to efficiently calculate repulsive forces, thus enabling real-time obstacle avoidance while mitigating issues commonly associated with traditional methods.
+
+\todocomment{Preveri pravilnost informacij, dodaj vire, ki potrjujejo informacije.}
+
+\todocomment{preveri podobnost naše metode z metodo: Vector Field Histogram (VFH) }
 
@@
 
+\subsection{Attractive Field}
+
+APF for EE task.
+
 \subsection{Convolution Techniques}
+
+Sobel-like directional kernels.
 
@@
 
+Null space.
+
 \section{Experiment and Results}
+
+The wall visualization.
+
+Line field strenght dx,dy,dz
 

### 2023-07-20 Articles/ERK/erk.tex
.
ERK Methodology

@@
 \usepackage{times}
+\usepackage{wrapfig}
 \usepackage{graphicx}
@@
 
-\newcommand\todocomment[1]{\textcolor{red}{||TODO:\\ #1\\||}}
+\usepackage{easyReview} % for \comment, \alert, etc.
+
+% to-do command
+\newcommand\todocomment[1]{\textcolor{red}{||\\ #1\\||}}
 
@@
 
+\todocomment{IMAGE OF ROBOT}
+
 This paper proposes a novel method for real-time obstacle avoidance in the Panda Emika 7-DOF manipulator.
@@
 \section{Background and Related Work}
+
+%The execution of the secondary task of obstacle avoidance necessitates an understanding of the spatial relationship between the manipulator and surrounding obstacles, usually quantified as a distance or interpreted as a repulsive force.
+By conceptualizing obstacles as sources of repulsive forces, the manipulator can be guided away from potential collisions, akin to how two magnets of the same polarity repel each other.
+%
+%Two commonly employed methods for obstacle avoidance include potential field methods and artificial neural networks.
+In potential field methods, the robot and obstacles are treated as charged entities.
+The robot is attracted to the target (goal) while being repelled by obstacles, resulting in a potential field.
+The robot navigates this field, moving along the path of steepest descent.
+While this method is intuitive and relatively easy to implement, it can sometimes result in local minima problems where the robot gets stuck in a position that is not the target but cannot find a path due to the repelling obstacles.
+%
+%Artificial neural networks (ANNs) are another prevalent approach, offering potential solutions to the local minima problem.
+By learning to map sensor readings to appropriate actions, ANNs can be trained to successfully navigate complex environments.
+However, they require extensive training and can be computically intensive.
+%
+%Our proposed approach seeks to combine the strengths of these methods, utilizing the obstacle grid and convolution techniques to efficiently calculate repulsive forces, thus enabling real-time obstacle avoidance while mitigating issues commonly associated with traditional methods.
 
@@
 
-Our proposed approach seeks to combine the strengths of these methods, utilizing the obstacle grid and convolution techniques to efficiently calculate repulsive forces, thus enabling real-time obstacle avoidance while mitigating issues commonly associated with traditional methods.
+An advanced version of ANNs in the context of 3D obstacle detection and avoidance is the use of 3D Convolutional Neural Networks (3D-CNNs).
+While not strictly a path planning method, 3D-CNNs can be trained to recognize obstacles and free space in 3D occupancy data.
+The convolutional layers in the network can learn spatial hierarchies and structures, and the final layers can output a direction or action to take.
+These methods, although effective, require extensive training data and computational resources, however it is capable to recognizing and avoiding some local minima.
+
+The Vector Field Histogram (VFH) method presents another approach.
+This method uses a 2D histogram grid as a spatial representation of the environment.
+The histogram cells are updated based on the proximity and angle of obstacles.
+Then a primary polar histogram is calculated from the grid cells, leading to a candidate valley for navigation.
+While originally developed for 2D, this method was later extended to 3D as VFH+.
+
+Our proposed approach uses a 3D occupancy or voxel grid and 3D directional kernels, akin to Sobel filters, to perform kernel convolution and calculate the direction in which to move the robot arm to avoid obstacles.
+This method marries the benefits of convolutional approaches with the intuitive spatial reasoning provided by a voxel grid representation.
+
+The primary strength of this method is its ability to swiftly calculate directional guidance for the robot arm in real-time.
+By using a voxel grid, we can efficiently represent the 3D environment, and the use of 3D directional kernels allows for robust calculations of obstacle proximity and directions of free space.
+This is beneficial in dynamic environments where quick responses are required.
+Moreover, this approach does not require extensive training as with neural networks, making it easier to deploy in new environments.
 
@@
 
+\subsection{Kinematics}
+
+The manipulation capabilities of a robot are fundamentally determined by its kinematics, the mathematical description of motion without considering the forces that cause it.
+In our experiments, we employed a 7 degree-of-freedom (DOF) robot.
+The seven joints provide a high level of versatility, enabling the robot to move and orient its end-effector in a broad range of positions and directions.
+To calculate the movements and positions of the robot, we use both geometric and differential kinematics.
+
+Geometric kinematics (also known as forward and inverse kinematics) focuses on the relationship between joint angles and the position and orientation of the end-effector.
+Forward kinematics involves calculating the position and orientation of the end-effector given the joint angles, while inverse kinematics is the process of determining the joint angles for a desired end-effector position and orientation.
+In the case or redundant manipulator, there isn't an unique mapping from EE coordinates to the joint coordinates.
+as there are exist an infinite number of different joint angles that can map to the same EE position.
+
+Differential kinematics, also known as the Jacobian analysis, forms the foundation of robot velocity analysis.
+The Jacobian matrix, often represented as $J$, encapsulates the relationship between the joint velocities and the end-effector linear and angular velocities.
+It's derived from the partial derivatives of the forward kinematics.
+
+The Jacobian matrix is a key element in the control of robotic manipulators.
+It can also be used for computing the inverse kinematics problem.
+However, for a 7-DOF manipulator, the Jacobian matrix is not square and its inverse does not exist in a traditional sense.
+In such cases, we resort to the pseudo-inverse of the Jacobian, calculated using the Moore-Penrose method.
+The Moore-Penrose pseudo-inverse, denoted as $J^+$, provides an approximation of the inverse for non-square matrices.
+
+The null space of the Jacobian corresponds to those joint motions which produce no movement in the end-effector.
+The primary task for our robot is the desired end-effector movement.
+The secondary task of obstacle avoidance in our research takes advantage of this null-space motion.
+Our method applies a repulsive velocity from detected obstacles in the environment.
+We calculate this velocities at certain points that are positioned on the manipulator arm segments.
+The robot arm is thus able to carry out the primary task while simultaneously avoiding obstacles by appropriately utilizing its redundant DOF.
+
+\todocomment{probably needs shortening \\
+			 add equations}
+
 \subsection{Obstacle Grid}
+
+In order to safely and efficiently navigate a robot through a complex environment, it is essential to have an appropriate representation of the robots work space.
+In our approach, we achieve 3D spatial understanding of the robot's surrounding environment by employing an obstacles grid, a method that transforms the environment into a three-dimensional grid of voxels (volumetric pixels).
+
+Each voxel within this grid corresponds to a specific volume in real-world space and contains information about the occupancy of that volume.
+This provides a structured and spatially efficient representation of the 3D environment, including the locations and extents of any obstacles present.
+This 3D voxel grid can be built from sensor data, such as that obtained from a LiDAR, 3D camera, or depth sensor.
+Each reading from the sensor provides a point in 3D space, and these points are used to update the occupancy probability of the corresponding voxels in the grid.
+
+\subsection{Avoidance velocities}
+
+\todocomment{IMAGE OF FILTERS}
+
+In our approach to robot arm obstacle avoidance, we apply a method of kernel convolution in conjunction with geometric transformations.
+This methodology allows us to calculate the necessary avoidance velocities for specific points of interest on the manipulator.
+
+The process initiates with kinematic transformations which are used to map the points of interest from the robot's configuration space to the cartesian space and than onto the 3D voxel grid.
+These points, strategically placed on the manipulator, serve to evaluate potential collision threats with obstacles and subsequently assist in computing the necessary obstacle avoidance velocities.
+
+Once the mapping of these points to the voxel grid is completed, we proceed to employ a specialized kernel convolution method.
+This method is tailored to calculate the constituent parts of the avoidance velocity vector in the Cartesian space, specifically for the x, y, and z directions.
+For each of these directions, we apply a separate 3D convolutional kernel.
+
+For each direction, we utilize its corresponding kernel and extract a segment of the obstacle grid of matching size, centered at the point of interest.
+This leads to two equivalent-sized 3D matrices - one from the obstacle grid and the other being the kernel.
+We then conduct an element-wise multiplication operation between these matrices, summing the resulting values.
+The result of this operation provides us with the avoidance velocity in each respective direction.
+Through this approach, we can dynamically adjust the robot's trajectory to effectively avoid obstacles.
+
+\subsection{Kernel Convolution}
+
+Our convolutional kernels are essentially discrete differential operators in 3D space, and their design reflects this.
+These kernels bear a resemblance to the Sobel kernels, which are widely used in computer vision for edge detection and approximation of spatial derivatives.
+Each kernel has a primary axis aligned with the direction in which the avoidance velocity is being calculated, with the other two axes orthogonal to it.
+The kernel values along the primary axis exhibit a symmetric pattern: the highest value lies in the center of the kernel, with the absolute values decreasing uniformly on either side.
+This decrease simulates the derivative-like behavior of the kernel.
+
+Contrary to the design of a classical Sobel operator, our custom kernels don't contain zeroes along the primary axis.
+Instead, we assign the maximum kernel values along this line.
+This choice is specifically designed to meet our goal of obstacle avoidance.
+By doing so, when our points of interest on the robot manipulator coincide with an obstacle - effectively being 'right on top' of the obstacle in the voxel grid - the convolution operation yields a maximal repulsive velocity.
+This encourages the robot to move away from the obstacle and ensures safer navigation in the presence of potential collisions.
+
+However, the kernel values are not merely mirror images on each side of the kernel.
+While the absolute values are mirrored, the signed values are positive on one side and negative on the other.
+This sign difference is critical for calculating directional gradients.
+
+In addition to the kernel values along the primary axis, the values along the orthogonal axes mimic the same pattern.
+These are derived from the primary axis values, decrease proportionally with the radial distance from the primary axis, and crucially, they maintain the same sign as dictated by the derivative along the primary axis.
+This symmetry in both magnitude and sign around the main axis is a fundamental characteristic of our kernels.
+
+By applying these specialized kernels to the voxel grid at the locations of the points of interest, we compute the gradients that guide the robot arm's movements.
+This approach enables the arm to continue its primary task while dynamically avoiding any potential collisions with obstacles.
+
+The dimensions of the kernel, both along the primary axis and orthogonal to it, play crucial roles in determining the effective detection range and the nature of the repulsive avoidance velocity.
+
+The length of the kernel along the detection axis is dependent on the desired detection range for potential obstacles.
+Longer kernels can detect obstacles further away from the robot, essentially extending the 'safety zone' around the robot.
+Additionally, the rate at which the absolute values decrease when moving away from the center of the kernel along the detection axis determines the avoidance velocity profile as we move away from the obstacle.
+We have experimented with both linear and Gaussian decay profiles.
+The Gaussian profile, with its bell-curve distribution, generates an avoidance velocity that decreases exponentially as we move away from the obstacle, creating a softer transition compared to the linear profile.
+
+Orthogonal to the primary axis, the size of the kernel dictates the width of the detection area.
+In most instances, an equal width and height around the primary axis are desired, creating a symmetrical detection area.
+Narrow kernels detect a focused, 'beam-like' area directly ahead, while wider radial kernels can detect obstacles not exactly aligned with the primary axis but still within the potential path of the robot.
+
+To optimize obstacle detection, we have explored the idea of reducing the kernel values orthogonal to the primary axis with increasing distance from it.
+The intent here is to minimize the impact of distant, off-axis obstacles while still ensuring they are detected.
+However, as the ideal kernel configuration depends on the grid resolution and the specific environment, further testing is required to refine this approach.
 
@@
 
-\subsection{Convolution Techniques}
-
-Sobel-like directional kernels.
-
-\subsection{Robot Segments Control}
-
-Null space.
+\todocomment{- we use APF for primary task\\
+			 - attractive field calculating \\
+		 	 - repulsive field is same as avoidance velocities}
 
@@
 
-The wall visualization.
-
-Line field strenght dx,dy,dz
+\todocomment{-The wall visualization. \\
+			 -Line field strenght dx,dy,dz}
 

### 2023-07-20 Articles/ERK/erk.tex
.

@@
 \usepackage[slovene,english]{babel}
-
-\usepackage{easyReview} % for \comment, \alert, etc.
 
@@
 \begin{abstract}{Abstract}
-For the papers written in Slovene language an English title and abstract is required. The abstract should be no more than 200 words.
-
-We accept only electronic submissions in pdf (without password protection) or Microsoft Word format. Submission form can be found on ERK web page \cite{ERK}.
+\todocomment{ADD}
 \end{abstract}
@@
 
-APF for EE task.
-
 \todocomment{- we use APF for primary task\\

### 2023-07-20 Articles/ERK/erk.tex
ERK main axis scheme

@@
 \usepackage[top=22.5mm, bottom=22.5mm, left=22.5mm, right=22.5mm]{geometry}
+\usepackage{float}  % add this in your preamble (before \begin{document})
 
@@
 Each kernel has a primary axis aligned with the direction in which the avoidance velocity is being calculated, with the other two axes orthogonal to it.
-The kernel values along the primary axis exhibit a symmetric pattern: the highest value lies in the center of the kernel, with the absolute values decreasing uniformly on either side.
+The kernel values along the primary axis exhibit a symmetric pattern: the highest value lies in the center of the kernel, with the absolute values decreasing uniformly on either side, with negative polarity on one side.
 This decrease simulates the derivative-like behavior of the kernel.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.85\linewidth]{kernel-mainvalues.png}
+	\caption{The kernel values along the primary axis exhibit a pattern of symmetrical distribution with negative polarity on one side.} 
+	\label{Kernel main axis}
+\end{figure}
 
@@
 
+\todocomment{maybe add element-wise convolution equation}
+
 \subsection{Attractive Field}
+
+\todocomment{Tukaj bi lahko uporabil D* metodo. Ampak realno je za 4 strani snovi že dovolj.}
 

### 2023-07-21 Articles/ERK/erk.tex
ERK article
ERK added color main axis scheme
Update erk.tex
ERK minor
ERK citat 1

@@
 
-Robotic manipulators, specifically those with a high degree of freedom (DOF), possess a crucial characteristic known as kinematic redundancy.
+Robotic manipulators, specifically those with a high degree of freedom (DOF), possess a crucial characteristic known as kinematic redundancy~\cite{Nguyen1991}.
 Kinematic redundancy refers to having more joints (i.e., degrees of freedom - DOF) than required to perform a task, thus allowing the robot to adopt multiple configurations to accomplish the same end-effector position and orientation.
@@
 Forward kinematics involves calculating the position and orientation of the end-effector given the joint angles, while inverse kinematics is the process of determining the joint angles for a desired end-effector position and orientation.
-In the case or redundant manipulator, there isn't an unique mapping from EE coordinates to the joint coordinates.
-as there are exist an infinite number of different joint angles that can map to the same EE position.
+In the case or redundant manipulator, there isn't an unique mapping from EE coordinates to the joint coordinates as there exist an infinite number of different joint angles configurations that map to the same EE position.
 
-Differential kinematics, also known as the Jacobian analysis, forms the foundation of robot velocity analysis.
+Differential kinematics, also known as the robots Jacobian, forms the foundation of robot velocity analysis.
 The Jacobian matrix, often represented as $J$, encapsulates the relationship between the joint velocities and the end-effector linear and angular velocities.
@@
 
-\todocomment{IMAGE OF FILTERS}
-
 In our approach to robot arm obstacle avoidance, we apply a method of kernel convolution in conjunction with geometric transformations.
@@
 	\centering
-	\includegraphics[width=0.85\linewidth]{kernel-mainvalues.png}
+	\includegraphics[width=0.8\linewidth]{kernel-mainvalues-colored.png}
 	\caption{The kernel values along the primary axis exhibit a pattern of symmetrical distribution with negative polarity on one side.} 
@@
 This encourages the robot to move away from the obstacle and ensures safer navigation in the presence of potential collisions.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.85\linewidth]{kernel-matrix.png}
+	\caption{In most instances, an equal width and height around the primary axis are desired, creating a symmetrical detection area.} 
+	\label{Kernel matrix}
+\end{figure}
+
+\todocomment{while not the most important fix, currently values in matrix don't follow values used in experiments}
 
@@
 
+\bibitem{Nguyen1991} Nguyen, Charles C., Zhou, Zhen-Lei, Mosier, Gary E.
+(1991) Analysis and control of a kinematically redundant manipulator.
+\textit{Computers \& Electrical Engineering}, 17(3), pp.
+147–161.
+doi:10.1016/0045-7906(91)90031-T.
 \bibitem{ERK} ERK, http://www.ieee.si/erk/index.html 

### 2023-07-21 Articles/ERK/erk.tex
ERK citations

@@
 
-In this research, we specifically consider a 7-DOF Panda Emika manipulator.
+In this research, we specifically consider a 7-DOF Franka Panda Emika~\cite{Franka2021} manipulator.
 When controlling both the position and orientation of the end-effector in a 3D space, which requires six degrees of freedom (three for position and three for orientation), a 7-DOF manipulator such as the Panda Emika has one redundant degree.
@@
 
-This paper proposes a novel method for real-time obstacle avoidance in the Panda Emika 7-DOF manipulator.
+This paper proposes a novel method for real-time obstacle avoidance of the manipulator arm.
 We use an obstacle grid combined with convolution techniques to assign avoidance directions to each manipulator segment, ensuring efficient and safe maneuvering within complex environments.
@@
 
-Two commonly employed methods for obstacle avoidance include potential field methods and artificial neural networks.
+Two commonly employed methods for obstacle avoidance include artificial potential field methods (APF)~\cite{Khatib1985} and artificial neural networks.
 In potential field methods, the robot and obstacles are treated as charged entities.
@@
 
-An advanced version of ANNs in the context of 3D obstacle detection and avoidance is the use of 3D Convolutional Neural Networks (3D-CNNs).
+An advanced version of ANNs in the context of 3D obstacle detection and avoidance is the use of 3D Convolutional Neural Networks (3D-CNNs)~\cite{Duguleana2011}.
 While not strictly a path planning method, 3D-CNNs can be trained to recognize obstacles and free space in 3D occupancy data.
@@
 
-The Vector Field Histogram (VFH) method presents another approach.
+The Vector Field Histogram~\cite{Borenstein1991} (VFH) method presents another approach.
 This method uses a 2D histogram grid as a spatial representation of the environment.
@@
 Then a primary polar histogram is calculated from the grid cells, leading to a candidate valley for navigation.
-While originally developed for 2D, this method was later extended to 3D as VFH+.
+While originally developed for 2D, this method was later extended to 3D as VFH+~\cite{Ulrich1998}.
 
-Our proposed approach uses a 3D occupancy or voxel grid and 3D directional kernels, akin to Sobel filters, to perform kernel convolution and calculate the direction in which to move the robot arm to avoid obstacles.
+Our proposed approach uses a 3D occupancy or voxel grid and 3D directional kernels, akin to 2D Sobel filters, to perform kernel convolution and calculate the direction in which to move the robot arm to avoid obstacles.
 This method marries the benefits of convolutional approaches with the intuitive spatial reasoning provided by a voxel grid representation.
@@
 doi:10.1016/0045-7906(91)90031-T.
-\bibitem{ERK} ERK, http://www.ieee.si/erk/index.html 
-\bibitem{Zbornik} B. Zajc, A. Trost: Zbornik triindvajsete mednarodne Elektrotehniške in računalniške konference ERK 2014, 22. - 24. September 2014, Portorož, Slovenija
+
+\bibitem{Franka2021} Franka Emika. (2021). Panda: Specification. \textit{Available at: https://www.franka.de/panda}
+
+\bibitem{Khatib1985} Khatib, O.
+(1985) Real-time obstacle avoidance for manipulators and mobile robots.
+In: \textit{1985 IEEE International Conference on Robotics and Automation Proceedings}, 2, pp.
+500–505.
+doi:10.1109/ROBOT.1985.1087247.
+
+\bibitem{Duguleana2011} 
+M. Duguleana, F. G. Barbuceanu, A. Teirelbar, and G. Mogan,
+"Obstacle avoidance of redundant manipulators using neural networks based reinforcement learning,"
+\textit{Robotics and Computer-Integrated Manufacturing}, 
+DOI: 10.1016/j.rcim.2011.07.004, Sep. 2011.
+
+\bibitem{Borenstein1991} 
+J. Borenstein and Y. Koren,
+"The vector field histogram-fast obstacle avoidance for mobile robots,"
+\textit{IEEE Transactions on Robotics and Automation}, 
+vol. 7, no. 3, pp. 278–288, Jun. 1991.
+DOI: 10.1109/70.88137
+
+\bibitem{Ulrich1998} 
+I. Ulrich and J. Borenstein,
+"VFH+: reliable obstacle avoidance for fast mobile robots,"
+In \textit{Proceedings. 1998 IEEE International Conference on Robotics and Automation}, 
+vol. 2, pp. 1572–1577, May 1998.
+DOI: 10.1109/ROBOT.1998.677362
 

### 2023-07-21 Articles/ERK/erk.tex
ERK kinematic equations

@@
 
-\todocomment{Preveri pravilnost informacij, dodaj vire, ki potrjujejo informacije.}
-
-\todocomment{preveri podobnost naše metode z metodo: Vector Field Histogram (VFH) }
-
 \section{Methodology }
@@
 
+\begin{equation}
+	J^{\dagger} = (J^{T}J)^{-1}J^{T}
+	\label{eq:moore_penrose}
+\end{equation}
+
 The null space of the Jacobian corresponds to those joint motions which produce no movement in the end-effector.
@@
 
+\begin{equation}
+	N = (I - J^{\dagger}(q)J(q))
+	\label{eq:Null-Space}
+\end{equation}
+
+\begin{equation}
+	\dot{q} = J^{\dagger}(q) \dot{x}_{ee} + N J^{\dagger}_{poi} \dot{x}_{poi}
+	\label{eq:manipulator_kinematics}
+\end{equation}
+
 \todocomment{probably needs shortening \\
-			 add equations}
+			 add kinematics equations}
 
@@
 
+However, the kernel values are not merely mirror images on each side of the kernel.
+While the absolute values are mirrored, the signed values are positive on one side and negative on the other.
+This sign difference is critical for calculating directional gradients.
+
 \begin{figure}[H]
@@
 \todocomment{while not the most important fix, currently values in matrix don't follow values used in experiments}
-
-However, the kernel values are not merely mirror images on each side of the kernel.
-While the absolute values are mirrored, the signed values are positive on one side and negative on the other.
-This sign difference is critical for calculating directional gradients.
 
@@
 
-\todocomment{maybe add element-wise convolution equation}
+\todocomment{add element-wise convolution equation}
 
 \subsection{Attractive Field}
-
-\todocomment{Tukaj bi lahko uporabil D* metodo. Ampak realno je za 4 strani snovi že dovolj.}
 
@@
 
-\todocomment{-The wall visualization. \\
-			 -Line field strenght dx,dy,dz}
+\todocomment{- The wall visualization with robot and wall and "dy beam" and dy arrow. \\
+			 - Line field strenght 2D plot dx,dy,dz \\
+		 	 - }
 

### 2023-07-21 Articles/ERK/erk.tex
ERK convolution equations

@@
 
+Damping, represented by the term $k_{damp}$ in the Moore-Penrose pseudo-inverse equation, is used to introduce stability and prevent potential numerical issues when computing the pseudo-inverse of the Jacobian for non-square matrices.
+It acts as a regularization term to ensure a well-conditioned solution.
+
 \begin{equation}   	
-	J^{\dagger} = (J^{T}J)^{-1}J^{T}
+	J^{\dagger} = (J^{T}J + k_{damp} ~I)^{-1}J^{T}
 	\label{eq:moore_penrose}
@@
 
+\todocomment{cite Tadej?}
+
 \begin{equation}
@@
 \end{equation}
+
+\todocomment{add references to equations in text}
 
@@
 
-\subsection{Avoidance velocities}
+%\subsection{Avoidance velocities}
 
-In our approach to robot arm obstacle avoidance, we apply a method of kernel convolution in conjunction with geometric transformations.
+%In our approach to robot arm obstacle avoidance, we apply a method of kernel convolution in conjunction with geometric transformations.
 This methodology allows us to calculate the necessary avoidance velocities for specific points of interest on the manipulator.
-
-The process initiates with kinematic transformations which are used to map the points of interest from the robot's configuration space to the cartesian space and than onto the 3D voxel grid.
+%
+%The process initiates with kinematic transformations which are used to map the points of interest from the robot's configuration space to the cartesian space and than onto the 3D voxel grid.
 These points, strategically placed on the manipulator, serve to evaluate potential collision threats with obstacles and subsequently assist in computing the necessary obstacle avoidance velocities.
-
-Once the mapping of these points to the voxel grid is completed, we proceed to employ a specialized kernel convolution method.
+%
+%Once the mapping of these points to the voxel grid is completed, we proceed to employ a specialized kernel convolution method.
 This method is tailored to calculate the constituent parts of the avoidance velocity vector in the Cartesian space, specifically for the x, y, and z directions.
@@
 
-For each direction, we utilize its corresponding kernel and extract a segment of the obstacle grid of matching size, centered at the point of interest.
+%\todocomment{convolution equation}
+
+%For each direction, we utilize its corresponding kernel and extract a segment of the obstacle grid of matching size, centered at the point of interest.
 This leads to two equivalent-sized 3D matrices - one from the obstacle grid and the other being the kernel.
@@
 Through this approach, we can dynamically adjust the robot's trajectory to effectively avoid obstacles.
+
+\subsection{Avoidance velocities}
+
+In our approach to robot arm obstacle avoidance, we apply a method of kernel convolution in conjunction with geometric transformations.
+This methodology allows us to calculate the necessary avoidance velocities for specific points of interest located on the manipulator.
+
+The process commences with kinematic transformations, which facilitate the mapping of points of interest from the robot's configuration space to the Cartesian space, followed by their projection onto the 3D voxel grid.
+
+%These points, strategically placed on the manipulator, serve to evaluate potential collision threats with obstacles and subsequently assist in computing the necessary obstacle avoidance velocities.
+
+Once the mapping of these points to the voxel grid is completed, we proceed to employ a specialized kernel convolution method.
+This method is tailored to calculate the constituent parts of the avoidance velocity vector in the Cartesian space, specifically for the \(x\), \(y\), and \(z\) directions.
+For each direction, we generate the corresponding kernel and extract a segment of the obstacle grid of matching size, centered at the point of interest, resulting in two comparable 3D matrices—one from the obstacle grid $A_d$ and the other representing the kernel $K_d$.
+
+By employing the Hadamard (element-wise) product between the segment of the obstacle grid \(A_d\) and the corresponding 3D convolutional kernel \(K_d\) for direction \(d\), we derive the resultant matrix \(C_d\), which can be expressed as:
+
+\[
+C_d = A_d \odot K_d
+\]
+
+Subsequently, the avoidance velocity in the Cartesian coordinate system is obtained by summing all the values in the resulting matrix, which can be represented as:
+
+\[
+v = \sum_{i}\sum_{j}\sum_{k} a_{ijk} 
+\]
+
+%\[
+%v = \sum_{x}\sum_{y}\sum_{z} a_{xyz} ~c_{xyz}
+%\]
 
@@
 This decrease simulates the derivative-like behavior of the kernel.
+
+\todocomment{add references to images in text}
 

### 2023-07-22 Code/ObstacleGrid_RTConvolution_OnePoint.m
fixed constants

@@ -4,11 +4,18 @@ function [joints_positions, EE_positions, goal_distances, q_velocities, ee_veloc
 p = inputParser;
 addOptional(p, 'mid_joints', true); % Default value is false
 addOptional(p, 'avoid_task', true); % Default value is true
+addOptional(p, 'task_constants', [1 4 0.5]);
 
 parse(p, varargin{:});
 
+% select which goals
 mid_joints = p.Results.mid_joints;
 avoid_task = p.Results.avoid_task;
+
+% weights for different tasks
+wp = p.Results.task_constants(1); % primary task
+wm = p.Results.task_constants(2); % mid-joints task
+wa = p.Results.task_constants(3); % obstacle avoidance task
 
 % -----------------------------------------------------------
 
@@ -30,11 +37,6 @@ q_range = [2.8973 -2.8973;
 
 % number of points per segment for obstacle avoidance task
 points_per_segment = 1*[1 1 1 1 1 1 1];
-
-% weights for different tasks
-wp = 5; % primary task
-wm = 5; % mid-joints task
-wa = 0.5; % obstacle avoidance task
 
 % -----------------------------------------------------------
 

### 2023-07-24 Code/ObstacleGrid_RTConvolution_OnePoint.m
ERK plots

@@ -61,7 +61,7 @@ values_APF = [];
 ee_velocities = [];
 
 % set directional kernels
-sigma = 41;
+sigma = 11;
 
 kernel_length = 61;
 kernel_width = 9;
@@ -149,32 +149,33 @@ while current_dist > goal_dist && Niter < Nmax
 
     % APF Avoidance Task
     % --------------------------------------------------
- 	if avoid_task
+    avoid_vel = [];
 
-        % GET JOINT 4 POSE
-        % --------------------
+    % GET JOINT 4 POSE
+    % --------------------
 
-        % select transformation for joint 4
-        robot_tree = transformations_list(4).tree;
+    % select transformation for joint 4
+    robot_tree = transformations_list(4).tree;
+
+    % set configuration to a robot tree
+    config = homeConfiguration(robot_tree); % Create an empty struct
     
-        % set configuration to a robot tree
-        config = homeConfiguration(robot_tree); % Create an empty struct
+    % Assign values to joints, there will be 1-7 joints for Panda
+    for i = 1:1:robot_tree.NumBodies
+        config(i).JointPosition  = q(i); % set i-th joint
+    end
+
+    % get transform
+    transform = getTransform(robot_tree,config,'body'+string(robot_tree.NumBodies));
+
+    % add base transformation
+    transform = Tbase * transform;
+
+    % get x,y,z values
+    xyz = ceil(transform(1:3,4)*grid.resolution);
         
-        % Assign values to joints, there will be 1-7 joints for Panda
-        for i = 1:1:robot_tree.NumBodies
-            config(i).JointPosition  = q(i); % set i-th joint
-        end
-    
-        % get transform
-        transform = getTransform(robot_tree,config,'body'+string(robot_tree.NumBodies));
-    
-        % add base transformation
-        transform = Tbase * transform;
-    
-        % get x,y,z values
-        xyz = ceil(transform(1:3,4)*grid.resolution);
-        
-        
+    if avoid_task
+
         % GET VELOCITIES USING DIRECTIONAL KERNELS
         % --------------------
 
@@ -228,6 +229,7 @@ while current_dist > goal_dist && Niter < Nmax
         % calculate avoidance joints velocities
 %         q_vel = q_vel + pinv_J0*(avoid_vel - J0*pinv_J*ee_vel);
 %         q_vel = q_vel + pinv_J0*(avoid_vel);
+
 
     % APPROXIMATE SOLUTION
     % -----------------------

### 2023-07-24 Articles/ERK/erk.tex
ERK equations format
fixed constants
ERK todocomments
ERK plots
ERK more plots 7-DOF_Robot_Positions_Standard_vs_Avoidance.eps

@@
 
-\todocomment{IMAGE OF ROBOT}
+\todocomment{IMAGE OF THE TASK}
 
@@
 
+\todocomment{check if "exact equation" gives better results}
+
 \todocomment{add references to equations in text}
@@
 			 add kinematics equations}
+
+\todocomment{želena sredinska lega sklepov}
 
@@
 
-\[
+\begin{equation}
 	C_d = A_d \odot K_d
-\]
+	\label{eq:hadamard}
+\end{equation}
 
@@
 
-\[
+\begin{equation}
 v = \sum_{i}\sum_{j}\sum_{k} a_{ijk} 
-\]
+\label{summation}
+\end{equation}
 
@@
 
-\subsection{Kernel Convolution}
+\subsection{Convolution Kernel}
 
@@
 
-\todocomment{add element-wise convolution equation}
+\subsection{Attractive Field}
 
-\subsection{Attractive Field}
+In our end-effector motion planning approach, we leverage derivatives of the attractive field to facilitate efficient trajectory generation.
+Prior to commencing motion planning, we construct a distance matrix, analogous in size to the matrix representing the obstacles.
+This distance matrix encompasses distance values from our designated goal point, corresponding to each voxel in the workspace.
+Subsequently, we employ interpolation techniques to estimate the value at the current end-effector (EE) point by considering the distance field values of the eight neighboring voxels and the distances of the current EE pose relative to these eight values.
+By performing interpolation between two points, we effectively calculate the numerical derivative of distance change, thereby obtaining the EE velocity.
+This approach enables us to navigate the robotic manipulator smoothly towards the desired goal while considering the surrounding obstacle space.
+
+Additionally, we can augment our velocity component with repulsive fields, which are calculated using the convolutional kernel technique.
+This enables us to guide the end-effector around obstacles smoothly, ensuring safe and collision-free navigation while dynamically adapting to the environment.
+By integrating both the attractive and repulsive field contributions, our motion planning approach ensures efficient and effective trajectory generation, enabling the robotic manipulator to reach its desired goal while avoiding potential obstacles in the workspace.
 
@@
 \todocomment{- The wall visualization with robot and wall and "dy beam" and dy arrow. \\
-			 - Line field strenght 2D plot dx,dy,dz \\
-		 	 - }
+				(with and without avoidance 3 to 5 keypoints) \\
+		 	  }
 
-\section{Pisanje prispevka}
-
-Pri pisanju uporabljajte sloge, ki so definirani v datoteki erk.sty in ne spreminjajte velikosti ali sloga pisave. Ta dokument lahko vzamete kot vzorec za oblikovanje prispevka v orodju \LaTeX.
-Tabela \ref{tab1} povzema osnovne sloge in velikosti pisave.
-
-\begin{table}[h]
-\caption{Osnovni slogi pri oblikovanju prispevka.} \label{tab1}
-\smallskip
-\begin{center}
-\begin{tabular}{ | r | c | c | }
-\hline  
-  \textbf{Text Style} & \textbf{Font Size} & \textbf{Attributes}\\ 
-\hline  
-  Paper title & 17pt & bold\\
-  Author names & 12pt & bold\\
-  Affiliation & 10pt & italic\\
-  Heading & 12pt & bold \\
-  Subheading & 10pt & bold\\
-  Regular text & 10pt &\\
-  Captions \& ref. & 9pt &\\
-\hline  
-\end{tabular}
-\end{center}
-\end{table}
-
-\subsection{Enačbe in slike}
-
-Enačbe in formule se pišejo s poševnimi črkami in so zaporedno oštevilčene. V besedilu naj bodo reference na enačbe zapisane v oklepajih (\ref{eq1}).
-
-\begin{equation}
- \int^{r_{2}}_{0}F(r,\varphi)\; dr\; d\varphi= [\sigma r_{2} / (2\mu_{0})]
-    \label{eq1}
-\end{equation}
-
-Prispevki za ERK se ne tiskajo v barvah, zato pripravite črno-bele slike.
-Slike naj bodo postavljene v okvirje širine enega ali dveh stolpcev in označene s številko in podnapisom.
-V besedilu morajo biti reference na slike (slika \ref{slika}).
-
-\begin{figure}[!htb]
-    \begin{center}
-        \includegraphics[scale=1]{field1.eps}
-        \caption{Podajte kratko razlago slike.} \label{slika}
-    \end{center}
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=1.0\linewidth]{Repulsive-Velocities-Along-the-Y-Axis-wide.eps}
+	\caption{Repulsive Velocities Along the Y Axis}
 \end{figure}
 
-\section{Oddaja prispevka}
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=1.0\linewidth]{Y-Positions.eps}
+	\caption{Repulsive Velocities Along the Y Axis}
+\end{figure}
 
-Prispevek oddajte v elektronski obliki preko formularja na spletni strani \cite{ERK}.
-Sprejemamo prispevke v formatu pdf (brez nastavljenih zaščitnih gesel) ali Microsoft Word.
-Pred oddajo prispevka prosimo preverite dodatne informacije za avtorje, ki so na spletnih straneh.
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[width=1.0\linewidth]{7-DOF_Robot_Positions_Standard_vs_Avoidance.eps}
+	\caption{Repulsive Velocities Along the Y Axis}
+\end{figure}
 
-Formular za elektronsko oddajo prispevkov bo aktiviran v juniju, rok za oddajo je \textbf{13. julij 2014}. Avtorje bomo obvestili o izboru konec avgusta.
+\todocomment{Maybe add joint positions plot, however might me too much}
+
+\section{Conclusions and further development}
+
+\todocomment{ADD}
 

### 2023-07-24 Articles/ERK/erk.tex
ERK added results text

@@
 \usepackage[top=22.5mm, bottom=22.5mm, left=22.5mm, right=22.5mm]{geometry}
+
 \usepackage{float}  % add this in your preamble (before \begin{document})
@@
 The primary task for our robot is the desired end-effector movement.
-The secondary task of obstacle avoidance in our research takes advantage of this null-space motion.
+The secondary task of obstacle avoidance in our research takes advantage of this null-space motion~\cite{lajpah_Petri_2012}.
 Our method applies a repulsive velocity from detected obstacles in the environment.
@@
 \todocomment{add references to equations in text}
-
-\todocomment{probably needs shortening \\
-			 add kinematics equations}
 
@@
 
-\todocomment{- we use APF for primary task\\
-			 - attractive field calculating \\
-			 	 - repulsive field is same as avoidance velocities}
+\todocomment{ - attractive field calculating equation }
 
 \section{Experiment and Results}
-
-\todocomment{- The wall visualization with robot and wall and "dy beam" and dy arrow. \\
-				(with and without avoidance 3 to 5 keypoints) \\
-		 	  }
 
@@
 	\includegraphics[width=1.0\linewidth]{Repulsive-Velocities-Along-the-Y-Axis-wide.eps}
-	\caption{Repulsive Velocities Along the Y Axis}
+	\caption{Plot depicting avoidance velocities (x, y, z) as calculated by our kernel along the y-axis, showing the robot's adaptive movement away from the wall obstacle.}
 \end{figure}
@@
 	\includegraphics[width=1.0\linewidth]{Y-Positions.eps}
-	\caption{Repulsive Velocities Along the Y Axis}
+	\caption{Comparative visualization of the y-position of the robot's fourth joint with and without the secondary avoidance task, highlighting the impact of the repulsive force.}
 \end{figure}
@@
 	\includegraphics[width=1.0\linewidth]{7-DOF_Robot_Positions_Standard_vs_Avoidance.eps}
-	\caption{Repulsive Velocities Along the Y Axis}
+	\caption{3D model displaying the final position of the robot, contrasting the scenarios with and without the implementation of joint 4 obstacle avoidance.}
 \end{figure}
 
-\todocomment{Maybe add joint positions plot, however might me too much}
+Our proposed algorithm was tested using a kinematic model of the 7 Degrees of Freedom (DoF) Panda Emika robot, we implemented within the MATLAB environment.
+This robot was selected for its operational versatility, thus serving to underscore the robustness and flexibility of our approach.
+
+The end-effector's motion was controlled using the attractive field method, serving as the primary task.
+This was complemented by the repulsive field method implemented as a secondary task, with the specific aim of maneuvering a point on the fourth joint of the robot as far away from the obstacle - in this case, a simple wall - as possible.
+
+To compute the joint positions, we used joint velocities derived through inverse kinematics, incorporating the aforementioned secondary tasks.
+We then applied numerical integration to these velocities to model the robot's motion in a simulated environment.
+
+An additional secondary task was also introduced, aimed at ensuring the manipulator's mid-joints remained in an upright position.
+This secondary task, aimed at maintaining the robot's "elbow" at the top, further reinforced the robot's upright posture while executing the primary task and the obstacle avoidance task.
+
+Figure 3 illustrates the avoidance velocities in the x, y, and z directions, computed by our kernel, when its center aligns with various points along the y-axis.
+Here, the y-coordinates are presented on the x-axis, and the corresponding velocities are plotted on the y-axis.
+As expected, the y-velocity is the only component exhibiting non-zero values.
+It effectively propels the robot away from the wall, moving in either a positive or negative direction, depending on the relative position to the wall.
+Conversely, the x and z velocities remain zero, reflecting the absence of obstacles along these axes.
+This aligns with our expectations given the presence of the wall obstacle is only along the y direction.
+
+In Figure 4, we observe the variation in the y-position of the fourth joint, both with and without the implementation of the secondary avoidance task.
+The corresponding plot beneath the image illustrates the diminishing repulsive force as the robot moves further away from the wall obstacle.
+The rate and function of this decrement can be modulated by altering the dimensions and values within the kernel.
+
+Figure 5 presents a 3D model depicting the final position of the robot, showcasing both scenarios—with the implementation of the joint 4 obstacle avoidance (illustrated in green) and without it (depicted in red).
+
+The figure also highlights how the application of the avoidance algorithm prevents the fourth joint from moving closer to the wall, as compared to when the avoidance algorithm is not employed.
+The position of the joint, upon task completion, is noted to be approximately 20 cm further away from the wall.
+
+Although these preliminary results demonstrate the efficacy of our proposed approach, further testing of the proposed method is needed to analyse  its adaptability in more complex operational environments.
 
@@
 
+\bibitem{lajpah_Petri_2012}
+\textit{Obstacle Avoidance for Redundant Manipulators as Control Problem},
+Leon Lajpah and Tadej Petri,
+In: \textit{Serial and Parallel Robot Manipulators - Kinematics, Dynamics, Control and Optimization},
+edited by Serdar Kucuk,
+InTech, March 2012,
+ISBN: 978-953-51-0437-7,
+DOI: 10.5772/32651,
+Language: English.
+
 \end{thebibliography}

### 2023-07-25 Articles/ERK/erk.tex
ERK conclusion, abstract

@@
 \begin{abstract}{Abstract}
-\todocomment{ADD}
+
+This study proposes a method for real-time obstacle avoidance in robotic arm control, leveraging a 3D voxel grid and 3D directional kernels.
+The approach integrates the benefits of convolutional techniques with the spatial understanding provided by voxel grid representation to efficiently calculate movement directions while evading obstacles.
+By focusing on kinematics and obstacle grid representation, the method efficiently utilizes the versatility of a 7-degree-of-freedom (DOF) robot arm, employing geometric and differential kinematics for movement calculations.
+The strategy doesn't require exhaustive training as with neural networks, making it deployable in various environments.
+
 \end{abstract}
@@
 Operating in a real-world environment often requires considering secondary tasks, such as real-time obstacle avoidance, to ensure safe and uninterrupted operation of the manipulator.
-
-\todocomment{IMAGE OF THE TASK}
 
@@
 
-The primary strength of this method is its ability to swiftly calculate directional guidance for the robot arm in real-time.
+The primary strength of the proposed method is its ability to swiftly calculate directional guidance for the robot arm in real-time.
 By using a voxel grid, we can efficiently represent the 3D environment, and the use of 3D directional kernels allows for robust calculations of obstacle proximity and directions of free space.
@@
 \end{equation}
-
-%\[
-%v = \sum_{x}\sum_{y}\sum_{z} a_{xyz} ~c_{xyz}
-%\]
 
@@
 
-\todocomment{ADD}
+%This paper presented a novel approach to robotic motion planning, combining the application of kernel convolution for obstacle avoidance velocity calculations with an integration of attractive and repulsive field velocities for end-effector trajectory generation and robot joint obstacle avoidance.
+The method was validated using the kinematic model of a 7 DoF Panda Emika robot, demonstrating its operational versatility and adaptability in varying workspace conditions.
+%
+%Notable advantages of our approach include its simplicity and the fact that we work directly in the obstacle space grid, negating the necessity for calculating the spatial primitives of shapes.
+Moreover, when implemented in parallel, our method exhibits high computational efficiency, making it a viable solution for real-time applications.
+The experiments conducted effectively showcased that the end-effector could move efficiently towards the target while successfully avoiding obstacles, specifically, a wall in this instance.
+%
+%Our results affirmed the robustness of the proposed approach, as evidenced by the robot's fourth joint being approximately 20 cm further from the wall upon task completion when the avoidance algorithm was employed.
+%
+%Despite the promising initial results, our approach does present some limitations.
+Particularly, the possibility of encountering local minima traps poses a significant challenge.
+Although these traps could be potentially mitigated by incorporating multiple convolutional kernels of varying sizes, it would result in an approach similar to convolutional neural networks (CNNs), adding to the complexity and computational cost of the algorithm.
+%
+%Additionally, we acknowledge that our study was conducted under simplified conditions with a single type of obstacle.
+As such, future work should focus on validating the proposed method in more complex environments, with different obstacle types, shapes, and densities, to further ascertain its adaptability and robustness.
+It may also be beneficial to explore the integration of other optimization strategies and machine learning techniques to further enhance the overall efficiency and effectiveness of the trajectory planning process.
+%
+%In conclusion, our work serves as a solid foundation for the development of more advanced and dynamic motion planning algorithms.
+
+This paper presented a novel approach to robotic motion planning, combining the application of kernel convolution for obstacle avoidance velocity calculations with an integration of attractive and repulsive field velocities for end-effector trajectory generation and robot joint obstacle avoidance.
+The method was validated using the kinematic model of a 7 DoF Panda Emika robot, demonstrating its operational versatility and adaptability in varying workspace conditions.
+
+Notable advantages of our approach include its simplicity and the fact that we work directly in the obstacle space grid, negating the necessity for calculating the spatial primitives of shapes.
+Moreover, when implemented in parallel, our method exhibits high computational efficiency, making it a viable solution for real-time applications.
+The experiments conducted effectively showcased that the end-effector could move efficiently towards the target while successfully avoiding obstacles, specifically, a wall in this instance.
+
+Our results affirmed the robustness of the proposed approach, as evidenced by the robot's fourth joint being approximately 20 cm further from the wall upon task completion when the avoidance algorithm was employed.
+
+Despite its effectiveness, the approach could potentially fall into local minima traps.
+This limitation could be addressed by incorporating multiple convolutional kernels, though this would increase complexity and resemble convolutional neural networks.
+
+Additionally, we acknowledge that our study was conducted under simplified conditions with a single type of obstacle.
+As such, future work should focus on validating the proposed method in more complex environments, with different obstacle types, shapes, and densities, to further ascertain its adaptability and robustness.
+It may also be beneficial to explore the integration of other optimization strategies and machine learning techniques to further enhance the overall efficiency and effectiveness of the trajectory planning process.
 

### 2023-07-27 Articles/ERK/erk.tex
ERK .
.
ERK references, eq descriptions

@@
 \usepackage{xcolor}
+% \usepackage{easyReview}
 \usepackage[top=22.5mm, bottom=22.5mm, left=22.5mm, right=22.5mm]{geometry}
@@
 
-\todocomment{Dodaj opis poglavij.}
-
 \section{Background and Related Work}
@@
 By conceptualizing obstacles as sources of repulsive forces, the manipulator can be guided away from potential collisions, akin to how two magnets of the same polarity repel each other.
+
+\todocomment{check that repulsive - avoidance is well changed (READ TEXT)}
 
@@
 
-\section{Methodology }
+\section{Methodology}
 
@@
 The Jacobian matrix is a key element in the control of robotic manipulators.
-It can also be used for computing the inverse kinematics problem.
+It can also be used for computing the inverse differential kinematics problem.
 However, for a 7-DOF manipulator, the Jacobian matrix is not square and its inverse does not exist in a traditional sense.
-In such cases, we resort to the pseudo-inverse of the Jacobian, calculated using the Moore-Penrose method.
+In such cases, we resort to the pseudo-inverse of the Jacobian, calculated using the Moore-Penrose method (eq.~\ref{eq:moore_penrose}).
 The Moore-Penrose pseudo-inverse, denoted as $J^+$, provides an approximation of the inverse for non-square matrices.
 
-Damping, represented by the term $k_{damp}$ in the Moore-Penrose pseudo-inverse equation, is used to introduce stability and prevent potential numerical issues when computing the pseudo-inverse of the Jacobian for non-square matrices.
+Damping, represented by the constant term $k_{damp}$ in the Moore-Penrose pseudo-inverse equation, is used to introduce stability and prevent potential numerical issues when computing the pseudo-inverse of the Jacobian for non-square matrices.
 It acts as a regularization term to ensure a well-conditioned solution.
@@
 
-The null space of the Jacobian corresponds to those joint motions which produce no movement in the end-effector.
+The null space (eq.~\ref{eq:Null-Space}) of the Jacobian corresponds to those joint motions which produce no movement in the end-effector.
 The primary task for our robot is the desired end-effector movement.
 The secondary task of obstacle avoidance in our research takes advantage of this null-space motion~\cite{lajpah_Petri_2012}.
-Our method applies a repulsive velocity from detected obstacles in the environment.
-We calculate this velocities at certain points that are positioned on the manipulator arm segments.
-The robot arm is thus able to carry out the primary task while simultaneously avoiding obstacles by appropriately utilizing its redundant DOF.
-
-\todocomment{cite Tadej?}
+Our method applies an avoidance velocity from detected obstacles in the environment.
+We calculate this velocities at certain points of interest (POI) that are positioned on the manipulator arm segments.
+We transform avoidance velocities into the joint space velocities using $J^{\dagger}_{poi}$ jacobian calculated in the specific POI.
+These velocities we than transform into the null-space using matrix $N$.
+The robot arm is thus able to carry out the primary task while simultaneously avoiding obstacles by appropriately utilizing its redundant DOF (eq.~\ref{eq:manipulator_kinematics}).
 
@@
 \end{equation}
-
-\todocomment{check if "exact equation" gives better results}
-
-\todocomment{add references to equations in text}
 
@@
 
-By employing the Hadamard (element-wise) product between the segment of the obstacle grid \(A_d\) and the corresponding 3D convolutional kernel \(K_d\) for direction \(d\), we derive the resultant matrix \(C_d\), which can be expressed as:
+By employing the Hadamard (element-wise) product (eq.~\ref{eq:hadamard}) between the cutout segment of the obstacle grid \(A_d\) and the corresponding 3D convolutional kernel \(K_d\) for direction \(d\), we derive the resultant matrix \(C_d\), which can be expressed as:
 
@@
 
-Subsequently, the avoidance velocity in the Cartesian coordinate system is obtained by summing all the values in the resulting matrix, which can be represented as:
+Subsequently, the avoidance velocity in the Cartesian coordinate system is obtained by summing all the values in the resulting matrix (eq.~\ref{eq:summation}), which can be represented as:
 
 \begin{equation}
-v = \sum_{i}\sum_{j}\sum_{k} a_{ijk} 
-\label{summation}
+	\dot{x}_{poi} = \sum_{i}\sum_{j}\sum_{k} c_{dijk} 
+	\label{eq:summation}
 \end{equation}
@@
 Each kernel has a primary axis aligned with the direction in which the avoidance velocity is being calculated, with the other two axes orthogonal to it.
-The kernel values along the primary axis exhibit a symmetric pattern: the highest value lies in the center of the kernel, with the absolute values decreasing uniformly on either side, with negative polarity on one side.
+The kernel values along the primary axis exhibit a symmetric pattern (Figure~\ref{Kernel main axis}): the highest value lies in the center of the kernel, with the absolute values decreasing uniformly on either side, with negative polarity on one side.
 This decrease simulates the derivative-like behavior of the kernel.
-
-\todocomment{add references to images in text}
 
@@
 This choice is specifically designed to meet our goal of obstacle avoidance.
-By doing so, when our points of interest on the robot manipulator coincide with an obstacle - effectively being 'right on top' of the obstacle in the voxel grid - the convolution operation yields a maximal repulsive velocity.
+By doing so, when our points of interest on the robot manipulator coincide with an obstacle - effectively being 'right on top' of the obstacle in the voxel grid - the convolution operation yields a maximal avoidance velocity.
 This encourages the robot to move away from the obstacle and ensures safer navigation in the presence of potential collisions.
@@
 
-\todocomment{while not the most important fix, currently values in matrix don't follow values used in experiments}
-
-In addition to the kernel values along the primary axis, the values along the orthogonal axes mimic the same pattern.
+In addition to the kernel values along the primary axis, the values along the orthogonal axes mimic the same pattern (Figure~\ref{Kernel matrix}).
 These are derived from the primary axis values, decrease proportionally with the radial distance from the primary axis, and crucially, they maintain the same sign as dictated by the derivative along the primary axis.
@@
 
-The dimensions of the kernel, both along the primary axis and orthogonal to it, play crucial roles in determining the effective detection range and the nature of the repulsive avoidance velocity.
+The dimensions of the kernel, both along the primary axis and orthogonal to it, play crucial roles in determining the effective detection range and the nature of the avoidance velocity.
 
@@
 
-Additionally, we can augment our velocity component with repulsive fields, which are calculated using the convolutional kernel technique.
+Additionally, we can augment our EE velocity with obstacles avoidance component, which is calculated using the convolutional kernel technique.
 This enables us to guide the end-effector around obstacles smoothly, ensuring safe and collision-free navigation while dynamically adapting to the environment.
-By integrating both the attractive and repulsive field contributions, our motion planning approach ensures efficient and effective trajectory generation, enabling the robotic manipulator to reach its desired goal while avoiding potential obstacles in the workspace.
+By integrating both the attractive and avoidance field contributions, our motion planning approach ensures efficient and effective trajectory generation, enabling the robotic manipulator to reach its desired goal while avoiding potential obstacles in the workspace.
 
@@
 	\caption{Plot depicting avoidance velocities (x, y, z) as calculated by our kernel along the y-axis, showing the robot's adaptive movement away from the wall obstacle.}
+	\label{y axis avoidance values}
 \end{figure}
@@
 	\includegraphics[width=1.0\linewidth]{Y-Positions.eps}
-	\caption{Comparative visualization of the y-position of the robot's fourth joint with and without the secondary avoidance task, highlighting the impact of the repulsive force.}
+	\caption{Comparative visualization of the y-position of the robot's fourth joint with and without the secondary avoidance task, highlighting the impact of the avoidance velocity component.}
+	\label{4 th joint plot}
 \end{figure}
@@
 	\caption{3D model displaying the final position of the robot, contrasting the scenarios with and without the implementation of joint 4 obstacle avoidance.}
+	\label{3D model}
 \end{figure}
@@
 The end-effector's motion was controlled using the attractive field method, serving as the primary task.
-This was complemented by the repulsive field method implemented as a secondary task, with the specific aim of maneuvering a point on the fourth joint of the robot as far away from the obstacle - in this case, a simple wall - as possible.
+This was complemented by the avoidance field method implemented as a secondary task, with the specific aim of maneuvering a point on the fourth joint of the robot as far away from the obstacle - in this case, a simple wall - as possible.
 
@@
 
-Figure 3 illustrates the avoidance velocities in the x, y, and z directions, computed by our kernel, when its center aligns with various points along the y-axis.
+Figure~\ref{y axis avoidance values} illustrates the avoidance velocities in the x, y, and z directions, computed by our kernel, when its center aligns with various points along the y-axis.
 Here, the y-coordinates are presented on the x-axis, and the corresponding velocities are plotted on the y-axis.
@@
 
-In Figure 4, we observe the variation in the y-position of the fourth joint, both with and without the implementation of the secondary avoidance task.
-The corresponding plot beneath the image illustrates the diminishing repulsive force as the robot moves further away from the wall obstacle.
+In Figure~\ref{4 th joint plot}, we observe the variation in the y-position of the fourth joint, both with and without the implementation of the secondary avoidance task.
+The corresponding plot beneath the image illustrates the diminishing avoidance velocity as the robot moves further away from the wall obstacle.
 The rate and function of this decrement can be modulated by altering the dimensions and values within the kernel.
 
-Figure 5 presents a 3D model depicting the final position of the robot, showcasing both scenarios—with the implementation of the joint 4 obstacle avoidance (illustrated in green) and without it (depicted in red).
+Figure~\ref{3D model} presents a 3D model depicting the final position of the robot, showcasing both scenarios—with the implementation of the joint 4 obstacle avoidance (illustrated in green) and without it (depicted in red).
 
@@
 
-This paper presented a novel approach to robotic motion planning, combining the application of kernel convolution for obstacle avoidance velocity calculations with an integration of attractive and repulsive field velocities for end-effector trajectory generation and robot joint obstacle avoidance.
+This paper presented a novel approach to robotic motion planning, combining the application of kernel convolution for obstacle avoidance velocity calculations with an integration of attractive and avoidance velocities components for end-effector trajectory generation and robot joint obstacle avoidance.
 The method was validated using the kinematic model of a 7 DoF Panda Emika robot, demonstrating its operational versatility and adaptability in varying workspace conditions.
@@
 It may also be beneficial to explore the integration of other optimization strategies and machine learning techniques to further enhance the overall efficiency and effectiveness of the trajectory planning process.
+
+\todocomment{add a note about picking up multiple POI}
+
+\todocomment{add note that this is just a part of algorithm development}
 

### 2023-07-27 Articles/ERK/erk.tex
ERK mid-joints

@@
 \begin{equation}
-	\dot{q} = J^{\dagger}(q) \dot{x}_{ee} + N J^{\dagger}_{poi} \dot{x}_{poi}
+	\dot{q} = J^{\dagger}(q) \dot{x}_{ee} + N J^{\dagger}_{poi} \dot{x}_{poi} + N \dot{q}_{mid}
 	\label{eq:manipulator_kinematics}
@@
 
-\todocomment{želena sredinska lega sklepov}
+We integrated a supplementary secondary task aiming to maintain the robot manipulator's joints as close to the midpoint of their range ( between $q(i)_{max}$ and $q(i)_{min}$ for the i-th joint ) as possible ( eq.~\ref{eq:mid-joints}).
+This auxiliary task, designed to elevate the robot's "elbow", reinforced the robot's upright posture during the execution of the primary task and the obstacle avoidance task.
+
+\begin{equation}
+	\dot{q}_{mid}(i) = w_m \left( \frac{q(i)_{max}-q(i)_{min}}{2} - q(i)\right) 
+	\label{eq:mid-joints}
+\end{equation}
 
@@
 
-An additional secondary task was also introduced, aimed at ensuring the manipulator's mid-joints remained in an upright position.
-This secondary task, aimed at maintaining the robot's "elbow" at the top, further reinforced the robot's upright posture while executing the primary task and the obstacle avoidance task.
-
 Figure~\ref{y axis avoidance values} illustrates the avoidance velocities in the x, y, and z directions, computed by our kernel, when its center aligns with various points along the y-axis.

### 2023-07-27 Articles/ERK/erk.tex
ERK beta

@@
 \usepackage{xcolor}
-% \usepackage{easyReview}
+ \usepackage{easyReview}
 \usepackage[top=22.5mm, bottom=22.5mm, left=22.5mm, right=22.5mm]{geometry}
@@
 In this research, we specifically consider a 7-DOF Franka Panda Emika~\cite{Franka2021} manipulator.
-When controlling both the position and orientation of the end-effector in a 3D space, which requires six degrees of freedom (three for position and three for orientation), a 7-DOF manipulator such as the Panda Emika has one redundant degree.
+When controlling both the position and orientation of the end-effector in a 3D space, which requires six degrees of freedom (three for position and three for orientation), a 7-DOF manipulator has one redundant degree.
 However, when controlling only the position, which requires three degrees of freedom, the manipulator possesses four redundant degrees.
@@
 In our approach, we distinguish between primary and secondary tasks.
-The primary task of the manipulator is usually defined as reaching a desired EE position and orientation or as following a trajectory made out of such points.
+The primary task of the manipulator is usually defined as reaching a desired EE position and orientation or as following a path made out of such points.
 In this work we instead use APF (Artificial Potential Field), to guide the EE towards the goal position, while avoiding the obstacles in the path of the EE.
@@
 
-\todocomment{check that repulsive - avoidance is well changed (READ TEXT)}
-
 Two commonly employed methods for obstacle avoidance include artificial potential field methods (APF)~\cite{Khatib1985} and artificial neural networks.
@@
 The convolutional layers in the network can learn spatial hierarchies and structures, and the final layers can output a direction or action to take.
-These methods, although effective, require extensive training data and computational resources, however it is capable to recognizing and avoiding some local minima.
+These methods, although effective, require extensive training data and computational resources, however they are capable of learning to an extent to recognize and avoid some local minima.
 
@@
 We integrated a supplementary secondary task aiming to maintain the robot manipulator's joints as close to the midpoint of their range ( between $q(i)_{max}$ and $q(i)_{min}$ for the i-th joint ) as possible ( eq.~\ref{eq:mid-joints}).
+$w(i)_m$ weights the importance of each joints position.
 This auxiliary task, designed to elevate the robot's "elbow", reinforced the robot's upright posture during the execution of the primary task and the obstacle avoidance task.
@@
 \begin{equation}
-	\dot{q}_{mid}(i) = w_m \left( \frac{q(i)_{max}-q(i)_{min}}{2} - q(i)\right) 
+	\dot{q}_{mid}(i) = w(i)_m \left( \frac{q(i)_{max}-q(i)_{min}}{2} - q(i)\right) 
 	\label{eq:mid-joints}
@@
 
-Subsequently, the avoidance velocity in the Cartesian coordinate system is obtained by summing all the values in the resulting matrix (eq.~\ref{eq:summation}), which can be represented as:
+Subsequently, the avoidance velocity in the Cartesian coordinate system is obtained (eq.~\ref{eq:summation}) by summing all the values in the resulting matrix $C_d$, which can be represented as:
 
@@
 
-In our end-effector motion planning approach, we leverage derivatives of the attractive field to facilitate efficient trajectory generation.
+In our end-effector motion planning approach, we leverage derivatives of the attractive field to facilitate efficient path generation.
 Prior to commencing motion planning, we construct a distance matrix, analogous in size to the matrix representing the obstacles.
@@
 This enables us to guide the end-effector around obstacles smoothly, ensuring safe and collision-free navigation while dynamically adapting to the environment.
-By integrating both the attractive and avoidance field contributions, our motion planning approach ensures efficient and effective trajectory generation, enabling the robotic manipulator to reach its desired goal while avoiding potential obstacles in the workspace.
-
-\todocomment{ - attractive field calculating equation }
+By integrating both the attractive and avoidance field contributions, our motion planning approach ensures efficient and effective path generation, enabling the robotic manipulator to reach its desired goal while avoiding potential obstacles in the workspace.
 
@@
 
-This paper presented a novel approach to robotic motion planning, combining the application of kernel convolution for obstacle avoidance velocity calculations with an integration of attractive and avoidance velocities components for end-effector trajectory generation and robot joint obstacle avoidance.
-The method was validated using the kinematic model of a 7 DoF Panda Emika robot, demonstrating its operational versatility and adaptability in varying workspace conditions.
+This paper presented an approach to robotic motion planning, combining the application of kernel convolution for obstacle avoidance velocity calculations with an integration of attractive and avoidance velocities components for end-effector path generation and manipulator body poi obstacle avoidance.
+The method was validated using the kinematic model of a 7-DOF Panda Emika robot, demonstrating its operational versatility and adaptability in varying workspace conditions.
 
-Notable advantages of our approach include its simplicity and the fact that we work directly in the obstacle space grid, negating the necessity for calculating the spatial primitives of shapes.
+Notable advantages of our approach include its simplicity and the fact that we work directly in the obstacle space grid, negating the necessity for calculating the spatial primitives of obstacles shapes.
 Moreover, when implemented in parallel, our method exhibits high computational efficiency, making it a viable solution for real-time applications.
@@
 
-Our results affirmed the robustness of the proposed approach, as evidenced by the robot's fourth joint being approximately 20 cm further from the wall upon task completion when the avoidance algorithm was employed.
+Our results affirmed the working of the proposed approach, as evidenced by the robot's fourth joint being approximately 20 cm further from the wall upon task completion when the avoidance algorithm was employed.
 
@@
 Additionally, we acknowledge that our study was conducted under simplified conditions with a single type of obstacle.
-As such, future work should focus on validating the proposed method in more complex environments, with different obstacle types, shapes, and densities, to further ascertain its adaptability and robustness.
-It may also be beneficial to explore the integration of other optimization strategies and machine learning techniques to further enhance the overall efficiency and effectiveness of the trajectory planning process.
-
-\todocomment{add a note about picking up multiple POI}
-
-\todocomment{add note that this is just a part of algorithm development}
+This is just a beginning of our work on this project and as such, future work will focus on validating the proposed method in more complex environments, with different obstacle types, shapes, and densities, to further ascertain its adaptability and robustness.
+It may also be beneficial to explore the integration of other optimization strategies and machine learning techniques to further enhance the overall efficiency and effectiveness of the path planning process.
+We intend to investigate the optimal strategy for choosing the best points on the robot for obstacle avoidance.
+Our future work will consider whether it is more advantageous to task the robot with multiple secondary avoidance velocities concurrently, or to only select the closest to the obstacle (largest) avoidance velocity, or to utilize a combination of both.
 
@@
 \textit{Obstacle Avoidance for Redundant Manipulators as Control Problem},
-Leon Lajpah and Tadej Petri,
+Leon Žlajpah and Tadej Petrič,
 In: \textit{Serial and Parallel Robot Manipulators - Kinematics, Dynamics, Control and Optimization},

### 2023-07-27 Articles/ERK/erk.tex
ERK SUBMISSION

@@
 The approach integrates the benefits of convolutional techniques with the spatial understanding provided by voxel grid representation to efficiently calculate movement directions while evading obstacles.
-By focusing on kinematics and obstacle grid representation, the method efficiently utilizes the versatility of a 7-degree-of-freedom (DOF) robot arm, employing geometric and differential kinematics for movement calculations.
+By focusing on kinematics and obstacle grid representation, the method efficiently utilizes the versatility of a 7-degree-of-freedom (DOF) robot arm, employing positional and differential kinematics for movement calculations.
 The strategy doesn't require exhaustive training as with neural networks, making it deployable in various environments.
@@
 
-Robotic manipulators, specifically those with a high degree of freedom (DOF), possess a crucial characteristic known as kinematic redundancy~\cite{Nguyen1991}.
+Robotic manipulators with a high degree of freedom (DOF), possess a crucial characteristic known as kinematic redundancy~\cite{Nguyen1991}.
 Kinematic redundancy refers to having more joints (i.e., degrees of freedom - DOF) than required to perform a task, thus allowing the robot to adopt multiple configurations to accomplish the same end-effector position and orientation.
@@
 The primary task of the manipulator is usually defined as reaching a desired EE position and orientation or as following a path made out of such points.
-In this work we instead use APF (Artificial Potential Field), to guide the EE towards the goal position, while avoiding the obstacles in the path of the EE.
+In this work we instead use a variation of APF (Artificial Potential Field), to guide the EE towards the goal position, while avoiding the obstacles in the path of the EE.
 Operating in a real-world environment often requires considering secondary tasks, such as real-time obstacle avoidance, to ensure safe and uninterrupted operation of the manipulator.
@@
 This paper proposes a novel method for real-time obstacle avoidance of the manipulator arm.
-We use an obstacle grid combined with convolution techniques to assign avoidance directions to each manipulator segment, ensuring efficient and safe maneuvering within complex environments.
+We use an obstacle grid combined with convolution techniques to assign avoidance velocities to points of interest (POI) on manipulator segments, ensuring efficient and safe maneuvering within complex environments.
 
@@
 
-The execution of the secondary task of obstacle avoidance necessitates an understanding of the spatial relationship between the manipulator and surrounding obstacles, usually quantified as a distance or interpreted as a repulsive force.
+The execution of the secondary task of obstacle avoidance necessitates an understanding of the spatial relationship between the manipulator and surrounding obstacles, usually quantified as a distance or interpreted as an avoidance or repulsive force.
 By conceptualizing obstacles as sources of repulsive forces, the manipulator can be guided away from potential collisions, akin to how two magnets of the same polarity repel each other.
@@
 While this method is intuitive and relatively easy to implement, it can sometimes result in local minima problems where the robot gets stuck in a position that is not the target but cannot find a path due to the repelling obstacles.
+Another problem of the classical APF method is that it is focused on simple objects shapes and as such it is complicated to calculate repulsive field for more complex shapes of obstacles.
 
@@
 The seven joints provide a high level of versatility, enabling the robot to move and orient its end-effector in a broad range of positions and directions.
-To calculate the movements and positions of the robot, we use both geometric and differential kinematics.
+To calculate the movements and positions of the robot, we use both positional and differential kinematics.
 
-Geometric kinematics (also known as forward and inverse kinematics) focuses on the relationship between joint angles and the position and orientation of the end-effector.
+Positional kinematics focuses on the relationship between joint angles and the position and orientation of the end-effector.
 Forward kinematics involves calculating the position and orientation of the end-effector given the joint angles, while inverse kinematics is the process of determining the joint angles for a desired end-effector position and orientation.
@@
 We calculate this velocities at certain points of interest (POI) that are positioned on the manipulator arm segments.
-We transform avoidance velocities into the joint space velocities using $J^{\dagger}_{poi}$ jacobian calculated in the specific POI.
+We transform avoidance velocities into the joint space velocities using $J^{\dagger}_{poi}$ Jacobian calculated for the specific POI.
 These velocities we than transform into the null-space using matrix $N$.
@@
 
-We integrated a supplementary secondary task aiming to maintain the robot manipulator's joints as close to the midpoint of their range ( between $q(i)_{max}$ and $q(i)_{min}$ for the i-th joint ) as possible ( eq.~\ref{eq:mid-joints}).
+We integrated a supplementary secondary task aiming to maintain the robot manipulator's joints as close to the midpoint of their range (centered between $q(i)_{max}$ and $q(i)_{min}$ for the i-th joint ) as possible ( eq.~\ref{eq:mid-joints}).
 $w(i)_m$ weights the importance of each joints position.
-This auxiliary task, designed to elevate the robot's "elbow", reinforced the robot's upright posture during the execution of the primary task and the obstacle avoidance task.
+This auxiliary task, designed to elevate the robot's "elbow", reinforces the robot's upright posture during the execution of the primary task and the obstacle avoidance task.
 
@@
 
-%\subsection{Avoidance velocities}
-
-%In our approach to robot arm obstacle avoidance, we apply a method of kernel convolution in conjunction with geometric transformations.
-This methodology allows us to calculate the necessary avoidance velocities for specific points of interest on the manipulator.
-%
-%The process initiates with kinematic transformations which are used to map the points of interest from the robot's configuration space to the cartesian space and than onto the 3D voxel grid.
-These points, strategically placed on the manipulator, serve to evaluate potential collision threats with obstacles and subsequently assist in computing the necessary obstacle avoidance velocities.
-%
-%Once the mapping of these points to the voxel grid is completed, we proceed to employ a specialized kernel convolution method.
-This method is tailored to calculate the constituent parts of the avoidance velocity vector in the Cartesian space, specifically for the x, y, and z directions.
-For each of these directions, we apply a separate 3D convolutional kernel.
-
-%\todocomment{convolution equation}
-
-%For each direction, we utilize its corresponding kernel and extract a segment of the obstacle grid of matching size, centered at the point of interest.
-This leads to two equivalent-sized 3D matrices - one from the obstacle grid and the other being the kernel.
-We then conduct an element-wise multiplication operation between these matrices, summing the resulting values.
-The result of this operation provides us with the avoidance velocity in each respective direction.
-Through this approach, we can dynamically adjust the robot's trajectory to effectively avoid obstacles.
-
 \subsection{Avoidance velocities}
 
-In our approach to robot arm obstacle avoidance, we apply a method of kernel convolution in conjunction with geometric transformations.
+In our approach to robot arm obstacle avoidance, we apply a method of kernel convolution in conjunction with kinematics.
 This methodology allows us to calculate the necessary avoidance velocities for specific points of interest located on the manipulator.
 
-The process commences with kinematic transformations, which facilitate the mapping of points of interest from the robot's configuration space to the Cartesian space, followed by their projection onto the 3D voxel grid.
+The process commences with positional kinematic transformations, which facilitate the mapping of points of interest from the robot's configuration space to the Cartesian space, followed by their projection onto the 3D voxel grid.
 
-%These points, strategically placed on the manipulator, serve to evaluate potential collision threats with obstacles and subsequently assist in computing the necessary obstacle avoidance velocities.
+These points, strategically placed on the manipulator, serve to evaluate potential collision threats with obstacles and subsequently assist in computing the necessary obstacle avoidance velocities.
 
@@
 Conversely, the x and z velocities remain zero, reflecting the absence of obstacles along these axes.
-This aligns with our expectations given the presence of the wall obstacle is only along the y direction.
+This aligns with our expectations given the changing in presence of the wall obstacle is only along the y direction.
 
@@
 The position of the joint, upon task completion, is noted to be approximately 20 cm further away from the wall.
+This distance depends on the specific choice of constants and kernel values.
 
@@
 \section{Conclusions and further development}
-
-%This paper presented a novel approach to robotic motion planning, combining the application of kernel convolution for obstacle avoidance velocity calculations with an integration of attractive and repulsive field velocities for end-effector trajectory generation and robot joint obstacle avoidance.
-The method was validated using the kinematic model of a 7 DoF Panda Emika robot, demonstrating its operational versatility and adaptability in varying workspace conditions.
-%
-%Notable advantages of our approach include its simplicity and the fact that we work directly in the obstacle space grid, negating the necessity for calculating the spatial primitives of shapes.
-Moreover, when implemented in parallel, our method exhibits high computational efficiency, making it a viable solution for real-time applications.
-The experiments conducted effectively showcased that the end-effector could move efficiently towards the target while successfully avoiding obstacles, specifically, a wall in this instance.
-%
-%Our results affirmed the robustness of the proposed approach, as evidenced by the robot's fourth joint being approximately 20 cm further from the wall upon task completion when the avoidance algorithm was employed.
-%
-%Despite the promising initial results, our approach does present some limitations.
-Particularly, the possibility of encountering local minima traps poses a significant challenge.
-Although these traps could be potentially mitigated by incorporating multiple convolutional kernels of varying sizes, it would result in an approach similar to convolutional neural networks (CNNs), adding to the complexity and computational cost of the algorithm.
-%
-%Additionally, we acknowledge that our study was conducted under simplified conditions with a single type of obstacle.
-As such, future work should focus on validating the proposed method in more complex environments, with different obstacle types, shapes, and densities, to further ascertain its adaptability and robustness.
-It may also be beneficial to explore the integration of other optimization strategies and machine learning techniques to further enhance the overall efficiency and effectiveness of the trajectory planning process.
-%
-%In conclusion, our work serves as a solid foundation for the development of more advanced and dynamic motion planning algorithms.
 
@@
 Despite its effectiveness, the approach could potentially fall into local minima traps.
-This limitation could be addressed by incorporating multiple convolutional kernels, though this would increase complexity and resemble convolutional neural networks.
+This limitation could be addressed by incorporating multiple convolutional kernels, though this would increase complexity and resemble a structure of a convolutional neural networks.
 

### 2023-07-27 Articles/ERK/erk.tex
.

@@
 Despite its effectiveness, the approach could potentially fall into local minima traps.
-This limitation could be addressed by incorporating multiple convolutional kernels, though this would increase complexity and resemble a structure of a convolutional neural networks.
+This limitation could be addressed by incorporating multiple convolutional kernels, though this would increase complexity and resemble a structure of convolutional neural networks.
 

### 2023-11-07 Code/REP_field_calculation.m
interp not yet working

@@ -25,7 +25,7 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
     % --------------------------------------------------------------
 
     % use interpolation
-    interpolation_mode = false; % true / false
+    interpolation_mode = true; % true / false
 
     % set values outside the known grid to
     box_value = 0;
@@ -36,35 +36,35 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
 
     if interpolation_mode
     %% USE INTERPOLATION
-
-    center_point = point * grid.resolution;
+        
+        center_point = point * grid.resolution
          
         % get nearby grid cells indexes
-        X = [floor(center_point(1)) ceil(center_point(1))];
-        Y = [floor(center_point(2)) ceil(center_point(2))];
-        Z = [floor(center_point(3)) ceil(center_point(3))];
+        X = [floor(center_point(1)) ceil(center_point(1))]
+        Y = [floor(center_point(2)) ceil(center_point(2))]
+        Z = [floor(center_point(3)) ceil(center_point(3))]
 
         % get values of nearby cells
-        V = ones(2,2,2);
+        Vx = ones(2,2,2);
+        Vy = ones(2,2,2);
+        Vz = ones(2,2,2);
+
         for x = 1:1:2 
             for y = 1:1:2
                 for z = 1:1:2               
                     
-                    % check if voxel location is in the range of the distance grid
-                    if (X(x) <= 0 || Y(y) <= 0 || Z(z) <= 0 || X(x) > size(grid.grid,1) || Y(y) > size(grid.grid,2) || Z(z) > size(grid.grid,3)) 
-                        % points beyond grid are set as obstacles during interpolation
-                        V(y,x,z) = box_value;
-                    else
-                        % points in grid are set to correct voxels
-                        V(y,x,z) = grid_distance(X(x)+1,Y(y)+1,Z(z)+1);
-                    end
-        
+                        % set value of the point
+                        interp_val = calculateFieldVector([X(x),Y(y),Z(z)]);
+                        Vx(y,x,z) = interp_val(1);
+                        Vy(y,x,z) = interp_val(2);
+                        Vz(y,x,z) = interp_val(3);
+                           
                 end
             end
         end
 
         % interpolation
[…]

### 2023-11-07 Code/REP_field_calculation.m
1d interpolation z-direction

@@ -44,27 +44,70 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
         Y = [floor(center_point(2)) ceil(center_point(2))]
         Z = [floor(center_point(3)) ceil(center_point(3))]
 
-        % get values of nearby cells
-        Vx = ones(2,2,2);
-        Vy = ones(2,2,2);
-        Vz = ones(2,2,2);
+        % SKIP INTERPOLATION IF TWO POINTS ARE LOCATED EXACTLY IN THE
+        % CENTER OF GRID CELLS
+        % --------------------------------------------------------------
+        no_interp_dist = 0.0001; 
+ 
+        % if x and y are in center
+        if mod(center_point(1),1) < no_interp_dist && mod(center_point(2),1) < no_interp_dist
+            % 1D interpolate - only z
 
-        for x = 1:1:2 
-            for y = 1:1:2
-                for z = 1:1:2               
-                    
-                        % set value of the point
-                        interp_val = calculateFieldVector([X(x),Y(y),Z(z)]);
-                        Vx(y,x,z) = interp_val(1);
-                        Vy(y,x,z) = interp_val(2);
-                        Vz(y,x,z) = interp_val(3);
-                           
+            Vx = ones(1,2);
+            Vy = ones(1,2);
+            Vz = ones(1,2);
+
+            % get vector field values for x,y,z
+
+            for j = 1:1:length(Z)
+                vect_val = calculateFieldVector([center_point(1),center_point(2),Z(j)]);
+                Vx(j) = vect_val(1);
+                Vy(j) = vect_val(2);
+                Vz(j) = vect_val(3);
+            end
+
+            % interpolate
+            rep_values = [interp1(Z,Vx,center_point(3)), interp1(Z,Vy,center_point(3)), interp1(Z,Vz,center_point(3))]
+        
+        % if y and z are in center
+        elseif mod(center_point(2),1) < no_interp_dist && mod(center_point(3),1) < no_interp_dist
+            % 1D interpolate - only x
+    
+    
+        % if x and z are in exact center cells
+        elseif mod(center_point(1),1) < no_interp_dist && mod(center_point(3),1) < no_interp_dist
+            % 1D interpolate - only y
+
+
+        else
+            % FULL 3D INTERPOLATION
+
+            % get values of nearby cells
+            Vx = ones(2,2,2);
+            Vy = ones(2,2,2);
+            Vz = ones(2,2,2);
+    
+            for x = 1:1:2 
[…]

### 2023-11-07 Code/REP_field_calculation.m
ADDED INTERPOLATION

@@ -25,7 +25,7 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
     % --------------------------------------------------------------
 
     % use interpolation
-    interpolation_mode = true; % true / false
+    interpolation_mode = false; % true / false
 
     % set values outside the known grid to
     box_value = 0;
@@ -37,14 +37,14 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
     if interpolation_mode
     %% USE INTERPOLATION
         
-        center_point = point * grid.resolution
+        center_point = point * grid.resolution;
          
         % get nearby grid cells indexes
-        X = [floor(center_point(1)) ceil(center_point(1))]
-        Y = [floor(center_point(2)) ceil(center_point(2))]
-        Z = [floor(center_point(3)) ceil(center_point(3))]
+        X = [floor(center_point(1)) ceil(center_point(1))];
+        Y = [floor(center_point(2)) ceil(center_point(2))];
+        Z = [floor(center_point(3)) ceil(center_point(3))];
 
-        % SKIP INTERPOLATION IF TWO POINTS ARE LOCATED EXACTLY IN THE
+        % 1D INTERPOLATION IF TWO POINTS ARE LOCATED EXACTLY IN THE
         % CENTER OF GRID CELLS
         % --------------------------------------------------------------
         no_interp_dist = 0.0001; 
@@ -58,7 +58,6 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
             Vz = ones(1,2);
 
             % get vector field values for x,y,z
-
             for j = 1:1:length(Z)
                 vect_val = calculateFieldVector([center_point(1),center_point(2),Z(j)]);
                 Vx(j) = vect_val(1);
@@ -67,20 +66,49 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
             end
 
             % interpolate
-            rep_values = [interp1(Z,Vx,center_point(3)), interp1(Z,Vy,center_point(3)), interp1(Z,Vz,center_point(3))]
+            rep_values = [interp1(Z,Vx,center_point(3)), interp1(Z,Vy,center_point(3)), interp1(Z,Vz,center_point(3))];
         
         % if y and z are in center
         elseif mod(center_point(2),1) < no_interp_dist && mod(center_point(3),1) < no_interp_dist
             % 1D interpolate - only x
-    
-    
+
+            Vx = ones(1,2);
+            Vy = ones(1,2);
+            Vz = ones(1,2);
+
[…]

### 2023-11-07 Code/REP_field_calculation.m
minor

@@ -25,17 +25,16 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
     % --------------------------------------------------------------
 
     % use interpolation
-    interpolation_mode = false; % true / false
+    interpolation_mode = true; % true / false
 
     % set values outside the known grid to
     box_value = 0;
 
 
     % --------------------------------------------------------------    
-
-
-    if interpolation_mode
+    
     %% USE INTERPOLATION
+    if interpolation_mode
         
         center_point = point * grid.resolution;
          
@@ -65,9 +64,16 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
                 Vz(j) = vect_val(3);
             end
 
+            % move sampled points coordinates to the mid of the cells
+            Z = [floor(center_point(3)) ceil(center_point(3))]+0.5;
+
             % interpolate
             rep_values = [interp1(Z,Vx,center_point(3)), interp1(Z,Vy,center_point(3)), interp1(Z,Vz,center_point(3))];
-        
+            
+            % check if value is NaN (probably because both interp. value
+            % are 0) and set it to 0
+            rep_values(isnan(rep_values)) = 0;
+
         % if y and z are in center
         elseif mod(center_point(2),1) < no_interp_dist && mod(center_point(3),1) < no_interp_dist
             % 1D interpolate - only x
@@ -84,8 +90,15 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
                 Vz(j) = vect_val(3);
             end
 
+            % move sampled points coordinates to the mid of the cells
+            X = [floor(center_point(1)) ceil(center_point(1))] + 0.5;
+
             % interpolate
             rep_values = [interp1(X,Vx,center_point(1)), interp1(X,Vy,center_point(1)), interp1(X,Vz,center_point(1))];
+
+            % check if value is NaN (probably because both interp. value
+            % are 0) and set it to 0
+            rep_values(isnan(rep_values)) = 0;
        
         % if x and z are in exact center cells
         elseif mod(center_point(1),1) < no_interp_dist && mod(center_point(3),1) < no_interp_dist
@@ -103,8 +116,15 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
                 Vz(j) = vect_val(3);
             end
 
+            % move sampled points coordinates to the mid of the cells
+            Y = [floor(center_point(2)) ceil(center_point(2))] + 0.5;
+
[…]

### 2023-11-07 Code/REP_kernels.m

@@ -21,8 +21,8 @@ function [kernels] = REP_kernels()
     %    kernels - A cell array containing the generated directional kernels.
 
     % set directional kernels
-    sigma = 13;
-    kernel_length = 25;
+    sigma = 21;
+    kernel_length = 31;
     kernel_width = 9;
     kernel_height = kernel_width;
     

### 2023-11-07 Code/REP_field_calculation.m
bug fix

@@ -36,7 +36,7 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
     %% USE INTERPOLATION
     if interpolation_mode
         
-        center_point = point * grid.resolution;
+        center_point = point * grid.resolution
          
         % get nearby grid cells indexes
         X = [floor(center_point(1)) ceil(center_point(1))];
@@ -47,9 +47,13 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
         % CENTER OF GRID CELLS
         % --------------------------------------------------------------
         no_interp_dist = 0.0001; 
+
+        % if x and y and z are center
+        if mod(center_point(1),1) < no_interp_dist && mod(center_point(2),1) < no_interp_dist && mod(center_point(3),1) < no_interp_dist
+            rep_values = calculateFieldVector(center_point)
  
         % if x and y are in center
-        if mod(center_point(1),1) < no_interp_dist && mod(center_point(2),1) < no_interp_dist
+        elseif mod(center_point(1),1) < no_interp_dist && mod(center_point(2),1) < no_interp_dist
             % 1D interpolate - only z
 
             Vx = ones(1,2);
@@ -186,15 +190,15 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
             half_size = floor(kernel_size / 2);
     
             % Calculate x, y, and z ranges for the window while ensuring they are positive and within grid boundaries.
-            x_cells = (grid_point(2) - half_size(1)) : (grid_point(2) + half_size(1)); % grid goes y-x-z
-            y_cells = (grid_point(1) - half_size(2)) : (grid_point(1) + half_size(2));
-            z_cells = (grid_point(3) - half_size(3)) : (grid_point(3) + half_size(3));
+            x_cells = round((grid_point(2) - half_size(1)) : (grid_point(2) + half_size(1))) % grid goes y-x-z
+            y_cells = round((grid_point(1) - half_size(2)) : (grid_point(1) + half_size(2)));
+            z_cells = round((grid_point(3) - half_size(3)) : (grid_point(3) + half_size(3)));
 
             % init every cells in window to occupued by default
             window = ones(kernel_size)*box_value;
 
             % create indexes for cutout
-            x_valid_index = find(x_cells>=1,1):find(x_cells<=size(grid.grid,3),1,'last');
[…]

### 2023-11-07 Code/Experiments APF/MANIP_EXPERIMENTS.m
added more points to look at the same time, removed scalling based on the primary velocity (didnt do anything, as it is always 1)

@@ -5,7 +5,7 @@ close all
 % ------------------------------------------------------------
 
 % select map
-map_selection = 'column'
+map_selection = 'corridor'
 
 [grid] = MAPS(map_selection);
 

### 2023-11-07 Code/Full_RTConvolution_Multiple_Points.m

@@ -13,7 +13,7 @@ timestep_primary_gain_change = 0 % if selected, primary task will start with lit
 points_per_segment = 1*[2 1 5 2 6 2 1];
 
 % the number of points taken into account and weighting factors
-weights_avoidance = [1] %  ones(1, sum(points_per_segment));
+weights_avoidance = [1 1 1 1 1 1 1 1]
 weights_avoidance = weights_avoidance / norm(weights_avoidance,1);
 
 % -----------------------------------------------------------
@@ -135,11 +135,8 @@ while current_dist > goal_dist && Niter <= Nmax
     % REPULSIVE 
     ee_vel_rep = REP_field_calculation(grid, rep_kernels, ee_point);
     
-    % scale ee_rep to not interfere with ee_att
-    scaled_ee_vel_rep = norm(ee_vel_att) * ee_vel_rep';
-
     % TOTAL : SUM   
-    ee_vel = (wp_att * ee_vel_att + wp_rep * avoid_task * scaled_ee_vel_rep);
+    ee_vel = (wp_att * ee_vel_att + wp_rep * avoid_task * ee_vel_rep');
 
     if timestep_primary_gain_change
         ee_vel = Niter * wp .* tanh([ee_vel ; 0 ; 0 ; 0]/sigm_factor_primary);

### 2023-11-07 Code/REP_field_calculation.m

@@ -190,7 +190,7 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
             half_size = floor(kernel_size / 2);
     
             % Calculate x, y, and z ranges for the window while ensuring they are positive and within grid boundaries.
-            x_cells = round((grid_point(2) - half_size(1)) : (grid_point(2) + half_size(1))) % grid goes y-x-z
+            x_cells = round((grid_point(2) - half_size(1)) : (grid_point(2) + half_size(1))); % grid goes y-x-z
             y_cells = round((grid_point(1) - half_size(2)) : (grid_point(1) + half_size(2)));
             z_cells = round((grid_point(3) - half_size(3)) : (grid_point(3) + half_size(3)));
 
@@ -198,7 +198,7 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
             window = ones(kernel_size)*box_value;
 
             % create indexes for cutout
-            x_valid_index = find(x_cells>=1,1):find(x_cells<=size(grid.grid,3),1,'last')
+            x_valid_index = find(x_cells>=1,1):find(x_cells<=size(grid.grid,3),1,'last');
             y_valid_index = find(y_cells>=1,1):find(y_cells<=size(grid.grid,3),1,'last');
             z_valid_index = find(z_cells>=1,1):find(z_cells<=size(grid.grid,3),1,'last');
 

### 2023-11-07 Code/Full_RTConvolution_Multiple_Points.m
added slowdown added, removed redundant tanh functions

@@ -7,6 +7,7 @@ mid_joints = 0
 avoid_task = 1
 kinematics_solution = 'exact-reduced' % OPTIONS: exact-reduced , exact , approximate
 timestep_primary_gain_change = 0 % if selected, primary task will start with little gain and grow with time
+secondary_exec_stop_k = 0.10 % primary task will slow down (>0) or stop executing if secondary task has big velocities 
 
 % -----------------------------------------------------------
 % number of points per segment for obstacle avoidance taskmanipulability_avoidance
@@ -19,7 +20,7 @@ weights_avoidance = weights_avoidance / norm(weights_avoidance,1);
 % -----------------------------------------------------------
 
 Tstep = 0.1 % time step
-Nmax = 200 % max number of iterations
+Nmax = 2000 % max number of iterations
 space_resolution = grid.resolution; % resolution of the obstacles grid
 
 % weights for different tasks
@@ -27,14 +28,9 @@ wp = 0.2 % primary task
 wp_att = 4 % primary task - attractive component
 wp_rep = 1 % primary task - repulsive component
 wm = 0.5 % mid-joints task
-wa = 0.2 % obstacle avoidance task
-wa_i = 1 % obstacle avoidance 
+wa = 0.1 % obstacle avoidance task
 
-% factor that controls sigmoid function (tanh) for primary task
-sigm_factor_primary = 1
 
-% factor that controls sigmoid function (tanh) for avoidance task
-sigm_factor_avoidance = 1
 
 % function parameters
 goal_dist = 0.02 % distance which satisfies ending of optimization
@@ -137,12 +133,8 @@ while current_dist > goal_dist && Niter <= Nmax
     
     % TOTAL : SUM   
     ee_vel = (wp_att * ee_vel_att + wp_rep * avoid_task * ee_vel_rep');
-
-    if timestep_primary_gain_change
-        ee_vel = Niter * wp .* tanh([ee_vel ; 0 ; 0 ; 0]/sigm_factor_primary);
-    else
-        ee_vel = wp .* tanh([ee_vel ; 0 ; 0 ; 0]/sigm_factor_primary);
-    end
+    ee_vel = wp .* [ee_vel ; 0 ; 0 ; 0];
+    
 
     
     %% CALCULATE SECONDARY TASKS (AVOIDANCE + MID-JOINTS)
@@ -246,8 +238,8 @@ while current_dist > goal_dist && Niter <= Nmax
             % avoidance direction - unit vector
[…]

### 2023-11-07 Code/REP_field_calculation.m

@@ -36,7 +36,7 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
     %% USE INTERPOLATION
     if interpolation_mode
         
-        center_point = point * grid.resolution
+        center_point = point * grid.resolution;
          
         % get nearby grid cells indexes
         X = [floor(center_point(1)) ceil(center_point(1))];

### 2023-11-08 Code/Experiments APF/MANIP_EXPERIMENTS.m
PRIMARY TASK EXECUTION SLOWDOWN

@@ -5,7 +5,7 @@ close all
 % ------------------------------------------------------------
 
 % select map
-map_selection = 'column'
+map_selection = 'corridor'
 
 [grid] = MAPS(map_selection);
 
@@ -55,8 +55,6 @@ elseif (matches(map_selection, 'column'))
     0.497594642234695
     0.654743179774447
     -0.0801486263893128];
-
-
     
 
 elseif (matches(map_selection, 'corridor'))     

### 2023-11-08 Code/Full_RTConvolution_Multiple_Points.m

@@ -7,14 +7,14 @@ mid_joints = 0
 avoid_task = 1
 kinematics_solution = 'exact-reduced' % OPTIONS: exact-reduced , exact , approximate
 timestep_primary_gain_change = 0 % if selected, primary task will start with little gain and grow with time
-secondary_exec_stop_k = 0.10 % primary task will slow down (>0) or stop executing if secondary task has big velocities 
+secondary_exec_stop_k = 0.25 % primary task will slow down (>0) or stop executing if secondary task has big velocities 
 
 % -----------------------------------------------------------
 % number of points per segment for obstacle avoidance taskmanipulability_avoidance
 points_per_segment = 1*[2 1 5 2 6 2 1];
 
 % the number of points taken into account and weighting factors
-weights_avoidance = [1 1 1 1 1 1 1 1]
+weights_avoidance = [1 1 1 1 1 1]
 weights_avoidance = weights_avoidance / norm(weights_avoidance,1);
 
 % -----------------------------------------------------------
@@ -236,7 +236,11 @@ while current_dist > goal_dist && Niter <= Nmax
             rep_magnitude = poi_sizes(selected_poi);        
     
             % avoidance direction - unit vector
-            rep_direction = rep_vel' ./ rep_magnitude;
+            if(rep_magnitude > 10^-12)
+                rep_direction = rep_vel' ./ rep_magnitude;
+            else
+                rep_direction = [0 0 0];
+            end
     
             % scalling
             avoid_vel = wa * rep_vel;
@@ -309,7 +313,10 @@ while current_dist > goal_dist && Niter <= Nmax
         % -----------------------
         q_avoid_total = q_avoid_list * weights_avoidance';
 
- 
+        
+        if(isnan(q_avoid_list(1)))
+            a=1
+        end
     end
 
     %% COMBINE TASKS

### 2023-11-08 Code/Experiments APF/MANIP_EXPERIMENTS.m
constants

@@ -5,7 +5,7 @@ close all
 % ------------------------------------------------------------
 
 % select map
-map_selection = 'corridor'
+map_selection = 'column'
 
 [grid] = MAPS(map_selection);
 

### 2023-11-08 Code/Full_RTConvolution_Multiple_Points.m

@@ -14,7 +14,7 @@ secondary_exec_stop_k = 0.25 % primary task will slow down (>0) or stop executin
 points_per_segment = 1*[2 1 5 2 6 2 1];
 
 % the number of points taken into account and weighting factors
-weights_avoidance = [1 1 1 1 1 1]
+weights_avoidance = [1 1 1 1 1 1 1 1 1]
 weights_avoidance = weights_avoidance / norm(weights_avoidance,1);
 
 % -----------------------------------------------------------
@@ -26,14 +26,14 @@ space_resolution = grid.resolution; % resolution of the obstacles grid
 % weights for different tasks
 wp = 0.2 % primary task
 wp_att = 4 % primary task - attractive component
-wp_rep = 1 % primary task - repulsive component
+wp_rep = 4 % primary task - repulsive component
 wm = 0.5 % mid-joints task
 wa = 0.1 % obstacle avoidance task
 
 
 
 % function parameters
-goal_dist = 0.02 % distance which satisfies ending of optimization
+goal_dist = 0.01 % distance which satisfies ending of optimization
 
 % damping factor to avoid inverse Jacobain matrix singularities
 damping_factor_primary = 0.01

### 2023-11-08 Code/REP_field_calculation.m
updated header comment

@@ -11,15 +11,19 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
     %   - POINT: A 1x3 vector specifying the coordinates of the point in the grid
     %            at which to compute the representative values.
     %
+    % Optional In-Function Settings:
+    %   - interpolation_mode: A boolean (true or false) to enable or disable interpolation.
+    %                         If enabled, the function interpolates APF values based on the
+    %                         nearest grid cells. 
+    %
+    %   - box_value: Use this parameter to set occupation of the cells that
+    %     are beyond the inputed grid space.
+    %
     %   Outputs:
     %   - REP_VALUES: A 1xN vector of representative values, where N is the number
     %                 of kernels. The i-th element of REP_VALUES is the representative
     %                 value computed using the i-th kernel in KERNELS.
 
-    % TODO: set input values beyond grid as occupied or free using optional
-    % argument value
-    %
-    % TODO: box_mode add walls around the grid space
 
     %% SETTINGS:
     % --------------------------------------------------------------

### 2023-11-08 Code/Full_RTConvolution_Multiple_Points.m
constants

@@ -11,7 +11,7 @@ secondary_exec_stop_k = 0.25 % primary task will slow down (>0) or stop executin
 
 % -----------------------------------------------------------
 % number of points per segment for obstacle avoidance taskmanipulability_avoidance
-points_per_segment = 1*[2 1 5 2 6 2 1];
+points_per_segment = 1*[2 1 5 2 8 2 1];
 
 % the number of points taken into account and weighting factors
 weights_avoidance = [1 1 1 1 1 1 1 1 1]
@@ -27,13 +27,13 @@ space_resolution = grid.resolution; % resolution of the obstacles grid
 wp = 0.2 % primary task
 wp_att = 4 % primary task - attractive component
 wp_rep = 4 % primary task - repulsive component
-wm = 0.5 % mid-joints task
-wa = 0.1 % obstacle avoidance task
+wm = 0.15 % mid-joints task
+wa = 0.5 % obstacle avoidance task
 
 
 
 % function parameters
-goal_dist = 0.01 % distance which satisfies ending of optimization
+goal_dist = 0.02 % distance which satisfies ending of optimization
 
 % damping factor to avoid inverse Jacobain matrix singularities
 damping_factor_primary = 0.01

### 2023-11-08 Code/REP_kernels.m

@@ -21,15 +21,18 @@ function [kernels] = REP_kernels()
     %    kernels - A cell array containing the generated directional kernels.
 
     % set directional kernels
-    sigma = 21;
+    sigma = 5;
     kernel_length = 31;
     kernel_width = 9;
     kernel_height = kernel_width;
     
     % form kernels
-    kernel_x = directional_kernel_3d('x', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, 'gaussian')/1000;
-    kernel_y = directional_kernel_3d('y', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, 'gaussian')/1000;
-    kernel_z = directional_kernel_3d('z', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, 'gaussian')/1000;
+    kernel_x = directional_kernel_3d('x', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, 'gaussian')/200;
+    kernel_y = directional_kernel_3d('y', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, 'gaussian')/200;
+    kernel_z = directional_kernel_3d('z', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, 'gaussian')/200;
+
+    figure()
+    plot(kernel_x(5,:,5))
     
     kernels = {kernel_x, kernel_y, kernel_z};
 

### 2023-11-09 Code/Full_RTConvolution_Multiple_Points.m
.

@@ -20,7 +20,7 @@ weights_avoidance = weights_avoidance / norm(weights_avoidance,1);
 % -----------------------------------------------------------
 
 Tstep = 0.1 % time step
-Nmax = 2000 % max number of iterations
+Nmax = 200 % max number of iterations
 space_resolution = grid.resolution; % resolution of the obstacles grid
 
 % weights for different tasks

### 2023-11-09 Articles/Article 1/main.tex
article start

\documentclass[]{article}

%opening
\title{Robotic Arm Path Planning Utilizing Interpolated 3D Voxel-Based Artificial Potential Fields}
\author{}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{}

\end{document}

### 2023-11-09 Articles/Article 1/main.tex
article structure

@@
 
+ \usepackage{easyReview}
+
 %opening
-\title{Robotic Arm Path Planning Utilizing Interpolated 3D Voxel-Based Artificial Potential Fields}
-\author{}
+\title{Robotic Manipulator Path Planning Utilizing Interpolated 3D Voxel-Based Obstacle Avoidance Method}
+\author{Jakob Baumgartner, Gregor Klančar}
 
@@
 
-\section{}
+\section{Introduction}
+
+\section{Related Works}
+
+\add{
+	- method was inspired by Khatib APF (however, it evolved into a different method) \\
+	- different existing APF manipulator implementation articles \\
+	- VFH \\
+}
+
+\section{Method}
+
+\add{
+	- optimization algorithm \\
+	- robot kinematics (include the exact-reduced method) \\
+	- primary task of distance goal \\
+	- secondary task of repulsive field \\
+	- task slowdown option \\
+	- secondary task of manipulation measure \\
+	- attractive field calculation (normalization of attractive force) \\
+	- repulsive field calculation - matrix "convolution" method \\
+	- matrix size and shape selection \\
+	- PLOT: (ERK) kernel graphics \\
+	- PLOT: (ERK) linear kernel graphics \\
+	- PLOT: kernel field shape 	\\
+	- interpolation of the repulsive field \\
+}
+
+\section{Implementation}
+
+\section{Experiments and simulations}
+
+\add{
+	- include execution times}
 

### 2023-11-09 Articles/Article 1/main.tex
further article structure creation
working title manipulation

@@
 %opening
-\title{Robotic Manipulator Path Planning Utilizing Interpolated 3D Voxel-Based Obstacle Avoidance Method}
+\title{\comment{3D Voxel Grid Based Path Planning for Robotic Manipulators through Matrix Multiplication Techniques}{Alternatives:\\ - Matrix Multiplication-Driven Repulsive Fields for 3D Voxel-Based Robotic Manipulator Path Planning \\ - Robotic Manipulator Path Planning Optimization Using Matrix-Derived Repulsive Fields Based on 3D Voxel Grid}}
+
 \author{Jakob Baumgartner, Gregor Klančar}
@@
 
-\section{Related Works}
+\section{Background}
 
@@
 
-\section{Method}
+\section{Methodology}
 
@@
 
-\section{Experiments and simulations}
+\section{Results}
 
 \add{
-	- include execution times}
+	- include execution times \\
+	- PLOT: kernel on robot graphics \\
+}
+
+\section{Discussion}
+
+\add{
+	- add the limitations of such method \\
+	- the limitations of local search \\
+	- number of parameters that need to be tuned (are there actually that many?)
+}
+
+\section{Conclusion}
 

### 2023-11-10 Articles/Article 1/main.tex
structuring the article

--- 2023-11-09 Articles/Article 1/main.tex
@@
 
+ \usepackage{easyReview}
+
 %opening
-\title{Robotic Arm Path Planning Utilizing Interpolated 3D Voxel-Based Artificial Potential Fields}
-\author{}
+\title{\comment{3D Voxel Grid Based Path Planning for Robotic Manipulators using Matrix Multiplication Technique}{Alternatives:\\ - Matrix Multiplication-Driven Repulsive Fields for 3D Voxel-Based Robotic Manipulator Path Planning \\ - Robotic Manipulator Path Planning Optimization Using Matrix-Derived Repulsive Fields Based on 3D Voxel Grid}}
+
+\author{Jakob Baumgartner, Gregor Klančar}
 
@@
 
-\section{}
+\section{Introduction}
+
+\section{Background}
+
+\todo{1-2 PAGES}
+
+\add{
+	- method was inspired by Khatib APF (however, it evolved into a different method) \\
+	- different existing APF manipulator implementation articles \\
+	- VFH \\
+}
+
+\section{Methodology}
+
+\todo{3 PAGES}
+
+\subsection{Optimization Algorithm}
+
+\add{
+	- optimization algorithm \\
+	- robot kinematics (include the exact-reduced method) \\
+	- primary task of distance goal \\
+	- secondary task of repulsive field \\
+	- task slowdown option \\
+	- secondary task of manipulation measure \\	
+}
+
+\subsection{Task Constraints}
+
+\add{-equations of occupied and empty space}
+
+\subsection{Attractive Velocity}
+
+\add{	
+	- attractive field calculation (normalization of attractive force) 
+}
+
+\subsection{Repulsive Velocity}
+
+\add{	
+	- repulsive field calculation - matrix "convolution" method \\
+	- matrix size and shape selection \\
+	- PLOT: (ERK) kernel graphics \\
+	- PLOT: (ERK) linear kernel graphics \\
+	- PLOT: kernel field shape 	\\
+	- interpolation of the repulsive field \\
+}
+
+\section{Implementation}
+
+\section{Results}
+
+\todo{3 PAGES}
+
+\add{
+	- include execution times \\
+	- PLOT: kernel on robot graphics \\
+}
+
+\section{Discussion}
+
+\todo{1 PAGE}
+
+\add{
+	- add the limitations of such method (already mentioned by Khatib) \\
+	- the limitations of local search \\
+	- number of parameters that need to be tuned (are there actually that many?)
+}
+
+\section{Conclusion}
 

### 2023-11-13 Articles/Article 1/main.tex
more planning

@@
 %opening
-\title{\comment{3D Voxel Grid Based Path Planning for Robotic Manipulators using Matrix Multiplication Technique}{Alternatives:\\ - Matrix Multiplication-Driven Repulsive Fields for 3D Voxel-Based Robotic Manipulator Path Planning \\ - Robotic Manipulator Path Planning Optimization Using Matrix-Derived Repulsive Fields Based on 3D Voxel Grid}}
+\title{\comment{3D Voxel Grid Based Path Planning for Robotic Manipulators using Matrix Multiplication Technique}{Alternatives:\\ - Efficient distance calculation / Efficient repulsive field calculation technique \\   - Matrix Multiplication-Driven Repulsive Fields for 3D Voxel-Based Robotic Manipulator Path Planning \\ - Robotic Manipulator Path Planning Optimization Using Matrix-Derived Repulsive Fields Based on 3D Voxel Grid}}
 
@@
 \add{
+	- PRESENT DISTANCE CALCULATION FOR MANIPULATORS (sensors, lidar, ir, bounding boxes) \\
+	- PRESENT PATH PLANNING METHODS FOR MANIPULATOR (optimization, sampling, biological, learning) \\
+	- similar to Distance Transform (a kind of inverted distance transform) \\
 	- method was inspired by Khatib APF (however, it evolved into a different method) \\
@@
 	- secondary task of repulsive field \\
+	- damped least squares \\
 	- task slowdown option \\
@@
 \add{	
+	- object detection is done in task domain and not c-space (more logical) \\
 	- repulsive field calculation - matrix "convolution" method \\
 	- matrix size and shape selection \\
+	- equation for repulsive kernel values (non-linear) \\
 	- PLOT: (ERK) kernel graphics \\
@@
 	- interpolation of the repulsive field \\
+	- what if there are obstacles behind wall (usually not the case, depth sensors show only thin walls, some noise doesnt matter, non-linear kernel, possible additional pre-convolution to convert obstacle grid to edges) \\
+	- effecient calculation in dinamic environments, lacking prediction capabilities (MPC) \\
 }
+
+\subsubsection{Kernel Selection}
+
+\subsubsection{Interpolation}
 

### 2023-11-14 Articles/Article 1/main.tex
Article: started work on attractive velocity

@@
 
+In the context of our robotic manipulator, the primary task requires an attractive force that directs the end effector (EE) towards the desired position and orientation.
+In the seminal work of Khatib, this task is defined in the context of the joint coordinate system and joint error.
+However, we have chosen to define the error of the primary target task in Cartesian coordinate space as the orientation and position error of the EE.
+This approach allows us to guide the manipulator to its target position and orientation without prior knowledge of its final joint configuration.
+This is particularly important in scenarios where the manipulator is redundant and there are multiple or potentially infinite positions that meet the requirements of the target and secondary tasks.
+Our method proves beneficial in efficiently identifying an optimal position in such cases since the target joint configuration of the manipulator is unknown.
+
+To compute the positional error of the end effector (EE), we evaluate the gradient of the squared difference between the vectors representing the current position of the EE and its goal position.
+Mathematically, this is expressed as:
+
+\[ U_{{goalAttraction}} = -\nabla\left(\frac{(x - x_g)^2}{2}\right) \]
+
+Here, \(x\) denotes the vector of the current EE position, and \(x_g\) is the vector of the goal position.
+The term \((x - x_g)^2\) represents the squared Euclidean distance between these two points.
+The gradient of this squared distance, indicated by \(-\nabla\), provides the direction and magnitude of the error in the EE's position.
+
 \add{	

### 2023-11-14 Code/Experiments APF/MANIP_EXPERIMENTS.m
added sigmoid velocity

@@ -5,7 +5,7 @@ close all
 % ------------------------------------------------------------
 
 % select map
-map_selection = 'roof'
+map_selection = 'wall'
 
 [grid] = MAPS(map_selection);
 

### 2023-11-14 Code/Full_RTConvolution_Multiple_Points.m

@@ -33,7 +33,7 @@ wa = 0.5 % obstacle avoidance task
 
 
 % function parameters
-goal_dist = 0.02 % distance which satisfies ending of optimization
+goal_dist = 0.001 % distance which satisfies ending of optimization
 
 % damping factor to avoid inverse Jacobain matrix singularities
 damping_factor_primary = 0.01
@@ -126,7 +126,8 @@ while current_dist > goal_dist && Niter <= Nmax
     % --------------------------------------------------
 
     % ATTRACTIVE ( OPTION KINEMATICS CLASSIC END EFFECTOR )
-    ee_vel_att = (goal_point(1:3)' - ee_point)/norm(goal_point(1:3)' - ee_point); % direction only - normalised
+    ee_vel_att_magn = norm(goal_point(1:3)' - ee_point)
+    ee_vel_att = (goal_point(1:3)' - ee_point)/ee_vel_att_magn * atan(100*ee_vel_att_magn)/pi*2 % direction only - normalised
 
     % REPULSIVE 
     ee_vel_rep = REP_field_calculation(grid, rep_kernels, ee_point);

### 2023-11-14 Code/Full_RTConvolution_Multiple_Points.m
added sigmoid for slow goal approach

@@ -127,7 +127,7 @@ while current_dist > goal_dist && Niter <= Nmax
 
     % ATTRACTIVE ( OPTION KINEMATICS CLASSIC END EFFECTOR )
     ee_vel_att_magn = norm(goal_point(1:3)' - ee_point)
-    ee_vel_att = (goal_point(1:3)' - ee_point)/ee_vel_att_magn * atan(100*ee_vel_att_magn)/pi*2 % direction only - normalised
+    ee_vel_att = (goal_point(1:3)' - ee_point)/ee_vel_att_magn * atan(100*ee_vel_att_magn)/pi*2 % direction only - normalised - sigmoid
 
     % REPULSIVE 
     ee_vel_rep = REP_field_calculation(grid, rep_kernels, ee_point);

### 2023-11-14 Articles/Article 1/main.tex
part of att velocity chapter

--- 2023-11-13 Articles/Article 1/main.tex
@@
 \add{
-	- attractive field calculation (normalization of attractive force) 
+	-orientation error \\ 
+	- p = v w \\
+	- describe constants
+	- shorten
 }
+
+% FIRST DRAFT
+%In the context of our robotic manipulator, the primary task requires an attractive force that directs the end effector (EE) towards the desired position and orientation.
+In the seminal work of Khatib, this task is defined in the context of the joint coordinate system and joint error.
+However, we have chosen to define the error of the primary target task in Cartesian coordinate space as the orientation and position error of the EE.
+This approach allows us to guide the manipulator to its target position and orientation without prior knowledge of its final joint configuration.
+This is particularly important in scenarios where the manipulator is redundant and there are multiple or potentially infinite positions that meet the requirements of the target and secondary tasks.
+Our method proves beneficial in efficiently identifying an optimal position in such cases since the target joint configuration of the manipulator is unknown.
+%To compute the positional error of the end effector (EE), we evaluate the gradient of the squared difference between the vectors representing the current position of the EE and its goal position.
+Because we want the robot to move with constant velocity most of the way and only slow down when bigger precision is needed we calculate direction unit vector of this error and call it directional vector.
+This allows us to move at a speed that doesnt overwhelm repulsive secondary task, so that the robot has time to move around the obstacles.
+Just this however is not enoug, as the robot would move with a constant velocity and as such overshot the goal position.
+To solve this we multiply the velocity vector with a sigmoid function, specifically atan function.
+This allows us to have relatively constant speed most of the way and smoothly slow down when we approach the goal point and dont skip around the goal.
+
+% SHORT
+%In developing our method, an important part is a velocity profile that guides the end effector (EE) towards its target position and orientation Our methodology represents a deviation from the paradigm established by Khatib.
+We employ Cartesian coordinates to define the primary task's error, enabling the manipulator to reach its target without predetermined joint configurations.
+This is vital for redundant manipulators with multiple feasible configurations.
+%Traditional methods using the gradient of squared distance lead to high initial velocities and decreased velocity near the target, potentially hindering lower priority tasks like obstacle avoidance.
+Our aim is a constant velocity profile, decelerating only near the target to prevent overshooting.
+We achieve this by normalizing the direction vector towards the target and modulating its magnitude with a sigmoid function, specifically atan(velocity).
+This allows steady movement, decreasing velocity as the target nears.
+%The approach is encapsulated in the equation:
+%\begin{equation} \vec{v}_{ATT} = \frac{\vec{x}_{EE} - \vec{x}_g}{||\vec{x}_{EE} - \vec{x}_g||} \times \arctan(||\vec{x}_{EE} - \vec{x}_g||) \end{equation}
+
+%LONG
+In the development of our method, a pivotal element is the integration of a velocity profile that effectively guides the end effector (EE) toward its designated target position and orientation.
+Our methodology represents a deviation from the paradigm established by Khatib.
+Instead of quantifying the primary task's error within the joint coordinate framework, we opt for a Cartesian coordinate representation.
+This aspect is important when the manipulator exhibits redundancy, characterized by multiple or potentially infinite configurations that satisfy the primary task's requirements and constraints.
+In such scenarios, it can becomes impractical, if not impossible, to predetermine the optimal goal joint configuration.
+Utilizing solely the gradient of the squared distance would result in a high initial velocity when distant from the target, followed by a progressively decreasing velocity upon nearing the target.
+Such a velocity profile is suboptimal, as the excessive initial velocity could potentially compromise lower priority tasks, such as obstacle avoidance and in leads to unnecessary slow speeds when approaching the target.
+Our objective is to maintain a consistent velocity profile for the majority of the trajectory, with a gradual deceleration as the goal is approached to avoid overshooting.
+To this end, we initially compute the unit vector of the distance from the target, yielding a normalized velocity in the required direction.
+However, this would lead to overshooting and instability near the target.
+To counteract this, we modulate the magnitude of this vector using a sigmoid function, specifically the atan function.
+This modulation ensures a constant movement for the majority of the time towards the target, with the sigmoid function effectively reducing the velocity in inverse relation when in near proximity to the target.
+The following equation encapsulates our approach:
+\begin{equation} 
+	\vec{v}_{ATT} = \frac{\vec{x}_{EE} - \vec{x}_g}{||\vec{x}_{EE} - \vec{x}_g||} \times \frac{\arctan(k_{sigm} \; ||\vec{x}_{EE} - \vec{x}_g||) }{pi/2}
+\end{equation}
 

### 2023-11-14 Articles/Article 1/main.tex
improvements to chapter

@@
 
-%LONG
-In the development of our method, a pivotal element is the integration of a velocity profile that effectively guides the end effector (EE) toward its designated target position and orientation.
+%%LONG
+%In the development of our method, a pivotal element is the integration of a velocity profile that effectively guides the end effector (EE) toward its designated target position and orientation.
 Our methodology represents a deviation from the paradigm established by Khatib.
@@
 In such scenarios, it can becomes impractical, if not impossible, to predetermine the optimal goal joint configuration.
-Utilizing solely the gradient of the squared distance would result in a high initial velocity when distant from the target, followed by a progressively decreasing velocity upon nearing the target.
+%
+%Utilizing solely the gradient of the squared distance would result in a high initial velocity when distant from the target, followed by a progressively decreasing velocity upon nearing the target.
 Such a velocity profile is suboptimal, as the excessive initial velocity could potentially compromise lower priority tasks, such as obstacle avoidance and in leads to unnecessary slow speeds when approaching the target.
@@
 This modulation ensures a constant movement for the majority of the time towards the target, with the sigmoid function effectively reducing the velocity in inverse relation when in near proximity to the target.
-The following equation encapsulates our approach:
+%The following equation encapsulates our approach:
+%\begin{equation} 
+%	\vec{v}_{ATT} = \frac{\vec{x}_{EE} - \vec{x}_g}{||\vec{x}_{EE} - \vec{x}_g||} \times \frac{\arctan(k_{sigm} \; ||\vec{x}_{EE} - \vec{x}_g||) }{pi/2}
+%\end{equation}
+
+% JUST RIGHT
+Our method innovatively employs a velocity profile to direct the end effector (EE) towards its target, marking a departure from Khatib's joint coordinate approach in favor of a Cartesian coordinate framework.
+This is particularly beneficial in scenarios involving redundant manipulators, where determining an optimal goal joint configuration in advance is challenging.
+
+When calculating translational velocity, we avoid the conventional gradient of the squared distance approach, which leads to high initial velocities and subsequently slow speeds near the target.
+Our aim is a consistent velocity throughout the trajectory, with controlled deceleration near the goal.
+This is achieved by first calculating the unit vector towards the target for direction, then modulating its magnitude using a sigmoid function, specifically the arctangent function, to prevent overshooting and ensure stable approaching motion.
+
 \begin{equation} 
 	\vec{v}_{ATT} = \frac{\vec{x}_{EE} - \vec{x}_g}{||\vec{x}_{EE} - \vec{x}_g||} \times \frac{\arctan(k_{sigm} \; ||\vec{x}_{EE} - \vec{x}_g||) }{pi/2}
+	\label{eqq: v_att}
 \end{equation}
+
+In the above equation (eq.~\ref{eqq: v_att}), $\vec{v}_{ATT}$  represents the end effector's translational velocity towards the target, combining direction and magnitude.
+The terms $\vec{x}_{EE}$ and $\vec{x}_g$ denote the current and goal positions of the EE, respectively, in Cartesian coordinates.
+The unit vector calculation, $\frac{\vec{x}_{EE}-\vec{x}_g}{||\vec{x}_{EE}-\vec{x}_g||}$, ensures motion directed towards the target.
+Finally, the sigmoid function, particularly the arctangent component, modulates this velocity to avoid overshooting, balancing speed and precision.
+The constant $k_{sigm}$ allows us to set how close to the goal does the robot EE start slowing down.
 

### 2023-11-15 Articles/Article 1/main.tex
minor

@@
 	- p = v w \\
-	- describe constants
 	- shorten
@@
 % JUST RIGHT
-Our method innovatively employs a velocity profile to direct the end effector (EE) towards its target, marking a departure from Khatib's joint coordinate approach in favor of a Cartesian coordinate framework.
+Our method \remove{innovatively} employs inverse kinematics approach (IK) to guide the end effector (EE) towards its target, marking a departure from Khatib's joint coordinate approach in favor of a Cartesian coordinate framework.
 This is particularly beneficial in scenarios involving redundant manipulators, where determining an optimal goal joint configuration in advance is challenging.

### 2023-11-15 Articles/Article 1/main.tex
rotation primary task description

@@
+% !TeX spellcheck = en_US
 \documentclass[]{article}
+
+\usepackage{amsmath, xparse}
 
@@
 \subsection{Attractive Velocity}
-
-\add{
-	-orientation error \\ 
-	- p = v w \\
-	- shorten
-}
 
@@
 	\vec{v}_{ATT} = \frac{\vec{x}_{EE} - \vec{x}_g}{||\vec{x}_{EE} - \vec{x}_g||} \times \frac{\arctan(k_{sigm} \; ||\vec{x}_{EE} - \vec{x}_g||) }{pi/2}
-	\label{eqq: v_att}
+	\label{eq: v_att}
 \end{equation}
 
-In the above equation (eq.~\ref{eqq: v_att}), $\vec{v}_{ATT}$  represents the end effector's translational velocity towards the target, combining direction and magnitude.
+In the above equation (eq.~\ref{eq: v_att}), $\vec{v}_{ATT}$  represents the end effector's translational velocity towards the target, combining direction and magnitude.
 The terms $\vec{x}_{EE}$ and $\vec{x}_g$ denote the current and goal positions of the EE, respectively, in Cartesian coordinates.
@@
 The constant $k_{sigm}$ allows us to set how close to the goal does the robot EE start slowing down.
+
+The rotational velocity error of the EE is needed for ensuring goal orientation of the EE.
+In our approach, orientations are depicted using rotation matrices.
+Specifically, \( R \) represents the current EE orientation, while \( gR \) signifies the goal EE orientation.
+The disparity between these orientations is encapsulated by the relative rotation matrix \( dR \).
+This matrix is formulated by multiplying the goal orientation matrix \( gR \) with the transpose of the current orientation matrix \( cR^{T} \).
+To ensure that it represents a pure rotation without any scaling we than normalize the so gotten matrix .
+
+\begin{equation}
+	dR = \frac{gR \cdot cR^{T}}{||gR \cdot cR^{T}||}
+	\label{eq: rot_diff_mat}
+\end{equation}
+
+The relative rotation matrix value is converted into a quaternion, which is then logarithmically transformed to represent the rotational error vectorially:
+
+\begin{equation}
+	dR \mapsto dQ = a + b \, i + c \, j + d \, k
+	\label{eq: quat_mapsto}
+\end{equation}
+
+\begin{equation}
+	dQl = 2 \cdot \log(dQ) = al + bl \, i + cl \, j + dl \, k
+	\label{eq:quat_log}
+\end{equation}
+
+The components of this quaternion, excluding the real part, form the rotational error vector \( \omega \):
+
+\begin{equation}
+	\vec{\omega} =
+	\begin{bmatrix}
+		bl \\
+		cl \\
+		dl
+	\end{bmatrix}
+	\label{eq:rot_error_vector}
+\end{equation}
+
+To get the full velocity of the end effector (EE), \( \mathrm{velocityEE} \), we combines translational and rotational velocities, which we scale using proportional gains \( k_p \) and \( k_r \).
+
+\begin{equation}
+	\vec{\mathrm{velocityEE}} = 
+	\begin{bmatrix}
+		k_p \times \vec{v}   \\
+		k_{\omega} \times \vec{\omega}
+	\end{bmatrix}
+	\label{eq:ee_velocity}
+\end{equation}
 

### 2023-11-15 Articles/Article 1/main.tex
minor improvements

@@
 
+\add{
+	-equation references \\
+	- citations \\
+	- check if quat log equation is correct
+}
+
 % JUST RIGHT
-Our method \remove{innovatively} employs inverse kinematics approach (IK) to guide the end effector (EE) towards its target, marking a departure from Khatib's joint coordinate approach in favor of a Cartesian coordinate framework.
+Our method \remove{innovatively} employs inverse kinematics approach (IK) to guide the end effector (EE) towards its target, marking a departure from Khatib's joint coordinates approach in favor of a Cartesian coordinates framework.
 This is particularly beneficial in scenarios involving redundant manipulators, where determining an optimal goal joint configuration in advance is challenging.
@@
 \begin{equation} 
-	\vec{v}_{ATT} = \frac{\vec{x}_{EE} - \vec{x}_g}{||\vec{x}_{EE} - \vec{x}_g||} \times \frac{\arctan(k_{sigm} \; ||\vec{x}_{EE} - \vec{x}_g||) }{pi/2}
+	\vec{v} = \frac{\vec{x}_{EE} - \vec{x}_g}{||\vec{x}_{EE} - \vec{x}_g||} \times \frac{\arctan(k_{sigm} \; ||\vec{x}_{EE} - \vec{x}_g||) }{pi/2}
 	\label{eq: v_att}
@@
 
-In the above equation (eq.~\ref{eq: v_att}), $\vec{v}_{ATT}$  represents the end effector's translational velocity towards the target, combining direction and magnitude.
+In the above equation (eq.~\ref{eq: v_att}), $\vec{v}$  represents the end effector's translational velocity towards the target, combining direction and magnitude.
 The terms $\vec{x}_{EE}$ and $\vec{x}_g$ denote the current and goal positions of the EE, respectively, in Cartesian coordinates.

### 2023-11-15 Articles/Article 1/main.tex
task constraints chapter

@@
 
-\subsection{Task Constraints}
+\subsection{Constraints and Limits}
 
-\add{-equations of occupied and empty space}
+\add{
+	- what is the point of this part anyways \\
+	- but how do we limit this - this chapter only makes sense if we describe how the limits are applied
+}
+
+In robotic manipulators, effective path planning and motion control are subject to constraints related to free space, joint position limits, and joint velocity limits.
+These constraints are essential to ensure safety and practical functionality.
+
+\textbf{Free Space Constraints:} These constraints ensure the manipulator operates in unoccupied space, avoiding obstacles. They are defined as:
+\begin{equation}
+	\mathcal{C}_{free}(\vec{x}) = \begin{cases} 
+		1, & \text{if }\vec{x} \text{ is in free space} \\
+		0, & \text{otherwise}
+	\end{cases}
+	\label{eq:free_space_constraint}
+\end{equation}
+This binary representation indicates whether a point in the workspace is in free space or obstructed.
+During the execution of our algorithm, the repulsive field guides the entire manipulator into the empty task space, which in turn moves the manipulator in the obstacle free configutaion space.
+
+\textbf{Joint Position Limits:} Each joint of the manipulator has a range of motion determined by its mechanical design.
+These limits prevent the joints from exceeding their design capabilities, ensuring reliable operation:
+\begin{equation}
+	\theta_{i, min} \leq \theta_i \leq \theta_{i, max}
+	\label{eq:joint_position_constraint}
+\end{equation}
+Here, \(\theta_{i, min}\) and \(\theta_{i, max}\) represent the minimum and maximum positions for each joint.
+
+\textbf{Joint Velocity Limits:} Velocity limits are set for each joint to maintain controlled movements and prevent mechanical stress. These limits facilitate precise and stable motion:
+\begin{equation}
+	\dot{\theta}_{i, min} \leq \dot{\theta}_i \leq \dot{\theta}_{i, max}
+	\label{eq:joint_velocity_constraint}
+\end{equation}
+The values \(\dot{\theta}_{i, min}\) and \(\dot{\theta}_{i, max}\) specify the minimum and maximum velocities for each joint.
 
@@
 
-To get the full velocity of the end effector (EE), \( \mathrm{velocityEE} \), we combines translational and rotational velocities, which we scale using proportional gains \( k_p \) and \( k_r \).
+To get the full velocity of the end effector (EE), \( \mathrm{\vec{v}_{ATT}} \), we combines translational and rotational velocities, which we scale using proportional gains \( k_p \) and \( k_r \).
 
 \begin{equation}
-	\vec{\mathrm{velocityEE}} = 
+	\mathrm{\vec{v}_{ATT}} = 
 	\begin{bmatrix}

### 2023-11-15 Articles/Article 1/main.tex
minor

@@
 	- what is the point of this part anyways \\
-	- but how do we limit this - this chapter only makes sense if we describe how the limits are applied
+	- YES, BUT HOW DO WE APPLY THOSE - this chapter only makes sense if we describe how the limits are applied
 }

### 2023-11-15 Articles/Article 1/main.tex
comment

--- 2023-11-15 Articles/Article 1/main.tex
@@
 \maketitle
+
+\comment{}{for keywords use the SEO algorithm to find which terms are popular}
 
@@
 	- what is the point of this part anyways \\
-	- but how do we limit this - this chapter only makes sense if we describe how the limits are applied
+	- YES, BUT HOW DO WE APPLY THOSE - this chapter only makes sense if we describe how the limits are applied \\
+	- add $x_{start}$ and $x_{goal}$
 }

### 2023-11-17 Articles/Article 1/main.tex
part optimization algorithm

@@
 
+We use inverse kinematics algorithm to calculate joint velocities based on the primary goal, that is EE position and secondary task of obstacle avoidance.
+(INSERT: a sentence or two about ik algorithm, also a reference) We also use additional secondary tasks, such as mid-joints position and more importantly manipulability task, which allows us to avoid some singularity positions of the manipulator arm and some local minima, which are a consequence of the local optimization approach used.
+
+We start our algorithm in a start joint configuration, that can be previous goal position.
+We run optimization loop, that runs step by step from start to goal configuration, while avoiding obstacles / applieing constraints and limits.
+Each step we calculate the joint velocities and integrate them into new joint positions.
+Each of so gotten positions is one of the points on our joints trajectory.
+The priamry IK task is to take the manipulator EE to the desired goal position and orientation, we describe how to calculate cartesian velocities in the chapter ref:Attractive Velocity, which are than calculated into joint velocities.
+The size of this primary velociti is constant for the entire part of the optimization and is reduced only when approaching the goal.
+
+To avoid obstacles we calculate the avoidance velocities as descrubed in chapter ref:Repulsive Velocity.
+As one of the biggest problems of the APF Khatib method is local minima points, where the robot gets stuck in a point, where attractive and repulsive forces cancel each other out, we use transformation into the null space (INSERT: add citation of null space robotics manipulators) to only move the robot away from the obstacles in such a way as to not interfere with the main task.
+If not, the robot could not even approach the obstacles, which however is necessary in some situations, as it allows it to reach the goal points for which to reach the manipulator segments need to be positions in proximity to the obstacles.
+For this approach to work we need enough degree of manipulability left for the secondary task, so that the main / primary task doesnt overwhelm the secondary task.
+As described in the chapter bellow (ref: Attractive Velocity) we solve this problem by normalizing the primary attractive velocity.
+
+\add{this can still sometimes lead to loss of manipulability, add terciary, or secondary maipulability task}
+
+To add another level of certainity that the primary task wont overwhelm the secondary task we added the primary task execution slowdown that slows down the movement of the manipulator towards the primary goal and allows the secondary task more manipulability, so that it can better configure the arm in such a way, as to avoid the obstacles.
+
 \add{
@@
 	- secondary task of manipulation measure \\	
+	- different tasks merging equation \\
 }
@@
 
-\add{
+\alert{
 	- what is the point of this part anyways \\
 	- YES, BUT HOW DO WE APPLY THOSE - this chapter only makes sense if we describe how the limits are applied \\
-	- add $x_{start}$ and $x_{goal}$
+}
+
+\add{ 
+	- add $x_{start}$ and $x_{goal}$ \\
 }
@@
 
-The relative rotation matrix value is converted into a quaternion, which is then logarithmically transformed to represent the rotational error vectorially:
+The relative rotation matrix value is converted into a quaternion, which is then logarithmically transformed \remove{to represent the rotational error vectorially}.
+The components of this quaternion, excluding the real part, than form the rotational error vector \( \omega \).
 
@@
 
-The components of this quaternion, excluding the real part, form the rotational error vector \( \omega \):
+\alert{There needs to be an explanation why log, where is this from. Maybe a reference.}
 
@@
 \add{
+	- kinematic calculations \\
+	- null spaces calculations \\
+	- how to join multiple velocities - weights \\
+	- general overview	\\
+}
+
+***
+
+\add{	
 	- object detection is done in task domain and not c-space (more logical) \\
@@
 	- effecient calculation in dinamic environments, lacking prediction capabilities (MPC) \\
+	- good for working in "statistical, noisy" obstacle grids
 }

### 2023-11-22 Articles/Article 1/main.tex
starter repulsive

@@
 
+We compute obstacle distances in task space with our unique matrix kernel method.
+Focusing on task space is beneficial as it provides a more direct and realistic representation of the environment, leading to better spatial awareness, quicker responsiveness, easier sensor integration, and more effective obstacle avoidance.
+Our method is particularly effective when integrating noisy sensor inputs like LIDAR or depth cameras for obstacle avoidance, \alert{as it robustly filters and interprets spatial data.}
+
 \add{

### 2023-11-22 Articles/Article 1/main.tex
minor

@@
 Focusing on task space is beneficial as it provides a more direct and realistic representation of the environment, leading to better spatial awareness, quicker responsiveness, easier sensor integration, and more effective obstacle avoidance.
-Our method is particularly effective when integrating noisy sensor inputs like LIDAR or depth cameras for obstacle avoidance, \alert{as it robustly filters and interprets spatial data.}
+Our method is particularly effective when integrating noisy sensor inputs like LIDAR or depth cameras for obstacle avoidance, as it efficiently processes and utilizes this spatial information, even in the presence of noise, to accurately determine repulsive velocities.
 

### 2023-11-22 Articles/Article 1/main.tex
more of the repulsive chapter

--- 2023-11-17 Articles/Article 1/main.tex
@@
 
+%We compute obstacle distances in task space with our unique matrix kernel method.
+Focusing on task space is beneficial as it provides a more direct and realistic representation of the environment, leading to better spatial awareness, quicker responsiveness, easier sensor integration, and more effective obstacle avoidance.
+Our method is particularly effective when integrating noisy sensor inputs like LIDAR or depth cameras for obstacle avoidance, as it efficiently processes and utilizes this spatial information, even in the presence of noise, to accurately determine repulsive velocities.
+%
+% Odbojne sile so odvisne od oddaljenosti od ovir oz.
+objekotv v prostoru.
+Odbojno polje mora pasti na nič, ko smo oddaljeni od ovir za neko maksimalno razdaljo in narasti na maksimalno vrednost, ko smo tik ob oviri.
+% 
+% Izračunati želimo odbojne sile v različnih točkah interesa na manipulatorju.
+\add{citation for this approach} Te točke strategično porazdelimo po manipulatorju tako, da je celotna okolica manipulatorja pokrita z matrikami odbojnega polja.
+% 
+% Med izračunom odbojnega polja uporabimo kinematične enačbe, da pozicijo točk iz prostora notranjih spremenljivk (sklepov) robota transformiramo v prostor naloge, oz kartezični prostor. 
+% 
+% Ko imamo pozicijo točk na robotu v kartezičnem prostoru, predlagamo izračun vrednost odbojnega polja v posamezni točki na dva različna načina.
+Prostor ovir v katerem se nahaja naš robot je predstavljen kot mreža zasedenosti, v kateri imajo posamezni voxli vrednosti med 0, če se v prostoru ovira zagotovo ne nahaja in vrednost 1, če smo zagotovo prepričani, da tam je ovira, običajno pa je vrednost nekje med tema dvema skrajnostima.
+Prostor ovir ima svojo resolucijo, je diskreten, medtem ko je kartezični prostor zvezen.
+Ker za naš pristop potrebujemo poznati katere in kako zasedene celice se nahajajo v okolici našega robota predlagamo dva načina preslikave v prostor mreže zasedenosti.
+Enostavnejši pristop je, da točko na robotu v kateri želimo ugotoviti vrednost odbojnega polja enostavno preslikamo v center najbličje celice, glede na eulerjevo razdaljo.
+Včasih pa lahko zaradi te diskretizacije pride do nezveznosti.
+Zato predlagamo drugi pristop, linearno interpolacijo izračunanega odbojnega polja, da dobimo zvezno vrednost polja.
+% 
+% Ko smo pozicijo točke na manipulatorju z eno od metod preslikali v prostor ovir, lahko uporabimo naša odbojna konvolucijska jedra, za izračun odbojnih hitrosti.
+Za vsako od kartezičnih smeri (x,y,z) bomo generirali odbojno jedro, kot je opisano v poglavju "Odbojna konvolucijska jedra".
+Nato iz mreže ovir za vsako od treh smeri izrežemo matrično okno, ki je enake dimenzije kot je so jedra, z centrom v naši točki.
+Z množenjem isto ležnih elementov (Hadamardov produkt), izrezanega okna ovir in naše odbojne matrike, dobimo rezultančno matriko.
+Ko vse elemente matrike seštejemo, dobimo odbojno hitrost za posamezno karttezično smer v izbrani točki.
+% 
+% \subsubsection{Način najbližjega polja}
+% 
+% \subsubsection{Način interpolacije}
+
+In our method, we compute repulsive velocities within the task space using a novel matrix kernel multiplication approach.
+Concentrating on the task space is advantageous as it provides a more direct and realistic representation of the environment.
+This leads to improved spatial awareness, faster responsiveness, and is particularly well-suited for integrating noisy sensor inputs like LIDAR or depth cameras, efficiently processing and using this spatial information to accurately determine repulsive velocities, even amidst noise.
+
+Repulsive velocities tell the manipulator in which direction to move, so that it avoids nearby obstacles.
+These velocities drop to zero when the manipulator maintains a minimum safe distance from obstacles, and rise to their highest when it nears an obstacle, facilitating immediate evasive action.
+
+Our goal is to calculate repulsive forces at various points of interest on the manipulator.
+These points are strategically distributed across the manipulator to ensure that the entire area of the robot is covered by the matrices of the repulsive field.
+
+During the calculation of the repulsive field, kinematic equations are utilized to transform the position of points from the robot's internal joint space to the Cartesian global coordinate space.
+
+The space of obstacles in which our robot operates is represented as an occupancy grid, where individual voxels have values ranging from 0 (indicating no obstacle presence) to 1 (certainty of an obstacle), with typical values lying somewhere between these extremes.
+Since the obstacle space is discrete (has finite resolution), while the Cartesian space is continuous, we propose two methods for mapping from Cartesian space to the occupancy grid space.
+The simpler approach involves mapping the point directly to the center of the nearest occupancy grid voxel, based on Euclidean distance.
+However, this discretization can sometimes lead to discontinuities.
+Therefore, we propose a second approach: linear interpolation of the calculated repulsive field to achieve a continuous field value.
+
+Once the position of a point on the manipulator has been mapped to the obstacle space using one of these methods, we can employ our repulsive convolutional kernels to calculate repulsive velocities.
+For each Cartesian direction (x, y, z), we will generate a repulsive kernel, as described in the "Repulsive Convolutional Kernels" chapter.
+Then, from the obstacle grid, for each of the three directions, we extract a matrix cutout window of the same dimension as the kernels, centered on the point of interest.
+By performing the Hadamard (element-wise) product of the extracted obstacle grid window and our repulsive matrix, we obtain a resultant matrix.
+Summing all the elements of this matrix gives us the repulsive velocity for each of the three Cartesian directions at the selected point.
+
+ \add{
+ 	-there is an option of interpolation plot different matrixes, that are sligthely moved ... \\
+ 	-in results maybe compare smoothenss of interpolated and non-interpolated method
+ }
+
 \add{
@@
 	- the limitations of local search \\
+	- good for parellelization \\
 	- number of parameters that need to be tuned (are there actually that many?)

### 2023-11-24 Articles/Article 1/main.tex
better equation for poi, other

@@
 Concentrating on the task space is advantageous as it provides a more direct and realistic representation of the environment.
-This leads to improved spatial awareness, faster responsiveness, and is particularly well-suited for integrating noisy sensor inputs like LIDAR or depth cameras, efficiently processing and using this spatial information to accurately determine repulsive velocities, even amidst noise.
+\replace{This leads to improved spatial awareness, faster responsiveness, and is particularly well-suited for integrating noisy sensor inputs like LIDAR or depth cameras, efficiently processing and using this spatial information to accurately determine repulsive velocities, even amidst noise.}{add some references for this statement}
 
@@
 Our goal is to calculate repulsive forces at various points of interest on the manipulator.
-These points are strategically distributed across the manipulator to ensure that the entire area of the robot is covered by the matrices of the repulsive field.
+These points \add{add reference} are strategically distributed across the manipulator to ensure that the entire area of the robot is covered by the matrices of the repulsive field.
 
@@
 Summing all the elements of this matrix gives us the repulsive velocity for each of the three Cartesian directions at the selected point.
+
+\begin{equation}
+	\mathrm{\vec{v}_{poi}} = 
+	\begin{bmatrix}
+	\sum_{(i, j, k)} W_x \odot K_x \\
+	\sum_{(i, j, k)} W_y \odot K_y \\
+	\sum_{(i, j, k)} W_z \odot K_z	
+	\end{bmatrix}
+	\label{eq:poi_velocity}
+\end{equation}
 

### 2023-11-24 Articles/Article 1/main.tex
\subsubsection{Kernel Selection}

@@
 Once the position of a point on the manipulator has been mapped to the obstacle space using one of these methods, we can employ our repulsive convolutional kernels to calculate repulsive velocities.
-For each Cartesian direction (x, y, z), we will generate a repulsive kernel, as described in the "Repulsive Convolutional Kernels" chapter.
+For each Cartesian direction $(x, y, z)$, we will generate a repulsive kernel, as described in the section~\ref{chap:kernels}.
 Then, from the obstacle grid, for each of the three directions, we extract a matrix cutout window of the same dimension as the kernels, centered on the point of interest.
@@
 Summing all the elements of this matrix gives us the repulsive velocity for each of the three Cartesian directions at the selected point.
+
+\comment{}{PLOT: POI on the ROBOT + ONE OR MORE KERNELS ON THE ROBOT}
 
@@
 \subsubsection{Kernel Selection}
+\label{chap:kernels}
+
+The fundamental concept of our directional kernels lies in computing the repulsive field individually for each direction within the Cartesian coordinate system.
+Our filters can be likened to a variation of the Sobel \add{reference} operator, a 2D convolutional \add{reference} filter frequently utilized in computer vision for calculating image gradients at specific points.
+
+Our kernels are designed as three-dimensional structures with a primary axis aligned along a specific Cartesian direction, corresponding to the calculated repulsive velocity.
+The two secondary axes are orthogonal to this primary axis.
+The distribution of values along the primary axis is inversely symmetric, exhibiting positive values on one side and negative values on the other, with the magnitude reaching zero at the kernel's center.
+The nature of the increase in magnitude along the primary axis of the kernel defines the shape of the repulsive velocity field, determining how the repulsive velocity escalates as the manipulator approaches an obstacle.
+
+The length of the primary axis is critical, as it dictates the detection range for obstacles. Longer kernels can detect obstacles further away
+from the robot, essentially extending the ’safety zone’ around the robot.
+If the primary axis is too long, it can lead to extra calculations and may cause the robot to unnecessarily avoid obstacles that aren't in its immediate path, making its movement and path planning less efficient.
+A kernel with a primary axis that is too short might restrict the robot's ability to maneuver, detecting obstacles potentially too late, compromising the robots capacity to avoid obstacles effectively.
+Moreover, it is essential for the magnitudes at the kernel's periphery to be minimal, promoting a smooth increase in repulsive velocity rather than a sudden spike.
+
+In the orthogonal axes, the magnitude of values diminishes with increasing distance from the core of the kernel, nearing zero at the edges \add{using gaussian distribution}.
+This gradation is vital for a smooth velocity transition when obstacles emerge in the kernels range.
+The peak magnitude is located at the kernel's core, along the primary axis.
+The polarity of the values is determined by their position relative to the primary axis.
+The length of the kernel along the secondary axes influences the peripheral detection range for obstacles.
+Excessively large kernels may generate repulsive velocities unnecessarily, whereas too narrow kernels might only detect obstacles aligned directly with the Cartesian direction in the point of interest.
+Therefore, when selecting the width and height of the kernel, one must consider the density of the neighboring points of interest on the robot, ensuring that their collective fields adequately cover the manipulator's surrounding area.
+
+\comment{}{PLOT: kernel with gaussian functions}
+\comment{}{EQUATION: gaussian equations}
+
+\add{- choice of velocity profile (gaussian, linear ... maybe could move this part to implementation / experiments)}
 

### 2023-11-27 Articles/Article 1/main.tex
added sentence about sobel difference
Primary Task Slowdown

@@
 We use inverse kinematics algorithm to calculate joint velocities based on the primary goal, that is EE position and secondary task of obstacle avoidance.
-(INSERT: a sentence or two about ik algorithm, also a reference) We also use additional secondary tasks, such as mid-joints position and more importantly manipulability task, which allows us to avoid some singularity positions of the manipulator arm and some local minima, which are a consequence of the local optimization approach used.
+\add{(INSERT: a sentence or two about ik algorithm, also a reference)} We also use additional secondary tasks, such as mid-joints position and more importantly manipulability task, which allows us to avoid some singularity positions of the manipulator arm and some local minima, which are a consequence of the local optimization approach used.
 
@@
 
-To avoid obstacles we calculate the avoidance velocities as descrubed in chapter ref:Repulsive Velocity.
-As one of the biggest problems of the APF Khatib method is local minima points, where the robot gets stuck in a point, where attractive and repulsive forces cancel each other out, we use transformation into the null space (INSERT: add citation of null space robotics manipulators) to only move the robot away from the obstacles in such a way as to not interfere with the main task.
+To avoid obstacles we calculate the avoidance velocities as descrubed in section~\ref{chap:repulsive velocity}.
+As one of the biggest problems of the APF Khatib method is local minima points, where the robot gets stuck in a point, where attractive and repulsive forces cancel each other out, we use transformation into the null space \add{(INSERT: add citation of null space robotics manipulators)} to only move the robot away from the obstacles in such a way as to not interfere with the main task.
 If not, the robot could not even approach the obstacles, which however is necessary in some situations, as it allows it to reach the goal points for which to reach the manipulator segments need to be positions in proximity to the obstacles.
 For this approach to work we need enough degree of manipulability left for the secondary task, so that the main / primary task doesnt overwhelm the secondary task.
-As described in the chapter bellow (ref: Attractive Velocity) we solve this problem by normalizing the primary attractive velocity.
+As described in the section~\ref{chap:attractive velocity} we solve this problem by normalizing the primary attractive velocity.
 
@@
 
-To add another level of certainity that the primary task wont overwhelm the secondary task we added the primary task execution slowdown that slows down the movement of the manipulator towards the primary goal and allows the secondary task more manipulability, so that it can better configure the arm in such a way, as to avoid the obstacles.
+Another way we make sure that the primary task doesn't take over the avoidance task is using execution slowdown as described in section~\ref{chap:primary slowdown}.
 
@@
 	- damped least squares \\
-	- task slowdown option \\
 	- secondary task of manipulation measure \\	
@@
 }
+
+\subsubsection{Primary Task Slowdown}
+\label{chap:primary slowdown}
+
+To ensure the primary task doesn't overpower the secondary task, we've integrated a primary task execution slowdown.
+This mechanism can reduce the manipulator's velocity towards its primary goal, leaving more maneuverability space for the secondary tasks.
+
+\begin{equation}
+	\label{eq:slowdown}
+	\xi_{p}=
+	\frac{1}{1 + \kappa_{\text{sec}} \left(\frac{1}{\delta_{min}}\right)}
+\end{equation}
+
+The slowdown factor (eq.~\ref{eq:slowdown}) is influenced by the constant \( \kappa_{\text{sec}} \) and the robot's minimum distance from an obstacle \( \delta_{\text{min}} \).
+We calculate the minimal distance in the repulsive velocities phase of the algorithm, as explained in section~\ref{chap:repulsive velocity}.
+As per the equation, a large \( \delta_{\text{min}} \) minimizes the slowdown effect, allowing uninterrupted primary task execution.
+Conversely, a small \( \delta_{\text{min}} \) increases the slowdown, by making the $\xi_{p}$ factor smaller, giving the secondary task more time for corrective actions.
 
@@
 
-In robotic manipulators, effective path planning and motion control are subject to constraints related to free space, joint position limits, and joint velocity limits.
+%In robotic manipulators, effective path planning and motion control are subject to constraints related to free space, joint position limits, and joint velocity limits.
 These constraints are essential to ensure safety and practical functionality.
-
-\textbf{Free Space Constraints:} These constraints ensure the manipulator operates in unoccupied space, avoiding obstacles. They are defined as:
-\begin{equation}
-	\mathcal{C}_{free}(\vec{x}) = \begin{cases} 
-		1, & \text{if }\vec{x} \text{ is in free space} \\
-		0, & \text{otherwise}
-	\end{cases}
-	\label{eq:free_space_constraint}
-\end{equation}
-This binary representation indicates whether a point in the workspace is in free space or obstructed.
+%
+%\textbf{Free Space Constraints:} These constraints ensure the manipulator operates in unoccupied space, avoiding obstacles. They are defined as:
+%\begin{equation}
+%	\mathcal{C}_{free}(\vec{x}) = \begin{cases} 
+%		1, & \text{if }\vec{x} \text{ is in free space} \\
+%		0, & \text{otherwise}
+%	\end{cases}
+%	\label{eq:free_space_constraint}
+%\end{equation}
+%This binary representation indicates whether a point in the workspace is in free space or obstructed.
 During the execution of our algorithm, the repulsive field guides the entire manipulator into the empty task space, which in turn moves the manipulator in the obstacle free configutaion space.
-
-\textbf{Joint Position Limits:} Each joint of the manipulator has a range of motion determined by its mechanical design.
+%
+%\textbf{Joint Position Limits:} Each joint of the manipulator has a range of motion determined by its mechanical design.
 These limits prevent the joints from exceeding their design capabilities, ensuring reliable operation:
-\begin{equation}
-	\theta_{i, min} \leq \theta_i \leq \theta_{i, max}
-	\label{eq:joint_position_constraint}
-\end{equation}
-Here, \(\theta_{i, min}\) and \(\theta_{i, max}\) represent the minimum and maximum positions for each joint.
-
-\textbf{Joint Velocity Limits:} Velocity limits are set for each joint to maintain controlled movements and prevent mechanical stress. These limits facilitate precise and stable motion:
-\begin{equation}
-	\dot{\theta}_{i, min} \leq \dot{\theta}_i \leq \dot{\theta}_{i, max}
-	\label{eq:joint_velocity_constraint}
-\end{equation}
-The values \(\dot{\theta}_{i, min}\) and \(\dot{\theta}_{i, max}\) specify the minimum and maximum velocities for each joint.
+%\begin{equation}
+%	\theta_{i, min} \leq \theta_i \leq \theta_{i, max}
+%	\label{eq:joint_position_constraint}
+%\end{equation}
+%Here, \(\theta_{i, min}\) and \(\theta_{i, max}\) represent the minimum and maximum positions for each joint.
+%
+%\textbf{Joint Velocity Limits:} Velocity limits are set for each joint to maintain controlled movements and prevent mechanical stress. These limits facilitate precise and stable motion:
+%\begin{equation}
+%	\dot{\theta}_{i, min} \leq \dot{\theta}_i \leq \dot{\theta}_{i, max}
+%	\label{eq:joint_velocity_constraint}
+%\end{equation}
+%The values \(\dot{\theta}_{i, min}\) and \(\dot{\theta}_{i, max}\) specify the minimum and maximum velocities for each joint.
 
 \subsection{Attractive Velocity}
+\label{chap:attractive velocity}
 
@@
 \subsection{Repulsive Velocity}
+\label{chap:repulsive velocity}
 
@@
 Our filters can be likened to a variation of the Sobel \add{reference} operator, a 2D convolutional \add{reference} filter frequently utilized in computer vision for calculating image gradients at specific points.
+However contrary to the Sobel operator, our filter is not calculating the gradients of the matrix, as the inputs are weighted depending on the distance from the center of the kernel.
 

### 2023-11-28 Articles/Article 1/main.tex
minor
interpolation, comments

@@
 
+\comment{}{optimization algorithm diagram}
+
 We start our algorithm in a start joint configuration, that can be previous goal position.
@@
 \add{
-	- optimization algorithm \\
-	- robot kinematics (include the exact-reduced method) \\
-	- primary task of distance goal \\
-	- secondary task of repulsive field \\
-	- damped least squares \\
+	- robot kinematics the exact-reduced method \\
 	- secondary task of manipulation measure \\	
@@
 
-\subsubsection{Primary Task Slowdown}
-\label{chap:primary slowdown}
+\subsubsection{Exact Reduced Inverse Kinematics Method}
 
-To ensure the primary task doesn't overpower the secondary task, we've integrated a primary task execution slowdown.
-This mechanism can reduce the manipulator's velocity towards its primary goal, leaving more maneuverability space for the secondary tasks.
-
-\begin{equation}
-	\label{eq:slowdown}
-	\xi_{p}=
-	\frac{1}{1 + \kappa_{\text{sec}} \left(\frac{1}{\delta_{min}}\right)}
-\end{equation}
-
-The slowdown factor (eq.~\ref{eq:slowdown}) is influenced by the constant \( \kappa_{\text{sec}} \) and the robot's minimum distance from an obstacle \( \delta_{\text{min}} \).
-We calculate the minimal distance in the repulsive velocities phase of the algorithm, as explained in section~\ref{chap:repulsive velocity}.
-As per the equation, a large \( \delta_{\text{min}} \) minimizes the slowdown effect, allowing uninterrupted primary task execution.
-Conversely, a small \( \delta_{\text{min}} \) increases the slowdown, by making the $\xi_{p}$ factor smaller, giving the secondary task more time for corrective actions.
+\add{
+	- cite žlajpah article \\
+	- explain how it works \\
+	- equation \\
+	- scalar simplification	
+}
 
@@
 
+\subsubsection{Primary Task Slowdown}
+\label{chap:primary slowdown}
+
+To ensure the primary task doesn't overpower the secondary task, we've integrated a primary task execution slowdown.
+This mechanism can reduce the manipulator's velocity towards its primary goal, leaving more maneuverability space for the secondary tasks.
+
+\begin{equation}
+	\label{eq:slowdown}
+	\xi_{p}=
+	\frac{1}{1 + \kappa_{\text{sec}} \left(\frac{1}{\delta_{min}}\right)}
+\end{equation}
+
+\alert{Is this equation correct, maybe add a singularity damping factor.}
+
+The slowdown factor (eq.~\ref{eq:slowdown}) is influenced by the constant \( \kappa_{\text{sec}} \) and the robot's minimum distance from an obstacle \( \delta_{\text{min}} \).
+We calculate the minimal distance in the repulsive velocities phase of the algorithm, as explained in section~\ref{chap:repulsive velocity}.
+As per the equation, a large \( \delta_{\text{min}} \) minimizes the slowdown effect, allowing uninterrupted primary task execution.
+Conversely, a small \( \delta_{\text{min}} \) increases the slowdown, by making the $\xi_{p}$ factor smaller, giving the secondary task more time for corrective actions.
+
 \subsection{Repulsive Velocity}
@@
 	- PLOT: kernel field shape 	\\
-	- interpolation of the repulsive field \\
 	- what if there are obstacles behind wall (usually not the case, depth sensors show only thin walls, some noise doesnt matter, non-linear kernel, possible additional pre-convolution to convert obstacle grid to edges) \\
@@
 
+Želimo, da je gibanje robota zvezno in ne skokovito.
+Da lahko to dosežemo je potrebno, da vsi hitrostni prispevki, ki vplivajo na robota prehajajo med seboj zvezno.
+Ker imamo mrežo ovir diskretno definirano, to ni res.
+Z zvišanjem resolucije polja ovir in se lahko zveznosti teoretično neskončno približamo, vendar pa imamo v realnosti običajno opravka z končno resolucijo.
+Da je velikost hitrosti kljub temu zvezna, ko prehajamo z točko interesa POI iz ene celice mreže ovir v drugo, uporabimo trilinearno interpolacijo, s katero za vsako od hitrosti v treh kartezičnih smeri posebaj interpoliramo hitrosti in dobimo zvezno in neskokovito linearno aproksimacijo pri prehodu med celicami.
+
+\add{kako pridobimo vrednosti v okoliških celicah}
+\add{zakaj vrednosti definiramo v centrih voxlov (celicah)}
+\add{enačbe trilinearne interpolacije}
+
+\add{	- interpolation of the repulsive field \\}
+
 \section{Implementation}

### 2023-11-28 Articles/Article 1/main.tex
interpolation text

@@
 
-\subsubsection{Interpolation}
+\subsubsection{Interpolation of the Repulsive Velocity}
 
-Želimo, da je gibanje robota zvezno in ne skokovito.
+%Želimo, da je gibanje robota zvezno in ne skokovito.
 Da lahko to dosežemo je potrebno, da vsi hitrostni prispevki, ki vplivajo na robota prehajajo med seboj zvezno.
@@
 Da je velikost hitrosti kljub temu zvezna, ko prehajamo z točko interesa POI iz ene celice mreže ovir v drugo, uporabimo trilinearno interpolacijo, s katero za vsako od hitrosti v treh kartezičnih smeri posebaj interpoliramo hitrosti in dobimo zvezno in neskokovito linearno aproksimacijo pri prehodu med celicami.
+%
+%Ko poznamo koordinate naše POI izračunamo za vsako od točk x,y in z koordinate zaokrožene navzgor in navzdol, da dobimo 8 okoliških celic. 
+%
+%Ko imamo poznane celice, uporabimo našo poznano konvolucijsko metodo, da izračunamo x,y in z komponento vektorja odbojne hitrosti za vsako od celic.
+%
+%Potem izvedemo za vsako od smeri vektorja odbojne hitrosti posebaj trilinearno interpolacijo odbojne hitrosti.
+Pri tem upoštevamo, da se izraučunane vrednosti odbojnih hitrosti nahajajo v centrih celic, pri interpolaciji torej zamaknemo koordinate poznanih vrenodsti za polovico resolucije mreže ovir v vsaki od smeri.
+Rezultat je vektor odbojne hitrosti, ki zvezno prehaja med diskretnimi vrednostmi izračunanimi v diskretnih točkah mreže ovir.
+%
 
-\add{kako pridobimo vrednosti v okoliških celicah}
-\add{zakaj vrednosti definiramo v centrih voxlov (celicah)}
+It is essential that the velocity contributions affecting the robot change smoothly.
+However, since our obstacle grid is discretely defined, achieving perfect continuity can be challenging.
+Increasing the resolution of the obstacle field can theoretically bring us closer to continuous behavior, but in practice, we are constrained by finite resolution.
+To ensure that the velocity remains continuous when transitioning from one cell of the obstacle grid to another at a point of interest (POI), we employ trilinear interpolation.
+This technique allows for a smooth and continuous linear approximation of velocities in all three Cartesian directions (x, y, and z) as the POI moves between cells.
+
+Firstly, from the coordinates of our POI, we calculate the coordinates of the eight surrounding cells in the obstacle grid, by rounding the coordinates up and down in x, y, and z directions.
+
+Once we have identified these cells, we apply our established kernel matrix method to compute the x, y, and z components of the repulsive velocity vector for each of the cells.
+
+Subsequently, we separately perform trilinear interpolation for each direction of the repulsive velocity vector.
+It is important to consider that the calculated repulsive velocity values are located at the centers of the cells.
+Therefore, during interpolation, we shift the values of the cells coordinates by half the resolution of the obstacle grid for each direction.
+This results in a repulsive velocity vector that transitions smoothly between the discrete values calculated at distinct points in the obstacle grid.
+
 \add{enačbe trilinearne interpolacije}
-
-\add{	- interpolation of the repulsive field \\}
 

### 2023-11-29 Articles/Article 1/main.tex
.
added interpolation equations

@@
 \todo{1-2 PAGES}
+
+\alert{Maybe add a sentence about computer effeciency of the method, as it is not dependent on the number of obstacles.}
 
@@
 
-Firstly, from the coordinates of our POI, we calculate the coordinates of the eight surrounding cells in the obstacle grid, by rounding the coordinates up and down in x, y, and z directions.
+Firstly we use our kernel matrix multiplication method, to calculate the repulsive velocity values for the eight surrounding cells that are closest to the chosen POI.
+We get the indexes of the surrounding cells by first scaling the POI position by grid resolution and than rounding the position to the nearest lower and upper integer positions.
+
+\begin{equation}
+ \vec{P} = \vec{p}_{\mathrm{POI}} \times\Delta \mathrm{grid}
+\end{equation}
+
+%\begin{equation}
+% X = \left[ \lfloor p_{x} \rfloor, \lceil p_{x} \rceil \right] 
+%\end{equation}
+%
+%\begin{equation}
+% Y = \left[ \lfloor p_{y} \rfloor, \lceil p_{y} \rceil \right] 
+%\end{equation}
+%
+% \begin{equation}
+% Z = \left[ \lfloor p_{z} \rfloor, \lceil p_{z} \rceil \right] 
+%\end{equation}
+
+\begin{equation}
+	\vec{P} =
+	\begin{bmatrix}
+		X \\
+ 		Y \\
+  		Z \\
+	\end{bmatrix}
+	=
+	\begin{bmatrix}
+		\; \lfloor \vec{p}_{\mathrm{POI}}(1) \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(1) \rceil \;  \\
+		\; \lfloor \vec{p}_{\mathrm{POI}}(2) \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(2) \rceil \; \\
+		\; \lfloor \vec{p}_{\mathrm{POI}}(3) \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(3) \rceil \; 
+	\end{bmatrix}
+\end{equation}
+
+\begin{equation}
+	\vec{V}rep_{xyz} = \mathrm{calc\_rep\_vel}(X[i], Y[j], Z[k]) \quad \forall i, j, k \in \{1, 2\}
+\end{equation}
+
+%\begin{equation}
+%	\Delta X = \frac{\left( P_x - X(1) \right)}{\left( X(2) - X(1) \right)}
+%\end{equation}
+%
+%\begin{equation}
+%	\Delta Y = \frac{\left( P_y - Y(1) \right)}{\left( Y(2) - Y(1) \right)}
+%\end{equation}
+%
+%\begin{equation}
+%	\Delta Z = \frac{\left( P_z - Z(1) \right)}{\left( Z(2) - Z(1) \right)}
+%\end{equation}
+
+\begin{equation}
+	\Delta \vec{P} =
+	\begin{bmatrix}
+		\Delta x \\
+		\Delta y \\
+		\Delta z		
+	\end{bmatrix}
+	=
+	\begin{bmatrix}
+		 \frac{\left( P_x - X(1) \right)}{\left( X(2) - X(1) \right)} \\
+		 \frac{\left( P_y - Y(1) \right)}{\left( Y(2) - Y(1) \right)} \\
+		 \frac{\left( P_z - Z(1) \right)}{\left( Z(2) - Z(1) \right)} \\
+ 	\end{bmatrix}
+\end{equation}
+
+%\begin{equation}
+%	\vec{V}_{rep,00} = \vec{V}_{rep,000}(1 - x_d) + \vec{V}_{rep,100}x_d
+%\end{equation}
+%\begin{equation}
+%	\vec{V}_{rep,01} = \vec{V}_{rep,001}(1 - x_d) + \vec{V}_{rep,101}x_d
+%\end{equation}
+%\begin{equation}
+%	\vec{V}_{rep,10} = \vec{V}_{rep,010}(1 - x_d) + \vec{V}_{rep,110}x_d
+%\end{equation}
+%\begin{equation}
+%	\vec{V}_{rep,11} = \vec{V}_{rep,011}(1 - x_d) + \vec{V}_{rep,111}x_d
+%\end{equation}
+
+\begin{equation}
+	\vec{V}rep_{xyz,jk} = \vec{V}rep_{xyz,0jk}(1 - \Delta x) + \vec{V}rep_{xyz,1jk} \, \Delta x \quad \forall \, j, k \in \{1, 2\}
+\end{equation}
+
+%\begin{equation}
+%	\vec{V}rep_{xyz,0} = \vec{V}rep_{xyz,00}(1 - \Delta y) + \vec{V}rep_{xyz,10} \, \Delta y
+%\end{equation}
+%
+%\begin{equation}
+%	\vec{V}rep_{xyz,1} = \vec{V}rep_{xyz,01}(1 - \Delta y) + \vec{V}rep_{xyz,11} \, \Delta y
+%\end{equation}
+
+\begin{equation}
+	\vec{V}rep_{xyz,k} = \vec{V}rep_{xyz,0k}(1 - \Delta y) + \vec{V}rep_{xyz,1k} \, \Delta y \quad \forall \, k \in \{1, 2\}
+\end{equation}
+
+\begin{equation}
+	\vec{V}rep_{xyz} = \vec{V}rep_{xyz,0}(1 - \Delta z) + \vec{V}rep_{xyz,1} \, \Delta z 
+\end{equation}
+
+\comment{}{PLOT: surrounding cells}
 
@@
 Subsequently, we separately perform trilinear interpolation for each direction of the repulsive velocity vector.
+
 It is important to consider that the calculated repulsive velocity values are located at the centers of the cells.

### 2023-11-30 Articles/Article 1/main.tex
added interpolation

@@
 To ensure that the velocity remains continuous when transitioning from one cell of the obstacle grid to another at a point of interest (POI), we employ trilinear interpolation.
-This technique allows for a smooth and continuous linear approximation of velocities in all three Cartesian directions (x, y, and z) as the POI moves between cells.
+\remove{This technique allows for a smooth and continuous linear approximation of velocities in all three Cartesian directions (x, y, and z) as the POI moves between cells.}
 
-Firstly we use our kernel matrix multiplication method, to calculate the repulsive velocity values for the eight surrounding cells that are closest to the chosen POI.
-We get the indexes of the surrounding cells by first scaling the POI position by grid resolution and than rounding the position to the nearest lower and upper integer positions.
+We start by scalling the coordinates of POI into the grid koordinate system, by multiplying it by grid resolution (eq.~\ref{eq:scale grid}).
 
 \begin{equation}
+	\label{eq:scale grid}
 	\vec{P} = \vec{p}_{\mathrm{POI}} \times\Delta \mathrm{grid}
 \end{equation}
+
+We get the indexes of the surrounding cells by first scaling the POI position by grid resolution and than rounding the position to the nearest lower and upper integer positions (eq.~\ref{eq:floor and ceil}).
 
@@
 	\begin{equation}
+		\label{eq:floor and ceil}
 		\vec{P} =
@@
 
+Once we got the indexes of the eight surrounding cells of our POI, we use our kernel matrix multiplication method, to calculate the 3x1 repulsive velocity vectors for all the cells (eq.~\ref{eq: calc rep vel}).
+
 \begin{equation}
-	\vec{V}rep_{xyz} = \mathrm{calc\_rep\_vel}(X[i], Y[j], Z[k]) \quad \forall i, j, k \in \{1, 2\}
+	\label{eq: calc rep vel}
+	\vec{V}rep_{xyz,ijk} = \mathrm{calc\_rep\_vel}(X[i], Y[j], Z[k]) \quad \forall i, j, k \in \{1, 2\}
 \end{equation}
 
+Trilinear interpolation method works on a 3-dimensional regular grid.
+Before we can start with the interpolation we need to calculate the distance between POI and smaller coordinates of the cells where we calculated the repulsive velocities (eq.~\ref{eq: deltas interp}).
+\replace{Since the repulsive values we calculate for the cells are alligned with the centers of the cells, we need to move before the interpolation the positions of known grid points by half of the cell width.}{The calculated repulsive velocity values are located at the centers of the cells.
+Therefore, before interpolation, we shift the values of the cells coordinates by half the resolution of the obstacle grid for each direction.
+}
+
 %\begin{equation}
-%	\Delta X = \frac{\left( P_x - X(1) \right)}{\left( X(2) - X(1) \right)}
+%	\Delta X = \frac{\left( P_x - \left( X(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( X(2) - X(1) \right)}
 %\end{equation}
@@
 %\begin{equation}
-%	\Delta Y = \frac{\left( P_y - Y(1) \right)}{\left( Y(2) - Y(1) \right)}
+%	\Delta Y = \frac{\left( P_y - \left( Y(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Y(2) - Y(1) \right)}
 %\end{equation}
@@
 %\begin{equation}
-%	\Delta Z = \frac{\left( P_z - Z(1) \right)}{\left( Z(2) - Z(1) \right)}
+%	\Delta Z = \frac{\left( P_z - \left( Z(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Z(2) - Z(1) \right)}
 %\end{equation}
@@
 \begin{equation}
+	\label{eq: deltas interp}
 	\Delta \vec{P} =
@@
 	\begin{bmatrix}
-		 \frac{\left( P_x - X(1) \right)}{\left( X(2) - X(1) \right)} \\
-		 \frac{\left( P_y - Y(1) \right)}{\left( Y(2) - Y(1) \right)} \\
-		 \frac{\left( P_z - Z(1) \right)}{\left( Z(2) - Z(1) \right)} \\
+		 \frac{\left( P_x - \left( X(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( X(2) - X(1) \right)} \\
+		 \frac{\left( P_y - \left( Y(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Y(2) - Y(1) \right)} \\
+		 \frac{\left( P_z - \left( Z(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Z(2) - Z(1) \right)} \\
  	\end{bmatrix}
 \end{equation}
+
+The result of the interpolation is independent of the order of the operations. We first interpolate along the x-axis, followed by along the y-axis and finally along z-axis.
 
@@
 \begin{equation}
+	\label{eq: interp x}
 	\vec{V}rep_{xyz,jk} = \vec{V}rep_{xyz,0jk}(1 - \Delta x) + \vec{V}rep_{xyz,1jk} \, \Delta x \quad \forall \, j, k \in \{1, 2\}
@@
 \begin{equation}
+	\label{eq: interp y}
 	\vec{V}rep_{xyz,k} = \vec{V}rep_{xyz,0k}(1 - \Delta y) + \vec{V}rep_{xyz,1k} \, \Delta y \quad \forall \, k \in \{1, 2\}
@@
 \begin{equation}
+	\label{eq: interp z}
 	\vec{V}rep_{xyz} = \vec{V}rep_{xyz,0}(1 - \Delta z) + \vec{V}rep_{xyz,1} \, \Delta z 
@@
 
-\comment{}{PLOT: surrounding cells}
+The final result is a repulsive velocity vector that transitions smoothly between the discrete values calculated at distinct points in the obstacle grid.
 
-Once we have identified these cells, we apply our established kernel matrix method to compute the x, y, and z components of the repulsive velocity vector for each of the cells.
-
-Subsequently, we separately perform trilinear interpolation for each direction of the repulsive velocity vector.
-
-It is important to consider that the calculated repulsive velocity values are located at the centers of the cells.
-Therefore, during interpolation, we shift the values of the cells coordinates by half the resolution of the obstacle grid for each direction.
-This results in a repulsive velocity vector that transitions smoothly between the discrete values calculated at distinct points in the obstacle grid.
-
-\add{enačbe trilinearne interpolacije}
+\comment{}{PLOT: surrounding cells, interp grid}
 

### 2023-12-18 Articles/Article 1/main.tex
.
ALERT ABOUT BUG
2d platform extensions

@@
 
-\subsection{Optimization Algorithm}
+\subsection{Kinematic Control}
 
@@
 Conversely, a small \( \delta_{\text{min}} \) increases the slowdown, by making the $\xi_{p}$ factor smaller, giving the secondary task more time for corrective actions.
+
+\alert{THIS WILL TOTALLY STOP MOVEMENT IN TIGHT SPACES, NEED ADDITIONAL PART. (Maybe use some kind of derivative of Rep Field).}
 
@@
 	- PLOT: kernel on robot graphics \\
+	- IMPORTANT: maybe add 2D mobile platform results \\
 }

### 2023-12-18 Articles/Article 1/main.tex
more commetns

@@
 
-\alert{There needs to be an explanation why log, where is this from. Maybe a reference.}
+\alert{There needs to be an explanation why log, where is this from. Maybe a reference. cite: DMP Quaternions article Petrič, Žlajpah, Ude}
 
@@
 
-\alert{Is this equation correct, maybe add a singularity damping factor.}
+\alert{Is this equation correct, ADD a singularity damping factor. Maybe even some variation of a changing damping.}
 
@@
 
-\add{- choice of velocity profile (gaussian, linear ... maybe could move this part to implementation / experiments)}
+\add{- choice of velocity profile (gaussian, linear ... maybe could move this part to implementation / experiments) \\}
+\add{- maybe first debluring - line convolution operation (if thick walls) - however this might move wall back, which could cause crash }
 
@@
 	- IMPORTANT: maybe add 2D mobile platform results \\
+	- use pybullet or some engine that tells you distance from obstacles and run the same task, compare distances after execution, smoothness of paths (MAYBE FOR PUBLICATION, NOT CONFERENCE)
 }

### 2023-12-19 Code/Full_RTConvolution_Multiple_Points.m
Article1FieldVisualization

@@ -95,7 +95,7 @@ rep_kernels = REP_kernels();
 
 % obstacles distance kernel
 % dist_kernel = euclidian_kernel_3D(61, 61, 61);
-dist_kernel = gaussian_kernel_3d(61, 61, 61, 10);
+% dist_kernel = gaussian_kernel_3d(61, 61, 61, 10);
 
 
 

### 2023-12-19 Code/REP_field_calculation.m

@@ -54,7 +54,7 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
 
         % if x and y and z are center
         if mod(center_point(1),1) < no_interp_dist && mod(center_point(2),1) < no_interp_dist && mod(center_point(3),1) < no_interp_dist
-            rep_values = calculateFieldVector(center_point)
+            rep_values = calculateFieldVector(center_point);
  
         % if x and y are in center
         elseif mod(center_point(1),1) < no_interp_dist && mod(center_point(2),1) < no_interp_dist

### 2023-12-19 Code/REP_kernels.m

@@ -31,8 +31,8 @@ function [kernels] = REP_kernels()
     kernel_y = directional_kernel_3d('y', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, 'gaussian')/200;
     kernel_z = directional_kernel_3d('z', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, 'gaussian')/200;
 
-    figure()
-    plot(kernel_x(5,:,5))
+%     figure()
+%     plot(kernel_x(5,:,5))
     
     kernels = {kernel_x, kernel_y, kernel_z};
 

### 2023-12-20 Code/REP_field_calculation.m
removed incorrect interpolation grid coordinates bias

@@ -29,7 +29,7 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
     % --------------------------------------------------------------
 
     % use interpolation
-    interpolation_mode = true; % true / false
+    interpolation_mode = false; % true / false
 
     % set values outside the known grid to
     box_value = 0;
@@ -73,7 +73,7 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
             end
 
             % move sampled points coordinates to the mid of the cells
-            Z = [floor(center_point(3)) ceil(center_point(3))]+0.5;
+            Z = [floor(center_point(3)) ceil(center_point(3))];
 
             % interpolate
             rep_values = [interp1(Z,Vx,center_point(3)), interp1(Z,Vy,center_point(3)), interp1(Z,Vz,center_point(3))];
@@ -99,7 +99,7 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
             end
 
             % move sampled points coordinates to the mid of the cells
-            X = [floor(center_point(1)) ceil(center_point(1))] + 0.5;
+            X = [floor(center_point(1)) ceil(center_point(1))];
 
             % interpolate
             rep_values = [interp1(X,Vx,center_point(1)), interp1(X,Vy,center_point(1)), interp1(X,Vz,center_point(1))];
@@ -125,7 +125,7 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
             end
 
             % move sampled points coordinates to the mid of the cells
-            Y = [floor(center_point(2)) ceil(center_point(2))] + 0.5;
+            Y = [floor(center_point(2)) ceil(center_point(2))];
 
             % interpolate
             rep_values = [interp1(Y,Vx,center_point(2)), interp1(Y,Vy,center_point(2)), interp1(Y,Vz,center_point(2))];
@@ -162,6 +162,8 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
 
 
         end
+
+        
 
 
 

### 2023-12-20 Code/REP_kernels.m

@@ -21,9 +21,9 @@ function [kernels] = REP_kernels()
     %    kernels - A cell array containing the generated directional kernels.
 
     % set directional kernels
-    sigma = 5;
-    kernel_length = 31;
-    kernel_width = 9;
+    sigma = 1.5;
+    kernel_length = 11;
+    kernel_width = 5;
     kernel_height = kernel_width;
     
     % form kernels

### 2024-01-03 Code/REP_kernels.m
fixed another kernels bug, added kernel surf visualization, added 3D kernel visualization

@@ -1,9 +1,9 @@
-function [kernels] = REP_kernels()
+function [kernels] = REP_kernels(type)
 
     % REP_kernels - Generate directional kernels for a 3D image processing task
     %
     % Usage:
-    %    [kernels] = REP_kernels()
+    %    [kernels] = REP_kernels(type)
     %
     % Description:
     %    This function generates directional kernels for a 3D image processing task.
@@ -15,21 +15,23 @@ function [kernels] = REP_kernels()
     %   directional_kernel_3d(direction, length, sigma1, width, sigma2, height, sigma3, type)
     %
     % Inputs:
-    %    None
+    %    type: 'linear' or 'gaussian'
     %
     % Outputs:
     %    kernels - A cell array containing the generated directional kernels.
 
+    type
+
     % set directional kernels
     sigma = 2
-    kernel_length = 10; % floored to even number
-    kernel_width = 5;
-    kernel_height = kernel_width;
+    kernel_length = 10 % floored to even number
+    kernel_width = 5
+    kernel_height = kernel_width
     
     % form kernels
-    kernel_x = directional_kernel_3d('x', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, 'gaussian');
-    kernel_y = directional_kernel_3d('y', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, 'gaussian');
-    kernel_z = directional_kernel_3d('z', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, 'gaussian');
+    kernel_x = directional_kernel_3d('x', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, type);
+    kernel_y = directional_kernel_3d('y', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, type);
+    kernel_z = directional_kernel_3d('z', kernel_length, sigma, kernel_width, sigma, kernel_height, sigma, type);
 
 %     figure()
 %     scatter(1:1:(length(kernel_x(5,:,5))),kernel_x(5,:,5))

### 2024-01-03 Code/directional_kernel_3d.m

@@ -12,14 +12,18 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
 
         % extend weights to kernel width
         center_width = floor(kernel_width/2);
-        width_extender = [1:1:center_width center_width flip(1:1:center_width)]/(center_width);
+%         width_extender = [1:1:center_width center_width flip(1:1:center_width)]/(center_width);
+        width_extender = [1:1:center_width flip(1:1:center_width)]/(center_width);
+
 
         % generate 2D kernel
         kernel2D = width_extender' * kernel;
 
         % extend weight to kernel height
         center_height = floor(kernel_height/2);
-        height_extender = [1:1:center_height center_height flip(1:1:center_height)]/(center_height);
+%         height_extender = [1:1:center_height center_height flip(1:1:center_height)]/(center_height);
+        height_extender = [1:1:center_height flip(1:1:center_height)]/(center_height);
+
         
         % create each layer
         kernel3D = repmat(kernel2D, 1, 1, length(height_extender)) .* reshape(height_extender, 1, 1, []);
@@ -36,7 +40,9 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         % prepare gaussian weights for width
         center_width = floor(kernel_width/2);
         width_extender = exp(-(1:center_width).^2/(2*sigma_width^2)); 
-        width_extender = [flip(width_extender) exp(-(center_width).^2/(2*sigma_width^2)) width_extender];
+%         width_extender = [flip(width_extender) exp(-(center_width).^2/(2*sigma_width^2)) width_extender];
+        width_extender = [flip(width_extender) width_extender];
+
         
         % generate 2D kernel
         kernel2D = width_extender' * kernel;
@@ -44,7 +50,9 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         % prepare gaussian weights for height
         center_height = floor(kernel_height/2);
         height_extender = exp(-(1:center_height).^2/(2*sigma_height^2));
-        height_extender = [flip(height_extender) exp(-(center_height).^2/(2*sigma_height^2)) height_extender];
+%         height_extender = [flip(height_extender) exp(-(center_height).^2/(2*sigma_height^2)) height_extender];
+        height_extender = [flip(height_extender) height_extender];
+
 
         % create each layer
         kernel3D = repmat(kernel2D, 1, 1, length(height_extender)) .* reshape(height_extender, 1, 1, []);

### 2024-01-03 Code/Full_RTConvolution_Multiple_Points.m
improvements to robot kernel visualization

@@ -7,12 +7,12 @@ mid_joints = 0
 avoid_task = 1
 kinematics_solution = 'exact-reduced' % OPTIONS: exact-reduced , exact , approximate
 timestep_secondary_gain_change = 0 % if selected, secondary task will start with normal gain and fall with time
-secondary_exec_stop_k = 10 % primary task will slow down (>0) or stop executing if secondary task has big velocities 
+secondary_exec_stop_k = 1 % primary task will slow down (>0) or stop executing if secondary task has big velocities 
 min_exec_slowdown_size = 0 % if poi is closer than this value, primary task will slow down
 
 % -----------------------------------------------------------
 % number of points per segment for obstacle avoidance taskmanipulability_avoidance
-points_per_segment = 1*[1 1 5 2 8 1 1];
+points_per_segment = 1*[1 1 2 1 3 1 1];
 
 % the number of points taken into account and weighting factors
 weights_avoidance = [1 1 1 1 1 1];
@@ -88,7 +88,7 @@ manipulability_avoidance = [];
 % -----------------------------------------------------------
 
 % repulsive kernels
-rep_kernels = REP_kernels();
+rep_kernels = REP_kernels('linear');
 
 % obstacles distance kernel
 % dist_kernel = euclidian_kernel_3D(61, 61, 61);

### 2024-01-04 Articles/Article 1/main.tex
added manipulator-kernel-visualizations, added some text

@@
 \usepackage{amsmath, xparse}
+
+\usepackage{graphicx}
 
@@
 
-\comment{}{PLOT: POI on the ROBOT + ONE OR MORE KERNELS ON THE ROBOT}
+\includegraphics[width=1\textwidth]{manipulator-kernels-visualization-grid-small.eps}
 
@@
 
-The fundamental concept of our directional kernels lies in computing the repulsive field individually for each direction within the Cartesian coordinate system.
-Our filters can be likened to a variation of the Sobel \add{reference} operator, a 2D convolutional \add{reference} filter frequently utilized in computer vision for calculating image gradients at specific points.
-However contrary to the Sobel operator, our filter is not calculating the gradients of the matrix, as the inputs are weighted depending on the distance from the center of the kernel.
+The fundamental concept of our directional kernels lies in computing the repulsive field individually for each direction within the Cartesian coordinate system.\remove{ Our filters share a passing similarity to the Sobel  operator, a 2D convolutional filter frequently utilized in computer vision for calculating image gradients at specific points.}
 
-Our kernels are designed as three-dimensional structures with a primary axis aligned along a specific Cartesian direction, corresponding to the calculated repulsive velocity.
-The two secondary axes are orthogonal to this primary axis.
-The distribution of values along the primary axis is inversely symmetric, exhibiting positive values on one side and negative values on the other, with the magnitude reaching zero at the kernel's center.
-The nature of the increase in magnitude along the primary axis of the kernel defines the shape of the repulsive velocity field, determining how the repulsive velocity escalates as the manipulator approaches an obstacle.
+Our kernels are designed as three-dimensional structures with a primary kernel axis aligned along a specific Cartesian direction, corresponding to the calculated repulsive velocity.
+The two secondary kernel axes are orthogonal to this primary axis.
+The distribution of values along the primary axis is inversely symmetric, exhibiting positive values on one side and negative values on the other, with the jump between max positive and max negative magnitude at the kernel's center.
+The function of the increase in magnitude along the primary axis of the kernel defines the shape of the repulsive velocity field, determining how the repulsive velocity changes as the point on the manipulator approaches an obstacle.
 
@@
 If the primary axis is too long, it can lead to extra calculations and may cause the robot to unnecessarily avoid obstacles that aren't in its immediate path, making its movement and path planning less efficient.
-A kernel with a primary axis that is too short might restrict the robot's ability to maneuver, detecting obstacles potentially too late, compromising the robots capacity to avoid obstacles effectively.
-Moreover, it is essential for the magnitudes at the kernel's periphery to be minimal, promoting a smooth increase in repulsive velocity rather than a sudden spike.
+A kernel with a primary axis that is too short might restrict the robot's ability to maneuver, detecting obstacles potentially too late, compromising the robots capacity to avoid obstacles effectively (eq.
+\ref{eq: detect range}).
+Moreover, it is essential for the magnitudes at the kernel's periphery to be minimal, promoting a smooth increase in repulsive velocity when approaching the obstacle rather than a sudden spike.
 
-In the orthogonal axes, the magnitude of values diminishes with increasing distance from the core of the kernel, nearing zero at the edges \add{using gaussian distribution}.
+\begin{equation}
+	\label{eq: detect range}
+	len_{primary} = \frac{2 \times \mathrm{range}}{\Delta \mathrm{grid}}\\
+\end{equation}
+
+%\remove{In the orthogonal axes, the magnitude of values diminishes with increasing distance from the core of the kernel, nearing zero at the edges.
 This gradation is vital for a smooth velocity transition when obstacles emerge in the kernels range.
@@
 The polarity of the values is determined by their position relative to the primary axis.
-The length of the kernel along the secondary axes influences the peripheral detection range for obstacles.
-Excessively large kernels may generate repulsive velocities unnecessarily, whereas too narrow kernels might only detect obstacles aligned directly with the Cartesian direction in the point of interest.
-Therefore, when selecting the width and height of the kernel, one must consider the density of the neighboring points of interest on the robot, ensuring that their collective fields adequately cover the manipulator's surrounding area.
+}
+
+The length of the orthogonal axes influences the peripheral detection range for obstacles.
+Excessively wide kernels may generate repulsive velocities for objects that are not in the path of the robot, whereas too narrow kernels might only detect obstacles aligned directly with the Cartesian direction in the point of interest.
+When selecting the width and height of the kernel, we must consider the density of the neighboring points of interest on the robot, ensuring that the \replace{collective fields}{combination of kernels} adequately cover the entire manipulator's surrounding area.
 

### 2024-01-04 Articles/Article 1/main.tex
added plot 3D field, with/without interpoaltion

@@
 	\label{eq: detect range}
-	len_{primary} = \frac{2 \times \mathrm{range}}{\Delta \mathrm{grid}}\\
+	num_{primary} = \frac{2 \times \mathrm{range}}{\Delta \mathrm{grid}}\\
 \end{equation}
@@
 \add{- choice of velocity profile (gaussian, linear ... maybe could move this part to implementation / experiments) \\}
-\add{- maybe first debluring - line convolution operation (if thick walls) - however this might move wall back, which could cause crash }
+%\add{- maybe first debluring - line convolution operation (if thick walls) - however this might move wall back, which could cause crash }
 
@@
 
+\includegraphics[width=0.75\textwidth]{field3D_visualization-3figures.eps}
+
 \section{Implementation}

### 2024-01-04 Code/REP_field_calculation.m
added constant secondary axis values, created 3D plots of field

@@ -1,6 +1,6 @@
-function [rep_values] = REP_field_calculation(grid, kernels, point)
+function [rep_values] = REP_field_calculation(grid, kernels, point, varargin)
 
-    %   REP_VALUES = REP_FIELD_CALCULATION(GRID, KERNELS, POINT) computes a set of
+    %   REP_VALUES = REP_FIELD_CALCULATION(GRID, KERNELS, POINT, VARARGIN) computes a set of
     %   representative values for a 3D grid at a specified point, using a set of 3D kernels.
     %
     %   Inputs:
@@ -10,14 +10,10 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
     %   - KERNELS: A cell array of 3D matrices, each representing a kernel.
     %   - POINT: A 1x3 vector specifying the coordinates of the point in the grid
     %            at which to compute the representative values.
-    %
-    % Optional In-Function Settings:
-    %   - interpolation_mode: A boolean (true or false) to enable or disable interpolation.
-    %                         If enabled, the function interpolates APF values based on the
-    %                         nearest grid cells. 
-    %
-    %   - box_value: Use this parameter to set occupation of the cells that
-    %     are beyond the inputed grid space.
+    %   - VARARGIN: Optional input arguments:
+    %       - 'interpolation_mode': A boolean (true or false) to enable or disable interpolation.
+    %       - 'box_value': Use this parameter to set occupation of the cells that
+    %         are beyond the inputed grid space.
     %
     %   Outputs:
     %   - REP_VALUES: A 1xN vector of representative values, where N is the number
@@ -28,12 +24,15 @@ function [rep_values] = REP_field_calculation(grid, kernels, point)
     %% SETTINGS:
     % --------------------------------------------------------------
 
-    % use interpolation
-    interpolation_mode = true; % true / false
+    % parse optional input arguments
+    p = inputParser;
+    addParameter(p, 'interpolation_mode', false);
+    addParameter(p, 'box_value', 0);
+    parse(p, varargin{:});
 
-    % set values outside the known grid to
-    box_value = 0;
-
+    % assign values from input parser
+    interpolation_mode = p.Results.interpolation_mode;
+    box_value = p.Results.box_value;
 
     % --------------------------------------------------------------    
     

### 2024-01-04 Code/REP_kernels.m

@@ -1,9 +1,9 @@
-function [kernels] = REP_kernels(type)
+function [kernels] = REP_kernels(varargin)
 
     % REP_kernels - Generate directional kernels for a 3D image processing task
     %
     % Usage:
-    %    [kernels] = REP_kernels(type)
+    %    [kernels] = REP_kernels(varargin)
     %
     % Description:
     %    This function generates directional kernels for a 3D image processing task.
@@ -15,31 +15,48 @@ function [kernels] = REP_kernels(type)
     %   directional_kernel_3d(direction, length, sigma1, width, sigma2, height, sigma3, type)
     %
     % Inputs:
-    %    type: 'linear' or 'gaussian'
+    %    varargin: optional input arguments
+    %       - type: 'linear' or 'gaussian'
+    %       - sigma: optional, default is 2
+    %       - kernel_length: optional, default is 10
+    %       - kernel_width: optional, default is 5
+    %       - kernel_height: optional, default is kernel_width
     %
     % Outputs:
     %    kernels - A cell array containing the generated directional kernels.
 
-    type
 
-    % set directional kernels
-    sigma = 2
-    kernel_length = 10 % floored to even number
-    kernel_width = 5
-    kernel_height = kernel_width
-    
+    % Create an input parser object
+    p = inputParser;
+
+    % Define the default values for the optional parameters
+    defaultSigma = 2;
+    defaultKernelLength = 10;
+    defaultKernelWidth = 5;
+    defaultKernelHeight = 5;
+
+    % Add the required and optional parameters to the input parser
+    addParameter(p, 'type', 'linear', @ischar);
+    addParameter(p, 'sigma', defaultSigma, @isnumeric);
+    addParameter(p, 'kernel_length', defaultKernelLength, @isnumeric);
+    addParameter(p, 'kernel_width', defaultKernelWidth, @isnumeric);
+    addParameter(p, 'kernel_height', defaultKernelHeight, @isnumeric);
+
+    % Parse the input arguments
+    parse(p, varargin{:});
+
+    % Extract the values from the input parser
+    type = p.Results.type;
+    sigma = p.Results.sigma;
+    kernel_length = p.Results.kernel_length;
+    kernel_width = p.Results.kernel_width;
+    kernel_height = p.Results.kernel_height;
+
     % form kernels
[…]

### 2024-01-04 Code/directional_kernel_3d.m

@@ -13,7 +13,13 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         % extend weights to kernel width
         center_width = floor(kernel_width/2);
 %         width_extender = [1:1:center_width center_width flip(1:1:center_width)]/(center_width);
-        width_extender = [1:1:center_width flip(1:1:center_width)]/(center_width);
+        
+        % if sigma_width is 0, then use a constant weight for the width
+        if (sigma_width == 0)
+            width_extender = ones(1,2*center_width);
+        else
+            width_extender = [1:1:center_width flip(1:1:center_width)]/(center_width);
+        end
 
 
         % generate 2D kernel
@@ -21,8 +27,13 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
 
         % extend weight to kernel height
         center_height = floor(kernel_height/2);
-%         height_extender = [1:1:center_height center_height flip(1:1:center_height)]/(center_height);
-        height_extender = [1:1:center_height flip(1:1:center_height)]/(center_height);
+
+        % if sigma_height is 0, then use a constant weight for the height
+        if (sigma_height == 0)
+            height_extender = ones(1,2*center_height);
+        else
+            height_extender = [1:1:center_height flip(1:1:center_height)]/(center_height);
+        end
 
         
         % create each layer
@@ -39,19 +50,29 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         
         % prepare gaussian weights for width
         center_width = floor(kernel_width/2);
-        width_extender = exp(-(1:center_width).^2/(2*sigma_width^2)); 
-%         width_extender = [flip(width_extender) exp(-(center_width).^2/(2*sigma_width^2)) width_extender];
-        width_extender = [flip(width_extender) width_extender];
 
+        % if sigma_width is 0, then use a constant weight for the width
+        if (sigma_width == 0)
+            width_extender = ones(1,2*center_width);
+        else
+            width_extender = exp(-(1:center_width).^2/(2*sigma_width^2)); 
+            %         width_extender = [flip(width_extender) exp(-(center_width).^2/(2*sigma_width^2)) width_extender];
+            width_extender = [flip(width_extender) width_extender];
+        end
         
         % generate 2D kernel
         kernel2D = width_extender' * kernel;
         
         % prepare gaussian weights for height
         center_height = floor(kernel_height/2);
[…]

### 2024-01-09 Articles/Article1-Conference/root.tex
added conference thing

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

%In case you encounter the following error:
%Error 1010 The PDF file may be corrupt (unable to open PDF file) OR
%Error 1000 An error occurred while parsing a contents stream. Unable to analyze the PDF file.
%This is a known problem with pdfLaTeX conversion filter. The file cannot be opened with acrobat reader
%Please use one of the alternatives below to circumvent this error by uncommenting one or the other
%\pdfobjcompresslevel=0
%\pdfminorversion=4

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

 \usepackage{easyReview}

\title{\LARGE \bf
Preparation of Papers for IEEE Sponsored Conferences \& Symposia*
}

\author{Albert Author$^{1}$ and Bernard D. Researcher$^{2}$% <-this % stops a space
\thanks{*This work was not supported by any organization}% <-this % stops a space
\thanks{$^{1}$Albert Author is with Faculty of Electrical Engineering, Mathematics and Computer Science,
        University of Twente, 7500 AE Enschede, The Netherlands
        {\tt\small albert.author@papercept.net}}%
\thanks{$^{2}$Bernard D. Researcheris with the Department of Electrical Engineering, Wright State University,
        Dayton, OH 45435, USA
        {\tt\small b.d.researcher@ieee.org}}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

This electronic document is a Olive template. The various components of your paper [title, text, heads, etc.] are already defined on the style sheet, as illustrated by the portions given in this document.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

\add{
V uvodu želimo predstaviti prvo uporabo kinematične redundance in kdo je delal na tem. Imena so Khatib, Maciejewki, Zlajpah ... \\
\\
Nato želimo omeniti, da so ti ljudje se problemu približali iz vidika izračuna oddaljenosti objektov. Omeniti je potrebno pristope z "bouding boxes AABB itd" in potem opisati "voxel grids". Končno je potrebno omeniti pristope z vektorskimi polji in nato pristope izračuna oddaljenosti.
}

\replace{

% PREDSTAVITEV PRISTOPOV PLANIRANJA POTI V ROBOTIKI
Sampling based, Kinematic Optimizazion (LS, LQP), GCS, Learning-based, DMP(Hoffman), Trajectory Optimization.

% PREDSTAVITEV REDUNDANCE
Kinematična redundanca je je pomembna značilnost modernih robotov. Kinematična redundanca je karakterizirana kot presežna količina prostorskih stopenj (DOF), sicer potrebnih za izvajanje določene naloge. Due to the redundant DOF the robot has the ability to move the end-effector along the same trajectory using different configurations of the mechanical structure. This provides means for solving sophisticated motion tasks, such as avoiding obstacles [4, 5], avoiding singularities [6], optimizing manipulability [7], minimizing joint torques [8], etc As a consequence, the dexterity of the system is significantly increased. This is important to accomplish complex tasks. Additionally, the redundancy can also have an important influence on the dynamic behavior of the robotic system.

% PREDSTAVITEV PRISTOPOV

}{replacement}

\section{PROCEDURE FOR PAPER SUBMISSION}

\subsection{Selecting a Template (Heading 2)}

First, confirm that you have the correct template for your paper size. This template has been tailored for output on the US-letter paper size. 
It may be used for A4 paper size if the paper size setting is suitably modified.

\subsection{Maintaining the Integrity of the Specifications}

The template is used to format your paper and style the text. All margins, column widths, line spaces, and text fonts are prescribed; please do not alter them. You may note peculiarities. For example, the head margin in this template measures proportionately more than is customary. This measurement and others are deliberate, using specifications that anticipate your paper as one part of the entire proceedings, and not as an independent document. Please do not revise any of the current designations

\section{MATH}

Before you begin to format your paper, first write and save the content as a separate text file. Keep your text and graphic files separate until after the text has been formatted and styled. Do not use hard tabs, and limit use of hard returns to only one return at the end of a paragraph. Do not add any kind of pagination anywhere in the paper. Do not number text heads-the template will do that for you.

Finally, complete content and organizational editing before formatting. Please take note of the following items when proofreading spelling and grammar:

\subsection{Abbreviations and Acronyms} Define abbreviations and acronyms the first time they are used in the text, even after they have been defined in the abstract. Abbreviations such as IEEE, SI, MKS, CGS, sc, dc, and rms do not have to be defined. Do not use abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}

\begin{itemize}

\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as O3.5-inch disk drive .
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: OWb/m2  or  webers per square meter , not  webers/m2 .  Spell out units when they appear in text:  . . . a few henries , not  . . . a few H .
\item Use a zero before decimal points:  0.25 , not  .25 . Use  cm3 , not  cc . (bullet list)

\end{itemize}

\subsection{Equations}

The equations are an exception to the prescribed specifications of this template. You will need to determine whether or not your equation should be typed using either the Times New Roman or the Symbol font (please no other font). To create multileveled equations, it may be necessary to treat the equation as a graphic and insert it into the text after your paper is styled. Number equations consecutively. Equation numbers, within parentheses, are to position flush right, as in (1), using a right tab stop. To make your equations more compact, you may use the solidus ( / ), the exp function, or appropriate exponents. Italicize Roman symbols for quantities and variables, but not Greek symbols. Use a long dash rather than a hyphen for a minus sign. Punctuate equations with commas or periods when they are part of a sentence, as in

$$
\alpha + \beta = \chi \eqno{(1)}
$$

Note that the equation is centered using a center tab stop. Be sure that the symbols in your equation have been defined before or immediately following the equation. Use  (1) , not  Eq. (1)  or  equation (1) , except at the beginning of a sentence:  Equation (1) is . . .

\subsection{Some Common Mistakes}
\begin{itemize}

\item The word  data  is plural, not singular.
\item The subscript for the permeability of vacuum ?0, and other common scientific constants, is zero with subscript formatting, not a lowercase letter  o .
\item In American English, commas, semi-/colons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an  inset , not an  insert . The word alternatively is preferred to the word  alternately  (unless you really mean something that alternates).
\item Do not use the word  essentially  to mean  approximately  or  effectively .
\item In your paper title, if the words  that uses  can accurately replace the word  using , capitalize the  u ; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones  affect  and  effect ,  complement  and  compliment ,  discreet  and  discrete ,  principal  and  principle .
\item Do not confuse  imply  and  infer .
\item The prefix  non  is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the  et  in the Latin abbreviation  et al. .
\item The abbreviation  i.e.  means  that is , and the abbreviation  e.g.  means  for example .

\end{itemize}

\section{USING THE TEMPLATE}

Use this sample document as your LaTeX source file to create your document. Save this file as {\bf root.tex}. You have to make sure to use the cls file that came with this distribution. If you use a different style file, you cannot expect to get required margins. Note also that when you are creating your out PDF file, the source file is only part of the equation. {\it Your \TeX\ $\rightarrow$ PDF filter determines the output file size. Even if you make all the specifications to output a letter file in the source - if your filter is set to produce A4, you will only get A4 output. }

It is impossible to account for all possible situation, one would encounter using \TeX. If you are using multiple \TeX\ files you must make sure that the ``MAIN`` source file is called root.tex - this is particularly important if your conference is using PaperPlaza's built in \TeX\ to PDF conversion tool.

\subsection{Headings, etc}

Text heads organize the topics on a relational, hierarchical basis. For example, the paper title is the primary text head because all subsequent material relates and elaborates on this one topic. If there are two or more sub-topics, the next level head (uppercase Roman numerals) should be used and, conversely, if there are not at least two sub-topics, then no subheads should be introduced. Styles named  Heading 1 ,  Heading 2 ,  Heading 3 , and  Heading 4  are prescribed.

\subsection{Figures and Tables}

Positioning Figures and Tables: Place figures and tables at the top and bottom of columns. Avoid placing them in the middle of columns. Large figures and tables may span across both columns. Figure captions should be below the figures; table heads should appear above the tables. Insert figures and tables after they are cited in the text. Use the abbreviation  Fig. 1 , even at the beginning of a sentence.

\begin{table}[h]
\caption{An Example of a Table}
\label{table_example}
\begin{center}
\begin{tabular}{|c||c|}
\hline
One & Two\\
\hline
Three & Four\\
\hline
\end{tabular}
\end{center}
\end{table}

   \begin{figure}[thpb]
      \centering
      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
}}
      %\includegraphics[scale=1.0]{figurefile}
      \caption{Inductance of oscillation winding on amorphous
       magnetic core versus DC bias magnetic field}
      \label{figurelabel}
   \end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words rather than symbols or abbreviations when writing Figure axis labels to avoid confusing the reader. As an example, write the quantity  Magnetization , or  Magnetization, M , not just  M . If including units in the label, present them within parentheses. Do not label axes only with units. In the example, write  Magnetization (A/m)  or  Magnetization {A[m(1)]} , not just  A/m . Do not label axes with a ratio of quantities and units. For example, write  Temperature (K) , not  Temperature/K.

\section{CONCLUSIONS}

A conclusion section is not required. Although a conclusion may review the main points of the paper, do not replicate the abstract as the conclusion. A conclusion might elaborate on the importance of the work or suggest applications and extensions.

\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{APPENDIX}

Appendixes should appear before the acknowledgment.

\section*{ACKNOWLEDGMENT}

The preferred spelling of the word  acknowledgment  in America is without an  e  after the  g . Avoid the stilted expression,  One of us (R. B. G.) thanks . . .   Instead, try  R. B. G. thanks . Put sponsor acknowledgments in the unnumbered footnote on the first page.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.

\begin{thebibliography}{99}

\bibitem{c1} G. O. Young,  Synthetic structure of industrial plastics (Book style with paper title and editor),  	in Plastics, 2nd ed. vol. 3, J. Peters, Ed.  New York: McGraw-Hill, 1964, pp. 15 64.
\bibitem{c2} W.-K. Chen, Linear Networks and Systems (Book style).	Belmont, CA: Wadsworth, 1993, pp. 123 135.
\bibitem{c3} H. Poor, An Introduction to Signal Detection and Estimation.   New York: Springer-Verlag, 1985, ch. 4.
\bibitem{c4} B. Smith,  An approach to graphs of linear forms (Unpublished work style),  unpublished.
\bibitem{c5} E. H. Miller,  A note on reflector arrays (Periodical styleÑAccepted for publication),  IEEE Trans. Antennas Propagat., to be publised.
\bibitem{c6} J. Wang,  Fundamentals of erbium-doped fiber amplifiers arrays (Periodical styleÑSubmitted for publication),  IEEE J. Quantum Electron., submitted for publication.
\bibitem{c7} C. J. Kaufman, Rocky Mountain Research Lab., Boulder, CO, private communication, May 1995.
\bibitem{c8} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa,  Electron spectroscopy studies on magneto-optical media and plastic substrate interfaces(Translation Journals style),  IEEE Transl. J. Magn.Jpn., vol. 2, Aug. 1987, pp. 740 741 [Dig. 9th Annu. Conf. Magnetics Japan, 1982, p. 301].
\bibitem{c9} M. Young, The Techincal Writers Handbook.  Mill Valley, CA: University Science, 1989.
\bibitem{c10} J. U. Duncombe,  Infrared navigationÑPart I: An assessment of feasibility (Periodical style),  IEEE Trans. Electron Devices, vol. ED-11, pp. 34 39, Jan. 1959.
\bibitem{c11} S. Chen, B. Mulgrew, and P. M. Grant,  A clustering technique for digital communications channel equalization using radial basis function networks,  IEEE Trans. Neural Networks, vol. 4, pp. 570 578, July 1993.
\bibitem{c12} R. W. Lucky,  Automatic equalization for digital communication,  Bell Syst. Tech. J., vol. 44, no. 4, pp. 547 588, Apr. 1965.
\bibitem{c13} S. P. Bingulac,  On the compatibility of adaptive controllers (Published Conference Proceedings style),  in Proc. 4th Annu. Allerton Conf. Circuits and Systems Theory, New York, 1994, pp. 8 16.
\bibitem{c14} G. R. Faulhaber,  Design of service systems with priority reservation,  in Conf. Rec. 1995 IEEE Int. Conf. Communications, pp. 3 8.
\bibitem{c15} W. D. Doyle,  Magnetization reversal in films with biaxial anisotropy,  in 1987 Proc. INTERMAG Conf., pp. 2.2-1 2.2-6.
\bibitem{c16} G. W. Juette and L. E. Zeffanella,  Radio noise currents n short sections on bundle conductors (Presented Conference Paper style),  presented at the IEEE Summer power Meeting, Dallas, TX, June 22 27, 1990, Paper 90 SM 690-0 PWRS.
\bibitem{c17} J. G. Kreifeldt,  An analysis of surface-detected EMG as an amplitude-modulated noise,  presented at the 1989 Int. Conf. Medicine and Biological Engineering, Chicago, IL.
\bibitem{c18} J. Williams,  Narrow-band analyzer (Thesis or Dissertation style),  Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, 1993. 
\bibitem{c19} N. Kawasaki,  Parametric study of thermal and chemical nonequilibrium nozzle flow,  M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
\bibitem{c20} J. P. Wilkinson,  Nonlinear resonant circuit devices (Patent style),  U.S. Patent 3 624 12, July 16, 1990.

\end{thebibliography}

\end{document}

### 2024-01-09 Articles/Article1-Conference/root.tex
Article1-Conference part of introduction

@@
 %\usepackage{amssymb}  % assumes amsmath package installed
+\usepackage{hyperref}
 
@@
 
+% PREDSTAVITEV REDUNDANCE
+
+Kinematic redundancy enables a kinematically redundant manipulator to follow a predefined task space trajectory using the endeffector (EE), while simultaneously, executing an additional task with the remaining movement capacity without impacting the trajectory adherence.
+This is possible because the robot's degrees of freedom (DOF) go beyond what is required to perform the primary task.
+Consequently, the robot can adopt different joint configurations optimised according to the secondary task while performing the primary task.
+Common secondary tasks include avoiding singularities, optimising the manipulability measure, minimising joint torques and avoiding obstacles in the operating space.
+
+\add{reference za posamezne pristope, dela \\}
+
+% PREDSTAVITEV PRISTOPOV PLANIRANJA POTI V ROBOTIKI
+\add{Sampling based, Kinematic Optimizazion (LS, LQP), GCS, Learning-based, DMP(Hoffman), Trajectory Optimization.
+\\}
+
+Motion planning is a fundamental problem in robotics \cite{IDEASLab2023}.
+It consists of finding a sequence of joint configurations for a robot so that the robot can move along this path from its initial configuration to the goal configuration without colliding with static obstacles or other robots in the environment.
+In addition to collision avoidance, motion planning for manipulators can optionally take into account various constraints, such as position, velocity, acceleration or jerk constraints for the joint angle or end effector, precision of the end effector with respect to position and orientation, stability of the manipulator, avoidance of singularities, or any number of other criteria.
+
+%
+% PREDSTAVITEV PRISTOPOV RAČUNANJA ODDALJENOSTI OVIRAM
+
 \add{
-V uvodu želimo predstaviti prvo uporabo kinematične redundance in kdo je delal na tem. Imena so Khatib, Maciejewki, Zlajpah ... \\
-\\
 	Nato želimo omeniti, da so ti ljudje se problemu približali iz vidika izračuna oddaljenosti objektov.
 Omeniti je potrebno pristope z "bouding boxes AABB itd" in potem opisati "voxel grids".
-Končno je potrebno omeniti pristope z vektorskimi polji in nato pristope izračuna oddaljenosti.
-}
+Končno je potrebno omeniti pristope z vektorskimi polji in nato pristope izračuna oddaljenosti.}
 
-\replace{
-
-% PREDSTAVITEV PRISTOPOV PLANIRANJA POTI V ROBOTIKI
-Sampling based, Kinematic Optimizazion (LS, LQP), GCS, Learning-based, DMP(Hoffman), Trajectory Optimization.
-
-% PREDSTAVITEV REDUNDANCE
-Kinematična redundanca je je pomembna značilnost modernih robotov.
-Kinematična redundanca je karakterizirana kot presežna količina prostorskih stopenj (DOF), sicer potrebnih za izvajanje določene naloge.
-Due to the redundant DOF the robot has the ability to move the end-effector along the same trajectory using different configurations of the mechanical structure.
-This provides means for solving sophisticated motion tasks, such as avoiding obstacles [4, 5], avoiding singularities [6], optimizing manipulability [7], minimizing joint torques [8], etc As a consequence, the dexterity of the system is significantly increased.
-This is important to accomplish complex tasks.
-Additionally, the redundancy can also have an important influence on the dynamic behavior of the robotic system.
-
-% PREDSTAVITEV PRISTOPOV
-
-}{replacement}
-
-\section{PROCEDURE FOR PAPER SUBMISSION}
-
-\subsection{Selecting a Template (Heading 2)}
-
-First, confirm that you have the correct template for your paper size. This template has been tailored for output on the US-letter paper size. 
-It may be used for A4 paper size if the paper size setting is suitably modified.
-
-\subsection{Maintaining the Integrity of the Specifications}
-
-The template is used to format your paper and style the text.
-All margins, column widths, line spaces, and text fonts are prescribed; please do not alter them.
-You may note peculiarities.
-For example, the head margin in this template measures proportionately more than is customary.
-This measurement and others are deliberate, using specifications that anticipate your paper as one part of the entire proceedings, and not as an independent document.
-Please do not revise any of the current designations
-
-\section{MATH}
-
-Before you begin to format your paper, first write and save the content as a separate text file.
-Keep your text and graphic files separate until after the text has been formatted and styled.
-Do not use hard tabs, and limit use of hard returns to only one return at the end of a paragraph.
-Do not add any kind of pagination anywhere in the paper.
-Do not number text heads-the template will do that for you.
-
-Finally, complete content and organizational editing before formatting. Please take note of the following items when proofreading spelling and grammar:
-
-\subsection{Abbreviations and Acronyms} Define abbreviations and acronyms the first time they are used in the text, even after they have been defined in the abstract.
-Abbreviations such as IEEE, SI, MKS, CGS, sc, dc, and rms do not have to be defined.
-Do not use abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
-
-\begin{itemize}
-
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as O3.5-inch disk drive .
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: OWb/m2  or  webers per square meter , not  webers/m2 .
-Spell out units when they appear in text:  .
-.
-.
-a few henries , not  .
-.
-.
-a few H .
-\item Use a zero before decimal points:  0.25 , not  .25 . Use  cm3 , not  cc . (bullet list)
-
-\end{itemize}
-
-\subsection{Equations}
-
-The equations are an exception to the prescribed specifications of this template.
-You will need to determine whether or not your equation should be typed using either the Times New Roman or the Symbol font (please no other font).
-To create multileveled equations, it may be necessary to treat the equation as a graphic and insert it into the text after your paper is styled.
-Number equations consecutively.
-Equation numbers, within parentheses, are to position flush right, as in (1), using a right tab stop.
-To make your equations more compact, you may use the solidus ( / ), the exp function, or appropriate exponents.
-Italicize Roman symbols for quantities and variables, but not Greek symbols.
-Use a long dash rather than a hyphen for a minus sign.
-Punctuate equations with commas or periods when they are part of a sentence, as in
-
-$$
-\alpha + \beta = \chi \eqno{(1)}
-$$
-
-Note that the equation is centered using a center tab stop.
-Be sure that the symbols in your equation have been defined before or immediately following the equation.
-Use  (1) , not  Eq.
-(1)  or  equation (1) , except at the beginning of a sentence:  Equation (1) is .
-.
-.
-
-\subsection{Some Common Mistakes}
-\begin{itemize}
-
-\item The word  data  is plural, not singular.
-\item The subscript for the permeability of vacuum ?0, and other common scientific constants, is zero with subscript formatting, not a lowercase letter  o .
-\item In American English, commas, semi-/colons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an  inset , not an  insert . The word alternatively is preferred to the word  alternately  (unless you really mean something that alternates).
-\item Do not use the word  essentially  to mean  approximately  or  effectively .
-\item In your paper title, if the words  that uses  can accurately replace the word  using , capitalize the  u ; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones  affect  and  effect ,  complement  and  compliment ,  discreet  and  discrete ,  principal  and  principle .
-\item Do not confuse  imply  and  infer .
-\item The prefix  non  is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the  et  in the Latin abbreviation  et al. .
-\item The abbreviation  i.e.  means  that is , and the abbreviation  e.g.  means  for example .
-
-\end{itemize}
+\clearpage
 
@@
 
+	\bibitem{IDEASLab2023}
+	IDEAS Lab, "Motion and Path Planning," presented at Purdue University, 2023. [Online]. Available: \url{https://ideas.cs.purdue.edu/research/robotics/planning/}. Accessed on: Jan. 9, 2024.
+
 \bibitem{c1} G. O. Young,  Synthetic structure of industrial plastics (Book style with paper title and editor),  	in Plastics, 2nd ed. vol. 3, J. Peters, Ed.  New York: McGraw-Hill, 1964, pp. 15 64.

### 2024-01-10 Articles/Article1-Conference/root.tex
added short summary of existing methods

@@
 
-Kinematic redundancy enables a kinematically redundant manipulator to follow a predefined task space trajectory using the endeffector (EE), while simultaneously, executing an additional task with the remaining movement capacity without impacting the trajectory adherence.
+Kinematic redundancy enables a manipulator to follow a predefined task space trajectory using the endeffector (EE), while simultaneously, optimising for an additional task with the remaining movement capacity without impacting the trajectory adherence.
 This is possible because the robot's degrees of freedom (DOF) go beyond what is required to perform the primary task.
@@
 
-\add{reference za posamezne pristope, dela \\}
+\add{citations}
 
 % PREDSTAVITEV PRISTOPOV PLANIRANJA POTI V ROBOTIKI
-\add{Sampling based, Kinematic Optimizazion (LS, LQP), GCS, Learning-based, DMP(Hoffman), Trajectory Optimization.
-\\}
 
-Motion planning is a fundamental problem in robotics \cite{IDEASLab2023}.
+Motion planning~\cite{IDEASLab2023} is a fundamental problem in robotics.
 It consists of finding a sequence of joint configurations for a robot so that the robot can move along this path from its initial configuration to the goal configuration without colliding with static obstacles or other robots in the environment.
 In addition to collision avoidance, motion planning for manipulators can optionally take into account various constraints, such as position, velocity, acceleration or jerk constraints for the joint angle or end effector, precision of the end effector with respect to position and orientation, stability of the manipulator, avoidance of singularities, or any number of other criteria.
+
+There are numerous methods for planning manipulator movements.
+Sampling-based methods such as PRM, RRT, RRT*, RRT-Connect, Informerd RRT*, BIT* and others offer a solution based on a global search in configuration space.
+However, the generated trajectories are not always smooth or optimal, and the performance of the methods may be insufficient for real-time operation.
+Recently, a number of learning-based methods using data-driven techniques have been proposed to improve or accelerate the functionality of sampling-based methods.
+\add{examples} Trajectory optimisation methods such as CHOMP, STOMP and TrajOpt, on the other hand, use optimization to improve an initial seed trajectory.
+Consequently, the optimality of the solution is highly dependent on this initial trajectory.
+Nevertheless, these methods are capable of generating smooth trajectories, and although they can be too computationally intensive for dynamic real-time environments, they are generally effective in finding constrained motion plans.
+The inverse kinematics with task prioritisation, which is based on finding a locally optimal least squares solution, and linear quadratic programming are two of the local optimization methods.
+Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
+However, since they do not plan further than one step ahead, they tend to get stuck in local minima.
+Therefore, they are often integrated with a higher-level planner, usually based on sampling, for global path search, while local optimisation takes dynamic environment changes into account.
+
+\add{citations}
 

### 2024-01-10 Articles/Article1-Conference/root.tex
environment representation methods

@@
 
-%
 % PREDSTAVITEV PRISTOPOV RAČUNANJA ODDALJENOSTI OVIRAM
+To be able to generate collision free trajectory we need to have a representation of the environment.
+One common approach is to represent environment with geomertic primitives such as bounding boxes or spheres.
+These representations are often combined into hierarhical trees, such as octatree, to make calculations more efficient.
+Such representations are memory effective and offer simple and computationally undemanding distance calculation and collision detection.
+These representations are widely used in computer simulations and games, however they can be suboptimal for , where sensor measurements arent always accurate and contain only part of the surrounding space.
+This can lead to approximation errors, which are additionally probelmatic when we are dealing with irregular shaped objects.
 
-\add{
-	Nato želimo omeniti, da so ti ljudje se problemu približali iz vidika izračuna oddaljenosti objektov.
-Omeniti je potrebno pristope z "bouding boxes AABB itd" in potem opisati "voxel grids".
-Končno je potrebno omeniti pristope z vektorskimi polji in nato pristope izračuna oddaljenosti.}
+Point-cloud representation captures the environment by directly using a collection of points obtained from sensors like LiDAR, providing a more precise depiction of space than bounding boxes or spheres.
+This method can more accurately represent the complex and irregular shapes of real-world objects, although it typically requires more memory and computational power to process, which can make it chllenging fore real-time representation.
+
+Voxel grid representation, on the other hand, divides the space into a regular grid of volumetric elements, or voxels, which can be used to create a more manageable approximation of the environment.
+While this approach offers a balance between detail and computational efficiency, it can still introduce discretization errors, particularly when modeling objects with smooth surfaces or intricate details.
+The fidelity of the representation is dependent on the size of the voxels: smaller voxels can capture more detail but require more memory and computation, while larger voxels result in coarser approximations but are more memory-efficient, adaptive voxel grids have been explored, where the voxel size can vary throughout the space to provide higher resolution in regions of interest while conserving resources in less critical areas.
+Voxel grids can incorporate probabilistic information, such as in occupancy grid maps, where each voxel holds a probability indicating the chance of an obstacle's presence.
+The occupancy probability of a voxel can be updated dynamically using sensor measurements and Bayesian updating methods.
+As new sensor data is collected, the probabilities are revised to reflect the increased or decreased likelihood of the presence of an obstacle in the voxel space.
+
+Additionally, techniques such as ESDF and TSDF can be utilized in conjunction with voxel grids to encode distance information, thereby enhancing the grid's utility for navigation and path planning by providing gradient information for obstacle avoidance and facilitating smoother trajectories
+
+\add{Gradient based voxel field methods. Voxel fields including distance information.}
 

### 2024-01-11 Articles/Article1-Conference/root.tex
idk

@@
 % PREDSTAVITEV PRISTOPOV PLANIRANJA POTI V ROBOTIKI
-
-Motion planning~\cite{IDEASLab2023} is a fundamental problem in robotics.
-It consists of finding a sequence of joint configurations for a robot so that the robot can move along this path from its initial configuration to the goal configuration without colliding with static obstacles or other robots in the environment.
+The task of finding the joint trajectories of a manipulator is called motion planning~\cite{IDEASLab2023}.
+It consists of finding a sequence of joint configurations for a robot so that the robot can move along this path from its initial configuration to the goal configuration without colliding with itself, static obstacles or other agents in the environment.
 In addition to collision avoidance, motion planning for manipulators can optionally take into account various constraints, such as position, velocity, acceleration or jerk constraints for the joint angle or end effector, precision of the end effector with respect to position and orientation, stability of the manipulator, avoidance of singularities, or any number of other criteria.
@@
 Consequently, the optimality of the solution is highly dependent on this initial trajectory.
-Nevertheless, these methods are capable of generating smooth trajectories, and although they can be too computationally intensive for dynamic real-time environments, they are generally effective in finding constrained motion plans.
-The inverse kinematics with task prioritisation, which is based on finding a locally optimal least squares solution, and linear quadratic programming are two of the local optimization methods.
+Nevertheless, these methods are capable of generating smooth trajectories, and although they can be too computationally intensive for high DOF dynamic real-time environments, they are generally effective in finding constrained motion plans.
+The inverse kinematics with task prioritisation, which is based on finding a least squares solution, and linear quadratic programming are two of the local optimization methods.
 Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
@@
 % PREDSTAVITEV PRISTOPOV RAČUNANJA ODDALJENOSTI OVIRAM
-To be able to generate collision free trajectory we need to have a representation of the environment.
+To be able to generate collision free trajectory we need to have a representation of the static environment.
 One common approach is to represent environment with geomertic primitives such as bounding boxes or spheres.
-These representations are often combined into hierarhical trees, such as octatree, to make calculations more efficient.
+Another way is to divide the occupied space into hierarhical trees, such as octatree, to make calculations more efficient.
 Such representations are memory effective and offer simple and computationally undemanding distance calculation and collision detection.
-These representations are widely used in computer simulations and games, however they can be suboptimal for , where sensor measurements arent always accurate and contain only part of the surrounding space.
-This can lead to approximation errors, which are additionally probelmatic when we are dealing with irregular shaped objects.
+These representations are widely used in computer simulations and games, however they can be suboptimal for real environment, where sensor measurements arent always accurate and contain only part of the surrounding space.
+This can lead to approximation errors, which are additionally probelmatic when we are dealing with irregularly shaped objects.
 
-Point-cloud representation captures the environment by directly using a collection of points obtained from sensors like LiDAR, providing a more precise depiction of space than bounding boxes or spheres.
-This method can more accurately represent the complex and irregular shapes of real-world objects, although it typically requires more memory and computational power to process, which can make it chllenging fore real-time representation.
+Point-cloud representation captures the environment by directly using a collection of points obtained from sensors like LiDAR, radar of RGBD camera, providing a more precise depiction of space than bounding boxes or spheres.
+This method can more accurately represent the complex and irregular shapes of real-world objects, although it typically requires more memory and computational power to process, which can make it chllenging for real-time representation.
+Point cloud data is often further converted into voxel grids.
 
-Voxel grid representation, on the other hand, divides the space into a regular grid of volumetric elements, or voxels, which can be used to create a more manageable approximation of the environment.
+Voxel grid representation divides the space into a regular grid of volumetric elements, or voxels, which can be used to create a more manageable approximation of the environment.
 While this approach offers a balance between detail and computational efficiency, it can still introduce discretization errors, particularly when modeling objects with smooth surfaces or intricate details.
@@
 As new sensor data is collected, the probabilities are revised to reflect the increased or decreased likelihood of the presence of an obstacle in the voxel space.
+Additionally, techniques such as ESDF and TSDF can be utilized in conjunction with voxel grids to encode signed distance information, that describes what is the distance between the selected voxel field and the nearest occupied field.
 
-Additionally, techniques such as ESDF and TSDF can be utilized in conjunction with voxel grids to encode distance information, thereby enhancing the grid's utility for navigation and path planning by providing gradient information for obstacle avoidance and facilitating smoother trajectories
-
-\add{Gradient based voxel field methods. Voxel fields including distance information.}
+\add{What are some ways of calculating ESDF in a voxel grid.}
 

### 2024-01-22 Articles/Article1-Conference/root.tex
some introduction progress

@@
 However, the generated trajectories are not always smooth or optimal, and the performance of the methods may be insufficient for real-time operation.
-Recently, a number of learning-based methods using data-driven techniques have been proposed to improve or accelerate the functionality of sampling-based methods.
-\add{examples} Trajectory optimisation methods such as CHOMP, STOMP and TrajOpt, on the other hand, use optimization to improve an initial seed trajectory.
+
+Recently, a number of learning-based methods using data-driven techniques have been proposed to improve or accelerate the functionality of sampling-based methods. \add{examples}
+
+Trajectory optimisation methods such as CHOMP, STOMP and TrajOpt, on the other hand, use optimization to improve an initial seed trajectory.
 Consequently, the optimality of the solution is highly dependent on this initial trajectory.
 Nevertheless, these methods are capable of generating smooth trajectories, and although they can be too computationally intensive for high DOF dynamic real-time environments, they are generally effective in finding constrained motion plans.
-The inverse kinematics with task prioritisation, which is based on finding a least squares solution, and linear quadratic programming are two of the local optimization methods.
+
+Local motion planning approaches employ optimization techniques, two common ones are inverse kinematics, that is based on finding a least squares solution of the manipulator joint velocities, and quadratic programming (QP)~\cite{c21}, \cite{c22}, \cite{c23}.
+Different articles have proposed the control schemes based on joint velocity control, joint acceleration control and torque control.
+
+\add{-add a few different options of IK \\
+ ->-using task priority or combined jacobian \\
+ -used for obstacle avoidance \\
+ -used for task manip, ... , ...
+}
+
+ The inverse kinematics with task prioritisation,  Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
+However, since they do not plan further than one step ahead, they tend to get stuck in local minima.
+Therefore, they are often combined with a higher-level planner, for global static environment based \replace{motion}{path} planning, while local optimisation takes dynamic environment changes into account.
+
+\add{Maybe add optimization - global intiger programming based approaches, by Tedrake.}
+
+%The inverse kinematics with task prioritisation, which is based on finding a least squares solution of the manipulator joint velocities, and linear quadratic programming (LQR) are two of the local optimization methods.
 Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
 However, since they do not plan further than one step ahead, they tend to get stuck in local minima.
-Therefore, they are often integrated with a higher-level planner, usually based on sampling, for global path search, while local optimisation takes dynamic environment changes into account.
+Therefore, they are often combined with a higher-level planner, for global static environment based \replace{motion}{path} planning, while local optimisation takes dynamic environment changes into account.
 
@@
 As new sensor data is collected, the probabilities are revised to reflect the increased or decreased likelihood of the presence of an obstacle in the voxel space.
-Additionally, techniques such as ESDF and TSDF can be utilized in conjunction with voxel grids to encode signed distance information, that describes what is the distance between the selected voxel field and the nearest occupied field.
+Additionally, techniques such as ESDF (Euclidian Signed Distance Field) and TSDF (Truncated Signed Distance Field) can be utilized in conjunction with voxel grids to encode signed distance information, that describes what is the distance between the selected voxel field and the nearest occupied field.
 
-\add{What are some ways of calculating ESDF in a voxel grid.}
+There has been a number of proposed algorithms for calculating ESDF.
+One common method is the Brushfire method, that spreads from obstacles until it calculates the distance for every field on a grid.
+\add{distance transforms} Jump Flooding Algorithm (JFA) is a similar method, that can be implemented on a GPU for faster parallel distance calculation.
+\add{\\-Wavefront \\
+	 -D* Klančar\\
+ 	 -Euclidian distnace algorithms\\
+ 	 -distance transforms\\
+  	 -voronoi diagrams\\
+   	 -voxblox, FIESTA \\
+     -ray casting, bullet casting ...}
 
@@
 
-\bibitem{c1} G. O. Young,  Synthetic structure of industrial plastics (Book style with paper title and editor),  	in Plastics, 2nd ed. vol. 3, J. Peters, Ed.  New York: McGraw-Hill, 1964, pp. 15 64.
-\bibitem{c2} W.-K. Chen, Linear Networks and Systems (Book style).	Belmont, CA: Wadsworth, 1993, pp. 123 135.
-\bibitem{c3} H. Poor, An Introduction to Signal Detection and Estimation.   New York: Springer-Verlag, 1985, ch. 4.
-\bibitem{c4} B. Smith,  An approach to graphs of linear forms (Unpublished work style),  unpublished.
-\bibitem{c5} E. H. Miller,  A note on reflector arrays (Periodical styleÑAccepted for publication),  IEEE Trans. Antennas Propagat., to be publised.
-\bibitem{c6} J. Wang,  Fundamentals of erbium-doped fiber amplifiers arrays (Periodical styleÑSubmitted for publication),  IEEE J. Quantum Electron., submitted for publication.
-\bibitem{c7} C. J. Kaufman, Rocky Mountain Research Lab., Boulder, CO, private communication, May 1995.
-\bibitem{c8} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa,  Electron spectroscopy studies on magneto-optical media and plastic substrate interfaces(Translation Journals style),  IEEE Transl.
-J.
-Magn.Jpn., vol.
-2, Aug.
-1987, pp.
-740 741 [Dig.
-9th Annu.
-Conf.
-Magnetics Japan, 1982, p.
-301].
-\bibitem{c9} M. Young, The Techincal Writers Handbook.  Mill Valley, CA: University Science, 1989.
-\bibitem{c10} J. U. Duncombe,  Infrared navigationÑPart I: An assessment of feasibility (Periodical style),  IEEE Trans. Electron Devices, vol. ED-11, pp. 34 39, Jan. 1959.
-\bibitem{c11} S.
-Chen, B.
-Mulgrew, and P.
-M.
-Grant,  A clustering technique for digital communications channel equalization using radial basis function networks,  IEEE Trans.
-Neural Networks, vol.
-4, pp.
-570 578, July 1993.
-\bibitem{c12} R. W. Lucky,  Automatic equalization for digital communication,  Bell Syst. Tech. J., vol. 44, no. 4, pp. 547 588, Apr. 1965.
-\bibitem{c13} S.
-P.
-Bingulac,  On the compatibility of adaptive controllers (Published Conference Proceedings style),  in Proc.
-4th Annu.
-Allerton Conf.
-Circuits and Systems Theory, New York, 1994, pp.
-8 16.
-\bibitem{c14} G. R. Faulhaber,  Design of service systems with priority reservation,  in Conf. Rec. 1995 IEEE Int. Conf. Communications, pp. 3 8.
-\bibitem{c15} W. D. Doyle,  Magnetization reversal in films with biaxial anisotropy,  in 1987 Proc. INTERMAG Conf., pp. 2.2-1 2.2-6.
-\bibitem{c16} G.
-W.
-Juette and L.
-E.
-Zeffanella,  Radio noise currents n short sections on bundle conductors (Presented Conference Paper style),  presented at the IEEE Summer power Meeting, Dallas, TX, June 22 27, 1990, Paper 90 SM 690-0 PWRS.
-\bibitem{c17} J. G. Kreifeldt,  An analysis of surface-detected EMG as an amplitude-modulated noise,  presented at the 1989 Int. Conf. Medicine and Biological Engineering, Chicago, IL.
-\bibitem{c18} J. Williams,  Narrow-band analyzer (Thesis or Dissertation style),  Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, 1993. 
-\bibitem{c19} N. Kawasaki,  Parametric study of thermal and chemical nonequilibrium nozzle flow,  M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
-\bibitem{c20} J. P. Wilkinson,  Nonlinear resonant circuit devices (Patent style),  U.S. Patent 3 624 12, July 16, 1990.
+\bibitem{c21} E.
+A.
+Basso and K.
+Y.
+Pettersen, “Task-Priority Control of Redundant Robotic Systems using Control Lyapunov and Control Barrier Function based Quadratic Programs,” in IFAC-PapersOnLine, vol.
+53, no.
+2, pp.
+9037–9044, 2020.
+doi: 10.1016/j.ifacol.2020.12.2024.
+
+\bibitem{c22} H.
+Toshani and M.
+Farrokhi, “Real-time inverse kinematics of redundant manipulators using neural networks and quadratic programming: A Lyapunov-based approach,” Robotics and Autonomous Systems, vol.
+62, no.
+6, pp.
+766–781, Jun.
+2014.
+doi: 10.1016/j.robot.2014.02.005.
+
+\bibitem{c23} Y.
+Zhang, S.
+S.
+Ge, and T.
+H.
+Lee, “A Unified Quadratic-Programming-Based Dynamical System Approach to Joint Torque Optimization of Physically Constrained Redundant Manipulators,” IEEE Trans.
+Syst., Man, Cybern.
+B, vol.
+34, no.
+5, pp.
+2126–2132, Oct.
+2004.
+doi: 10.1109/TSMCB.2004.830347.
 

### 2024-01-23 Articles/Article1-Conference/root.tex
added seminal works in IK

@@
 
-There are numerous methods for planning manipulator movements.
+There are numerous methods for planning manipulator movements, they can rougly be seperated into global and local approaches.
+
 Sampling-based methods such as PRM, RRT, RRT*, RRT-Connect, Informerd RRT*, BIT* and others offer a solution based on a global search in configuration space.
@@
 
-Local motion planning approaches employ optimization techniques, two common ones are inverse kinematics, that is based on finding a least squares solution of the manipulator joint velocities, and quadratic programming (QP)~\cite{c21}, \cite{c22}, \cite{c23}.
-Different articles have proposed the control schemes based on joint velocity control, joint acceleration control and torque control.
+Local motion planning approaches employ optimization techniques, two common ones are inverse kinematics~\cite{c29,c38}, that is based on finding a least squares solution of the manipulator joint velocities, and quadratic programming (QP)~\cite{c21,c22,c23}.
+In the following text we will focus on inverse kinematics based approaches.
 
-\add{-add a few different options of IK \\
- ->-using task priority or combined jacobian \\
+\remove{Different control schemes have been proposed.
+Acceleration-based control excels in precise handling of motion changes, velocity-based control offers consistent and smooth movement, while force and torque-based control provides direct control over joint forces for robust physical interactions.
+(CITATIONS in COMMENT BELLOW)}
+%\remove{Different control schemes have been proposed~\cite{c24}.
+Acceleration-based control excels in precise handling of motion changes, velocity-based control offers consistent and smooth movement, while force and torque-based control~\cite{c25, c26, c27, c28, c31} provides direct control over joint forces for robust physical interactions.}
+
+% ----
+
+%In our work we use velocity control strategy for obstacle avoidance using the manipulator.
+naša primarna naloga premik vrha robota - end efektorja (EE) v želeno pozo, to je točko in orientacijo v prostoru naloge.
+Naša sekundarna naloga se tako izvaja v preostalem prostoru gibanja sklepov redundance robota (DOF, DOR).
+???
+je predlagal razširitev jakobijeve matrike end-efektorja z dodatno matriko transformacije izogibne hitrosti iz kartzezičnega prostora v prostor sklepov.
+Če nalogi nista popolnoma neodvisni ena od druge, in njune sklepne hitrosti niso medsebojno ortagonalne, lahko to pripelje do medsebojnega oviranja izbajanja nalog.
+Zato je ???
+predlagal uporabo transformacijo hitrosti sekundarne naloge v  ničelnega prostora hitrosti primarne nalgoe, to je v prostor ortagonalnih hitrosti.
+Posledično izvajanje sekundarne naloge ne vpliva neposredno v izvajanje primarne naloge, gledano le v enem trenutku izvajanja.
+%
+%Različni razisklovalci so se lotili problema kinematičnega izogibanja na različne načine, kot je opisano v \cite{c30, c39}. 
+%
+%Colbaugh in Glass  1989 \cite{c31, c32} sta se problema lotila v dveh korakih.
+V prvem koraku sta izračunala trajektorijo vrha robota, v drugem pa sta uporabila optimizacijo za izboljšanje dinamičnega odziva robota med izvajanjem trajektorije in za druge naloge, kot so izogibanje ovir.
+%
+%Sciavicco and Siciliano \cite{c34, c35} and Egeland~\cite{c36} independetly introduced the concept ot task-space augmentation.
+It was later revisited by searji~\cite{c37}.
+The Solution is to stack the remaining rows in Jacobian with secondary task Jacobian into square matrix, that returns only one solution.
+%
+%Macijewski in Klein (Maciejewski & Klein, 1985) \cite{c29}, Nakamura in Hanafusa 1987 \cite{c38} predstavijo koncept prioritete nalog in ničelnega prostora za izogibanje, pri čemer se sekundarna naloga lahko giblje le s hitrostmi, ki ne vplivajo neposredno na gibanje primarne naloge.
+pri čemer sta izogibne sklepne hitrosti izračunala funkcijo najmanjše razdalje med točko na manipulatorju in točko izogibanja / objekta / ovire.
+%
+%Omenjeni pristopi se ne osredotočajo na okolje, tega pogosto prikažejo kot geometrijske primitive in nato razdaljo izračunajo kot razdaljo med primitivom in premicami, ki opisujejo manipulator. 
+%
+%Khatib \cite{c33} proposes the concept of Artificial Potential Field.
+Polje sestavljata odbojna komponenta, ki end effektor odbija od ovir, ki so predstavljene kot geometrijski primitivi in privlačna komponenta, ki manipulator privlači proti cilju.
+%
+%
+%
+
+Various researchers have adopted different approaches to kinematic avoidance, as detailed in~\cite{c30, c39}.
+
+Colbaugh and Glass (1989)~\cite{c31, c32} tackled this problem in a two-step process.
+Initially, they calculated the robot's end-effector trajectory.
+Subsequently, they used optimization to enhance the robot's dynamic response during trajectory execution and for tasks like obstacle avoidance.
+
+Sciavicco and Siciliano~\cite{c34, c35}, as well as Egeland~\cite{c36}, independently introduced the concept of task-space augmentation, later revisited by Searji~\cite{c37}.
+This method involves integrating the remaining rows of the Jacobian with the secondary task Jacobian into a square matrix, yielding a singular solution.
+
+Maciejewski and Klein (1985)~\cite{c29}, and Nakamura and Hanafusa (1987)~\cite{c38}, presented the concept of task prioritization and a null space for avoidance.
+In this framework, the secondary task is restricted to velocities that do not directly affect the primary task's movement.
+Their method calculates avoidance joint velocities based on the minimum distance between a point on the manipulator and an avoidance point, object, or obstacle.
+
+These approaches typically do not focus on the environmental context, often representing it as geometric primitives and then calculating the distance between these primitives and the lines describing the manipulator.
+
+Khatib~\cite{c33} proposes the concept of an Artificial Potential Field.
+This field consists of a repulsive component that deflects the end effector away from obstacles, depicted as geometric primitives, and an attractive component that draws the manipulator towards its target.
+
+\add{\\
+THESE SEMINAL WORKS HAVE BEEN FURTHER DEVELOPED / APPLIED IN ... add 5-10 modern citations
+\\
+\\
+}
+
+\add{
  -used for obstacle avoidance \\
  -used for task manip, ... , ...
+ -fix citation styles
 }
@@
 
+\bibitem{c24} J.
+Nakanishi, R.
+Cory, M.
+Mistry, J.
+Peters, and S.
+Schaal, “Comparative experiments on task space control with redundancy resolution,” in Proc.
+2005 IEEE/RSJ Int.
+Conf.
+on Intelligent Robots and Systems, Edmonton, Alta., Canada, 2005, pp.
+3901–3908.
+doi: 10.1109/IROS.2005.1545203.
+
+\bibitem{c25} M.
+H.
+Raibert and J.
+J.
+Craig, “Hybrid Position/Force Control of Manipulators,” Journal of Dynamic Systems, Measurement, and Control, vol.
+103, no.
+2, pp.
+126–133, Jun.
+1981.
+doi: 10.1115/1.3139652.
+
+\bibitem{c26} T.
+Yoshikawa, “Dynamic hybrid position/force control of robot manipulators--Description of hand constraints and calculation of joint driving force,” IEEE J.
+Robot.
+Automat., vol.
+3, no.
+5, pp.
+386–392, Oct.
+1987.
+doi: 10.1109/JRA.1987.1087120.
+
+\bibitem{c27} O.
+Khatib, “A unified approach for motion and force control of robot manipulators: The operational space formulation,” IEEE J.
+Robot.
+Automat., vol.
+3, no.
+1, pp.
+43–53, Feb.
+1987.
+doi: 10.1109/JRA.1987.1087068.
+
+\bibitem{c28} N. Hogan, “Impedance Control: An Approach to Manipulation,” in Proc. 1984 American Control Conf., San Diego, CA, USA, Jul. 1984, pp. 304–313. doi: 10.23919/ACC.1984.4788393.
+
+\bibitem{c29} A.
+A.
+Maciejewski and C.
+A.
+Klein, “Obstacle Avoidance for Kinematically Redundant Manipulators in Dynamically Varying Environments,” The International Journal of Robotics Research, vol.
+4, no.
+3, pp.
+109–117, Sep.
+1985.
+doi: 10.1177/027836498500400308.
+
+\bibitem{c30} L.
+Lajpah and T.
+Petri, “Obstacle Avoidance for Redundant Manipulators as Control Problem,” in Serial and Parallel Robot Manipulators - Kinematics, Dynamics, Control and Optimization, S.
+Kucuk, Ed.
+InTech, 2012.
+doi: 10.5772/32651.
+
+\bibitem{c31} R. Colbaugh and K. Glass, “Cartesian control of redundant robots,” J. Robotic Syst., vol. 6, no. 4, pp. 427–459, Aug. 1989. doi: 10.1002/rob.4620060409.
+
+\bibitem{c32} K.
+Glass, R.
+Colbaugh, D.
+Lim, and H.
+Seraji, “Real-time collision avoidance for redundant manipulators,” IEEE Trans.
+Robot.
+Automat., vol.
+11, no.
+3, pp.
+448–457, Jun.
+1995.
+doi: 10.1109/70.388789.
+
+\bibitem{c33} O.
+Khatib, “Real-time obstacle avoidance for manipulators and mobile robots,” in 1985 IEEE International Conference on Robotics and Automation Proceedings, Mar.
+1985, pp.
+500–505.
+doi: 10.1109/ROBOT.1985.1087247.
+
+\bibitem{c34} L.
+Sciavicco and B.
+Siciliano, “A solution algorithm to the inverse kinematic problem for redundant manipulators,” IEEE J.
+Robot.
+Automat., vol.
+4, no.
+4, pp.
+403–410, Aug.
+1988.
+doi: 10.1109/56.804.
+
+\bibitem{c35} L.
+Sciavicco and B.
+Siciliano, “Solving the Inverse Kinematic Problem for Robotic Manipulators,” in RoManSy 6, A.
+Morecki, G.
+Bianchi, and K.
+Kedzior, Eds., Boston, MA: Springer US, 1987, pp.
+107–114.
+doi: 10.1007/978-1-4684-6915-8\_9.
+
+\bibitem{c36} O. Egeland, “Task-space tracking with redundant manipulators,” IEEE J. Robot. Automat., vol. 3, no. 5, pp. 471–475, Oct. 1987. doi: 10.1109/JRA.1987.1087118.
+
+\bibitem{c37} H. Seraji, “Configuration control of redundant manipulators: theory and implementation,” IEEE Trans. Robot. Automat., vol. 5, no. 4, pp. 472–490, Aug. 1989. doi: 10.1109/70.88062.
+
+\bibitem{c38} Y.
+Nakamura, H.
+Hanafusa, and T.
+Yoshikawa, “Task-Priority Based Redundancy Control of Robot Manipulators,” The International Journal of Robotics Research, vol.
+6, no.
+2, pp.
+3–15, Jun.
+1987.
+doi: 10.1177/027836498700600201.
+
+\bibitem{c39} B. Siciliano and O. Khatib, Eds., Springer Handbook of Robotics. in Springer Handbooks. Cham: Springer International Publishing, 2016. doi: 10.1007/978-3-319-32552-1.
+
 \end{thebibliography}

### 2024-01-24 Articles/Article1-Conference/root.tex
more APF citati

@@
 Sciavicco and Siciliano~\cite{c34, c35}, as well as Egeland~\cite{c36}, independently introduced the concept of task-space augmentation, later revisited by Searji~\cite{c37}.
-This method involves integrating the remaining rows of the Jacobian with the secondary task Jacobian into a square matrix, yielding a singular solution.
+This method involves extending the primary task Jacobian with the secondary task Jacobian into a square matrix, yielding a singular solution.
 
@@
 Their method calculates avoidance joint velocities based on the minimum distance between a point on the manipulator and an avoidance point, object, or obstacle.
+Žlajpah~\cite{c41} improved the concept by proposing a reduced operational space formulation, that reshapes the Jacobain of the avoidance task from three cartesian axis to only the direction of obstacle avoidance.
+Petrič~\cite{c42} suggested a way to smoothly transition between avoidance and trajectory following tasks.
 
@@
 This field consists of a repulsive component that deflects the end effector away from obstacles, depicted as geometric primitives, and an attractive component that draws the manipulator towards its target.
+In the following years many different modifications and improvement of the original APF idea have been proposed, often focused on removal of local minimas in the potential field.
+Khim and Khosla~\cite{c40} suggested the use of harmonic functions to solve the problem of local minima.
+Pinto et al.
+\cite{c43} proposes to vary the field based on the distance of robot from obstacles to fill the local minima.
+
+Many of the recent works focus on use of a variation of artificial potential field to plan motion of the manipulator.
+Long~\cite{c44} suggests creating motion plan of the manipulator using APF, he extends it using RRT to calculate virtual attractive point for the robot to move towards in case of local minima.
+\add{cite 1 2 3 4 5}
 
@@
 
+\bibitem{c40} J.-O. Kim and P. Khosla, “Real-Time Obstacle Avoidance Using Harmonic Potential Functions,” 1992. doi: 10.1109/70.143352.
+
+\bibitem{c41} L.
+Zlajpah and B.
+Nemec, “Kinematic control algorithms for on-line obstacle avoidance for redundant manipulators,” in Proc.
+IEEE/RSJ International Conference on Intelligent Robots and Systems, Lausanne, Switzerland, 2002, pp.
+1898–1903.
+doi: 10.1109/IRDS.2002.1044033.
+
+\bibitem{c42} T.
+Petrič and L.
+Žlajpah, “Smooth continuous transition between tasks on a kinematic control level: Obstacle avoidance as a control problem,” Robotics and Autonomous Systems, vol.
+61, no.
+9, Art.
+no.
+9, Sep.
+2013.
+doi: 10.1016/j.robot.2013.04.019.
+
+\bibitem{c43} M.
+F.
+Pinto, T.
+R.
+F.
+Mendonça, L.
+R.
+Olivi, E.
+B.
+Costa, and A.
+L.
+M.
+Marcato, “Modified approach using variable charges to solve inherent limitations of potential fields method,” in Proc.
+2014 11th IEEE/IAS International Conference on Industry Applications, Dec.
+2014, pp.
+1–6.
+doi: 10.1109/INDUSCON.2014.7059414.
+
+\bibitem{c44} Z.
+Long, “Virtual target point-based obstacle-avoidance method for manipulator systems in a cluttered environment,” Engineering Optimization, vol.
+52, no.
+11, Art.
+no.
+11, Nov.
+2020.
+doi: 10.1080/0305215X.2019.1681986.
+
 \end{thebibliography}

### 2024-01-25 Articles/Article1-Conference/root.tex
Introduction: mostly rounded up Motion planning, kinematic avoidance and APF parts

@@
 
+\subsection{Motion Planning}
+
 % PREDSTAVITEV REDUNDANCE
 
-Kinematic redundancy enables a manipulator to follow a predefined task space trajectory using the endeffector (EE), while simultaneously, optimising for an additional task with the remaining movement capacity without impacting the trajectory adherence.
+Kinematic redundancy~\cite{siciliano1990kinematic, siciliano2016springer} enables a manipulator to follow a predefined task space trajectory using the endeffector (EE), while simultaneously, optimising for an additional task with the remaining movement capacity without impacting the trajectory adherence.
 This is possible because the robot's degrees of freedom (DOF) go beyond what is required to perform the primary task.
 Consequently, the robot can adopt different joint configurations optimised according to the secondary task while performing the primary task.
-Common secondary tasks include avoiding singularities, optimising the manipulability measure, minimising joint torques and avoiding obstacles in the operating space.
-
-\add{citations}
+Common secondary tasks~\cite{siciliano2010robot} include avoiding singularities, optimising the manipulability measure, minimising joint torques and avoiding obstacles in the operating space.
 
@@
 
-There are numerous methods for planning manipulator movements, they can rougly be seperated into global and local approaches.
-
-Sampling-based methods such as PRM, RRT, RRT*, RRT-Connect, Informerd RRT*, BIT* and others offer a solution based on a global search in configuration space.
+There are numerous methods for planning manipulator movements~\cite{c51, c52}, they can rougly be seperated into global and local approaches.
+Global, sampling-based, methods such as PRM~\cite{vsvestka1997motion}, RRT*~\cite{lavalle1998rapidly, karaman2010incremental}, RRT-Connect~\cite{kuffner2000rrt}, Informerd RRT*~\cite{gammell2014informed}, BIT*~\cite{gammell2015batch} and others offer a globally optimal solution based on a global search in configuration space.
 However, the generated trajectories are not always smooth or optimal, and the performance of the methods may be insufficient for real-time operation.
 
-Recently, a number of learning-based methods using data-driven techniques have been proposed to improve or accelerate the functionality of sampling-based methods. \add{examples}
+%Global planning methods are typically more comprehensive, aiming to find a globally optimal trajectory for the manipulator.
+These methods often involve mapping out the entire path before movement begins, taking into account the known environment, which usually includes static obstacles.
+However, the downside of global planning methods is their computational intensity, which often renders them too slow for real-time applications.
+This limitation is particularly pronounced in dynamic or unpredictable environments where conditions change rapidly and require quick responses.
+Examples of global planning methods include graph-based algorithms like A* and Dijkstra's algorithm, which are known for their effectiveness in pathfinding but also for their computational demands [(Lavalle, 2006)].
 
-Trajectory optimisation methods such as CHOMP, STOMP and TrajOpt, on the other hand, use optimization to improve an initial seed trajectory.
+%Sampling-based methods such as PRM, RRT, RRT*, RRT-Connect, Informerd RRT*, BIT* and others offer a solution based on a global search in configuration space.
+However, the generated trajectories are not always smooth or optimal, and the performance of the methods may be insufficient for real-time operation.
+
+\remove{Recently, a number of learning-based methods using data-driven techniques have been proposed to improve or accelerate the functionality of sampling-based methods. }
+
+\remove{Trajectory optimisation methods such as CHOMP, STOMP and TrajOpt, on the other hand, use optimization to improve an initial seed trajectory.
 Consequently, the optimality of the solution is highly dependent on this initial trajectory.
 Nevertheless, these methods are capable of generating smooth trajectories, and although they can be too computationally intensive for high DOF dynamic real-time environments, they are generally effective in finding constrained motion plans.
+}
 
 Local motion planning approaches employ optimization techniques, two common ones are inverse kinematics~\cite{c29,c38}, that is based on finding a least squares solution of the manipulator joint velocities, and quadratic programming (QP)~\cite{c21,c22,c23}.
+Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
+However, since they do not plan further than one step ahead, they tend to get stuck in local minima.
+Therefore, they are often combined with a higher-level planner, for global static environment based \replace{motion}{path} planning, while local optimisation takes dynamic environment changes into account.
 In the following text we will focus on inverse kinematics based approaches.
+
+\subsection{Kinematic Obstacle Avoidance}
 
@@
 
+\subsection{Artificial Potential Field}
+
 These approaches typically do not focus on the environmental context, often representing it as geometric primitives and then calculating the distance between these primitives and the lines describing the manipulator.
@@
 \cite{c43} proposes to vary the field based on the distance of robot from obstacles to fill the local minima.
+Many researchers tried using APF as a heuristic to better guide sampling based approaches~\cite{c45, c46, c47}.
 
 Many of the recent works focus on use of a variation of artificial potential field to plan motion of the manipulator.
+Xia et al.~\cite{c49} uses a variation of APF for manipulator motion~\cite{c49}.
+Chen et al.
+proposes an application using APF in joint space and a variable kinematic optimization step~\cite{c50}.
 Long~\cite{c44} suggests creating motion plan of the manipulator using APF, he extends it using RRT to calculate virtual attractive point for the robot to move towards in case of local minima.
-\add{cite 1 2 3 4 5}
-
-\add{\\
-THESE SEMINAL WORKS HAVE BEEN FURTHER DEVELOPED / APPLIED IN ... add 5-10 modern citations
-\\
-\\
-}
+Zhu et al.
+proposes use of APF in combination with MPC, to plan in environment with dynamic obstacles~\cite{c48}.
 
 \add{
- -used for obstacle avoidance \\
- -used for task manip, ... , ...
- -fix citation styles
+ -fix citation styles (sometimes et al, sometimes not, sometimes names, sometimes years ... not same format)
 }
-
- The inverse kinematics with task prioritisation,  Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
-However, since they do not plan further than one step ahead, they tend to get stuck in local minima.
-Therefore, they are often combined with a higher-level planner, for global static environment based \replace{motion}{path} planning, while local optimisation takes dynamic environment changes into account.
-
-\add{Maybe add optimization - global intiger programming based approaches, by Tedrake.}
 
@@
 
-\add{citations}
+\subsection{Environment representation}
 
@@
 
+\bibitem{c45} A. H. Qureshi and Y. Ayaz, “Potential Functions based Sampling Heuristic For Optimal Path Planning,” Auton Robot, vol. 40, no. 6, Art. no. 6, Aug. 2016. doi: 10.1007/s10514-015-9518-0.
+
+\bibitem{c46} A.
+H.
+Qureshi et al., “Adaptive Potential guided directional-RRT*,” in Proc.
+2013 IEEE International Conference on Robotics and Biomimetics (ROBIO), Shenzhen, China, Dec.
+2013, pp.
+1887–1892.
+doi: 10.1109/ROBIO.2013.6739744.
+
+\bibitem{c47} J.
+Yi, Q.
+Yuan, R.
+Sun, and H.
+Bai, “Path planning of a manipulator based on an improved P\_RRT* algorithm,” Complex Intell.
+Syst., vol.
+8, no.
+3, pp.
+2227–2245, Jun.
+2022.
+doi: 10.1007/s40747-021-00628-y.
+
+\bibitem{c48} T.
+Zhu, J.
+Mao, L.
+Han, C.
+Zhang, and J.
+Yang, “Real-Time Dynamic Obstacle Avoidance for Robot Manipulators Based on Cascaded Nonlinear MPC With Artificial Potential Field,” IEEE Trans.
+Ind.
+Electron., pp.
+1–11, 2023.
+doi: 10.1109/TIE.2023.3306405.
+
+\bibitem{c49} X. Xia et al., “Path Planning for Obstacle Avoidance of Robot Arm Based on Improved Potential Field Method,” Sensors, vol. 23, no. 7, Art. no. 7, Apr. 2023. doi: 10.3390/s23073754.
+
+\bibitem{c50} Y.
+Chen, L.
+Chen, J.
+Ding, and Y.
+Liu, “Research on Real-Time Obstacle Avoidance Motion Planning of Industrial Robotic Arm Based on Artificial Potential Field Method in Joint Space,” Applied Sciences, vol.
+13, no.
+12, p.
+6973, Jan.
+2023.
+doi: 10.3390/app13126973.
+
+\bibitem{c51} S. M. LaValle, Planning Algorithms. Cambridge: Cambridge University Press, 2006.
+
+\bibitem{c52} M.
+G.
+Tamizi, M.
+Yaghoubi, and H.
+Najjaran, “A review of recent trend in motion planning of industrial robots,” Int J Intell Robot Appl, vol.
+7, no.
+2, Art.
+no.
+2, Jun.
+2023.
+doi: 10.1007/s41315-023-00274-2.
+
+\bibitem{vsvestka1997motion} P.
+{\v{S}}vestka and M.
+H.
+Overmars, “Motion planning for carlike robots using a probabilistic learning approach,” The International Journal of Robotics Research, vol.
+16, no.
+2, pp.
+119–143, 1997.
+
+\bibitem{lavalle1998rapidly} S. LaValle, “Rapidly-exploring random trees: A new tool for path planning,” Research Report 9811, 1998.
+
+\bibitem{gammell2015batch} J.
+D.
+Gammell, S.
+S.
+Srinivasa, and T.
+D.
+Barfoot, “Batch Informed Trees (BIT*): Sampling-based Optimal Planning via the Heuristically Guided Search of Implicit Random Geometric Graphs,” in Proc.
+of the 2015 IEEE International Conference on Robotics and Automation (ICRA), May 2015, pp.
+3067–3074.
+doi: 10.1109/ICRA.2015.7139620.
+
+\bibitem{karaman2010incremental} S. Karaman and E. Frazzoli, "Incremental sampling-based algorithms for optimal motion planning," in Proc. Robotics: Science and Systems (RSS), 2010.
+
+\bibitem{gammell2014informed} J.
+D.
+Gammell, S.
+S.
+Srinivasa, and T.
+D.
+Barfoot, “Informed RRT*: Optimal sampling-based path planning focused via direct sampling of an admissible ellipsoidal heuristic,” in Proc.
+of the 2014 IEEE/RSJ International Conference on Intelligent Robots and Systems, Chicago, IL, USA, Sep.
+2014, pp.
+2997–3004.
+doi: 10.1109/IROS.2014.6942976.
+
+\bibitem{kuffner2000rrt} J.
+J.
+Kuffner and S.
+M.
+LaValle, “RRT-connect: An efficient approach to single-query path planning,” in Proceedings of the 2000 ICRA.
+Millennium Conference.
+IEEE International Conference on Robotics and Automation.
+Symposia Proceedings (Cat.
+No.00CH37065), Apr.
+2000, pp.
+995–1001 vol.2.
+doi: 10.1109/ROBOT.2000.844730.
+
+\bibitem{siciliano1990kinematic} B. Siciliano, “Kinematic control of redundant robot manipulators: A tutorial,” J. Intell. Robot. Syst., vol. 3, no. 3, Art. no. 3, 1990, doi: 10.1007/BF00126069.
+
+\bibitem{siciliano2016springer} B. Siciliano and O. Khatib, Eds., Springer Handbook of Robotics, in Springer Handbooks. Cham: Springer International Publishing, 2016. doi: 10.1007/978-3-319-32552-1.
+
+\bibitem{siciliano2010robot} B.
+Siciliano, L.
+Sciavicco, L.
+Villani, and G.
+Oriolo, Robot.
+Model.
+Plan.
+Control, 2010, pp.
+161–189.
+[Online].
+Available: http://link.springer.com/10.1007/978-1-84628-642-1\_4
+
 \end{thebibliography}

### 2024-01-26 Articles/Article1-Conference/root.tex
minor todo
Introduction: Environment representation approaches

@@
 
-\add{
+\alert{
  -fix citation styles (sometimes et al, sometimes not, sometimes names, sometimes years ... not same format)
@@
 % PREDSTAVITEV PRISTOPOV RAČUNANJA ODDALJENOSTI OVIRAM
+%To be able to generate collision free trajectory we need to have a representation of the static environment.
+One common approach is to enclose the obstacles around the manipulator into a collection of simple geometric primitives and then construct a tree-like structure to achieve the goal~\cite{dai2022review}.
+Some of the commonly used methods are AABB (axis-aligned bounding boxes)~\cite{vandenbergen1997efficient, chen2018path}, OBB (oriented bounding boxes)~\cite{gottschalk1996obbtree, puiu2011realtime}.
+
 To be able to generate collision free trajectory we need to have a representation of the static environment.
-One common approach is to represent environment with geomertic primitives such as bounding boxes or spheres.
-Another way is to divide the occupied space into hierarhical trees, such as octatree, to make calculations more efficient.
-Such representations are memory effective and offer simple and computationally undemanding distance calculation and collision detection.
-These representations are widely used in computer simulations and games, however they can be suboptimal for real environment, where sensor measurements arent always accurate and contain only part of the surrounding space.
-This can lead to approximation errors, which are additionally probelmatic when we are dealing with irregularly shaped objects.
+One common approach involves enclosing the obstacles around the manipulator within a collection of simple geometric primitives and then constructing a tree-like structure to facilitate efficient navigation and path planning~\cite{dai2022review}.
+Two of the commonly used methods for this purpose include AABB (axis-aligned bounding boxes)~\cite{vandenbergen1997efficient, chen2018path} and OBB (oriented bounding boxes)~\cite{gottschalk1996obbtree, puiu2011realtime}\remove{and FDH (fixed directions hulls).}.
+Another significant method worth mentioning is the use of Octomaps~\cite{wurmOctoMap}.
+Octomaps employ an octree data structure to represent 3D environments efficiently, making them particularly suitable for areas with large open spaces.
 
-Point-cloud representation captures the environment by directly using a collection of points obtained from sensors like LiDAR, radar of RGBD camera, providing a more precise depiction of space than bounding boxes or spheres.
-This method can more accurately represent the complex and irregular shapes of real-world objects, although it typically requires more memory and computational power to process, which can make it chllenging for real-time representation.
-Point cloud data is often further converted into voxel grids.
+In real environment we usually collect depth data with LiDAR, radar of RGBD camera sensors, that usually return clouds of points, that tell us the distance of objects in robots environment.
+While approaches have been proposed to use the point-cloud data directly to find obstacle free areas for robot operation~\cite{gao2019flying}, the data is often further converted into voxel grids~\cite{xu2021voxel, oleynikova2017voxblox, han2019fiesta}.
 
-Voxel grid representation divides the space into a regular grid of volumetric elements, or voxels, which can be used to create a more manageable approximation of the environment.
-While this approach offers a balance between detail and computational efficiency, it can still introduce discretization errors, particularly when modeling objects with smooth surfaces or intricate details.
-The fidelity of the representation is dependent on the size of the voxels: smaller voxels can capture more detail but require more memory and computation, while larger voxels result in coarser approximations but are more memory-efficient, adaptive voxel grids have been explored, where the voxel size can vary throughout the space to provide higher resolution in regions of interest while conserving resources in less critical areas.
-Voxel grids can incorporate probabilistic information, such as in occupancy grid maps, where each voxel holds a probability indicating the chance of an obstacle's presence.
+Voxel grid~\cite{xu2021voxel, elfes1989using} representation divides the space into a regular grid of volumetric elements, or voxels, which can be used to create a more manageable approximation of the environment.
+While this approach offers a balance between detail and computational efficiency, it can introduce discretization errors, particularly when modeling objects with smooth surfaces or intricate details.
+The fidelity of the representation is dependent on the size of the voxels: smaller voxels can capture more detail but require more memory and computation, while larger voxels result in coarser approximations but are more memory and computation efficient.
+Adaptive voxel grids have been explored~\cite{xu2021voxel}, where the voxel size can vary throughout the space to provide higher resolution in regions of interest while conserving resources in less critical areas.
+Nießner et al.~\cite{niessner2013realtime} proposed Voxel hashing, for more efficient memory management in sparse environments.
+
+Voxel grids can incorporate probabilistic information~\cite{thrun2002probabilistic, dryanovski2010multivolume}, such as in occupancy grid maps, where each voxel holds a probability indicating the chance of an obstacle's presence.
 The occupancy probability of a voxel can be updated dynamically using sensor measurements and Bayesian updating methods.
 As new sensor data is collected, the probabilities are revised to reflect the increased or decreased likelihood of the presence of an obstacle in the voxel space.
-Additionally, techniques such as ESDF (Euclidian Signed Distance Field) and TSDF (Truncated Signed Distance Field) can be utilized in conjunction with voxel grids to encode signed distance information, that describes what is the distance between the selected voxel field and the nearest occupied field.
 
-There has been a number of proposed algorithms for calculating ESDF.
-One common method is the Brushfire method, that spreads from obstacles until it calculates the distance for every field on a grid.
+Voxel grids can represent ESDF (Euclidian Signed Distance Field) and TSDF (Truncated Signed Distance Field)~\cite{oleynikova2017voxblox}, in which case each of the voxels contains information how far nearest obstacle is in its vicinity.
+
+ESDF grids can be generated using sensor measurements.
+Oleynikova et al.~\cite{oleynikova2017voxblox} proposed a method for calculating the ESDF from TSDF.
+Han et al.~\cite{han2019fiesta} proposed a way to integrate point-cloud data into ESDF using ray-casting.
+
+Another way is to generate ESDF from occupancy grids.
+One common approach is the Brushfire method, that spreads from obstacles until it calculates the distance for every field on a grid.
 \add{distance transforms} Jump Flooding Algorithm (JFA) is a similar method, that can be implemented on a GPU for faster parallel distance calculation.
+
 \add{\\-Wavefront \\
@@
   	 -voronoi diagrams\\
-   	 -voxblox, FIESTA \\
      -ray casting, bullet casting ...}
 
+ \alert{10-ish citatov iz seznama ni klicanih v tekstu !}
+
 \clearpage
-
-\section{USING THE TEMPLATE}
-
-Use this sample document as your LaTeX source file to create your document.
-Save this file as {\bf root.tex}.
-You have to make sure to use the cls file that came with this distribution.
-If you use a different style file, you cannot expect to get required margins.
-Note also that when you are creating your out PDF file, the source file is only part of the equation.
-{\it Your \TeX\ $\rightarrow$ PDF filter determines the output file size.
-Even if you make all the specifications to output a letter file in the source - if your filter is set to produce A4, you will only get A4 output.
-}
-
-It is impossible to account for all possible situation, one would encounter using \TeX.
-If you are using multiple \TeX\ files you must make sure that the ``MAIN`` source file is called root.tex - this is particularly important if your conference is using PaperPlaza's built in \TeX\ to PDF conversion tool.
-
-\subsection{Headings, etc}
-
-Text heads organize the topics on a relational, hierarchical basis.
-For example, the paper title is the primary text head because all subsequent material relates and elaborates on this one topic.
-If there are two or more sub-topics, the next level head (uppercase Roman numerals) should be used and, conversely, if there are not at least two sub-topics, then no subheads should be introduced.
-Styles named  Heading 1 ,  Heading 2 ,  Heading 3 , and  Heading 4  are prescribed.
-
-\subsection{Figures and Tables}
-
-Positioning Figures and Tables: Place figures and tables at the top and bottom of columns.
-Avoid placing them in the middle of columns.
-Large figures and tables may span across both columns.
-Figure captions should be below the figures; table heads should appear above the tables.
-Insert figures and tables after they are cited in the text.
-Use the abbreviation  Fig.
-1 , even at the beginning of a sentence.
-
-\begin{table}[h]
-\caption{An Example of a Table}
-\label{table_example}
-\begin{center}
-\begin{tabular}{|c||c|}
-\hline
-One & Two\\
-\hline
-Three & Four\\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-   \begin{figure}[thpb]
-      \centering
-      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
-}}
-      %\includegraphics[scale=1.0]{figurefile}
-      \caption{Inductance of oscillation winding on amorphous
-       magnetic core versus DC bias magnetic field}
-      \label{figurelabel}
-   \end{figure}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels.
-Use words rather than symbols or abbreviations when writing Figure axis labels to avoid confusing the reader.
-As an example, write the quantity  Magnetization , or  Magnetization, M , not just  M .
-If including units in the label, present them within parentheses.
-Do not label axes only with units.
-In the example, write  Magnetization (A/m)  or  Magnetization {A[m(1)]} , not just  A/m .
-Do not label axes with a ratio of quantities and units.
-For example, write  Temperature (K) , not  Temperature/K.
-
-\section{CONCLUSIONS}
-
-A conclusion section is not required.
-Although a conclusion may review the main points of the paper, do not replicate the abstract as the conclusion.
-A conclusion might elaborate on the importance of the work or suggest applications and extensions.
 
@@
                                   % sure that you do not shorten the textheight too much.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{APPENDIX}
-
-Appendixes should appear before the acknowledgment.
-
-\section*{ACKNOWLEDGMENT}
-
-The preferred spelling of the word  acknowledgment  in America is without an  e  after the  g .
-Avoid the stilted expression,  One of us (R.
-B.
-G.) thanks .
-.
-.
-Instead, try  R.
-B.
-G.
-thanks .
-Put sponsor acknowledgments in the unnumbered footnote on the first page.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.
 
@@
 
+\bibitem{dai2022review} Y.
+Dai, C.
+Xiang, Y.
+Zhang, Y.
+Jiang, W.
+Qu, and Q.
+Zhang, “A Review of Spatial Robotic Arm Trajectory Planning,” Aerospace, vol.
+9, p.
+361, Jul.
+2022, doi: 10.3390/aerospace9070361.
+
+\bibitem{gottschalk1996obbtree} S.
+Gottschalk, M.
+C.
+Lin, and D.
+Manocha, “OBBTree: a hierarchical structure for rapid interference detection,” in Proceedings of the 23rd Annual Conference on Computer Graphics and Interactive Techniques, ACM, Aug.
+1996, pp.
+171–180.
+doi: 10.1145/237170.237244.
+
+\bibitem{vandenbergen1997efficient} G.
+van den Bergen, “Efficient Collision Detection of Complex Deformable Models using AABB Trees,” Journal of Graphics Tools, vol.
+2, no.
+4, pp.
+1–13, 1997.
+doi: 10.1080/10867651.1997.10487480.
+
+\bibitem{chen2018path} G.
+Chen, D.
+Liu, Y.
+Wang, Q.
+Jia, and X.
+Zhang, “Path planning method with obstacle avoidance for manipulators in dynamic environment,” International Journal of Advanced Robotic Systems, vol.
+15, no.
+6, Art.
+no.
+1729881418820223, Nov.
+2018, doi: 10.1177/1729881418820223.
+
+\bibitem{puiu2011realtime} D.
+Puiu and F.
+Moldoveanu, “Real-time collision avoidance for redundant manipulators,” in Proc.
+of the 2011 6th IEEE International Symposium on Applied Computational Intelligence and Informatics (SACI), Timisoara, Romania, 2011, pp.
+403–408, doi: 10.1109/SACI.2011.5873037.
+
+\bibitem{oleynikova2017voxblox} H.
+Oleynikova, Z.
+Taylor, M.
+Fehr, R.
+Siegwart, and J.
+Nieto, “Voxblox: Incremental 3D Euclidean Signed Distance Fields for on-board MAV planning,” in Proc.
+of the 2017 IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS), Vancouver, BC, Sep.
+2017, pp.
+1366–1373.
+doi: 10.1109/IROS.2017.8202315.
+
+\bibitem{wurmOctoMap} K.
+M.
+Wurm, A.
+Hornung, M.
+Bennewitz, C.
+Stachniss, and W.
+Burgard, “OctoMap: A Probabilistic, Flexible, and Compact 3D Map Representation for Robotic Systems,” [Details of publication, e.g., in Proc.
+of the Conference/Journal Name, Year, pp.
+Page numbers].
+[DOI or URL if available].
+
+\bibitem{gao2019flying} F.
+Gao, W.
+Wu, W.
+Gao, and S.
+Shen, “Flying on point clouds: Online trajectory generation and autonomous navigation for quadrotors in cluttered environments,” Journal of Field Robotics, vol.
+36, no.
+4, pp.
+710–733, 2019, doi: 10.1002/rob.21842.
+
+\bibitem{elfes1989using} A. Elfes, “Using occupancy grids for mobile robot perception and navigation,” Computer, vol. 22, no. 6, pp. 46–57, Jun. 1989, doi: 10.1109/2.30720.
+
+\bibitem{han2019fiesta} L.
+Han, F.
+Gao, B.
+Zhou, and S.
+Shen, “FIESTA: Fast Incremental Euclidean Distance Fields for Online Motion Planning of Aerial Robots,” arXiv, Jul.
+26, 2019.
+Accessed: Jan.
+11, 2024.
+[Online].
+Available: http://arxiv.org/abs/1903.02144
+
+\bibitem{xu2021voxel} Y.
+Xu, X.
+Tong, and U.
+Stilla, “Voxel-based representation of 3D point clouds: Methods, applications, and its potential use in the construction industry,” Automation in Construction, vol.
+126, p.
+103675, Jun.
+2021, doi: 10.1016/j.autcon.2021.103675.
+
+\bibitem{niessner2013realtime} M.
+Nießner, M.
+Zollhöfer, S.
+Izadi, and M.
+Stamminger, “Real-time 3D reconstruction at scale using voxel hashing,” ACM Trans.
+Graph., vol.
+32, no.
+6, pp.
+1–11, Nov.
+2013, doi: 10.1145/2508363.2508374.
+
+\bibitem{dryanovski2010multivolume} I.
+Dryanovski, W.
+Morris, and J.
+Xiao, “Multi-volume occupancy grids: An efficient probabilistic 3D mapping model for micro aerial vehicles,” in Proc.
+of the 2010 IEEE/RSJ International Conference on Intelligent Robots and Systems, Taipei, Oct.
+2010, pp.
+1553–1559.
+doi: 10.1109/IROS.2010.5652494.
+
+\bibitem{thrun2002probabilistic} S. Thrun, Probabilistic robotics, vol. 45, 2002. Accessed: Jun. 14, 2023. [Online]. Available: https://dl.acm.org/doi/10.1145/504729.504754
+
 \end{thebibliography}

### 2024-01-26 Articles/Article1-Conference/root.tex
more

@@
 One common approach involves enclosing the obstacles around the manipulator within a collection of simple geometric primitives and then constructing a tree-like structure to facilitate efficient navigation and path planning~\cite{dai2022review}.
-Two of the commonly used methods for this purpose include AABB (axis-aligned bounding boxes)~\cite{vandenbergen1997efficient, chen2018path} and OBB (oriented bounding boxes)~\cite{gottschalk1996obbtree, puiu2011realtime}\remove{and FDH (fixed directions hulls).}.
+Two of the commonly used methods for this purpose include AABB (axis-aligned bounding boxes)~\cite{vandenbergen1997efficient, chen2018path, luo2018collisionfree} and OBB (oriented bounding boxes)~\cite{gottschalk1996obbtree, puiu2011realtime}\remove{and FDH (fixed directions hulls).}.
 Another significant method worth mentioning is the use of Octomaps~\cite{wurmOctoMap}.
@@
 
-Voxel grids can represent ESDF (Euclidian Signed Distance Field) and TSDF (Truncated Signed Distance Field)~\cite{oleynikova2017voxblox}, in which case each of the voxels contains information how far nearest obstacle is in its vicinity.
+Another type of data that voxel fields can hold are ESDF (Euclidian Signed Distance Field) or TSDF (Truncated Signed Distance Field)~\cite{oleynikova2017voxblox}, in which case each of the voxels contains information how far nearest obstacle is in its vicinity.
+
+\subsection{ESDF creation}
 
@@
 Another way is to generate ESDF from occupancy grids.
-One common approach is the Brushfire method, that spreads from obstacles until it calculates the distance for every field on a grid.
-\add{distance transforms} Jump Flooding Algorithm (JFA) is a similar method, that can be implemented on a GPU for faster parallel distance calculation.
+When generating ESDF from occupancy grid, common approach is the Brushfire method~\cite{lau2010improved}, that spreads from obstacles until it calculates the distance for every field on a grid.
+Jump Flooding Algorithm (JFA) is a similar method, that can be implemented on a GPU for faster parallel distance calculation.
 
@@
 
+\bibitem{lau2010improved} B.
+Lau, C.
+Sprunk, and W.
+Burgard, “Improved updating of Euclidean distance maps and Voronoi diagrams,” in Proc.
+of the 2010 IEEE/RSJ International Conference on Intelligent Robots and Systems, Taipei, Oct.
+2010, pp.
+281–286.
+doi: 10.1109/IROS.2010.5650794.
+
+\bibitem{luo2018collisionfree} L.
+Luo et al., “Collision-Free Path-Planning for Six-DOF Serial Harvesting Robot Based on Energy Optimal and Artificial Potential Field,” Complexity, vol.
+2018, pp.
+1–12, Nov.
+2018, doi: 10.1155/2018/3563846.
+
 \end{thebibliography}

### 2024-01-29 Articles/Article1-Conference/root.tex
added reference

@@
 Xia et al.~\cite{c49} uses a variation of APF for manipulator motion~\cite{c49}.
+Zhang et al.~\cite{zhang2021obstacle} proposes dynamic repulsive field based on direction and speed between point on robot and obstacle, it also suggests decision making force that moves the robot away from certain local minima.
 Chen et al.
-proposes an application using APF in joint space and a variable kinematic optimization step~\cite{c50}.
+proposes\remove{ an application }using APF in joint space and a variable kinematic optimization step~\cite{c50}.
 Long~\cite{c44} suggests creating motion plan of the manipulator using APF, he extends it using RRT to calculate virtual attractive point for the robot to move towards in case of local minima.
-Zhu et al.
-proposes use of APF in combination with MPC, to plan in environment with dynamic obstacles~\cite{c48}.
+Zhu et al.~\cite{c48} proposes use of APF in combination with MPC, to plan in environment with dynamic obstacles.
 
@@
 
-ESDF grids can be generated using sensor measurements.
+ESDF grids can be generated directly using sensor measurements.
 Oleynikova et al.~\cite{oleynikova2017voxblox} proposed a method for calculating the ESDF from TSDF.
@@
 
-Another way is to generate ESDF from occupancy grids.
+Another way is to generate ESDF from occupancy grids, that can be previously generated using sensors.
 When generating ESDF from occupancy grid, common approach is the Brushfire method~\cite{lau2010improved}, that spreads from obstacles until it calculates the distance for every field on a grid.
@@
 
+\bibitem{zhang2021obstacle} W.
+Zhang, H.
+Cheng, L.
+Hao, X.
+Li, M.
+Liu, and X.
+Gao, “An obstacle avoidance algorithm for robot manipulators based on decision-making force,” Robotics and Computer-Integrated Manufacturing, vol.
+71, p.
+102114, Oct.
+2021, doi: 10.1016/j.rcim.2020.102114.
+
 \end{thebibliography}

### 2024-01-30 Articles/Article1-Conference/root.tex
more citations

@@
 Xia et al.~\cite{c49} uses a variation of APF for manipulator motion~\cite{c49}.
+Park et al.~\cite{park2020trajectory} used a numerical Jacobian in combination with APF for motion planning.
 Zhang et al.~\cite{zhang2021obstacle} proposes dynamic repulsive field based on direction and speed between point on robot and obstacle, it also suggests decision making force that moves the robot away from certain local minima.
@@
 Zhu et al.~\cite{c48} proposes use of APF in combination with MPC, to plan in environment with dynamic obstacles.
-
-\alert{
- -fix citation styles (sometimes et al, sometimes not, sometimes names, sometimes years ... not same format)
-}
 
@@
 Two of the commonly used methods for this purpose include AABB (axis-aligned bounding boxes)~\cite{vandenbergen1997efficient, chen2018path, luo2018collisionfree} and OBB (oriented bounding boxes)~\cite{gottschalk1996obbtree, puiu2011realtime}\remove{and FDH (fixed directions hulls).}.
+The problem is generation of such hulls based on partial sensor measurements of the environment.
+Han et al.~\cite{han2018dynamic} used a complicated pipeline to convert point-cloud sensor measurements to octree, than to voxel grid and finally into convex hulls, used for collision detections.
 Another significant method worth mentioning is the use of Octomaps~\cite{wurmOctoMap}.
@@
 
-Another type of data that voxel fields can hold are ESDF (Euclidian Signed Distance Field) or TSDF (Truncated Signed Distance Field)~\cite{oleynikova2017voxblox}, in which case each of the voxels contains information how far nearest obstacle is in its vicinity.
+Another type of data that voxel \replace{grids}{fields} can hold are ESDF (Euclidian Signed Distance Field) or TSDF (Truncated Signed Distance Field)~\cite{oleynikova2017voxblox}, in which case each of the voxels contains information how far nearest obstacle is in its vicinity.
 
@@
 
-Another way is to generate ESDF from occupancy grids, that can be previously generated using sensors.
+Another way is to generate ESDF from occupancy grids, that can be previously generated using sensors or based on predefined maps.
 When generating ESDF from occupancy grid, common approach is the Brushfire method~\cite{lau2010improved}, that spreads from obstacles until it calculates the distance for every field on a grid.
-Jump Flooding Algorithm (JFA) is a similar method, that can be implemented on a GPU for faster parallel distance calculation.
+Jump Flooding Algorithm (JFA)~\cite{rong2006jump} is a similar method, that can be implemented on a GPU for faster parallelized distance calculation.
+
+%Euclidean Signed Distance Fields (ESDFs) are valuable in robotic navigation and 3D reconstruction, providing a measure of the shortest distance to the nearest obstacle for each point in space.
+The creation of ESDF grids can leverage sensor measurements directly or be derived from other data structures like Truncated Signed Distance Fields (TSDFs) or occupancy grids.
+%
+%One prominent method for generating ESDFs is converting TSDFs into ESDFs.
+Oleynikova et al.
+[(2017)] introduced a technique that efficiently computes the ESDF from TSDFs, utilizing incremental updates for dynamic environments.
+This approach is significant for real-time applications in robotics, where rapid updates to the environment model are crucial.
+Similarly, Han et al.
+[(2019)] developed a method that integrates point-cloud data into ESDF using ray-casting, offering a way to construct ESDFs directly from sensor data.
+%
+%Alternatively, ESDFs can be generated from occupancy grids, which themselves are often derived from sensor measurements.
+The Brushfire algorithm, as improved by Lau et al.
+[(2010)], is a commonly used technique in this context.
+It operates by spreading from obstacles within the grid, calculating the minimum distance to an obstacle for each grid cell.
+This approach is particularly effective for static environments where the layout of obstacles does not change frequently.
+%
+%Furthermore, the Jump Flooding Algorithm (JFA) offers another method for generating ESDFs from occupancy grids.
+JFA is notable for its suitability for implementation on GPUs, enabling faster parallel distance calculations.
+This advantage is critical in applications requiring rapid processing of large datasets, such as in high-resolution 3D mapping and complex robotic navigation tasks.
+%
+%In conclusion, the generation of ESDFs from sensor measurements or derived data structures like TSDFs and occupancy grids is a crucial aspect of robotic navigation and 3D environment modeling.
+Techniques like those proposed by Oleynikova et al.
+[(2017)], Han et al.
+[(2019)], and the improvements to the Brushfire algorithm by Lau et al.
+[(2010)], as well as the application of JFA, highlight the ongoing advancements in efficient and accurate ESDF creation methodologies.
 
@@
  \alert{10-ish citatov iz seznama ni klicanih v tekstu !}
+
+ \comment{}{
+ 	-fix citation styles (sometimes et al, sometimes not, sometimes names, sometimes years ... not same format)
+ }
 
@@
 
+\bibitem{park2020trajectory} S.-O.
+Park, M.
+C.
+Lee, and J.
+Kim, “Trajectory Planning with Collision Avoidance for Redundant Robots Using Jacobian and Artificial Potential Field-based Real-time Inverse Kinematics,” Int.
+J.
+Control Autom.
+Syst., vol.
+18, no.
+8, Art.
+no.
+8, Aug.
+2020, doi: 10.1007/s12555-019-0076-7.
+
+\bibitem{han2018dynamic} D.
+Han, H.
+Nie, J.
+Chen, and M.
+Chen, “Dynamic obstacle avoidance for manipulators using distance calculation and discrete detection,” Robotics and Computer-Integrated Manufacturing, vol.
+49, pp.
+98–104, Feb.
+2018, doi: 10.1016/j.rcim.2017.05.013.
+
+\bibitem{rong2006jump} G.
+Rong and T.-S.
+Tan, “Jump flooding in GPU with applications to Voronoi diagram and distance transform,” in Proceedings of the 2006 Symposium on Interactive 3D Graphics and Games - SI3D ’06, Redwood City, California, 2006, p.
+109.
+doi: 10.1145/1111411.1111431.
+
 \end{thebibliography}

### 2024-02-13 Articles/Article1-Conference/root.tex
UNFILTERED REFERENCES: added all references

@@
 
-Local motion planning approaches employ optimization techniques, two common ones are inverse kinematics~\cite{c29,c38}, that is based on finding a least squares solution of the manipulator joint velocities, and quadratic programming (QP)~\cite{c21,c22,c23}.
+Local motion planning approaches employ optimization techniques, two common ones are inverse kinematics~\cite{c29,c38}, that is based on finding a least squares solution of the manipulator joint velocities, and quadratic programming (QP)~\cite{c21,c22, haviland2021neo, c23}.
 Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
@@
 Nießner et al.~\cite{niessner2013realtime} proposed Voxel hashing, for more efficient memory management in sparse environments.
+Dryanovski et al.~\cite{dryanovski2010multivolume} proposed variable voxel height and saving information about occupied and about free measurements, as well as the information about number of measurements for a specific cell.
 
@@
 Jump Flooding Algorithm (JFA)~\cite{rong2006jump} is a similar method, that can be implemented on a GPU for faster parallelized distance calculation.
+
+Methods that allow for the dynamical distance field generation are rare.
+Zhou et al.~\cite{zhou2021egoplanner} propose use of pairs of points on trajectory and obstacles in their quadcopter trajectory optimization algorithm.
+
+%\remove{The field of image processing invented many different morphological algorithms, that work on binary images and calculate a value for every pixel in image, that represents the distance from nearest occupied binary pixel~\cite{szeliski2021computer, felzenszwalb2012distance, rosenfeld1968distance}.
+Approaches can be seperated based on the distance metric that they use.
+}
 
@@
 doi: 10.1016/j.robot.2014.02.005.
+
+\bibitem{haviland2021neo} J.
+Haviland and P.
+Corke, “NEO: A Novel Expeditious Optimisation Algorithm for Reactive Motion Control of Manipulators,” IEEE Robot.
+Autom.
+Lett., vol.
+6, no.
+2, Art.
+no.
+2, Apr.
+2021, doi: 10.1109/LRA.2021.3056060.
 
@@
 
+\bibitem{szeliski2021computer} R. Szeliski, “Computer Vision: Algorithms and Applications, 2nd Edition,” Springer, 2021.
+
+\bibitem{rosenfeld1968distance} A. Rosenfeld and J. L. Pfaltz, “Distance functions on digital pictures,” Pattern Recognition, vol. 1, no. 1, pp. 33–61, Jul. 1968, doi: 10.1016/0031-3203(68)90013-7.
+
+\bibitem{felzenszwalb2012distance} P.
+F.
+Felzenszwalb and D.
+P.
+Huttenlocher, “Distance Transforms of Sampled Functions,” Theory of Comput., vol.
+8, no.
+1, pp.
+415–428, 2012, doi: 10.4086/toc.2012.v008a019.
+
+\bibitem{zhou2021egoplanner} X.
+Zhou, Z.
+Wang, H.
+Ye, C.
+Xu, and F.
+Gao, “EGO-Planner: An ESDF-Free Gradient-Based Local Planner for Quadrotors,” IEEE Robot.
+Autom.
+Lett., vol.
+6, no.
+2, pp.
+478–485, Apr.
+2021, doi: 10.1109/LRA.2020.3047728.
+
+\bibitem{dryanovski2010multivolume} I.
+Dryanovski, W.
+Morris, and J.
+Xiao, “Multi-volume occupancy grids: An efficient probabilistic 3D mapping model for micro aerial vehicles,” in 2010 IEEE/RSJ International Conference on Intelligent Robots and Systems, Taipei: IEEE, Oct.
+2010, pp.
+1553–1559, doi: 10.1109/IROS.2010.5652494.
+
 \end{thebibliography}

### 2024-02-13 Code/Full_RTConvolution_Multiple_Points.m
fixed slowdown velocity, added gaussian kernel stretching to 1

@@ -7,8 +7,8 @@ mid_joints = 0
 avoid_task = 1
 kinematics_solution = 'exact-reduced' % OPTIONS: exact-reduced , exact , approximate
 timestep_secondary_gain_change = 0 % if selected, secondary task will start with normal gain and fall with time
-secondary_exec_stop_k = 1 % primary task will slow down (>0) or stop executing if secondary task has big velocities 
-min_exec_slowdown_size = 0 % if poi is closer than this value, primary task will slow down
+secondary_exec_stop_k = 5 % primary task will slow down (>0) or stop executing if secondary task has big velocities 
+min_exec_slowdown_size = 1 % if poi is closer than this value, primary task will slow down
 
 % -----------------------------------------------------------
 % number of points per segment for obstacle avoidance taskmanipulability_avoidance
@@ -21,7 +21,7 @@ weights_avoidance = weights_avoidance / norm(weights_avoidance,1) / 10;
 % -----------------------------------------------------------
 
 Tstep = 0.1 % time step
-Nmax = 500 % max number of iterations
+Nmax = 250 % max number of iterations
 space_resolution = grid.resolution; % resolution of the obstacles grid
 
 % weights for different tasks
@@ -88,13 +88,7 @@ manipulability_avoidance = [];
 % -----------------------------------------------------------
 
 % repulsive kernels
-rep_kernels = REP_kernels('linear');
-
-% obstacles distance kernel
-% dist_kernel = euclidian_kernel_3D(61, 61, 61);
-% dist_kernel = gaussian_kernel_3d(61, 61, 61, 10);
-
-
+rep_kernels = REP_kernels('type', 'linear');
 
 %% OPTIMIZATION LOOP
 % -----------------------------------------------------------
@@ -219,7 +213,8 @@ while current_dist > goal_dist && Niter <= Nmax
 
         % SORT POI BASED ON DISTANCE FROM OBSTACLES
         % --------------------
-        [poi_sizes, poi_indeces] = sort(vecnorm(rep_values));            
+        [poi_sizes, poi_indeces] = sort(vecnorm(rep_values));
+        poi_sizes            
 
 
         %% FOR EVERY POI CALCULATE AVOIDANCE JOINT SPEEDS (USING INVERSE KINEMATICS SOLUTION)
@@ -326,12 +321,12 @@ while current_dist > goal_dist && Niter <= Nmax
[…]

### 2024-02-13 Code/REP_kernels.m

@@ -57,6 +57,6 @@ function [kernels] = REP_kernels(varargin)
     kernel_y = directional_kernel_3d('y', kernel_length, sigma, kernel_width, 0, kernel_height, 0, type);
     kernel_z = directional_kernel_3d('z', kernel_length, sigma, kernel_width, 0, kernel_height, 0, type);
 
-    kernels = {kernel_x, kernel_y, kernel_z};
+    kernels = {kernel_x/max(kernel_x(:)), kernel_y/max(kernel_y(:)), kernel_z/max(kernel_z(:))};
 
 end

### 2024-02-26 Code/REP_field_calculation.m
move point by 0.5 of resolution step, to basically set potential field values to the middle of the grid cells

@@ -43,7 +43,15 @@ function [rep_values, logs_rep_values] = REP_field_calculation(grid, kernels, po
     %% USE INTERPOLATION
     if interpolation_mode
         
-        center_point = point * grid.resolution;
+        % PREPARE POINT COORDINATES
+        % --------------------------------------------------------------
+
+        % move point by 0.5 of resolution step, to basically set potential 
+        % field values to the middle of the grid cells
+        point_grid = (point + 1/grid.resolution/2);
+
+        % scale to grid resolution
+        center_point = point_grid * grid.resolution; 
          
         % get nearby grid cells indexes
         X = [floor(center_point(1)) ceil(center_point(1))];
@@ -121,9 +129,9 @@ function [rep_values, logs_rep_values] = REP_field_calculation(grid, kernels, po
             half_size = floor(kernel_size / 2);
     
             % Calculate x, y, and z ranges for the window while ensuring they are positive and within grid boundaries.
-            x_cells = round((grid_point(2) - half_size(1)) : (grid_point(2) + half_size(1))); % grid goes y-x-z
-            y_cells = round((grid_point(1) - half_size(2)) : (grid_point(1) + half_size(2)));
-            z_cells = round((grid_point(3) - half_size(3)) : (grid_point(3) + half_size(3)));
+            x_cells = floor((grid_point(2) - half_size(1)) : (grid_point(2) + half_size(1))); % grid goes y-x-z
+            y_cells = floor((grid_point(1) - half_size(2)) : (grid_point(1) + half_size(2)));
+            z_cells = floor((grid_point(3) - half_size(3)) : (grid_point(3) + half_size(3)));
 
             % init every cells in window to occupued by default
             window = ones(kernel_size)*box_value;

### 2024-02-26 Articles/Article 1/main.tex
merged article1 text drafts

@@
 \section{Introduction}
+
+\subsection{Motion Planning}
+
+% PREDSTAVITEV REDUNDANCE
+
+Kinematic redundancy~\cite{siciliano1990kinematic, siciliano2016springer} enables a manipulator to follow a predefined task space trajectory using the endeffector (EE), while simultaneously, optimising for an additional task with the remaining movement capacity without impacting the trajectory adherence.
+This is possible because the robot's degrees of freedom (DOF) go beyond what is required to perform the primary task.
+Consequently, the robot can adopt different joint configurations optimised according to the secondary task while performing the primary task.
+Common secondary tasks~\cite{siciliano2010robot} include avoiding singularities, optimising the manipulability measure, minimising joint torques and avoiding obstacles in the operating space.
+
+% PREDSTAVITEV PRISTOPOV PLANIRANJA POTI V ROBOTIKI
+The task of finding the joint trajectories of a manipulator is called motion planning~\cite{IDEASLab2023}.
+It consists of finding a sequence of joint configurations for a robot so that the robot can move along this path from its initial configuration to the goal configuration without colliding with itself, static obstacles or other agents in the environment.
+In addition to collision avoidance, motion planning for manipulators can optionally take into account various constraints, such as position, velocity, acceleration or jerk constraints for the joint angle or end effector, precision of the end effector with respect to position and orientation, stability of the manipulator, avoidance of singularities, or any number of other criteria.
+
+There are numerous methods for planning manipulator movements~\cite{c51, c52}, they can rougly be seperated into global and local approaches.
+Global, sampling-based, methods such as PRM~\cite{vsvestka1997motion}, RRT*~\cite{lavalle1998rapidly, karaman2010incremental}, RRT-Connect~\cite{kuffner2000rrt}, Informerd RRT*~\cite{gammell2014informed}, BIT*~\cite{gammell2015batch} and others offer a globally optimal solution based on a global search in configuration space.
+However, the generated trajectories are not always smooth or optimal, and the performance of the methods may be insufficient for real-time operation.
+
+%Global planning methods are typically more comprehensive, aiming to find a globally optimal trajectory for the manipulator.
+These methods often involve mapping out the entire path before movement begins, taking into account the known environment, which usually includes static obstacles.
+However, the downside of global planning methods is their computational intensity, which often renders them too slow for real-time applications.
+This limitation is particularly pronounced in dynamic or unpredictable environments where conditions change rapidly and require quick responses.
+Examples of global planning methods include graph-based algorithms like A* and Dijkstra's algorithm, which are known for their effectiveness in pathfinding but also for their computational demands [(Lavalle, 2006)].
+
+%Sampling-based methods such as PRM, RRT, RRT*, RRT-Connect, Informerd RRT*, BIT* and others offer a solution based on a global search in configuration space.
+However, the generated trajectories are not always smooth or optimal, and the performance of the methods may be insufficient for real-time operation.
+
+\remove{Recently, a number of learning-based methods using data-driven techniques have been proposed to improve or accelerate the functionality of sampling-based methods. }
+
+\remove{Trajectory optimisation methods such as CHOMP, STOMP and TrajOpt, on the other hand, use optimization to improve an initial seed trajectory.
+Consequently, the optimality of the solution is highly dependent on this initial trajectory.
+Nevertheless, these methods are capable of generating smooth trajectories, and although they can be too computationally intensive for high DOF dynamic real-time environments, they are generally effective in finding constrained motion plans.
+}
+
+Local motion planning approaches employ optimization techniques, two common ones are inverse kinematics~\cite{c29,c38}, that is based on finding a least squares solution of the manipulator joint velocities, and quadratic programming (QP)~\cite{c21,c22, haviland2021neo, c23}.
+Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
+However, since they do not plan further than one step ahead, they tend to get stuck in local minima.
+Therefore, they are often combined with a higher-level planner, for global static environment based \replace{motion}{path} planning, while local optimisation takes dynamic environment changes into account.
+In the following text we will focus on inverse kinematics based approaches.
+
+\subsection{Kinematic Obstacle Avoidance}
+
+\remove{Different control schemes have been proposed.
+Acceleration-based control excels in precise handling of motion changes, velocity-based control offers consistent and smooth movement, while force and torque-based control provides direct control over joint forces for robust physical interactions.
+(CITATIONS in COMMENT BELLOW)}
+%\remove{Different control schemes have been proposed~\cite{c24}.
+Acceleration-based control excels in precise handling of motion changes, velocity-based control offers consistent and smooth movement, while force and torque-based control~\cite{c25, c26, c27, c28, c31} provides direct control over joint forces for robust physical interactions.}
+
+% ----
+
+%In our work we use velocity control strategy for obstacle avoidance using the manipulator.
+naša primarna naloga premik vrha robota - end efektorja (EE) v želeno pozo, to je točko in orientacijo v prostoru naloge.
+Naša sekundarna naloga se tako izvaja v preostalem prostoru gibanja sklepov redundance robota (DOF, DOR).
+???
+je predlagal razširitev jakobijeve matrike end-efektorja z dodatno matriko transformacije izogibne hitrosti iz kartzezičnega prostora v prostor sklepov.
+Če nalogi nista popolnoma neodvisni ena od druge, in njune sklepne hitrosti niso medsebojno ortagonalne, lahko to pripelje do medsebojnega oviranja izbajanja nalog.
+Zato je ???
+predlagal uporabo transformacijo hitrosti sekundarne naloge v  ničelnega prostora hitrosti primarne nalgoe, to je v prostor ortagonalnih hitrosti.
+Posledično izvajanje sekundarne naloge ne vpliva neposredno v izvajanje primarne naloge, gledano le v enem trenutku izvajanja.
+%
+%Različni razisklovalci so se lotili problema kinematičnega izogibanja na različne načine, kot je opisano v \cite{c30, c39}. 
+%
+%Colbaugh in Glass  1989 \cite{c31, c32} sta se problema lotila v dveh korakih.
+V prvem koraku sta izračunala trajektorijo vrha robota, v drugem pa sta uporabila optimizacijo za izboljšanje dinamičnega odziva robota med izvajanjem trajektorije in za druge naloge, kot so izogibanje ovir.
+%
+%Sciavicco and Siciliano \cite{c34, c35} and Egeland~\cite{c36} independetly introduced the concept ot task-space augmentation.
+It was later revisited by searji~\cite{c37}.
+The Solution is to stack the remaining rows in Jacobian with secondary task Jacobian into square matrix, that returns only one solution.
+%
+%Macijewski in Klein (Maciejewski & Klein, 1985) \cite{c29}, Nakamura in Hanafusa 1987 \cite{c38} predstavijo koncept prioritete nalog in ničelnega prostora za izogibanje, pri čemer se sekundarna naloga lahko giblje le s hitrostmi, ki ne vplivajo neposredno na gibanje primarne naloge.
+pri čemer sta izogibne sklepne hitrosti izračunala funkcijo najmanjše razdalje med točko na manipulatorju in točko izogibanja / objekta / ovire.
+%
+%Omenjeni pristopi se ne osredotočajo na okolje, tega pogosto prikažejo kot geometrijske primitive in nato razdaljo izračunajo kot razdaljo med primitivom in premicami, ki opisujejo manipulator. 
+%
+%Khatib \cite{c33} proposes the concept of Artificial Potential Field.
+Polje sestavljata odbojna komponenta, ki end effektor odbija od ovir, ki so predstavljene kot geometrijski primitivi in privlačna komponenta, ki manipulator privlači proti cilju.
+%
+%
+%
+
+Various researchers have adopted different approaches to kinematic avoidance, as detailed in~\cite{c30, c39}.
+
+Colbaugh and Glass (1989)~\cite{c31, c32} tackled this problem in a two-step process.
+Initially, they calculated the robot's end-effector trajectory.
+Subsequently, they used optimization to enhance the robot's dynamic response during trajectory execution and for tasks like obstacle avoidance.
+
+Sciavicco and Siciliano~\cite{c34, c35}, as well as Egeland~\cite{c36}, independently introduced the concept of task-space augmentation, later revisited by Searji~\cite{c37}.
+This method involves extending the primary task Jacobian with the secondary task Jacobian into a square matrix, yielding a singular solution.
+
+Maciejewski and Klein (1985)~\cite{c29}, and Nakamura and Hanafusa (1987)~\cite{c38}, presented the concept of task prioritization and a null space for avoidance.
+In this framework, the secondary task is restricted to velocities that do not directly affect the primary task's movement.
+Their method calculates avoidance joint velocities based on the minimum distance between a point on the manipulator and an avoidance point, object, or obstacle.
+Žlajpah~\cite{c41} improved the concept by proposing a reduced operational space formulation, that reshapes the Jacobain of the avoidance task from three cartesian axis to only the direction of obstacle avoidance.
+Petrič~\cite{c42} suggested a way to smoothly transition between avoidance and trajectory following tasks.
+
+\subsection{Artificial Potential Field}
+
+These approaches typically do not focus on the environmental context, often representing it as geometric primitives and then calculating the distance between these primitives and the lines describing the manipulator.
+
+Khatib~\cite{c33} proposes the concept of an Artificial Potential Field.
+This field consists of a repulsive component that deflects the end effector away from obstacles, depicted as geometric primitives, and an attractive component that draws the manipulator towards its target.
+In the following years many different modifications and improvement of the original APF idea have been proposed, often focused on removal of local minimas in the potential field.
+Khim and Khosla~\cite{c40} suggested the use of harmonic functions to solve the problem of local minima.
+Pinto et al.
+\cite{c43} proposes to vary the field based on the distance of robot from obstacles to fill the local minima.
+Many researchers tried using APF as a heuristic to better guide sampling based approaches~\cite{c45, c46, c47}.
+
+Many of the recent works focus on use of a variation of artificial potential field to plan motion of the manipulator.
+Xia et al.~\cite{c49} uses a variation of APF for manipulator motion~\cite{c49}.
+Park et al.~\cite{park2020trajectory} used a numerical Jacobian in combination with APF for motion planning.
+Zhang et al.~\cite{zhang2021obstacle} proposes dynamic repulsive field based on direction and speed between point on robot and obstacle, it also suggests decision making force that moves the robot away from certain local minima.
+Chen et al.
+proposes\remove{ an application }using APF in joint space and a variable kinematic optimization step~\cite{c50}.
+Long~\cite{c44} suggests creating motion plan of the manipulator using APF, he extends it using RRT to calculate virtual attractive point for the robot to move towards in case of local minima.
+Zhu et al.~\cite{c48} proposes use of APF in combination with MPC, to plan in environment with dynamic obstacles.
+
+%The inverse kinematics with task prioritisation, which is based on finding a least squares solution of the manipulator joint velocities, and linear quadratic programming (LQR) are two of the local optimization methods.
+Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
+However, since they do not plan further than one step ahead, they tend to get stuck in local minima.
+Therefore, they are often combined with a higher-level planner, for global static environment based \replace{motion}{path} planning, while local optimisation takes dynamic environment changes into account.
+
+\subsection{Environment representation}
+
+% PREDSTAVITEV PRISTOPOV RAČUNANJA ODDALJENOSTI OVIRAM
+%To be able to generate collision free trajectory we need to have a representation of the static environment.
+One common approach is to enclose the obstacles around the manipulator into a collection of simple geometric primitives and then construct a tree-like structure to achieve the goal~\cite{dai2022review}.
+Some of the commonly used methods are AABB (axis-aligned bounding boxes)~\cite{vandenbergen1997efficient, chen2018path}, OBB (oriented bounding boxes)~\cite{gottschalk1996obbtree, puiu2011realtime}.
+
+To be able to generate collision free trajectory we need to have a representation of the static environment.
+One common approach involves enclosing the obstacles around the manipulator within a collection of simple geometric primitives and then constructing a tree-like structure to facilitate efficient navigation and path planning~\cite{dai2022review}.
+Two of the commonly used methods for this purpose include AABB (axis-aligned bounding boxes)~\cite{vandenbergen1997efficient, chen2018path, luo2018collisionfree} and OBB (oriented bounding boxes)~\cite{gottschalk1996obbtree, puiu2011realtime}\remove{and FDH (fixed directions hulls).}.
+The problem is generation of such hulls based on partial sensor measurements of the environment.
+Han et al.~\cite{han2018dynamic} used a complicated pipeline to convert point-cloud sensor measurements to octree, than to voxel grid and finally into convex hulls, used for collision detections.
+Another significant method worth mentioning is the use of Octomaps~\cite{wurmOctoMap}.
+Octomaps employ an octree data structure to represent 3D environments efficiently, making them particularly suitable for areas with large open spaces.
+
+In real environment we usually collect depth data with LiDAR, radar of RGBD camera sensors, that usually return clouds of points, that tell us the distance of objects in robots environment.
+While approaches have been proposed to use the point-cloud data directly to find obstacle free areas for robot operation~\cite{gao2019flying}, the data is often further converted into voxel grids~\cite{xu2021voxel, oleynikova2017voxblox, han2019fiesta}.
+
+Voxel grid~\cite{xu2021voxel, elfes1989using} representation divides the space into a regular grid of volumetric elements, or voxels, which can be used to create a more manageable approximation of the environment.
+While this approach offers a balance between detail and computational efficiency, it can introduce discretization errors, particularly when modeling objects with smooth surfaces or intricate details.
+The fidelity of the representation is dependent on the size of the voxels: smaller voxels can capture more detail but require more memory and computation, while larger voxels result in coarser approximations but are more memory and computation efficient.
+Adaptive voxel grids have been explored~\cite{xu2021voxel}, where the voxel size can vary throughout the space to provide higher resolution in regions of interest while conserving resources in less critical areas.
+Nießner et al.~\cite{niessner2013realtime} proposed Voxel hashing, for more efficient memory management in sparse environments.
+Dryanovski et al.~\cite{dryanovski2010multivolume} proposed variable voxel height and saving information about occupied and about free measurements, as well as the information about number of measurements for a specific cell.
+
+Voxel grids can incorporate probabilistic information~\cite{thrun2002probabilistic, dryanovski2010multivolume}, such as in occupancy grid maps, where each voxel holds a probability indicating the chance of an obstacle's presence.
+The occupancy probability of a voxel can be updated dynamically using sensor measurements and Bayesian updating methods.
+As new sensor data is collected, the probabilities are revised to reflect the increased or decreased likelihood of the presence of an obstacle in the voxel space.
+
+Another type of data that voxel \replace{grids}{fields} can hold are ESDF (Euclidian Signed Distance Field) or TSDF (Truncated Signed Distance Field)~\cite{oleynikova2017voxblox}, in which case each of the voxels contains information how far nearest obstacle is in its vicinity.
+
+\subsection{ESDF creation}
+
+ESDF grids can be generated directly using sensor measurements.
+Oleynikova et al.~\cite{oleynikova2017voxblox} proposed a method for calculating the ESDF from TSDF.
+Han et al.~\cite{han2019fiesta} proposed a way to integrate point-cloud data into ESDF using ray-casting.
+
+Another way is to generate ESDF from occupancy grids, that can be previously generated using sensors or based on predefined maps.
+When generating ESDF from occupancy grid, common approach is the Brushfire method~\cite{lau2010improved}, that spreads from obstacles until it calculates the distance for every field on a grid.
+Jump Flooding Algorithm (JFA)~\cite{rong2006jump} is a similar method, that can be implemented on a GPU for faster parallelized distance calculation.
+
+Methods that allow for the dynamical distance field generation are rare.
+Zhou et al.~\cite{zhou2021egoplanner} propose use of pairs of points on trajectory and obstacles in their quadcopter trajectory optimization algorithm.
+
+%\remove{The field of image processing invented many different morphological algorithms, that work on binary images and calculate a value for every pixel in image, that represents the distance from nearest occupied binary pixel~\cite{szeliski2021computer, felzenszwalb2012distance, rosenfeld1968distance}.
+Approaches can be seperated based on the distance metric that they use.
+}
+
+%Euclidean Signed Distance Fields (ESDFs) are valuable in robotic navigation and 3D reconstruction, providing a measure of the shortest distance to the nearest obstacle for each point in space.
+The creation of ESDF grids can leverage sensor measurements directly or be derived from other data structures like Truncated Signed Distance Fields (TSDFs) or occupancy grids.
+%
+%One prominent method for generating ESDFs is converting TSDFs into ESDFs.
+Oleynikova et al.
+[(2017)] introduced a technique that efficiently computes the ESDF from TSDFs, utilizing incremental updates for dynamic environments.
+This approach is significant for real-time applications in robotics, where rapid updates to the environment model are crucial.
+Similarly, Han et al.
+[(2019)] developed a method that integrates point-cloud data into ESDF using ray-casting, offering a way to construct ESDFs directly from sensor data.
+%
+%Alternatively, ESDFs can be generated from occupancy grids, which themselves are often derived from sensor measurements.
+The Brushfire algorithm, as improved by Lau et al.
+[(2010)], is a commonly used technique in this context.
+It operates by spreading from obstacles within the grid, calculating the minimum distance to an obstacle for each grid cell.
+This approach is particularly effective for static environments where the layout of obstacles does not change frequently.
+%
+%Furthermore, the Jump Flooding Algorithm (JFA) offers another method for generating ESDFs from occupancy grids.
+JFA is notable for its suitability for implementation on GPUs, enabling faster parallel distance calculations.
+This advantage is critical in applications requiring rapid processing of large datasets, such as in high-resolution 3D mapping and complex robotic navigation tasks.
+%
+%In conclusion, the generation of ESDFs from sensor measurements or derived data structures like TSDFs and occupancy grids is a crucial aspect of robotic navigation and 3D environment modeling.
+Techniques like those proposed by Oleynikova et al.
+[(2017)], Han et al.
+[(2019)], and the improvements to the Brushfire algorithm by Lau et al.
+[(2010)], as well as the application of JFA, highlight the ongoing advancements in efficient and accurate ESDF creation methodologies.
+
+\add{\\-Wavefront \\
+	 -D* Klančar\\
+ 	 -Euclidian distnace algorithms\\
+ 	 -distance transforms\\
+  	 -voronoi diagrams\\
+     -ray casting, bullet casting ...}
+
+ \alert{10-ish citatov iz seznama ni klicanih v tekstu !}
+
+ \comment{}{
+ 	-fix citation styles (sometimes et al, sometimes not, sometimes names, sometimes years ... not same format)
+ }
 

### 2024-02-28 Articles/Article1-Conference/root.tex
Revert "removed conference paper draft (was merged)"
This reverts commit 4e5f44f496969dd6f231785ab53d448d52aa4270.

--- 2024-01-29 Articles/Article1-Conference/root.tex
@@
 
-\subsection{Motion Planning}
-
-% PREDSTAVITEV REDUNDANCE
-
-Kinematic redundancy~\cite{siciliano1990kinematic, siciliano2016springer} enables a manipulator to follow a predefined task space trajectory using the endeffector (EE), while simultaneously, optimising for an additional task with the remaining movement capacity without impacting the trajectory adherence.
-This is possible because the robot's degrees of freedom (DOF) go beyond what is required to perform the primary task.
-Consequently, the robot can adopt different joint configurations optimised according to the secondary task while performing the primary task.
-Common secondary tasks~\cite{siciliano2010robot} include avoiding singularities, optimising the manipulability measure, minimising joint torques and avoiding obstacles in the operating space.
-
-% PREDSTAVITEV PRISTOPOV PLANIRANJA POTI V ROBOTIKI
-The task of finding the joint trajectories of a manipulator is called motion planning~\cite{IDEASLab2023}.
-It consists of finding a sequence of joint configurations for a robot so that the robot can move along this path from its initial configuration to the goal configuration without colliding with itself, static obstacles or other agents in the environment.
-In addition to collision avoidance, motion planning for manipulators can optionally take into account various constraints, such as position, velocity, acceleration or jerk constraints for the joint angle or end effector, precision of the end effector with respect to position and orientation, stability of the manipulator, avoidance of singularities, or any number of other criteria.
-
-There are numerous methods for planning manipulator movements~\cite{c51, c52}, they can rougly be seperated into global and local approaches.
-Global, sampling-based, methods such as PRM~\cite{vsvestka1997motion}, RRT*~\cite{lavalle1998rapidly, karaman2010incremental}, RRT-Connect~\cite{kuffner2000rrt}, Informerd RRT*~\cite{gammell2014informed}, BIT*~\cite{gammell2015batch} and others offer a globally optimal solution based on a global search in configuration space.
-However, the generated trajectories are not always smooth or optimal, and the performance of the methods may be insufficient for real-time operation.
-
-%Global planning methods are typically more comprehensive, aiming to find a globally optimal trajectory for the manipulator.
-These methods often involve mapping out the entire path before movement begins, taking into account the known environment, which usually includes static obstacles.
-However, the downside of global planning methods is their computational intensity, which often renders them too slow for real-time applications.
-This limitation is particularly pronounced in dynamic or unpredictable environments where conditions change rapidly and require quick responses.
-Examples of global planning methods include graph-based algorithms like A* and Dijkstra's algorithm, which are known for their effectiveness in pathfinding but also for their computational demands [(Lavalle, 2006)].
-
-%Sampling-based methods such as PRM, RRT, RRT*, RRT-Connect, Informerd RRT*, BIT* and others offer a solution based on a global search in configuration space.
-However, the generated trajectories are not always smooth or optimal, and the performance of the methods may be insufficient for real-time operation.
-
-\remove{Recently, a number of learning-based methods using data-driven techniques have been proposed to improve or accelerate the functionality of sampling-based methods. }
-
-\remove{Trajectory optimisation methods such as CHOMP, STOMP and TrajOpt, on the other hand, use optimization to improve an initial seed trajectory.
-Consequently, the optimality of the solution is highly dependent on this initial trajectory.
-Nevertheless, these methods are capable of generating smooth trajectories, and although they can be too computationally intensive for high DOF dynamic real-time environments, they are generally effective in finding constrained motion plans.
-}
-
-Local motion planning approaches employ optimization techniques, two common ones are inverse kinematics~\cite{c29,c38}, that is based on finding a least squares solution of the manipulator joint velocities, and quadratic programming (QP)~\cite{c21,c22,c23}.
-Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
-However, since they do not plan further than one step ahead, they tend to get stuck in local minima.
-Therefore, they are often combined with a higher-level planner, for global static environment based \replace{motion}{path} planning, while local optimisation takes dynamic environment changes into account.
-In the following text we will focus on inverse kinematics based approaches.
-
-\subsection{Kinematic Obstacle Avoidance}
-
-\remove{Different control schemes have been proposed.
-Acceleration-based control excels in precise handling of motion changes, velocity-based control offers consistent and smooth movement, while force and torque-based control provides direct control over joint forces for robust physical interactions.
-(CITATIONS in COMMENT BELLOW)}
-%\remove{Different control schemes have been proposed~\cite{c24}.
-Acceleration-based control excels in precise handling of motion changes, velocity-based control offers consistent and smooth movement, while force and torque-based control~\cite{c25, c26, c27, c28, c31} provides direct control over joint forces for robust physical interactions.}
-
-% ----
-
-%In our work we use velocity control strategy for obstacle avoidance using the manipulator.
-naša primarna naloga premik vrha robota - end efektorja (EE) v želeno pozo, to je točko in orientacijo v prostoru naloge.
-Naša sekundarna naloga se tako izvaja v preostalem prostoru gibanja sklepov redundance robota (DOF, DOR).
-???
-je predlagal razširitev jakobijeve matrike end-efektorja z dodatno matriko transformacije izogibne hitrosti iz kartzezičnega prostora v prostor sklepov.
-Če nalogi nista popolnoma neodvisni ena od druge, in njune sklepne hitrosti niso medsebojno ortagonalne, lahko to pripelje do medsebojnega oviranja izbajanja nalog.
-Zato je ???
-predlagal uporabo transformacijo hitrosti sekundarne naloge v  ničelnega prostora hitrosti primarne nalgoe, to je v prostor ortagonalnih hitrosti.
-Posledično izvajanje sekundarne naloge ne vpliva neposredno v izvajanje primarne naloge, gledano le v enem trenutku izvajanja.
-%
-%Različni razisklovalci so se lotili problema kinematičnega izogibanja na različne načine, kot je opisano v \cite{c30, c39}. 
-%
-%Colbaugh in Glass  1989 \cite{c31, c32} sta se problema lotila v dveh korakih.
-V prvem koraku sta izračunala trajektorijo vrha robota, v drugem pa sta uporabila optimizacijo za izboljšanje dinamičnega odziva robota med izvajanjem trajektorije in za druge naloge, kot so izogibanje ovir.
-%
-%Sciavicco and Siciliano \cite{c34, c35} and Egeland~\cite{c36} independetly introduced the concept ot task-space augmentation.
-It was later revisited by searji~\cite{c37}.
-The Solution is to stack the remaining rows in Jacobian with secondary task Jacobian into square matrix, that returns only one solution.
-%
-%Macijewski in Klein (Maciejewski & Klein, 1985) \cite{c29}, Nakamura in Hanafusa 1987 \cite{c38} predstavijo koncept prioritete nalog in ničelnega prostora za izogibanje, pri čemer se sekundarna naloga lahko giblje le s hitrostmi, ki ne vplivajo neposredno na gibanje primarne naloge.
-pri čemer sta izogibne sklepne hitrosti izračunala funkcijo najmanjše razdalje med točko na manipulatorju in točko izogibanja / objekta / ovire.
-%
-%Omenjeni pristopi se ne osredotočajo na okolje, tega pogosto prikažejo kot geometrijske primitive in nato razdaljo izračunajo kot razdaljo med primitivom in premicami, ki opisujejo manipulator. 
-%
-%Khatib \cite{c33} proposes the concept of Artificial Potential Field.
-Polje sestavljata odbojna komponenta, ki end effektor odbija od ovir, ki so predstavljene kot geometrijski primitivi in privlačna komponenta, ki manipulator privlači proti cilju.
-%
-%
-%
-
-Various researchers have adopted different approaches to kinematic avoidance, as detailed in~\cite{c30, c39}.
-
-Colbaugh and Glass (1989)~\cite{c31, c32} tackled this problem in a two-step process.
-Initially, they calculated the robot's end-effector trajectory.
-Subsequently, they used optimization to enhance the robot's dynamic response during trajectory execution and for tasks like obstacle avoidance.
-
-Sciavicco and Siciliano~\cite{c34, c35}, as well as Egeland~\cite{c36}, independently introduced the concept of task-space augmentation, later revisited by Searji~\cite{c37}.
-This method involves extending the primary task Jacobian with the secondary task Jacobian into a square matrix, yielding a singular solution.
-
-Maciejewski and Klein (1985)~\cite{c29}, and Nakamura and Hanafusa (1987)~\cite{c38}, presented the concept of task prioritization and a null space for avoidance.
-In this framework, the secondary task is restricted to velocities that do not directly affect the primary task's movement.
-Their method calculates avoidance joint velocities based on the minimum distance between a point on the manipulator and an avoidance point, object, or obstacle.
-Žlajpah~\cite{c41} improved the concept by proposing a reduced operational space formulation, that reshapes the Jacobain of the avoidance task from three cartesian axis to only the direction of obstacle avoidance.
-Petrič~\cite{c42} suggested a way to smoothly transition between avoidance and trajectory following tasks.
-
-\subsection{Artificial Potential Field}
-
-These approaches typically do not focus on the environmental context, often representing it as geometric primitives and then calculating the distance between these primitives and the lines describing the manipulator.
-
-Khatib~\cite{c33} proposes the concept of an Artificial Potential Field.
-This field consists of a repulsive component that deflects the end effector away from obstacles, depicted as geometric primitives, and an attractive component that draws the manipulator towards its target.
-In the following years many different modifications and improvement of the original APF idea have been proposed, often focused on removal of local minimas in the potential field.
-Khim and Khosla~\cite{c40} suggested the use of harmonic functions to solve the problem of local minima.
-Pinto et al.
-\cite{c43} proposes to vary the field based on the distance of robot from obstacles to fill the local minima.
-Many researchers tried using APF as a heuristic to better guide sampling based approaches~\cite{c45, c46, c47}.
-
-Many of the recent works focus on use of a variation of artificial potential field to plan motion of the manipulator.
-Xia et al.~\cite{c49} uses a variation of APF for manipulator motion~\cite{c49}.
-Zhang et al.~\cite{zhang2021obstacle} proposes dynamic repulsive field based on direction and speed between point on robot and obstacle, it also suggests decision making force that moves the robot away from certain local minima.
-Chen et al.
-proposes\remove{ an application }using APF in joint space and a variable kinematic optimization step~\cite{c50}.
-Long~\cite{c44} suggests creating motion plan of the manipulator using APF, he extends it using RRT to calculate virtual attractive point for the robot to move towards in case of local minima.
-Zhu et al.~\cite{c48} proposes use of APF in combination with MPC, to plan in environment with dynamic obstacles.
-
-\alert{
- -fix citation styles (sometimes et al, sometimes not, sometimes names, sometimes years ... not same format)
-}
-
-%The inverse kinematics with task prioritisation, which is based on finding a least squares solution of the manipulator joint velocities, and linear quadratic programming (LQR) are two of the local optimization methods.
-Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
-However, since they do not plan further than one step ahead, they tend to get stuck in local minima.
-Therefore, they are often combined with a higher-level planner, for global static environment based \replace{motion}{path} planning, while local optimisation takes dynamic environment changes into account.
-
-\subsection{Environment representation}
-
-% PREDSTAVITEV PRISTOPOV RAČUNANJA ODDALJENOSTI OVIRAM
-%To be able to generate collision free trajectory we need to have a representation of the static environment.
-One common approach is to enclose the obstacles around the manipulator into a collection of simple geometric primitives and then construct a tree-like structure to achieve the goal~\cite{dai2022review}.
-Some of the commonly used methods are AABB (axis-aligned bounding boxes)~\cite{vandenbergen1997efficient, chen2018path}, OBB (oriented bounding boxes)~\cite{gottschalk1996obbtree, puiu2011realtime}.
-
-To be able to generate collision free trajectory we need to have a representation of the static environment.
-One common approach involves enclosing the obstacles around the manipulator within a collection of simple geometric primitives and then constructing a tree-like structure to facilitate efficient navigation and path planning~\cite{dai2022review}.
-Two of the commonly used methods for this purpose include AABB (axis-aligned bounding boxes)~\cite{vandenbergen1997efficient, chen2018path, luo2018collisionfree} and OBB (oriented bounding boxes)~\cite{gottschalk1996obbtree, puiu2011realtime}\remove{and FDH (fixed directions hulls).}.
-Another significant method worth mentioning is the use of Octomaps~\cite{wurmOctoMap}.
-Octomaps employ an octree data structure to represent 3D environments efficiently, making them particularly suitable for areas with large open spaces.
-
-In real environment we usually collect depth data with LiDAR, radar of RGBD camera sensors, that usually return clouds of points, that tell us the distance of objects in robots environment.
-While approaches have been proposed to use the point-cloud data directly to find obstacle free areas for robot operation~\cite{gao2019flying}, the data is often further converted into voxel grids~\cite{xu2021voxel, oleynikova2017voxblox, han2019fiesta}.
-
-Voxel grid~\cite{xu2021voxel, elfes1989using} representation divides the space into a regular grid of volumetric elements, or voxels, which can be used to create a more manageable approximation of the environment.
-While this approach offers a balance between detail and computational efficiency, it can introduce discretization errors, particularly when modeling objects with smooth surfaces or intricate details.
-The fidelity of the representation is dependent on the size of the voxels: smaller voxels can capture more detail but require more memory and computation, while larger voxels result in coarser approximations but are more memory and computation efficient.
-Adaptive voxel grids have been explored~\cite{xu2021voxel}, where the voxel size can vary throughout the space to provide higher resolution in regions of interest while conserving resources in less critical areas.
-Nießner et al.~\cite{niessner2013realtime} proposed Voxel hashing, for more efficient memory management in sparse environments.
-
-Voxel grids can incorporate probabilistic information~\cite{thrun2002probabilistic, dryanovski2010multivolume}, such as in occupancy grid maps, where each voxel holds a probability indicating the chance of an obstacle's presence.
-The occupancy probability of a voxel can be updated dynamically using sensor measurements and Bayesian updating methods.
-As new sensor data is collected, the probabilities are revised to reflect the increased or decreased likelihood of the presence of an obstacle in the voxel space.
-
-Another type of data that voxel fields can hold are ESDF (Euclidian Signed Distance Field) or TSDF (Truncated Signed Distance Field)~\cite{oleynikova2017voxblox}, in which case each of the voxels contains information how far nearest obstacle is in its vicinity.
-
-\subsection{ESDF creation}
-
-ESDF grids can be generated directly using sensor measurements.
-Oleynikova et al.~\cite{oleynikova2017voxblox} proposed a method for calculating the ESDF from TSDF.
-Han et al.~\cite{han2019fiesta} proposed a way to integrate point-cloud data into ESDF using ray-casting.
-
-Another way is to generate ESDF from occupancy grids, that can be previously generated using sensors.
-When generating ESDF from occupancy grid, common approach is the Brushfire method~\cite{lau2010improved}, that spreads from obstacles until it calculates the distance for every field on a grid.
-Jump Flooding Algorithm (JFA) is a similar method, that can be implemented on a GPU for faster parallel distance calculation.
-
-\add{\\-Wavefront \\
-	 -D* Klančar\\
- 	 -Euclidian distnace algorithms\\
- 	 -distance transforms\\
-  	 -voronoi diagrams\\
-     -ray casting, bullet casting ...}
-
- \alert{10-ish citatov iz seznama ni klicanih v tekstu !}
-
 \clearpage
@@
 
+\bibitem{park2020trajectory} S.-O.
+Park, M.
+C.
+Lee, and J.
+Kim, “Trajectory Planning with Collision Avoidance for Redundant Robots Using Jacobian and Artificial Potential Field-based Real-time Inverse Kinematics,” Int.
+J.
+Control Autom.
+Syst., vol.
+18, no.
+8, Art.
+no.
+8, Aug.
+2020, doi: 10.1007/s12555-019-0076-7.
+
+\bibitem{han2018dynamic} D.
+Han, H.
+Nie, J.
+Chen, and M.
+Chen, “Dynamic obstacle avoidance for manipulators using distance calculation and discrete detection,” Robotics and Computer-Integrated Manufacturing, vol.
+49, pp.
+98–104, Feb.
+2018, doi: 10.1016/j.rcim.2017.05.013.
+
+\bibitem{rong2006jump} G.
+Rong and T.-S.
+Tan, “Jump flooding in GPU with applications to Voronoi diagram and distance transform,” in Proceedings of the 2006 Symposium on Interactive 3D Graphics and Games - SI3D ’06, Redwood City, California, 2006, p.
+109.
+doi: 10.1145/1111411.1111431.
+
+\bibitem{szeliski2021computer} R. Szeliski, “Computer Vision: Algorithms and Applications, 2nd Edition,” Springer, 2021.
+
+\bibitem{rosenfeld1968distance} A. Rosenfeld and J. L. Pfaltz, “Distance functions on digital pictures,” Pattern Recognition, vol. 1, no. 1, pp. 33–61, Jul. 1968, doi: 10.1016/0031-3203(68)90013-7.
+
+\bibitem{felzenszwalb2012distance} P.
+F.
+Felzenszwalb and D.
+P.
+Huttenlocher, “Distance Transforms of Sampled Functions,” Theory of Comput., vol.
+8, no.
+1, pp.
+415–428, 2012, doi: 10.4086/toc.2012.v008a019.
+
+\bibitem{zhou2021egoplanner} X.
+Zhou, Z.
+Wang, H.
+Ye, C.
+Xu, and F.
+Gao, “EGO-Planner: An ESDF-Free Gradient-Based Local Planner for Quadrotors,” IEEE Robot.
+Autom.
+Lett., vol.
+6, no.
+2, pp.
+478–485, Apr.
+2021, doi: 10.1109/LRA.2020.3047728.
+
+\bibitem{dryanovski2010multivolume} I.
+Dryanovski, W.
+Morris, and J.
+Xiao, “Multi-volume occupancy grids: An efficient probabilistic 3D mapping model for micro aerial vehicles,” in 2010 IEEE/RSJ International Conference on Intelligent Robots and Systems, Taipei: IEEE, Oct.
+2010, pp.
+1553–1559, doi: 10.1109/IROS.2010.5652494.
+
 \end{thebibliography}

### 2024-02-29 Articles/IROS/root.tex
titles, started IROS article

--- 2024-02-28 Articles/Article1-Conference/root.tex
@@
 \title{\LARGE \bf
-Preparation of Papers for IEEE Sponsored Conferences \& Symposia*
+Matrix Multiplication-Driven Repulsive Fields for 3D Voxel-Based Robotic Manipulator Path Planning
 }
 
-\author{Albert Author$^{1}$ and Bernard D. Researcher$^{2}$% <-this % stops a space
-\thanks{*This work was not supported by any organization}% <-this % stops a space
-\thanks{$^{1}$Albert Author is with Faculty of Electrical Engineering, Mathematics and Computer Science,
-        University of Twente, 7500 AE Enschede, The Netherlands
-        {\tt\small albert.author@papercept.net}}%
-\thanks{$^{2}$Bernard D. Researcheris with the Department of Electrical Engineering, Wright State University,
-        Dayton, OH 45435, USA
-        {\tt\small b.d.researcher@ieee.org}}%
+\author{
+	Jakob Baumgartner$^{1}$ and Gregor Klančar$^{2}$% <-this % stops a space
+%\thanks{*This work was not supported by any organization}% <-this % stops a space
+%\thanks{$^{1}$Albert Author is with Faculty of Electrical Engineering, Mathematics and Computer Science,
+%        University of Twente, 7500 AE Enschede, The Netherlands
+%        {\tt\small albert.author@papercept.net}}%
+%\thanks{$^{2}$Bernard D. Researcheris with the Department of Electrical Engineering, Wright State University,
+%        Dayton, OH 45435, USA
+%        {\tt\small b.d.researcher@ieee.org}}%
 }
@@
 
-This electronic document is a Olive template.
-The various components of your paper [title, text, heads, etc.] are already defined on the style sheet, as illustrated by the portions given in this document.
-
 \end{abstract}
@@
 
-\clearpage
+\section{RELATED WORKS}
+
+\section{REPULSIVE FIELD CALCULATION}
+
+We want the repulsive field to push in any point of the space away from the obstacles / towards the direction away from all surrounding obstacles.
+
+The operating environment is modelled by discrete voxels.
+As the robots environment can dynamically change, we propose a method that looks at the surrounding space of the robot and calculates these direction away from all the surrounding obstacles in real time.
+We only look in a predefined area / perimeter arround the robot.
+
+\subsection{AREA SELECTION}
+
+\subsection{CONVOLUTIONAL KERNELS}
+
+\subsection{3D INTERPOLATION}
+
+\section{INVERSE KINEMATICS}
+
+\section{REPULSIVE VELOCITIES CALCULATION}
+
+\section{SIMULATION RESULTS}
+
+\section{CONCLUSION}
 

### 2024-02-29 Articles/IROS/root.tex
rought draft or rep APF intro

@@
 
-We want the repulsive field to push in any point of the space away from the obstacles / towards the direction away from all surrounding obstacles.
-
 The operating environment is modelled by discrete voxels.
@@
 We only look in a predefined area / perimeter arround the robot.
+
+In 3D computer graphics, a voxel represents a value on a regular grid in three-dimensional space.
+Each of the voxels holds the probability value of its occupation.
+In case of voxel being empty it holds the value of 0, if the voxel is occupied it holds the value of non-zero, depending on our assurance of it being occupied.
+If it is definitly occupied it holds the value of 1.
+
+Mreža voxels je lahko predefinirana, glede na model / 3d zemljevid prostora.
+Zasedenost voxlov lahko spreminjamo glede na poznane pozicije in trajektorije preostalih agentov v prostoru.
+Kot omenjeno v uvodu, pa lahko zasedenost voxlov pridobimo tudi z senzorskimi sistemi.
+
+In our method, we compute repulsive velocities within the task space using a novel matrix kernel multiplication approach.
+Concentrating on the task space is advantageous as it provides a more direct and realistic representation of the environment.
+
+Naša metoda je posebno primerna za uporabo z senzorskimi sistemi kot so LIDAR ali globinske kamere, saj zaradi upoštevanja celotne okolice točke in ne le razdalje do najbližje točke v okolici efektivno filtriramo senzorski šum.
+
+Repulsive velocities tell the agent in which direction to move, so that it avoids nearby obstacles.
+These velocities drop to zero when the agent maintains a minimum safe distance from obstacles, and rise to their highest when it nears an obstacle, facilitating immediate evasive action.
+As the repulsive field calculation is locally based, it will also go to zero when the agent is surrounded by all directions, equally spaced from all sides.
+That is, it is in the best local minima away from all the obstacles.
+
+Since the obstacle space is discrete (has finite resolution), while the Cartesian space is continuous, we propose two methods for mapping from Cartesian space to the occupancy grid space.
+The simpler approach involves mapping the point directly to the center of the nearest occupancy grid voxel, based on Euclidean distance.
+However, this discretization can sometimes lead to discontinuities.
+Therefore, we propose a second approach: tri-linear interpolation of the calculated repulsive field to achieve a continuous repulsive field value.
 

### 2024-02-29 Articles/IROS/root.tex
rough draft of convolution and interpolation

@@
 %\usepackage{times} % assumes new font selection scheme installed
-%\usepackage{amsmath} % assumes amsmath package installed
-%\usepackage{amssymb}  % assumes amsmath package installed
+\usepackage{amsmath} % assumes amsmath package installed
+\usepackage{amssymb}  % assumes amsmath package installed
 \usepackage{hyperref}
@@
 
+\subsection{MAPPING}
+
 Since the obstacle space is discrete (has finite resolution), while the Cartesian space is continuous, we propose two methods for mapping from Cartesian space to the occupancy grid space.
-The simpler approach involves mapping the point directly to the center of the nearest occupancy grid voxel, based on Euclidean distance.
-However, this discretization can sometimes lead to discontinuities.
+The simpler approach involves mapping the point directly to the center of the nearest occupancy grid voxel, based on Euclidean distance.\add{is this clear, do i need equation?} However, this discretization can sometimes lead to discontinuities.
 Therefore, we propose a second approach: tri-linear interpolation of the calculated repulsive field to achieve a continuous repulsive field value.
 
-\subsection{AREA SELECTION}
+Once the mapping of these points to the voxel grid is completed, we proceed to employ a specialized kernel convolution method.
+This method is tailored to calculate the components of the avoidance velocity vector in the Cartesian space, we generate the corresponding kernel and extract a segment of the obstacle grid of matching size, centered at agent or the point of interest (POI), resulting in two same size 3D matrices—one is a "window" from the obstacle grid $A_d$ and the other representing the kernel $K_d$.
+
+If the agent is located near the edge of our known voxel grid we can set the eleemnts of the window would be located in the space beyond the matrix as empty in which case the robot might want to move towards this space or as occupied, which will prevent the robot from moving out of the known grid space.
+
+By employing the Hadamard (element-wise) product (eq.~\ref{eq:hadamard}) between the cutout segment of the obstacle grid \(A_d\) and the corresponding 3D convolutional kernel \(K_d\) for direction \(d\), we derive the resultant matrix \(C_d\), which can be expressed as:
+
+\begin{equation}
+	C_d = A_d \odot K_d
+	\label{eq:hadamard}
+\end{equation}
+
+Subsequently, the avoidance velocity in the Cartesian coordinate system is obtained (eq.~\ref{eq:summation}) by summing all the values in the resulting matrix $C_d$, which can be represented as:
+
+\begin{equation}
+	\dot{x}_{poi} = \sum_{i}\sum_{j}\sum_{k} c_{dijk} 
+	\label{eq:summation}
+\end{equation}
 
@@
 
+The fundamental concept of our directional kernels lies in computing the repulsive field individually for each direction within the Cartesian coordinate system.
+\remove{Our filters structure was inspired by the Sobel  operator, a 2D convolutional filter frequently utilized in computer vision for calculating image gradients at specific points.}
+
+Our kernels are designed as three-dimensional matrixes with a primary kernel axis aligned along a specific Cartesian direction, corresponding to the calculated repulsive velocity.
+The two secondary kernel axes are orthogonal to this primary axis.
+The distribution of values along the primary axis is inversely symmetric, exhibiting positive values on one side and negative values on the other, with the zero valued cell in the center of the kernel, where jump between max positive and max negative magnitude is.
+The function of the increase in magnitude along the primary axis of the kernel defines the shape of the repulsive velocity field, determining how the repulsive velocity changes as the agent approaches an obstacle.
+
+The length of the primary axis is critical, as it dictates the detection range for obstacles. Longer kernels can detect obstacles further away
+from the robot, essentially extending the ’safety zone’ around the robot.
+If the primary axis is too long, it can lead to extra calculations and may cause the robot to unnecessarily avoid obstacles that aren't in its immediate path, making its movement and path planning less efficient.
+A kernel with a primary axis that is too short might restrict the robot's ability to maneuver, detecting obstacles potentially too late, compromising the robots capacity to avoid obstacles effectively (eq.
+\ref{eq: detect range}).
+Moreover, it is essential for the magnitudes at the kernel's periphery to be minimal, promoting a smooth increase in repulsive velocity when approaching the obstacle rather than a sudden spike.
+
+\begin{equation}
+	\label{eq: detect range}
+	num_{primary} = \frac{2 \times \mathrm{range}}{\Delta \mathrm{grid}}\\
+\end{equation}
+
+%\remove{In the orthogonal axes, the magnitude of values diminishes with increasing distance from the core of the kernel, nearing zero at the edges.
+This gradation is vital for a smooth velocity transition when obstacles emerge in the kernels range.
+The peak magnitude is located at the kernel's core, along the primary axis.
+The polarity of the values is determined by their position relative to the primary axis.
+}
+
+The length of the orthogonal axes influences the peripheral detection range for obstacles.
+Excessively wide kernels may generate repulsive velocities for objects that are not in the path of the robot, whereas too narrow kernels might only detect obstacles aligned directly with the Cartesian direction in the point of interest.
+When selecting the width and height of the kernel, we must consider the density of the neighboring points of interest on the robot, ensuring that the \replace{collective fields}{combination of kernels} adequately cover the entire manipulator's surrounding area.
+
+Če imamo opravka s točkastim agentom, kot je recimo štirikopet - dron, potem je dobra oblika posameznih matrik enaka dimenzija dolžine, širine in višine.
+Tako dobimo enakomerno pokritost krogle prostora v okolici agenta in preprečimo mrtve kote, ki se sicer lahko pojavijo, ko se ovira nahaja v bližini agenta, vendar med posameznimi jedri.
+\add{ADD:primer,slika}
+
+\comment{}{PLOT: kernel 2D images}
+\comment{}{EQUATION: kernel generation / values equations}
+
+\add{choice of velocity profile (gaussian, linear ... maybe could move this part to implementation / experiments)}
+
 \subsection{3D INTERPOLATION}
+
+It is essential that the velocity contributions affecting the robot change smoothly.
+However, since our obstacle grid is discretely defined, achieving perfect continuity can be challenging.
+Increasing the resolution of the obstacle field can theoretically bring us closer to continuous behavior, but in practice, we are constrained by finite resolution.
+To ensure that the velocity remains continuous when transitioning from one cell of the obstacle grid to another at a point of interest (POI), we employ trilinear interpolation.
+\remove{This technique allows for a smooth and continuous linear approximation of velocities in all three Cartesian directions (x, y, and z) as the POI moves between cells.}
+
+We start by scalling the coordinates of POI into the grid koordinate system, by multiplying it by grid resolution (eq.~\ref{eq:scale grid}).
+
+\begin{equation}
+	\label{eq:scale grid}
+	\vec{P} = \vec{p}_{\mathrm{POI}} \times\Delta \mathrm{grid}
+\end{equation}
+
+We get the indexes of the surrounding cells by first scaling the POI position by grid resolution and than rounding the position to the nearest lower and upper integer positions (eq.~\ref{eq:floor and ceil}).
+
+%\begin{equation}
+% X = \left[ \lfloor p_{x} \rfloor, \lceil p_{x} \rceil \right] 
+%\end{equation}
+%
+%\begin{equation}
+% Y = \left[ \lfloor p_{y} \rfloor, \lceil p_{y} \rceil \right] 
+%\end{equation}
+%
+% \begin{equation}
+% Z = \left[ \lfloor p_{z} \rfloor, \lceil p_{z} \rceil \right] 
+%\end{equation}
+
+\begin{equation}
+	\label{eq:floor and ceil}
+	\vec{P} =
+	\begin{bmatrix}
+		X \\
+		Y \\
+		Z \\
+	\end{bmatrix}
+	=
+	\begin{bmatrix}
+		\; \lfloor \vec{p}_{\mathrm{POI}}(1) \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(1) \rceil \;  \\
+		\; \lfloor \vec{p}_{\mathrm{POI}}(2) \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(2) \rceil \; \\
+		\; \lfloor \vec{p}_{\mathrm{POI}}(3) \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(3) \rceil \; 
+	\end{bmatrix}
+\end{equation}
+
+Once we got the indexes of the eight surrounding cells of our POI, we use our kernel matrix multiplication method, to calculate the 3x1 repulsive velocity vectors for all the cells (eq.~\ref{eq: calc rep vel}).
+
+\begin{equation}
+	\label{eq: calc rep vel}
+	\vec{V}rep_{xyz,ijk} = \mathrm{calc\_rep\_vel}(X[i], Y[j], Z[k]) \quad \forall i, j, k \in \{1, 2\}
+\end{equation}
+
+Trilinear interpolation method works on a 3-dimensional regular grid.
+Before we can start with the interpolation we need to calculate the distance between POI and smaller coordinates of the cells where we calculated the repulsive velocities (eq.~\ref{eq: deltas interp}).
+\replace{Since the repulsive values we calculate for the cells are alligned with the centers of the cells, we need to move before the interpolation the positions of known grid points by half of the cell width.}{The calculated repulsive velocity values are located at the centers of the cells.
+Therefore, before interpolation, we shift the values of the cells coordinates by half the resolution of the obstacle grid for each direction.
+}
+
+%\begin{equation}
+%	\Delta X = \frac{\left( P_x - \left( X(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( X(2) - X(1) \right)}
+%\end{equation}
+%
+%\begin{equation}
+%	\Delta Y = \frac{\left( P_y - \left( Y(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Y(2) - Y(1) \right)}
+%\end{equation}
+%
+%\begin{equation}
+%	\Delta Z = \frac{\left( P_z - \left( Z(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Z(2) - Z(1) \right)}
+%\end{equation}
+
+\begin{equation}
+	\label{eq: deltas interp}
+	\Delta \vec{P} =
+	\begin{bmatrix}
+		\Delta x \\
+		\Delta y \\
+		\Delta z		
+	\end{bmatrix}
+	=
+	\begin{bmatrix}
+		\frac{\left( P_x - \left( X(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( X(2) - X(1) \right)} \\
+		\frac{\left( P_y - \left( Y(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Y(2) - Y(1) \right)} \\
+		\frac{\left( P_z - \left( Z(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Z(2) - Z(1) \right)} \\
+	\end{bmatrix}
+\end{equation}
+
+The result of the interpolation is independent of the order of the operations. We first interpolate along the x-axis, followed by along the y-axis and finally along z-axis.
+
+%\begin{equation}
+%	\vec{V}_{rep,00} = \vec{V}_{rep,000}(1 - x_d) + \vec{V}_{rep,100}x_d
+%\end{equation}
+%\begin{equation}
+%	\vec{V}_{rep,01} = \vec{V}_{rep,001}(1 - x_d) + \vec{V}_{rep,101}x_d
+%\end{equation}
+%\begin{equation}
+%	\vec{V}_{rep,10} = \vec{V}_{rep,010}(1 - x_d) + \vec{V}_{rep,110}x_d
+%\end{equation}
+%\begin{equation}
+%	\vec{V}_{rep,11} = \vec{V}_{rep,011}(1 - x_d) + \vec{V}_{rep,111}x_d
+%\end{equation}
+
+\begin{equation}
+	\label{eq: interp x}
+	\vec{V}rep_{xyz,jk} = \vec{V}rep_{xyz,0jk}(1 - \Delta x) + \vec{V}rep_{xyz,1jk} \, \Delta x \quad \forall \, j, k \in \{1, 2\}
+\end{equation}
+
+%\begin{equation}
+%	\vec{V}rep_{xyz,0} = \vec{V}rep_{xyz,00}(1 - \Delta y) + \vec{V}rep_{xyz,10} \, \Delta y
+%\end{equation}
+%
+%\begin{equation}
+%	\vec{V}rep_{xyz,1} = \vec{V}rep_{xyz,01}(1 - \Delta y) + \vec{V}rep_{xyz,11} \, \Delta y
+%\end{equation}
+
+\begin{equation}
+	\label{eq: interp y}
+	\vec{V}rep_{xyz,k} = \vec{V}rep_{xyz,0k}(1 - \Delta y) + \vec{V}rep_{xyz,1k} \, \Delta y \quad \forall \, k \in \{1, 2\}
+\end{equation}
+
+\begin{equation}
+	\label{eq: interp z}
+	\vec{V}rep_{xyz} = \vec{V}rep_{xyz,0}(1 - \Delta z) + \vec{V}rep_{xyz,1} \, \Delta z 
+\end{equation}
+
+The final result is a repulsive velocity vector that transitions smoothly between the discrete values calculated at distinct points in the obstacle grid.
 

### 2024-03-01 Code/directional_kernel_3d.m
removed zero cross bug
zero cross bug; there should only be one zero surface in the middle of directional kernel, not for other directions

@@ -15,9 +15,9 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         
         % if sigma_width is 0, then use a constant weight for the width
         if (sigma_width == 0)
-            width_extender = [ones(1,center_width) 0 ones(1,center_width)];
+            width_extender = [ones(1,center_width) center_width ones(1,center_width)];
         else
-            width_extender = [1:1:center_width 0 flip(1:1:center_width)]/(center_width);
+            width_extender = [1:1:center_width center_width flip(1:1:center_width)]/(center_width);
         end
 
 
@@ -29,9 +29,9 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
 
         % if sigma_height is 0, then use a constant weight for the height
         if (sigma_height == 0)
-            height_extender = [ones(1,center_width) 0 ones(1,center_width)];
+            height_extender = [ones(1,center_height) center_height ones(1,center_height)];
         else
-            height_extender = [1:1:center_height 0 flip(1:1:center_height)]/(center_height);
+            height_extender = [1:1:center_height center_height flip(1:1:center_height)]/(center_height);
         end
 
         
@@ -52,10 +52,10 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
 
         % if sigma_width is 0, then use a constant weight for the width
         if (sigma_width == 0)
-            width_extender = [ones(1,center_width) 0 ones(1,center_width)];
+            width_extender = [ones(1,center_width) center_width ones(1,center_width)];
         else
             width_extender = exp(-(1:center_width).^2/(2*sigma_width^2)); 
-            width_extender = [flip(width_extender) 0 width_extender];
+            width_extender = [flip(width_extender) center_width width_extender];
         end
         
         % generate 2D kernel
@@ -66,10 +66,10 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
 
         % if sigma_height is 0, then use a constant weight for the height
         if (sigma_height == 0)
-            height_extender = [ones(1,center_height) 0 ones(1,center_height)];
+            height_extender = [ones(1,center_height) center_height ones(1,center_height)];
         else
             height_extender = exp(-(1:center_height).^2/(2*sigma_height^2));
-            height_extender = [flip(height_extender) 0 height_extender];
+            height_extender = [flip(height_extender) center_height height_extender];
         end
 
 

### 2024-03-04 Articles/IROS/root.tex
text reordering
draft primary axis weights distribution description

@@
 
-\subsection{CONVOLUTIONAL KERNELS}
+\comment{}{TODO: add the indexes for x,y,z directions}
+
+\subsection{KERNELS SELECTION}
 
@@
 The function of the increase in magnitude along the primary axis of the kernel defines the shape of the repulsive velocity field, determining how the repulsive velocity changes as the agent approaches an obstacle.
+Moreover, it is essential for the magnitudes at the kernel's periphery to be minimal, promoting a smooth increase in repulsive velocity when approaching the obstacle rather than a sudden spike.
+
+% KERNEL WEIGHTS
+
+% PRIMARY AXIS
+
+We propose two different primary axis weights distributions, of course there is no reason why any other distribution of weights could not be used.
+The choice of the weights distribution should depend on the profile of the repulsive velocities we want to archieve for the APF.
+
+The first of the proposed functions is a mirrored normal / gaussian distribution. By changing the sigma we can control how fast or slow does the field value grow when we approach obstacles.
+
+\begin{equation}
+w_i = 
+\begin{cases} 
+	e^{-\frac{\Delta i^2}{2\sigma^2}} / (\sigma \sqrt{2\pi}) & \text{if } \Delta i > 0 \\
+	0 & \text{if } \Delta i = 0 \\
+	-e^{-\frac{\Delta i^2}{2\sigma^2}} / (\sigma \sqrt{2\pi}) & \text{if } \Delta i < 0 
+\end{cases}
+\end{equation}
+
+%\begin{equation}
+%w_i = 
+%\begin{cases} 
+%	\frac{e^{-\frac{\Delta i^2}{2\sigma^2}}}{\sigma \sqrt{2\pi}} & \text{if } \Delta i > 0 \\
+%	0 & \text{if } \Delta i = 0 \\
+%	-\frac{e^{-\frac{\Delta i^2}{2\sigma^2}}}{\sigma \sqrt{2\pi}} & \text{if } \Delta i < 0 
+%\end{cases}
+%\end{equation}
+
+Another distribution we used is mirrored linear, where the $l=\lceil width / 2 \rceil$ is the rounded up half length of the primary axis kernel.
+
+\begin{equation}
+	w_i = 
+	\begin{cases} 
+	 	\frac{l - \Delta i}{l} & \text{if } \Delta i > 0 \\
+		0 & \text{if } \Delta i = 0 \\
+		\frac{l + \Delta i}{l}& \text{if } \Delta i < 0 
+	\end{cases}
+\end{equation}
+
+% ORTHAGONAL AXIS
 
@@
 \ref{eq: detect range}).
-Moreover, it is essential for the magnitudes at the kernel's periphery to be minimal, promoting a smooth increase in repulsive velocity when approaching the obstacle rather than a sudden spike.
 

### 2024-03-04 Code/Experiments APF/MANIP_EXPERIMENTS.m
kernel fixes

@@ -5,7 +5,7 @@ close all
 % ------------------------------------------------------------
 
 % select map
-map_selection = 'column'
+map_selection = 'roof'
 
 [grid] = MAPS(map_selection);
 

### 2024-03-04 Code/Experiments APF/REP_EXPERIMENTS.m

@@ -4,7 +4,7 @@ close all
 point = [0.7,0.7,1]
 
 %% SELECT MAP
-map_selection = 'cross'
+map_selection = 'wall'
 % -----------------------------------------------------------
 
 [grid] = MAPS(map_selection)

### 2024-03-04 Code/Full_RTConvolution_Multiple_Points.m

@@ -3,7 +3,7 @@ function [output] = Full_RTConvolution_Multiple_Points(grid, goal_point, Tbase, 
 %% PARAMETERS
 
 % select which goals
-mid_joints = 0
+mid_joints = 1
 avoid_task = 1
 kinematics_solution = 'exact-reduced' % OPTIONS: exact-reduced , exact , approximate
 timestep_secondary_gain_change = 0 % if selected, secondary task will start with normal gain and fall with time
@@ -21,15 +21,15 @@ weights_avoidance = weights_avoidance / norm(weights_avoidance,1) / 10;
 % -----------------------------------------------------------
 
 Tstep = 0.1 % time step
-Nmax = 250 % max number of iterations
+Nmax = 500 % max number of iterations
 space_resolution = grid.resolution; % resolution of the obstacles grid
 
 % weights for different tasks
 wp = 0.75 % primary task
 wp_att = 1 % primary task - attractive component
 wp_rep = 0 % primary task - repulsive component
-wm = 0.15 % mid-joints task
-wa = 1 % obstacle avoidance task
+wm = 0.1 % mid-joints task
+wa = 0.1 % obstacle avoidance task
 
 
 
@@ -213,8 +213,7 @@ while current_dist > goal_dist && Niter <= Nmax
 
         % SORT POI BASED ON DISTANCE FROM OBSTACLES
         % --------------------
-        [poi_sizes, poi_indeces] = sort(vecnorm(rep_values));
-        poi_sizes            
+        [poi_sizes, poi_indeces] = sort(vecnorm(rep_values));          
 
 
         %% FOR EVERY POI CALCULATE AVOIDANCE JOINT SPEEDS (USING INVERSE KINEMATICS SOLUTION)
@@ -319,14 +318,13 @@ while current_dist > goal_dist && Niter <= Nmax
     % PRIMARY: position
 
     % exec slowdown - if secondary task has big velocities, primary task will slow down to give more time to collisions avoidance
-
     if poi_sizes(end) > min_exec_slowdown_size % if poi is too close
-        exec_slowdown = 1 / (1 + secondary_exec_stop_k * poi_sizes(end))
+        exec_slowdown = 1 / (1 + secondary_exec_stop_k * poi_sizes(end));
     else
         exec_slowdown = 1; % or set to the default value if no slowdown is needed
     end
     
-    q_vel_position = pinv_J * exec_slowdown * ee_vel
[…]

### 2024-03-04 Code/REP_kernels.m

@@ -53,9 +53,9 @@ function [kernels] = REP_kernels(varargin)
     kernel_height = p.Results.kernel_height;
 
     % form kernels
-    kernel_x = directional_kernel_3d('x', kernel_length, sigma, kernel_width, 2, kernel_height, 2, type);
-    kernel_y = directional_kernel_3d('y', kernel_length, sigma, kernel_width, 2, kernel_height, 2, type);
-    kernel_z = directional_kernel_3d('z', kernel_length, sigma, kernel_width, 2, kernel_height, 2, type);
+    kernel_x = directional_kernel_3d('x', kernel_length, sigma, kernel_width, 0, kernel_height, 0, type);
+    kernel_y = directional_kernel_3d('y', kernel_length, sigma, kernel_width, 0, kernel_height, 0, type);
+    kernel_z = directional_kernel_3d('z', kernel_length, sigma, kernel_width, 0, kernel_height, 0, type);
 
     kernels = {kernel_x/max(kernel_x(:)), kernel_y/max(kernel_y(:)), kernel_z/max(kernel_z(:))};
 

### 2024-03-04 Code/directional_kernel_3d.m

@@ -11,11 +11,11 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         kernel = [1:1:center 0 -flip(1:1:center)] / (center); 
 
         % extend weights to kernel width
-        center_width = floor(kernel_width/2);
+        center_width = ceil(kernel_width/2);
         
         % if sigma_width is 0, then use a constant weight for the width
         if (sigma_width == 0)
-            width_extender = [ones(1,center_width) center_width ones(1,center_width)];
+            width_extender = [ones(1,center_width) 1 ones(1,center_width)];
         else
             width_extender = [1:1:center_width center_width flip(1:1:center_width)]/(center_width);
         end
@@ -25,11 +25,11 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         kernel2D = width_extender' * kernel;
 
         % extend weight to kernel height
-        center_height = floor(kernel_height/2);
+        center_height = ceil(kernel_height/2);
 
         % if sigma_height is 0, then use a constant weight for the height
         if (sigma_height == 0)
-            height_extender = [ones(1,center_height) center_height ones(1,center_height)];
+            height_extender = [ones(1,center_height) 1 ones(1,center_height)];
         else
             height_extender = [1:1:center_height center_height flip(1:1:center_height)]/(center_height);
         end
@@ -48,11 +48,11 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         kernel = [flip(kernel) 0 -kernel];
         
         % prepare gaussian weights for width
-        center_width = floor(kernel_width/2);
+        center_width = ceil(kernel_width/2);
 
         % if sigma_width is 0, then use a constant weight for the width
         if (sigma_width == 0)
-            width_extender = [ones(1,center_width) center_width ones(1,center_width)];
+            width_extender = [ones(1,center_width) 1 ones(1,center_width)];
         else
             width_extender = exp(-(1:center_width).^2/(2*sigma_width^2)); 
             width_extender = [flip(width_extender) center_width width_extender];
@@ -62,11 +62,11 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         kernel2D = width_extender' * kernel;
         
         % prepare gaussian weights for height
-        center_height = floor(kernel_height/2);
+        center_height = ceil(kernel_height/2);
 
         % if sigma_height is 0, then use a constant weight for the height
[…]

### 2024-03-04 Code/showMovementPandaMultiplePoints.m

@@ -1,7 +1,7 @@
 function [f] = showMovementPandaMultiplePoints(grid, Tbase, control_points, output)
 
     % APF field vector length
-    arrow_length = 5; % adjust the length to your preference
+    arrow_length = 0.5; % adjust the length to your preference
 
     % show floor
     floor = false

### 2024-03-05 Code/REP_kernels.m
fixed kernels, shortened kernels, added sine kernel extensions

@@ -31,9 +31,9 @@ function [kernels] = REP_kernels(varargin)
 
     % Define the default values for the optional parameters
     defaultSigma = 2;
-    defaultKernelLength = 10;
-    defaultKernelWidth = 10;
-    defaultKernelHeight = 10;
+    defaultKernelLength = 7;
+    defaultKernelWidth = 7;
+    defaultKernelHeight = 7;
 
     % Add the required and optional parameters to the input parser
     addParameter(p, 'type', 'linear', @ischar);
@@ -53,9 +53,13 @@ function [kernels] = REP_kernels(varargin)
     kernel_height = p.Results.kernel_height;
 
     % form kernels
-    kernel_x = directional_kernel_3d('x', kernel_length, sigma, kernel_width, 0, kernel_height, 0, type);
-    kernel_y = directional_kernel_3d('y', kernel_length, sigma, kernel_width, 0, kernel_height, 0, type);
-    kernel_z = directional_kernel_3d('z', kernel_length, sigma, kernel_width, 0, kernel_height, 0, type);
+%     kernel_x = directional_kernel_3d('x', kernel_length, sigma, kernel_width, 2, kernel_height, 2, type);
+%     kernel_y = directional_kernel_3d('y', kernel_length, sigma, kernel_width, 2, kernel_height, 2, type);
+%     kernel_z = directional_kernel_3d('z', kernel_length, sigma, kernel_width, 2, kernel_height, 2, type);
+
+    kernel_x = directional_kernel_3d_sin('x', kernel_length, sigma, kernel_width, kernel_height, type);
+    kernel_y = directional_kernel_3d_sin('y', kernel_length, sigma, kernel_width, kernel_height, type);
+    kernel_z = directional_kernel_3d_sin('z', kernel_length, sigma, kernel_width, kernel_height, type);
 
     kernels = {kernel_x/max(kernel_x(:)), kernel_y/max(kernel_y(:)), kernel_z/max(kernel_z(:))};
 

### 2024-03-05 Code/directional_kernel_3d.m

@@ -11,7 +11,7 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         kernel = [1:1:center 0 -flip(1:1:center)] / (center); 
 
         % extend weights to kernel width
-        center_width = ceil(kernel_width/2);
+        center_width = floor(kernel_width/2);
         
         % if sigma_width is 0, then use a constant weight for the width
         if (sigma_width == 0)
@@ -25,7 +25,7 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         kernel2D = width_extender' * kernel;
 
         % extend weight to kernel height
-        center_height = ceil(kernel_height/2);
+        center_height = floor(kernel_height/2);
 
         % if sigma_height is 0, then use a constant weight for the height
         if (sigma_height == 0)
@@ -48,7 +48,7 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         kernel = [flip(kernel) 0 -kernel];
         
         % prepare gaussian weights for width
-        center_width = ceil(kernel_width/2);
+        center_width = floor(kernel_width/2);
 
         % if sigma_width is 0, then use a constant weight for the width
         if (sigma_width == 0)
@@ -62,7 +62,7 @@ function [kernel3D] = directional_kernel_3d(direction, kernel_length, sigma, ker
         kernel2D = width_extender' * kernel;
         
         % prepare gaussian weights for height
-        center_height = ceil(kernel_height/2);
+        center_height = floor(kernel_height/2);
 
         % if sigma_height is 0, then use a constant weight for the height
         if (sigma_height == 0)

### 2024-03-05 Code/directional_kernel_3d_sin.m

@@ -1,0 +1,167 @@
+function [kernel3D] = directional_kernel_3d_sin(direction, kernel_length, sigma, kernel_width, kernel_height, kernel_type)
+
+   % SETTINGS
+   % ---------------------------------------------------------------------
+   constant_kernel_extensions = 0;
+
+
+   % FORM KERNELS 
+   % ---------------------------------------------------------------------
+
+    center = floor(kernel_length/2); % mid of the kernel
+
+   if strcmp('linear',kernel_type)
+
+        % prepare directional weights
+        kernel = [1:1:center 0 -flip(1:1:center)] / (center); 
+
+%         % extend weights to kernel width
+%         center_width = floor(kernel_width/2);
+%         
+%         % if sigma_width is 0, then use a constant weight for the width
+%         if (sigma_width == 0)
+%             width_extender = [ones(1,center_width) 1 ones(1,center_width)];
+%         else
+%             width_extender = [1:1:center_width center_width flip(1:1:center_width)]/(center_width);
+%         end
+% 
+% 
+%         % generate 2D kernel
+%         kernel2D = width_extender' * kernel;
+% 
+%         % extend weight to kernel height
+%         center_height = floor(kernel_height/2);
+% 
+%         % if sigma_height is 0, then use a constant weight for the height
+%         if (sigma_height == 0)
+%             height_extender = [ones(1,center_height) 1 ones(1,center_height)];
+%         else
+%             height_extender = [1:1:center_height center_height flip(1:1:center_height)]/(center_height);
+%         end
+% 
+%         
+%         % create each layer
+%         kernel3D = repmat(kernel2D, 1, 1, length(height_extender)) .* reshape(height_extender, 1, 1, []);
+
+
+    elseif strcmp('gaussian',kernel_type)
+      
+        % generate gaussian weights
+        kernel = exp(-(1:center).^2/(2*sigma^2)) / (sigma * sqrt(2*pi)); 
+        
+        % generate symmetric kernel
+        kernel = [flip(kernel) 0 -kernel];
+        
+%         % prepare gaussian weights for width
+%         center_width = floor(kernel_width/2);
+% 
+%         % if sigma_width is 0, then use a constant weight for the width
+%         if (sigma_width == 0)
+%             width_extender = [ones(1,center_width) 1 ones(1,center_width)];
+%         else
+%             width_extender = exp(-(1:center_width).^2/(2*sigma_width^2)); 
[…]

### 2024-03-05 Code/Experiments APF/MANIP_EXPERIMENTS.m
removed non-movable POI, ran some experiments

@@ -5,7 +5,7 @@ close all
 % ------------------------------------------------------------
 
 % select map
-map_selection = 'roof'
+map_selection = 'column'
 
 [grid] = MAPS(map_selection);
 
@@ -179,22 +179,20 @@ toc()
 %% PLOT DISTANCE AND MANIPULABILITY
 % ------------------------- 
 
-% figure()
-% plot(output.goal_distances, 'g')
-% hold on
-% plot(output.manipulability_avoidance*10, 'r')
-% plot(output.manipulability_primary*10, 'b')
-% 
-% legend('Goal Distances', 'Manipulability Avoidance x10', 'Manipulability Primary x10')
-% xlabel('Iteration')
-% ylabel('Value')
-% title('Plot of Goal Distances and Manipulabilities')
-% grid on
+figure()
+plot(output.goal_distances, 'g')
+hold on
+plot(output.manipulability_avoidance*10, 'r')
+plot(output.manipulability_primary*10, 'b')
+
+legend('Goal Distances', 'Manipulability Avoidance x10', 'Manipulability Primary x10')
+xlabel('Iteration')
+ylabel('Value')
+title('Plot of Goal Distances and Manipulabilities')
+grid on
 
 
 %% PLOT VISUALIZATION
 % ------------------------- 
-
-% showMovementPandaRepulsiveOnePoint(grid, control_points, output.joints_positions, output.EE_positions, output.values_APF, Tbase)
 
 showMovementPandaMultiplePoints(grid, Tbase, control_points, output)

### 2024-03-05 Code/Experiments APF/REP_EXPERIMENTS.m

@@ -4,7 +4,7 @@ close all
 point = [0.7,0.7,1]
 
 %% SELECT MAP
-map_selection = 'wall'
+map_selection = 'column'
 % -----------------------------------------------------------
 
 [grid] = MAPS(map_selection)

### 2024-03-05 Code/Full_RTConvolution_Multiple_Points.m

@@ -3,15 +3,15 @@ function [output] = Full_RTConvolution_Multiple_Points(grid, goal_point, Tbase, 
 %% PARAMETERS
 
 % select which goals
-mid_joints = 1
+mid_joints = 0
 avoid_task = 1
 kinematics_solution = 'exact-reduced' % OPTIONS: exact-reduced , exact , approximate
 timestep_secondary_gain_change = 0 % if selected, secondary task will start with normal gain and fall with time
 secondary_exec_stop_k = 1 % primary task will slow down (>0) or stop executing if secondary task has big velocities 
-min_exec_slowdown_size = 1 % if poi is closer than this value, primary task will slow down
+min_exec_slowdown_size = 0 % if poi is closer than this value, primary task will slow down
 
 % -----------------------------------------------------------
-% number of points per segment for obstacle avoidance taskmanipulability_avoidance
+% number of points per segment for obstacle avoidance task manipulability_avoidance
 points_per_segment = 1*[1 1 2 1 3 1 1];
 
 % the number of points taken into account and weighting factors
@@ -21,20 +21,20 @@ weights_avoidance = weights_avoidance / norm(weights_avoidance,1) / 10;
 % -----------------------------------------------------------
 
 Tstep = 0.1 % time step
-Nmax = 500 % max number of iterations
+Nmax = 50 % max number of iterations
 space_resolution = grid.resolution; % resolution of the obstacles grid
 
 % weights for different tasks
-wp = 0.75 % primary task
+wp = 5 % primary task
 wp_att = 1 % primary task - attractive component
 wp_rep = 0 % primary task - repulsive component
-wm = 0.1 % mid-joints task
-wa = 0.1 % obstacle avoidance task
+wm = 0.5 % mid-joints task
+wa = 0.5 % obstacle avoidance task
 
 
 
 % function parameters
-goal_dist = 0.001 % distance which satisfies ending of optimization
+goal_dist = 0.01 % distance which satisfies ending of optimization
 
 % damping factor to avoid inverse Jacobain matrix singularities
 damping_factor_primary = 0.01
@@ -118,7 +118,7 @@ while current_dist > goal_dist && Niter <= Nmax
 
     % ATTRACTIVE ( OPTION KINEMATICS CLASSIC END EFFECTOR )
     ee_vel_att_magn = norm(goal_point(1:3)' - ee_point);
[…]

### 2024-03-06 Code/Full_RTConvolution_Multiple_Points.m
Update Full_RTConvolution_Multiple_Points.m

@@ -3,7 +3,7 @@ function [output] = Full_RTConvolution_Multiple_Points(grid, goal_point, Tbase, 
 %% PARAMETERS
 
 % select which goals
-mid_joints = 0
+mid_joints = 1
 avoid_task = 1
 kinematics_solution = 'exact-reduced' % OPTIONS: exact-reduced , exact , approximate
 timestep_secondary_gain_change = 0 % if selected, secondary task will start with normal gain and fall with time
@@ -21,7 +21,7 @@ weights_avoidance = weights_avoidance / norm(weights_avoidance,1) / 10;
 % -----------------------------------------------------------
 
 Tstep = 0.1 % time step
-Nmax = 50 % max number of iterations
+Nmax = 1000 % max number of iterations
 space_resolution = grid.resolution; % resolution of the obstacles grid
 
 % weights for different tasks
@@ -29,7 +29,7 @@ wp = 5 % primary task
 wp_att = 1 % primary task - attractive component
 wp_rep = 0 % primary task - repulsive component
 wm = 0.5 % mid-joints task
-wa = 0.5 % obstacle avoidance task
+wa = 1 % obstacle avoidance task
 
 
 
@@ -118,7 +118,7 @@ while current_dist > goal_dist && Niter <= Nmax
 
     % ATTRACTIVE ( OPTION KINEMATICS CLASSIC END EFFECTOR )
     ee_vel_att_magn = norm(goal_point(1:3)' - ee_point);
-    ee_vel_att = (goal_point(1:3)' - ee_point)/ee_vel_att_magn * atan(10*ee_vel_att_magn)/pi*2; % direction only - normalised - sigmoid
+    ee_vel_att = (goal_point(1:3)' - ee_point)/ee_vel_att_magn * atan(100*ee_vel_att_magn)/pi*2; % direction only - normalised - sigmoid
 
     % REPULSIVE 
     ee_vel_rep = REP_field_calculation(grid, rep_kernels, ee_point);

### 2024-03-07 Articles/IROS/root.tex
added field equations

@@
 
+\begin{equation}
+	S = \sum_{i,j,k} (W \odot A)_{ijk} = \sum_{i}\sum_{j}\sum_{k} w_{\Delta i \Delta j \Delta k} \times  a_{\Delta i \Delta j \Delta k}
+\end{equation}
+
 Subsequently, the avoidance velocity in the Cartesian coordinate system is obtained (eq.~\ref{eq:summation}) by summing all the values in the resulting matrix $C_d$, which can be represented as:
@@
 
+$\Delta i$, $\Delta j$ in $\Delta k$ predstavljajo število celic odmika od centralnega polja matrike, v katerem se nahaja naša točka na agentu v posamezno koordinatno smer.
+
 \begin{equation}
-w_i = 
+w_{\Delta i} = 
 \begin{cases} 
@@
 
-Another distribution we used is mirrored linear, where the $l=\lceil width / 2 \rceil$ is the rounded up half length of the primary axis kernel.
+Another distribution we used is mirrored linear, where the $l=\lfloor width / 2 \rfloor$ is the rounded down half length of the primary axis kernel.
 
 \begin{equation}
-	w_i = 
+	w_{\Delta i} = 
 	\begin{cases} 
@@
 % ORTHAGONAL AXIS
+
+If the matrix would be only 1 field width and height the field would work kind of as ray tracing in each of the main cartesian coordinate directions.
+Since our need is to detect also obstacles that dont align perfectly along the cartesian direction, it is important that our matrixes have width and height.
+However as we want bigger repulsive field when the obstacle is head on in the direction than when the obstacle is off the cardinal direction, we propose the following multiplicator, to account for the off direction obstacles.
 
@@
 Excessively wide kernels may generate repulsive velocities for objects that are not in the path of the robot, whereas too narrow kernels might only detect obstacles aligned directly with the Cartesian direction in the point of interest.
-When selecting the width and height of the kernel, we must consider the density of the neighboring points of interest on the robot, ensuring that the \replace{collective fields}{combination of kernels} adequately cover the entire manipulator's surrounding area.
+When selecting the width and height of the kernel, we must consider the density of the neighboring points of interest on the agent, ensuring that the \replace{collective fields}{combination of kernels} adequately cover the entire agent's surrounding area.
+
+For smooth transitions when approaching the obstacles we propose the following sinus based function for the orthagonal axis distributions.
+The proposed equation is the same for both of the orthagonal matrix directions / axis.
+That is of course while operating with axis the $r=\lfloor width / 2 \rfloor$ is the rounded down half length of the selected orthagonal axis kernel.
+
+\begin{equation}
+	w_{\Delta j} =
+	\begin{cases} 
+		sin(\left| \Delta j \right| \pi / (2 r)) & \text{if } \Delta j < 0 \text{~or }  \Delta j > 0 \\
+		1 & \text{if } \Delta j = 0 \\
+	\end{cases}
+\end{equation}
+
+Finally we get three matrix kernels, one for each of the three coordinate axis directions by multiplieing weights components for primary and orthagonal axis.
+
+\begin{equation}
+	w_{\Delta i \Delta j \Delta k} = w_{\Delta i} * w_{\Delta j} * w_{\Delta k}
+\end{equation}
 
@@
 \comment{}{EQUATION: kernel generation / values equations}
-
-\add{choice of velocity profile (gaussian, linear ... maybe could move this part to implementation / experiments)}
 

### 2024-03-07 Articles/IROS/root.tex
izbira primerov

@@
 \title{\LARGE \bf
-Matrix Multiplication-Driven Repulsive Fields for 3D Voxel-Based Robotic Manipulator Path Planning
+Matrix Multiplication-Driven Repulsive Fields for 3D Voxel-Based TSDF Calculation
 }
@@
 
-\begin{equation}
-	C_d = A_d \odot K_d
-	\label{eq:hadamard}
-\end{equation}
+%\begin{equation}
+%	C_d = A_d \odot K_d
+%	\label{eq:hadamard}
+%\end{equation}
 
@@
 
-$\Delta i$, $\Delta j$ in $\Delta k$ predstavljajo število celic odmika od centralnega polja matrike, v katerem se nahaja naša točka na agentu v posamezno koordinatno smer.
+$\Delta i = c_i - i$, $\Delta j = c_j - j$ in $\Delta k = c_k - k$ predstavljajo število celic odmika od centralnega polja matrike, v katerem se nahaja naša točka na agentu v posamezno koordinatno smer.
 
@@
 
+\add{mogoče dodaj še kak stavek o izbiri dimenzij matrik}
+
 %\remove{In the orthogonal axes, the magnitude of values diminishes with increasing distance from the core of the kernel, nearing zero at the edges.
@@
 		sin(\left| \Delta j \right| \pi / (2 r)) & \text{if } \Delta j < 0 \text{~or }  \Delta j > 0 \\
+		1 & \text{if } \Delta j = 0 \\
+	\end{cases}
+\end{equation}
+
+Another option is to use linearly falling weights.
+
+\begin{equation}
+	w_{\Delta j} =
+	\begin{cases} 
+		\frac{l - \Delta j }{l} & \text{if } \Delta j < 0 \text{~or }  \Delta j > 0 \\
 		1 & \text{if } \Delta j = 0 \\
@@
 
-\section{INVERSE KINEMATICS}
+\section{SIMULATION RESULTS}
 
-\section{REPULSIVE VELOCITIES CALCULATION}
+\subsection{Repulsive Field Visualization}
 
-\section{SIMULATION RESULTS}
+\add{Visualization of the repulsive field around more or less complicated obstacles.}
+
+\comment{}{we dont have the close obstacles bottlenecks}
+
+\subsection{Manipulator Examples}
+
+\add{EXAMPLE: MANIPULATOR AND COLUMN}
+
+\add{EXAMPLE: MANIPULATOR AND MOVING NOISY BALL}
+
+\subsection{Drone Example}
+
+\add{EXAMPLE: DRONE MOVING THROUGH THE MAP OF OBSTACLES USING APF}
 

### 2024-03-08 Code/REP_kernels.m
added vector field visualization

@@ -53,13 +53,13 @@ function [kernels] = REP_kernels(varargin)
     kernel_height = p.Results.kernel_height;
 
     % form kernels
-%     kernel_x = directional_kernel_3d('x', kernel_length, sigma, kernel_width, 2, kernel_height, 2, type);
-%     kernel_y = directional_kernel_3d('y', kernel_length, sigma, kernel_width, 2, kernel_height, 2, type);
-%     kernel_z = directional_kernel_3d('z', kernel_length, sigma, kernel_width, 2, kernel_height, 2, type);
-
-    kernel_x = directional_kernel_3d_sin('x', kernel_length, sigma, kernel_width, kernel_height, type);
-    kernel_y = directional_kernel_3d_sin('y', kernel_length, sigma, kernel_width, kernel_height, type);
-    kernel_z = directional_kernel_3d_sin('z', kernel_length, sigma, kernel_width, kernel_height, type);
+    kernel_x = directional_kernel_3d('x', kernel_length, sigma, kernel_width, 2, kernel_height, 2, type);
+    kernel_y = directional_kernel_3d('y', kernel_length, sigma, kernel_width, 2, kernel_height, 2, type);
+    kernel_z = directional_kernel_3d('z', kernel_length, sigma, kernel_width, 2, kernel_height, 2, type);
+% 
+%     kernel_x = directional_kernel_3d_sin('x', kernel_length, sigma, kernel_width, kernel_height, type);
+%     kernel_y = directional_kernel_3d_sin('y', kernel_length, sigma, kernel_width, kernel_height, type);
+%     kernel_z = directional_kernel_3d_sin('z', kernel_length, sigma, kernel_width, kernel_height, type);
 
     kernels = {kernel_x/max(kernel_x(:)), kernel_y/max(kernel_y(:)), kernel_z/max(kernel_z(:))};
 

### 2024-03-11 Articles/IROS/root.tex
comments
added field visualization

@@
 \end{equation}
+
+\alert{remove the above equation and add a vector of calculated velocities equation instead}
 
@@
 
-\comment{}{we dont have the close obstacles bottlenecks}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\linewidth]{demo-field1.eps} % Replace 'example-image' with the filename of your image
+	\caption{Example Image}
+	\label{fig:example-field}
+\end{figure}
+
+\comment{}{we dont have the close obstacles bottlenecks, we do have weird bottlenecks in the orthagonal directions - maybe}
 

### 2024-03-11 Articles/IROS/root.tex
added interpolation example

@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{demo-field1.eps} % Replace 'example-image' with the filename of your image
+	\includegraphics[width=0.95\linewidth]{demo-field1.eps} % Replace 'example-image' with the filename of your image
 	\caption{Example Image}
@@
 
+\add{add label, caption, every plot has different axis scale}
+
 \comment{}{we dont have the close obstacles bottlenecks, we do have weird bottlenecks in the orthagonal directions - maybe}
+
+\subsection{Repulsive Field Interpolation}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\linewidth]{non_and_interp_4_plots.eps} % Replace 'example-image' with the filename of your image
+	\caption{Example Image}
+	\label{fig:interp-experiment}
+\end{figure}
 
@@
 
-\subsection{Drone Example}
+\remove{Drone Example}
 
-\add{EXAMPLE: DRONE MOVING THROUGH THE MAP OF OBSTACLES USING APF}
+\remove{EXAMPLE: DRONE MOVING THROUGH THE MAP OF OBSTACLES USING APF}
 

### 2024-03-11 Code/Full_RTConvolution_Multiple_Points.m
constants

@@ -28,8 +28,8 @@ space_resolution = grid.resolution; % resolution of the obstacles grid
 wp = 5 % primary task
 wp_att = 1 % primary task - attractive component
 wp_rep = 0 % primary task - repulsive component
-wm = 0.5 % mid-joints task
-wa = 1 % obstacle avoidance task
+wm = 1 % mid-joints task
+wa = 1.5 % obstacle avoidance task
 
 
 
@@ -119,6 +119,9 @@ while current_dist > goal_dist && Niter <= Nmax
     % ATTRACTIVE ( OPTION KINEMATICS CLASSIC END EFFECTOR )
     ee_vel_att_magn = norm(goal_point(1:3)' - ee_point);
     ee_vel_att = (goal_point(1:3)' - ee_point)/ee_vel_att_magn * atan(100*ee_vel_att_magn)/pi*2; % direction only - normalised - sigmoid
+
+    % ORIENTATION
+    ee_vel_orient = 
 
     % REPULSIVE 
     ee_vel_rep = REP_field_calculation(grid, rep_kernels, ee_point);

### 2024-03-11 Code/Full_RTConvolution_Multiple_Points.m
added orientations

@@ -1,10 +1,27 @@
 function [output] = Full_RTConvolution_Multiple_Points(grid, goal_point, Tbase, q)
 
+%% POSSIBLE GOAL ORIENTATION OPTIONS
+
+R_x = [cos(pi/2)  0 sin(pi/2);
+       0         1 0;
+      -sin(pi/2) 0 cos(pi/2)];
+
+
+R_y = [1 0        0;
+       0 cos(-pi/2) -sin(-pi/2);
+       0 sin(-pi/2) cos(-pi/2)];
+
+R_z = [cos(pi) 0 sin(pi);
+       0       1 0;
+      -sin(pi) 0 cos(pi)];
+
+
 %% PARAMETERS
 
 % select which goals
 mid_joints = 1
 avoid_task = 1
+orientation_task = 1
 kinematics_solution = 'exact-reduced' % OPTIONS: exact-reduced , exact , approximate
 timestep_secondary_gain_change = 0 % if selected, secondary task will start with normal gain and fall with time
 secondary_exec_stop_k = 1 % primary task will slow down (>0) or stop executing if secondary task has big velocities 
@@ -31,7 +48,8 @@ wp_rep = 0 % primary task - repulsive component
 wm = 1 % mid-joints task
 wa = 1.5 % obstacle avoidance task
 
-
+% EE orientation goal 
+orientation_goal = R_z;
 
 % function parameters
 goal_dist = 0.01 % distance which satisfies ending of optimization
@@ -60,6 +78,10 @@ transformations_list = getPartialTransformationsListPanda('points_per_segment', 
 
 % set current ee point to start point
 ee_point = robot_transforms(1:3,4,8);
+
+% set current ee orientation to start orientation matrix
+ee_orient = robot_transforms(1:3,1:3,8);
+
 
 % calculate distance from goal
 current_dist = norm(ee_point'- goal_point(1:3));
@@ -121,7 +143,7 @@ while current_dist > goal_dist && Niter <= Nmax
     ee_vel_att = (goal_point(1:3)' - ee_point)/ee_vel_att_magn * atan(100*ee_vel_att_magn)/pi*2; % direction only - normalised - sigmoid
 
     % ORIENTATION
-    ee_vel_orient = 
+%     ee_vel_orient
 
     % REPULSIVE 
     ee_vel_rep = REP_field_calculation(grid, rep_kernels, ee_point);
@@ -349,9 +371,11 @@ while current_dist > goal_dist && Niter <= Nmax
[…]

### 2024-03-11 Code/Experiments APF/MANIP_EXPERIMENTS.m
added orientation task - control

@@ -191,6 +191,10 @@ ylabel('Value')
 title('Plot of Goal Distances and Manipulabilities')
 grid on
 
+% PLOT ORIENTATION ERROR
+figure()
+plot(output.goal_orientation)
+
 
 %% PLOT VISUALIZATION
 % ------------------------- 

### 2024-03-11 Code/Full_RTConvolution_Multiple_Points.m

@@ -2,18 +2,14 @@ function [output] = Full_RTConvolution_Multiple_Points(grid, goal_point, Tbase, 
 
 %% POSSIBLE GOAL ORIENTATION OPTIONS
 
-R_x = [cos(pi/2)  0 sin(pi/2);
-       0         1 0;
-      -sin(pi/2) 0 cos(pi/2)];
+Rot_z1 = [1 0 0 ; 
+         0 -1 0 ; 
+         0 0 -1 ];
 
+Rot_z2 = [-1 0 0 ;
+          0 1 0 ;
+          0 0 -1];
 
-R_y = [1 0        0;
-       0 cos(-pi/2) -sin(-pi/2);
-       0 sin(-pi/2) cos(-pi/2)];
-
-R_z = [cos(pi) 0 sin(pi);
-       0       1 0;
-      -sin(pi) 0 cos(pi)];
 
 
 %% PARAMETERS
@@ -38,18 +34,19 @@ weights_avoidance = weights_avoidance / norm(weights_avoidance,1) / 10;
 % -----------------------------------------------------------
 
 Tstep = 0.1 % time step
-Nmax = 1000 % max number of iterations
+Nmax = 50 % max number of iterations
 space_resolution = grid.resolution; % resolution of the obstacles grid
 
 % weights for different tasks
 wp = 5 % primary task
+wp_or = 20 % primary task - orientation component
 wp_att = 1 % primary task - attractive component
 wp_rep = 0 % primary task - repulsive component
 wm = 1 % mid-joints task
 wa = 1.5 % obstacle avoidance task
 
 % EE orientation goal 
-orientation_goal = R_z;
+orientation_goal = Rot_z1
 
 % function parameters
 goal_dist = 0.01 % distance which satisfies ending of optimization
@@ -86,6 +83,27 @@ ee_orient = robot_transforms(1:3,1:3,8);
 % calculate distance from goal
 current_dist = norm(ee_point'- goal_point(1:3));
 
+% starting orientation error
+    % rotation goal
+    rR = orientation_goal;    
+        
+    % current rotation
+    RR = ee_orient;
+
+    % calculate rotation
+    dR = rR*RR';
+                
+    dR = dR / norm(dR);
+        
+    eq = 2*log(quaternion(rotm2quat(dR)));
+        
+    [~, qB, qC, qD] = eq.parts;
+        
+    eR = wp_or * [qB qC qD]
+
+% calculate distance in orientation
[…]

### 2024-03-11 Code/generate_figure_keyframes.m
Create generate_figure_keyframes.m

@@ -1,0 +1,85 @@
+function [f] = generate_figure_keyframes(grid, Tbase, control_points, output)
+    % Define the keyframes
+    total_frames = size(output.joints_positions, 2);
+    keyframes = round(linspace(1, total_frames-1, 6));
+
+    % APF field vector length
+    arrow_length = 0.5; % adjust the length to your preference
+
+    %% COLORMAP for APF vectors
+    cMap = interp1(0:1,[0 1 0; 1 0 0],linspace(0,1,256));
+    cMap = cMap.^(1/2.4); % linearize
+
+    %% Create figure for subplots
+    f = figure('Position', [100, 100, 500, 300]);
+    set(f, 'WindowStyle', 'normal');
+
+    % Loop through each keyframe
+    for kf = 1:length(keyframes)
+        % Set the axes limits to ensure consistency across subplots
+        xlim([0 20]);   
+        ylim([0 20]);   
+        zlim([0 10]);
+%         axis equal;  
+
+        % Create subplot for each keyframe
+        subplot(2, 3, kf);
+        % Remove axis rulers
+        set(gca, 'XTick', [], 'YTick', [], 'ZTick', []);
+        box on; % Turn the box on
+%         set(gca, 'LineWidth', 1); % Set the box line width to 1 (adjust this value as needed)
+        hold on;
+%         axis equal;
+        
+        % Current keyframe index
+        j = keyframes(kf);
+        
+        % Display grid
+        grid.showGridVol3D(grid.grid, 'floor', false, 'height', true);
+        
+        % Display control points
+        for i = 1:size(control_points,1)
+            scatter3(control_points(i,1)*grid.resolution, control_points(i,2)*grid.resolution, control_points(i,3)*grid.resolution, 'r');
+        end
+        
+        % Draw EE trajectory up to the current keyframe
+        plot3(output.EE_positions(1,1:j)*grid.resolution, output.EE_positions(2,1:j)*grid.resolution, output.EE_positions(3,1:j)*grid.resolution, 'red');
+        
+        % Calculate transforms for the current keyframe
+        [transforms] = GeometricPandaMATLAB(output.joints_positions(:,j), Tbase);
+        
+        % Draw robot configuration for the current keyframe
+        for i = 1:length(transforms)
+            if i > 1
[…]

### 2024-03-11 Articles/IROS/root.tex
added 3d keyframes plot for column

@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{demo-field1.eps} % Replace 'example-image' with the filename of your image
+	\includegraphics[width=0.95\linewidth]{demo-field1c.eps} % Replace 'example-image' with the filename of your image
 	\caption{Example Image}
@@
 \add{EXAMPLE: MANIPULATOR AND COLUMN}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\linewidth]{keyframes_3D.eps} % Replace 'example-image' with the filename of your image
+	\caption{Example Image}
+	\label{fig:keyframes-3d-column}
+\end{figure}
 

### 2024-03-11 Code/generate_figure_keyframes.m
added 2D option

@@ -25,7 +25,7 @@ function [f] = generate_figure_keyframes(grid, Tbase, control_points, output)
         % Create subplot for each keyframe
         subplot(2, 3, kf);
         % Remove axis rulers
-        set(gca, 'XTick', [], 'YTick', [], 'ZTick', []);
+%         set(gca, 'XTick', [], 'YTick', [], 'ZTick', []);
         box on; % Turn the box on
 %         set(gca, 'LineWidth', 1); % Set the box line width to 1 (adjust this value as needed)
         hold on;
@@ -39,11 +39,11 @@ function [f] = generate_figure_keyframes(grid, Tbase, control_points, output)
         
         % Display control points
         for i = 1:size(control_points,1)
-            scatter3(control_points(i,1)*grid.resolution, control_points(i,2)*grid.resolution, control_points(i,3)*grid.resolution, 'r');
+            scatter3(control_points(i,1)*grid.resolution, control_points(i,2)*grid.resolution, control_points(i,3)*grid.resolution, 'xr');
         end
         
         % Draw EE trajectory up to the current keyframe
-        plot3(output.EE_positions(1,1:j)*grid.resolution, output.EE_positions(2,1:j)*grid.resolution, output.EE_positions(3,1:j)*grid.resolution, 'red');
+%         plot3(output.EE_positions(1,1:j)*grid.resolution, output.EE_positions(2,1:j)*grid.resolution, output.EE_positions(3,1:j)*grid.resolution, 'red');
         
         % Calculate transforms for the current keyframe
         [transforms] = GeometricPandaMATLAB(output.joints_positions(:,j), Tbase);
@@ -66,19 +66,24 @@ function [f] = generate_figure_keyframes(grid, Tbase, control_points, output)
         apf_magnitude_remapped = round(min(apf_magnitude,1)*255 + 1);
         apf_color = cMap(apf_magnitude_remapped,:);
         
-        quiver3(output.POI_locations{j}(1,i)*grid.resolution, output.POI_locations{j}(2,i)*grid.resolution, output.POI_locations{j}(3,i)*grid.resolution, output.POI_values{j}(1,i)*arrow_length, output.POI_values{j}(2,i)*arrow_length, output.POI_values{j}(3,i)*arrow_length, 'LineWidth', 2, 'MaxHeadSize', 1, 'Color', apf_color);
[…]

### 2024-03-11 Code/OctoGrid.m
added 2 columns view, fixed octogrid bug

@@ -103,7 +103,7 @@ classdef OctoGrid < handle
 
                         if heightColorMap
                             % Get the value of the current voxel
-                            voxel_value = grid(x,y,z);
+                            voxel_value = grid(y,x,z);
                             % Define the scalar value
                             value = z/size(grid,3);
 
@@ -115,7 +115,7 @@ classdef OctoGrid < handle
 
                         else
                             % Get the value of the current voxel
-                            voxel_value = grid(x,y,z);
+                            voxel_value = grid(y,x,z);
                         end
         
                         % Draw the voxel as a cube if its value is above the 

### 2024-03-11 Code/generate_figure_keyframes_2D.m

@@ -1,4 +1,4 @@
-function [f] = generate_figure_keyframes(grid, Tbase, control_points, output)
+function [f] = generate_figure_keyframes_2D(grid, Tbase, control_points, output)
     % Define the keyframes
     total_frames = size(output.joints_positions, 2);
     keyframes = round(linspace(1, total_frames-1, 6));
@@ -35,8 +35,8 @@ function [f] = generate_figure_keyframes(grid, Tbase, control_points, output)
         j = keyframes(kf);
         
         % Display grid
-        grid.showGridVol3D(grid.grid, 'floor', false, 'height', true);
-%         grid.showGridVoxel(grid.grid, 'floor', false, 'height', false);
+%         grid.showGridVol3D(grid.grid, 'floor', false, 'height', true);
+        grid.showGridVoxel(grid.grid, 'floor', false, 'height', false);
 
         
         % Display control points
@@ -81,10 +81,10 @@ function [f] = generate_figure_keyframes(grid, Tbase, control_points, output)
 
 
     % 3D view
-    view([-180 45.0]);
+%     view([-180 45.0]);
 
     % bird view (top-down view)
-%     view(0, 90);
+    view(0, 90);
 
     
     hold off;

### 2024-03-11 Code/generate_figure_keyframes_column.m

@@ -1,0 +1,83 @@
+function [f] = generate_figure_keyframes_column(grid, Tbase, control_points, output)
+    % Define the keyframes
+    total_frames = size(output.joints_positions, 2);
+    keyframes = round(linspace(1, total_frames-1, 4)); % Change number of keyframes to 4
+
+    % APF field vector length
+    arrow_length = 0.5; % adjust the length to your preference
+
+    %% COLORMAP for APF vectors
+    cMap = interp1(0:1,[0 1 0; 1 0 0],linspace(0,1,256));
+    cMap = cMap.^(1/2.4); % linearize
+
+    %% Create figure for subplots
+    f = figure('Position', [100, 100, 400, 600]); % Adjust figure size to accommodate two columns
+    set(f, 'WindowStyle', 'normal');
+
+    % Loop through each keyframe twice for 3D and 2D views
+    for kf = 1:length(keyframes)
+        for viewMode = 1:2 % 1 for 3D view, 2 for 2D view
+            % Create subplot for each keyframe with 3D and 2D views in two columns
+            subplot(4, 2, (kf-1)*2 + viewMode); % Adjust subplot indexing for two columns
+            box on; % Turn the box on
+            hold on;
+            
+            % Set the axes limits to ensure consistency across subplots
+            xlim([0 20]);
+            ylim([0 20]);
+            if viewMode == 1 % Only set zlim for 3D view
+                zlim([0 10]);
+            end
+
+            % Current keyframe index
+            j = keyframes(kf);
+            
+            % Display grid
+            if viewMode == 1
+                grid.showGridVol3D(grid.grid, 'floor', false, 'height', true);            
+            else
+                grid.showGridVoxel(grid.grid, 'floor', false, 'height', false);            
+            end
+            
+            % Display control points
+            for i = 1:size(control_points,1)
+                scatter3(control_points(i,1)*grid.resolution, control_points(i,2)*grid.resolution, control_points(i,3)*grid.resolution, 'xr');
+            end
+            
+            % Calculate transforms for the current keyframe
+            [transforms] = GeometricPandaMATLAB(output.joints_positions(:,j), Tbase);
+            
+            % Draw robot configuration for the current keyframe
+            for i = 1:length(transforms)
+                if i > 1
+                    prevPos = transforms(1:3, 4, i-1);
[…]

### 2024-03-12 Code/Full_RTConvolution_Multiple_Points.m
changed weights, to lower the impact of less important points

@@ -28,7 +28,7 @@ min_exec_slowdown_size = 0 % if poi is closer than this value, primary task will
 points_per_segment = 1*[1 1 2 1 3 1 1];
 
 % the number of points taken into account and weighting factors
-weights_avoidance = [1 1 1 1 1 1];
+weights_avoidance = [3 2 1 1 1 1];
 weights_avoidance = weights_avoidance / norm(weights_avoidance,1) / 10;
 
 % -----------------------------------------------------------

### 2024-03-12 Code/Full_RTConvolution_Moving_Ball.m
Create Full_RTConvolution_Moving_Ball.m

@@ -1,0 +1,473 @@
+function [output] = Full_RTConvolution_Moving_Ball(grid, goal_point, Tbase, q)
+
+%% POSSIBLE GOAL ORIENTATION OPTIONS
+
+Rot_z1 = [1 0 0 ; 
+         0 -1 0 ; 
+         0 0 -1 ];
+
+Rot_z2 = [-1 0 0 ;
+          0 1 0 ;
+          0 0 -1];
+
+
+
+%% PARAMETERS
+
+% select which goals
+mid_joints = 1
+avoid_task = 1
+orientation_task = 1
+kinematics_solution = 'exact-reduced' % OPTIONS: exact-reduced , exact , approximate
+timestep_secondary_gain_change = 0 % if selected, secondary task will start with normal gain and fall with time
+secondary_exec_stop_k = 1 % primary task will slow down (>0) or stop executing if secondary task has big velocities 
+min_exec_slowdown_size = 0 % if poi is closer than this value, primary task will slow down
+
+% -----------------------------------------------------------
+% number of points per segment for obstacle avoidance task manipulability_avoidance
+points_per_segment = 1*[1 1 2 1 3 1 1];
+
+% the number of points taken into account and weighting factors
+weights_avoidance = [1 1 1 1 1 1];
+weights_avoidance = weights_avoidance / norm(weights_avoidance,1) / 10;
+
+% -----------------------------------------------------------
+
+Tstep = 0.1 % time step
+Nmax = 50 % max number of iterations
+space_resolution = grid.resolution; % resolution of the obstacles grid
+
+% weights for different tasks
+wp = 5 % primary task
+wp_or = 20 % primary task - orientation component
+wp_att = 1 % primary task - attractive component
+wp_rep = 0 % primary task - repulsive component
+wm = 1 % mid-joints task
+wa = 1.5 % obstacle avoidance task
+
+% EE orientation goal 
+orientation_goal = Rot_z1
+
+% function parameters
+goal_dist = 0.01 % distance which satisfies ending of optimization
+
+% damping factor to avoid inverse Jacobain matrix singularities
+damping_factor_primary = 0.01
+damping_factor_avoidance = 0.01
+
+% joints range
+q_range = [2.8973 -2.8973;
+           1.7628 -1.7628;
+           2.8973 -2.8973;
+           -0.0698 -3.0718;
+           2.8973 -2.8973;
+           3.7525 -0.0175;
+           2.8973 -2.8973];
+
[…]

### 2024-03-12 Code/Experiments APF/MANIP_EXPERIMENTS_BALL.m
Create MANIP_EXPERIMENTS_BALL.m

@@ -1,0 +1,79 @@
+clear
+close all
+
+% PREPARE MAP & ROBOT
+% ------------------------------------------------------------
+
+    grid = OctoGrid(2,2,2,10);
+
+   % Generate goal points
+    % trajectory generation points (xyz position, xyz angles)
+    control_points = [1.5 1.05 0.25 0 0 0]; 
+
+    % goal
+    goal = control_points;
+    
+    % Robot Base position
+    % robot base transformation (move to x - 1m, y - 1m)
+    Tbase = eye(4);
+    Tbase(1:2,4) = [1 1]';
+    
+    % set joints starting pose
+    q = [0.0319913344656175
+0.164451369212247
+0.00933902434572278
+-2.38024192403430
+0.00518018325880462
+2.53266747724704
+0.0400646905027541];
+    
+
+
+
+%% RUN MOTION PLANNING - JOINT 4 ONLY POI
+% ------------------------------------------------------------
+
+% tic()
+% [output] = Full_RTConvolution(grid, goal, Tbase, q);
+% toc()
+
+%% RUN MOTION PLANNING - MULTIPLE POI (POINTS OF INTEREST)
+
+% tic()
+% [output] = Full_RTConvolution_Multiple_Points(grid, goal, Tbase, q);
+% toc()
+
+%%  RUN MOTION PLANNING - BALL AVOIDANCE
+
+[output] = Full_RTConvolution_Moving_Ball(grid, goal, Tbase, q);
+
+%% RUN MOTION PLANNING - AVOIDANCE FIRST PRIORITY
+% 
+% tic()
+% [output] = Full_RTConvolution_AvoidanceFirst(grid, goal, Tbase, q);
+% toc()
+
+%% PLOT DISTANCE AND MANIPULABILITY
+% ------------------------- 
+
+figure()
+plot(output.goal_distances, 'g')
+hold on
+plot(output.manipulability_avoidance*10, 'r')
+plot(output.manipulability_primary*10, 'b')
+
+legend('Goal Distances', 'Manipulability Avoidance x10', 'Manipulability Primary x10')
+xlabel('Iteration')
+ylabel('Value')
+title('Plot of Goal Distances and Manipulabilities')
+grid on
+
+% PLOT ORIENTATION ERROR
+figure()
+plot(output.goal_orientation)
+
+
+%% PLOT VISUALIZATION
+% ------------------------- 
+
+showMovementPandaMultiplePoints(grid, Tbase, control_points, output)

### 2024-03-13 Code/OctoGrid.m
display the changing grid

@@ -266,5 +266,81 @@ classdef OctoGrid < handle
 
         end
 
+        function addSphere(obj, centerX, centerY, centerZ, radius)
+
+            %ADDSPHERE Adds a spherical object to the occupancy grid.
+            %
+            % This method colors voxels within the specified sphere's radius based
+            % on their occupancy level. For voxels fully enclosed by the sphere, they
+            % are set to 1. For voxels partially covered by the sphere, a value
+            % proportional to the volume of occupancy is assigned. This function
+            % assumes a simplistic linear model for partial volume occupancy.
+            %
+            % Parameters:
+            % obj - The instance of the OctoGrid class
+            % centerX, centerY, centerZ - The center coordinates of the sphere in meters
+            % radius - The radius of the sphere in meters
+            %
+            % Example:
+            % grid.addSphere(0.5, 0.5, 0.5, 0.1)
+            % This example adds a sphere with a radius of 0.1 meters centered at
+            % coordinates (0.5, 0.5, 0.5) meters in the occupancy grid.
+            %
+            % Note: This method modifies the grid property of the OctoGrid object,
+            % updating the occupancy values of the voxels within the sphere's influence.
+        
+            % Convert center coordinates and radius to grid indices
+            centerX_idx = round(centerX * obj.resolution);
+            centerY_idx = round(centerY * obj.resolution);
+            centerZ_idx = round(centerZ * obj.resolution);
+            radius_idx = round(radius * obj.resolution);
+        
+            % Calculate the index range to iterate over, ensuring we stay within bounds
+            x_range = max(1, centerX_idx-radius_idx):min(size(obj.grid, 2), centerX_idx+radius_idx);
+            y_range = max(1, centerY_idx-radius_idx):min(size(obj.grid, 1), centerY_idx+radius_idx);
+            z_range = max(1, centerZ_idx-radius_idx):min(size(obj.grid, 3), centerZ_idx+radius_idx);
+        
+            for x = x_range
+                for y = y_range
+                    for z = z_range
+                        % Calculate the distance from the voxel center to the sphere center in grid indices
+                        dist = sqrt((x - centerX_idx)^2 + (y - centerY_idx)^2 + (z - centerZ_idx)^2);
+        
+                        if dist < radius_idx
+                            % Calculate approximate occupancy
+                            if dist <= radius_idx - 1
+                                occupancy = 1;  % Fully within the sphere
+                            else
+                                % Simple approximation for partial occupancy
[…]

### 2024-03-13 Code/showMovementPandaChangingGrid.m

@@ -1,0 +1,317 @@
+function [f] = showMovementPandaChangingGrid(Tbase, control_points, output)
+
+    % APF field vector length
+    arrow_length = 0.5; % adjust the length to your preference
+
+    % show floor
+    floor = false
+
+    %% COLORMAP
+    % ---------------------------------------------------------------
+    cMap = interp1(0:1,[0 1 0; 1 0 0],linspace(0,1,256));
+    cMap = cMap.^(1/2.4); % linearize
+
+    %% FIGURE AND CONTROLS
+    % ---------------------------------------------------------------
+
+    % create a figure
+    f = figure;
+
+    % move figure to external window
+    set(f, 'WindowStyle', 'normal');  % Change 'docked' to 'normal' to move to external window
+    
+    % add slider to the bottom of the figure
+    s = uicontrol('Style', 'slider',...
+        'Min',1,'Max',100,'Value',100,...
+        'Position', [150 20 300 20],...
+        'Callback', @sliderCallback); 
+
+    % Add checkbox for APF vectors
+    cv = uicontrol('Style', 'checkbox',...
+        'String', 'APF Field',...
+        'Position', [20 20 100 20],...
+        'Value', 1,...
+        'Callback', @checkboxCallbackAPF); 
+
+    % Add checkbox for Obstacles grid
+    co = uicontrol('Style', 'checkbox',...
+        'String', 'Obstacle field',...
+        'Position', [20 40 100 20],...
+        'Value', 1,...
+        'Callback', @checkboxCallbackObstacles); 
+
+    % Create a "Play" button
+    play_button = uicontrol('Style', 'pushbutton', 'String', 'Play', ...
+        'Position', [460, 20, 60, 20], 'Callback', @playButtonCallback);
+
+
+    % On/off APF vectors
+    show_arrows = 1;
+
+    % create grid object to get to the methods
+    grid = OctoGrid(2,2,2,10);
+    
+    % display grid
+    HObstacles = grid.showGridVol3D(output.grids{end},'floor',floor,'height',true);
+    hold on
+    axis equal
+
+    %% CONTROL POINTS AND TRAJECTORY
+    % ---------------------------------------------------------------
+
+    % display goal points 
+    for i = 1:1:size(control_points,1)
+        scatter3(control_points(i,1)*grid.resolution,control_points(i,2)*grid.resolution,control_points(i,3)*grid.resolution,'r')
+    end
+
[…]

### 2024-03-13 Code/Experiments APF/MANIP_EXPERIMENTS_BALL.m
avoid ball

@@ -20,13 +20,12 @@ close all
     
     % set joints starting pose
     q = [0.0319913344656175
-0.164451369212247
-0.00933902434572278
--2.38024192403430
-0.00518018325880462
-2.53266747724704
-0.0400646905027541];
-    
+        0.164451369212247
+        0.00933902434572278
+        -2.38024192403430
+        0.00518018325880462
+        2.53266747724704
+        0.0400646905027541];  
 
 
 
@@ -76,4 +75,9 @@ plot(output.goal_orientation)
 %% PLOT VISUALIZATION
 % ------------------------- 
 
-showMovementPandaMultiplePoints(grid, Tbase, control_points, output)
+% showMovementPandaMultiplePoints(grid, Tbase, control_points, output)
+
+%% PLOT VISUALIZATION - CHANGING GRID
+
+[f] = showMovementPandaChangingGrid(Tbase, control_points, output)
+

### 2024-03-13 Code/Full_RTConvolution_Moving_Ball.m

@@ -15,12 +15,12 @@ Rot_z2 = [-1 0 0 ;
 %% PARAMETERS
 
 % select which goals
-mid_joints = 1
+mid_joints = 0
 avoid_task = 1
-orientation_task = 1
+orientation_task = 0
 kinematics_solution = 'exact-reduced' % OPTIONS: exact-reduced , exact , approximate
 timestep_secondary_gain_change = 0 % if selected, secondary task will start with normal gain and fall with time
-secondary_exec_stop_k = 1 % primary task will slow down (>0) or stop executing if secondary task has big velocities 
+secondary_exec_stop_k = 0.5 % primary task will slow down (>0) or stop executing if secondary task has big velocities 
 min_exec_slowdown_size = 0 % if poi is closer than this value, primary task will slow down
 
 % -----------------------------------------------------------
@@ -34,16 +34,16 @@ weights_avoidance = weights_avoidance / norm(weights_avoidance,1) / 10;
 % -----------------------------------------------------------
 
 Tstep = 0.1 % time step
-Nmax = 50 % max number of iterations
+Nmax = 100 % max number of iterations
 space_resolution = grid.resolution; % resolution of the obstacles grid
 
 % weights for different tasks
-wp = 5 % primary task
-wp_or = 20 % primary task - orientation component
+wp = 2 % primary task
+wp_or = 1 % primary task - orientation component
 wp_att = 1 % primary task - attractive component
 wp_rep = 0 % primary task - repulsive component
 wm = 1 % mid-joints task
-wa = 1.5 % obstacle avoidance task
+wa = 3 % obstacle avoidance task
 
 % EE orientation goal 
 orientation_goal = Rot_z1
@@ -78,7 +78,6 @@ ee_point = robot_transforms(1:3,4,8);
 
 % set current ee orientation to start orientation matrix
 ee_orient = robot_transforms(1:3,1:3,8);
-
 
 % calculate distance from goal
 current_dist = norm(ee_point'- goal_point(1:3));
@@ -121,6 +120,7 @@ output.POI_locations = [];
 output.POI_values = [];
 output.EE_orientation = [];
 output.goal_orientation = [];
+output.grids = {};
 manipulability_primary = 0;
 manipulability_avoidance = [];
 
[…]

### 2024-03-13 Code/Full_RTConvolution_Moving_Ball.m
solid ball avoid example constants

@@ -20,7 +20,7 @@ avoid_task = 1
 orientation_task = 0
 kinematics_solution = 'exact-reduced' % OPTIONS: exact-reduced , exact , approximate
 timestep_secondary_gain_change = 0 % if selected, secondary task will start with normal gain and fall with time
-secondary_exec_stop_k = 0.5 % primary task will slow down (>0) or stop executing if secondary task has big velocities 
+secondary_exec_stop_k = 0 % primary task will slow down (>0) or stop executing if secondary task has big velocities 
 min_exec_slowdown_size = 0 % if poi is closer than this value, primary task will slow down
 
 % -----------------------------------------------------------
@@ -28,7 +28,7 @@ min_exec_slowdown_size = 0 % if poi is closer than this value, primary task will
 points_per_segment = 1*[1 1 2 1 3 1 1];
 
 % the number of points taken into account and weighting factors
-weights_avoidance = [1 1 1 1 1 1];
+weights_avoidance = [1 1 1 1 1];
 weights_avoidance = weights_avoidance / norm(weights_avoidance,1) / 10;
 
 % -----------------------------------------------------------
@@ -39,7 +39,7 @@ space_resolution = grid.resolution; % resolution of the obstacles grid
 
 % weights for different tasks
 wp = 2 % primary task
-wp_or = 1 % primary task - orientation component
+wp_or = 5 % primary task - orientation component
 wp_att = 1 % primary task - attractive component
 wp_rep = 0 % primary task - repulsive component
 wm = 1 % mid-joints task
@@ -146,7 +146,7 @@ while Niter <= Nmax
     %% UPDATE BALL POSITION
     x_ball = 1.25;
     y_ball = 0.4 + (1.4 - 0.4) * Niter / Nmax;
-    z_ball = 0.6;
+    z_ball = 0.7;
 
     % clear previous ball position
     grid.clearGrid();
@@ -181,7 +181,10 @@ while Niter <= Nmax
 
     % ATTRACTIVE ( OPTION KINEMATICS CLASSIC END EFFECTOR )
     ee_vel_att_magn = norm(goal_point(1:3)' - ee_point);
-    ee_vel_att = (goal_point(1:3)' - ee_point)/ee_vel_att_magn * atan(10*ee_vel_att_magn)/pi*2 % direction only - normalised - sigmoid
+
[…]

### 2024-03-13 Articles/IROS/root.tex
added 2D keyframes
added 2 columns view
added ball figures

@@
 
-\begin{figure}[h]
+\begin{figure}
 	\centering
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{keyframes_2D_a.eps} % Replace 'example-image' with the filename of your image
+	\caption{Example Image}
+	\label{fig:keyframes-2d-column}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{keyframes_2_columns.eps} % Replace 'example-image' with the filename of your image
+	\caption{Example Image}
+	\label{fig:keyframes-2-column}
+\end{figure}
+
 \add{EXAMPLE: MANIPULATOR AND MOVING NOISY BALL}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{keyframes_3D_ball.eps} % Replace 'example-image' with the filename of your image
+	\caption{Example Image}
+	\label{fig:keyframes-3d-ball}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{keyframes_2D_ball.eps} % Replace 'example-image' with the filename of your image
+	\caption{Example Image}
+	\label{fig:keyframes-2d-ball}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.75\linewidth]{keyframes_2_columns_ball.eps} % Replace 'example-image' with the filename of your image
+	\caption{Example Image}
+	\label{fig:keyframes-2-column-ball}
+\end{figure}
 

### 2024-03-13 Code/generate_figure_keyframes.m
generates figures for ball experiment

@@ -1,6 +1,6 @@
 function [f] = generate_figure_keyframes(grid, Tbase, control_points, output)
     % Define the keyframes
-    total_frames = size(output.joints_positions, 2);
+    total_frames = 80% size(output.joints_positions, 2);
     keyframes = round(linspace(1, total_frames-1, 6));
 
     % APF field vector length
@@ -16,11 +16,7 @@ function [f] = generate_figure_keyframes(grid, Tbase, control_points, output)
 
     % Loop through each keyframe
     for kf = 1:length(keyframes)
-        % Set the axes limits to ensure consistency across subplots
-        xlim([0 20]);   
-        ylim([0 20]);   
-        zlim([0 10]);
-%         axis equal;  
+
 
         % Create subplot for each keyframe
         subplot(2, 3, kf);
@@ -35,7 +31,7 @@ function [f] = generate_figure_keyframes(grid, Tbase, control_points, output)
         j = keyframes(kf);
         
         % Display grid
-        grid.showGridVol3D(grid.grid, 'floor', false, 'height', true);
+        grid.showGridVol3D(output.grids{j}, 'floor', false, 'height', true);
 %         grid.showGridVoxel(grid.grid, 'floor', false, 'height', false);
 
         
@@ -81,10 +77,19 @@ function [f] = generate_figure_keyframes(grid, Tbase, control_points, output)
 
 
     % 3D view
-    view([-180 45.0]);
+%     view([-180 45.0]);
+%     view(50,10)
+
+    % side-view
+    view(+90,0)
 
     % bird view (top-down view)
 %     view(0, 90);
+
+        xlim([5 20]);   
+        ylim([5 15]);   
+        zlim([0 10]);
+%         axis equal;  
 
     
     hold off;

### 2024-03-13 Code/generate_figure_keyframes_column.m

@@ -4,7 +4,7 @@ function [f] = generate_figure_keyframes_column(grid, Tbase, control_points, out
     keyframes = round(linspace(1, total_frames-1, 4)); % Change number of keyframes to 4
 
     % APF field vector length
-    arrow_length = 0.5; % adjust the length to your preference
+    arrow_length = 2 % adjust the length to your preference
 
     %% COLORMAP for APF vectors
     cMap = interp1(0:1,[0 1 0; 1 0 0],linspace(0,1,256));
@@ -34,9 +34,9 @@ function [f] = generate_figure_keyframes_column(grid, Tbase, control_points, out
             
             % Display grid
             if viewMode == 1
-                grid.showGridVol3D(grid.grid, 'floor', false, 'height', true);            
+                grid.showGridVol3D(output.grids{j}, 'floor', false, 'height', true);            
             else
-                grid.showGridVoxel(grid.grid, 'floor', false, 'height', false);            
+                grid.showGridVoxel(output.grids{j}, 'floor', false, 'height', false);            
             end
             
             % Display control points
@@ -71,10 +71,10 @@ function [f] = generate_figure_keyframes_column(grid, Tbase, control_points, out
             % Set the subplot title to indicate the keyframe and view
             if viewMode == 1
                 title([num2str(kf) ' - 3D']);
-                view([-180 45.0]); % 3D view
+                view(50,10); % 3D view
             else
                 title([num2str(kf) ' - Top Down']);
-                view(2); % Bird's eye view (top-down view)
+                view(90, 0); % Bird's eye view (top-down view)
             end
             
             hold off;

### 2024-03-13 Articles/IROS/root.tex
inverse kinematics control - attractive vel draft

--- 2024-03-11 Articles/IROS/root.tex
@@
 
+\section{MANIPULATOR KINEMATICS}
+
+\subsection{INVERSE KINEMATIC CONTROL}
+
+The desired movement of the end-effector is achieved by using inverse kinematics velocity control scheme with task prioritisation.
+
+\add{inverse kinematics equation with task prioritization}
+
+\subsection{END-EFFECTOR VELOCITY}
+
+Vodenje vrha manipulatorja (end-effector) je naloga z najvišjo prioriteto.
+
+\remove{Our method employs inverse kinematics approach (IK) to guide the end effector (EE) towards its target, marking a departure from Khatib's joint coordinates approach in favor of a Cartesian coordinates framework.
+This is particularly beneficial in scenarios involving redundant manipulators, where determining an optimal goal joint configuration in advance is challenging.}
+
+When calculating translational velocity, we avoid the conventional gradient of the squared distance approach, which leads to high initial velocities and subsequently slow speeds near the target.
+Our aim is a consistent velocity throughout the trajectory, with controlled deceleration near the goal.
+This is achieved by first calculating the unit vector towards the target for direction, then modulating its magnitude using a sigmoid function, specifically the arctangent function, to prevent overshooting and ensure stable approaching motion.
+
+\begin{equation} 
+	\vec{v} = \frac{\vec{x}_{EE} - \vec{x}_g}{||\vec{x}_{EE} - \vec{x}_g||} \times \frac{\arctan(k_{sigm} \; ||\vec{x}_{EE} - \vec{x}_g||) }{pi/2}
+	\label{eq: v_att}
+\end{equation}
+
+In the above equation (eq.~\ref{eq: v_att}), $\vec{v}$  represents the end effector's translational velocity towards the target, combining direction and magnitude.
+The terms $\vec{x}_{EE}$ and $\vec{x}_g$ denote the current and goal positions of the EE, respectively, in Cartesian coordinates.
+The unit vector calculation, $\frac{\vec{x}_{EE}-\vec{x}_g}{||\vec{x}_{EE}-\vec{x}_g||}$, ensures motion directed towards the target.
+Finally, the sigmoid function, particularly the arctangent component, modulates this velocity to avoid overshooting, balancing speed and precision.
+The constant $k_{sigm}$ allows us to set how close to the goal does the robot EE start slowing down.
+
+The rotational velocity error of the EE is needed for ensuring goal orientation of the EE.
+In our approach, orientations are depicted using rotation matrices.
+Specifically, \( R \) represents the current EE orientation, while \( gR \) signifies the goal EE orientation.
+The disparity between these orientations is encapsulated by the relative rotation matrix \( dR \).
+This matrix is formulated by multiplying the goal orientation matrix \( gR \) with the transpose of the current orientation matrix \( cR^{T} \).
+To ensure that it represents a pure rotation without any scaling we than normalize the so gotten matrix .
+
+\begin{equation}
+	dR = \frac{gR \cdot cR^{T}}{||gR \cdot cR^{T}||}
+	\label{eq: rot_diff_mat}
+\end{equation}
+
+The relative rotation matrix value is converted into a quaternion, which is then logarithmically transformed \remove{to represent the rotational error vectorially}.
+The components of this quaternion, excluding the real part, than form the rotational error vector \( \omega \).
+
+\begin{equation}
+	dR \mapsto dQ = a + b \, i + c \, j + d \, k
+	\label{eq: quat_mapsto}
+\end{equation}
+
+\begin{equation}
+	dQl = 2 \cdot \log(dQ) = al + bl \, i + cl \, j + dl \, k
+	\label{eq:quat_log}
+\end{equation}
+
+\alert{There needs to be an explanation why log, where is this from. Maybe a reference. cite: DMP Quaternions article Petrič, Žlajpah, Ude}
+
+\begin{equation}
+	\vec{\omega} =
+	\begin{bmatrix}
+		bl \\
+		cl \\
+		dl
+	\end{bmatrix}
+	\label{eq:rot_error_vector}
+\end{equation}
+
+To get the full velocity of the end effector (EE), \( \mathrm{\vec{v}_{ATT}} \), we combines translational and rotational velocities, which we scale using proportional gains \( k_p \) and \( k_r \).
+
+\begin{equation}
+	\mathrm{\vec{v}_{ATT}} = 
+	\begin{bmatrix}
+		k_p * \vec{v}   \\
+		k_{\omega} * \vec{\omega}
+	\end{bmatrix}
+	\label{eq:ee_velocity}
+\end{equation}
+
+Ker uporabljamo preslikavo v ničelni prostor primarne hitrosti (null space), lahko v primeru velikih hitrosti primarne naloge preostane premalo prostostnih stopenj, da bi se lahko varno izognili oviram.
+
+\label{chap:primary slowdown}
+
+To ensure the primary task doesn't overpower the secondary task, we've integrated a primary task execution slowdown.
+This mechanism can reduce the manipulator's velocity towards its primary goal, leaving more maneuverability space for the secondary tasks.
+
+\begin{equation}
+	\label{eq:slowdown}
+	\xi_{p}=
+	\frac{1}{1 + \kappa_{\text{sec}} \delta_{min}}
+\end{equation}
+
+\alert{Is this equation correct, ADD a singularity damping factor. Maybe even some variation of a changing damping.}
+
+The slowdown factor (eq.~\ref{eq:slowdown}) is influenced by the constant \( \kappa_{\text{sec}} \) and the robot's minimum distance from an obstacle \( \delta_{\text{min}} \).
+We calculate the minimal distance in the repulsive velocities phase of the algorithm, as explained in section~\ref{chap:repulsive velocity}.
+As per the equation, a large \( \delta_{\text{min}} \) minimizes the slowdown effect, allowing uninterrupted primary task execution.
+Conversely, a small \( \delta_{\text{min}} \) increases the slowdown, by making the $\xi_{p}$ factor smaller, giving the secondary task more time for corrective actions.
+
+\subsection{AVOIDANCE VELOCITIES}
+
 \section{SIMULATION RESULTS}
@@
 \add{Visualization of the repulsive field around more or less complicated obstacles.}
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.95\linewidth]{demo-field1.eps} % Replace 'example-image' with the filename of your image
-	\caption{Example Image}
-	\label{fig:example-field}
-\end{figure}
-
-\add{add label, caption, every plot has different axis scale}
-
-\comment{}{we dont have the close obstacles bottlenecks, we do have weird bottlenecks in the orthagonal directions - maybe}
-
-\subsection{Repulsive Field Interpolation}
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=1\linewidth]{non_and_interp_4_plots.eps} % Replace 'example-image' with the filename of your image
-	\caption{Example Image}
-	\label{fig:interp-experiment}
-\end{figure}
-
-\subsection{Manipulator Examples}
-
-\add{EXAMPLE: MANIPULATOR AND COLUMN}
-
-\add{EXAMPLE: MANIPULATOR AND MOVING NOISY BALL}
-
-\remove{Drone Example}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.95\linewidth]{demo-field1c.eps} % Replace 'example-image' with the filename of your image
+%	\caption{Example Image}
+%	\label{fig:example-field}
+%\end{figure}
+%
+%\add{add label, caption, every plot has different axis scale}
+%
+%\comment{}{we dont have the close obstacles bottlenecks, we do have weird bottlenecks in the orthagonal directions - maybe}
+%
+%\subsection{Repulsive Field Interpolation}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=1\linewidth]{non_and_interp_4_plots.eps} % Replace 'example-image' with the filename of your image
+%	\caption{Example Image}
+%	\label{fig:interp-experiment}
+%\end{figure}
+%
+%\subsection{Manipulator Examples}
+%
+%\add{EXAMPLE: MANIPULATOR AND COLUMN}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=1\linewidth]{keyframes_3D.eps} % Replace 'example-image' with the filename of your image
+%	\caption{Example Image}
+%	\label{fig:keyframes-3d-column}
+%\end{figure}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=1\linewidth]{keyframes_2D_a.eps} % Replace 'example-image' with the filename of your image
+%	\caption{Example Image}
+%	\label{fig:keyframes-2d-column}
+%\end{figure}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=1\linewidth]{keyframes_2_columns.eps} % Replace 'example-image' with the filename of your image
+%	\caption{Example Image}
+%	\label{fig:keyframes-2-column}
+%\end{figure}
+%
+%
+%\add{EXAMPLE: MANIPULATOR AND MOVING NOISY BALL}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=1\linewidth]{keyframes_3D_ball.eps} % Replace 'example-image' with the filename of your image
+%	\caption{Example Image}
+%	\label{fig:keyframes-3d-ball}
+%\end{figure}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=1\linewidth]{keyframes_2D_ball.eps} % Replace 'example-image' with the filename of your image
+%	\caption{Example Image}
+%	\label{fig:keyframes-2d-ball}
+%\end{figure}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.75\linewidth]{keyframes_2_columns_ball.eps} % Replace 'example-image' with the filename of your image
+%	\caption{Example Image}
+%	\label{fig:keyframes-2-column-ball}
+%\end{figure}
+%
+%\remove{Drone Example}
 

### 2024-03-13 Articles/IROS/root.tex
little bit about avoidance

@@
 
+To move the manipulator away from the obstacles in its environment, we cover the manipulator evenly with virtual points.
+Gostota virtualnih točk je odvisna od velikosti matrik.
+Če je $w$ širina matrike v smeri pravokotno na primarno smer, je dobro, če so točke medsebojno razmaknjene glede na potek segmentov za $\Delta = \frac{1 * w}{3}$.
+Tedaj se bodo posamezne matrike deloma pokrivale in dosežemo enakomerno porazdelitev točk (POI) po celotnem robotu.
+Med izvajanjem kinematične optimizacije izračunavamo izogibne / odbojne hitrosti za vsako od točk po postopku opisanem v zgornjem poglavju.
+\add{referenca poglavja} To dobimo tako, da izvedemo preslikavo vsake od točk v task space, kjer izvajamo izračun odbojne hitrosti.
+Pri tem je posamezna kartezična preslikava sestavljena iz kinematične preslikave iz baze robota do začetka segmenta na katerem se točka nahaja in dodane preslikave do izbrane točke (od začetka segmenta do dela kjer se nahaja točka).Za vsako od točk dobimo tri komponente kartezične hitrosti.
+
 \section{SIMULATION RESULTS}
@@
 %
-%\remove{Drone Example}
-
-\remove{EXAMPLE: DRONE MOVING THROUGH THE MAP OF OBSTACLES USING APF}
 

### 2024-03-14 Articles/IROS/root.tex
more progress

@@
 
-Naša metoda je posebno primerna za uporabo z senzorskimi sistemi kot so LIDAR ali globinske kamere, saj zaradi upoštevanja celotne okolice točke in ne le razdalje do najbližje točke v okolici efektivno filtriramo senzorski šum.
+\replace{Naša metoda je posebno primerna za uporabo z senzorskimi sistemi kot so LIDAR ali globinske kamere, saj zaradi upoštevanja celotne okolice točke in ne le razdalje do najbližje točke v okolici efektivno filtriramo senzorski šum.}{prestavi to v diskusijo}
 
@@
 
-\add{inverse kinematics equation with task prioritization}
+\begin{equation}
+	\dot{q} = J^+ \xi_{p} \mathrm{\vec{v}_{ATT}} + \dot{q}_{REP}
+\end{equation}
+
+The damped Moore-Penrose pseudo-inverse, \(J^+~=~J'(JJ' + \sigma I)^{-1}\), is utilized to mitigate singularity issues and improve numerical stability in inverse kinematics computations.
+$\xi_{p}$ is the primary task execution slowdown constant.
+Finally $\dot{q}_{REP}$ are the weighted sum of avoidance joint velocities.
+\remove{For redundant robots, it optimizes solution selection by minimizing the joint velocity norm, while damping is applied to limit excessive joint velocities near singular configurations, thus improving solution robustness.}
+
+\add{maybe add the middle joints component in the equation (it is not necessary, doesn't make much change in the chosen experiments)}
+
+We run the inverse kinematics algorithm once for every time step, until we reach the desired cost function or reach the selected time limit.
 
@@
 
-\alert{Is this equation correct, ADD a singularity damping factor. Maybe even some variation of a changing damping.}
-
 The slowdown factor (eq.~\ref{eq:slowdown}) is influenced by the constant \( \kappa_{\text{sec}} \) and the robot's minimum distance from an obstacle \( \delta_{\text{min}} \).
@@
 Gostota virtualnih točk je odvisna od velikosti matrik.
-Če je $w$ širina matrike v smeri pravokotno na primarno smer, je dobro, če so točke medsebojno razmaknjene glede na potek segmentov za $\Delta = \frac{1 * w}{3}$.
+Če je $w$ širina matrike v smeri pravokotno na primarno smer, je dobro, če so točke medsebojno razmaknjene glede na potek segmentov za $\Delta = \frac{w}{3}$.
 Tedaj se bodo posamezne matrike deloma pokrivale in dosežemo enakomerno porazdelitev točk (POI) po celotnem robotu.
@@
 \section{SIMULATION RESULTS}
+
+\add{describe inverse kinematics integration of velocities}
 

### 2024-03-14 Articles/IROS/root.tex
repulsive joint velocities

@@
 \begin{equation}
-	\dot{q} = J^+ \xi_{p} \mathrm{\vec{v}_{ATT}} + \dot{q}_{REP}
-\end{equation}
-
-The damped Moore-Penrose pseudo-inverse, \(J^+~=~J'(JJ' + \sigma I)^{-1}\), is utilized to mitigate singularity issues and improve numerical stability in inverse kinematics computations.
+	\dot{q} = J^+ \xi_{p} \mathrm{\vec{v}_{att}} + \dot{q}_{rep}
+\end{equation}
+
+The damped Moore-Penrose pseudo-inverse, \(J^+~=~J'(JJ' + \sigma_{ee} I)^{-1}\), is utilized to mitigate singularity issues and improve numerical stability in inverse kinematics computations.
 $\xi_{p}$ is the primary task execution slowdown constant.
-Finally $\dot{q}_{REP}$ are the weighted sum of avoidance joint velocities.
+Finally $\dot{q}_{REP}$ are the weighted sum of avoidance joint velocities, each transformed into the null space of primary velocities, as described in the chapter below.
 \remove{For redundant robots, it optimizes solution selection by minimizing the joint velocity norm, while damping is applied to limit excessive joint velocities near singular configurations, thus improving solution robustness.}
@@
 \begin{equation}
-	\mathrm{\vec{v}_{ATT}} = 
+	\mathrm{\vec{v}_{att}} = 
 	\begin{bmatrix}
@@
 Pri tem je posamezna kartezična preslikava sestavljena iz kinematične preslikave iz baze robota do začetka segmenta na katerem se točka nahaja in dodane preslikave do izbrane točke (od začetka segmenta do dela kjer se nahaja točka).Za vsako od točk dobimo tri komponente kartezične hitrosti.
+
+\add{EQUATION: transformation of position}
+
+Ko dobimo za vsako od točk odbojno hitrost (repulsive / avoidance velocity), izračunamo drugo normo vsake od hitrosti.
+Za $K$ točk, v katerih so izogibne hitrosti največje in so posledično najbližje oviri nato izračunamo sklepne hitrosti, ki povzročijo izogibanje oviri.
+
+%\begin{equation}
+%J_{d_0}^+ = N J_{d_0}' (J_{d_0} N J_{d_0}' + \sigma_{rep})^{-1};
+%\end{equation}
+
+\begin{equation}
+	J_{d_i}^{+} = N J_{d_i}' (J_{d_i} N J_{d_i}' + \sigma_{rep})^{-1}, \quad \text{for } i = 1, 2, \ldots, K;
+\end{equation}
+
+Za vsako od točk na robotu izračunamo Jacobijevo transformacijo hitrosti, ki preslika translacijske hitrosti izbrane točke na robotu iz kartezičnega prostora, v prostor sklepov na robotu, ki povezujejo bazo z izbrano točko.
+Transforamcije kotnih hitrosti v točki nas ne zanimajo.
+V resnici nas zanima samo komponenta translacijske hitrosti v smeri normale, ki kaže od najbližje točke na objektu proti izbrani točki interesa (POI) na manipulatorju.
+Naš prostor operacij se posledično skrči na eno dimenzijo in the Jacobian, which relates the joint space velocities q and the velocity in the direction of do, can be calculated as  $J_{d_i} = \vec{n}_i^T J_i$.
+\add{citat:žlajpah} Pri čemer je vektor normale želene avoidance velocity kar enotski vektor hitrosti, ki ga dobimo s pomočjo naših matrik $\vec{n}_{i} = \frac{\vec{v}_{poi i}}{||\vec{v}_{poi i}||}$.
+This method offers computational efficiency by eliminating complex matrix inversions, as the Jacobian changes from matrix into a vector dimensions $1 \times n$, where n is number of joints that are located from before the POI and simplifying repulsive velocity calculations.
+
+\begin{equation}
+	\dot{q}_{rep} = \sum_{i=1}^{K} \alpha_i J_{d_i}^{+} \left(v_i - J_{d_i} J^{+} \xi_{p} \vec{v}_{att}\right)
+\end{equation}
+
+Celotne izogibne sklepne hitrosti dobimo z obteženo vsoto izogibnih sklepnih hitrosti izbranih $K$ najbližjih točk, pri čemer upoštevamo vpliv primarne naloge gibanja vrha manipulatorja na sklepne hitrosti.
 

### 2024-03-14 Articles/IROS/root.tex
added POI transformations equation

@@
 \begin{equation}
-	S = \sum_{i,j,k} (W \odot A)_{ijk} = \sum_{i}\sum_{j}\sum_{k} w_{\Delta i \Delta j \Delta k} \times  a_{\Delta i \Delta j \Delta k}
+	S = \sum_{i,j,k} (W \odot A)_{ijk} = \sum_{i}\sum_{j}\sum_{k} w_{\Delta i \Delta j \Delta k} ~  a_{\Delta i \Delta j \Delta k}
 \end{equation}
@@
 \add{referenca poglavja} To dobimo tako, da izvedemo preslikavo vsake od točk v task space, kjer izvajamo izračun odbojne hitrosti.
-Pri tem je posamezna kartezična preslikava sestavljena iz kinematične preslikave iz baze robota do začetka segmenta na katerem se točka nahaja in dodane preslikave do izbrane točke (od začetka segmenta do dela kjer se nahaja točka).Za vsako od točk dobimo tri komponente kartezične hitrosti.
-
-\add{EQUATION: transformation of position}
+Pri tem je posamezna kartezična preslikava sestavljena iz kinematične preslikave iz baze robota do začetka segmenta na katerem se točka nahaja in dodane preslikave do izbrane točke (od začetka segmenta do dela kjer se nahaja točka).
+\(T_{\text{base} \rightarrow \text{POI}} = T_{\text{base} \rightarrow 1} \cdot T_{1 \rightarrow 2} \cdot \ldots \cdot T_{(j-1) \rightarrow j} \cdot T_{j \rightarrow \text{POI}}\) Za vsako od točk dobimo tri komponente kartezične hitrosti.
+
+\begin{equation}
+	T_{\text{base} \rightarrow \text{POI}} = T_{\text{base} \rightarrow 1} \cdot T_{1 \rightarrow 2} \cdot \ldots \cdot T_{(j-1) \rightarrow j} \cdot T_{j \rightarrow \text{POI}}	
+\end{equation}
 

### 2024-03-14 Articles/IROS/root.tex
.

@@
 Pri tem je posamezna kartezična preslikava sestavljena iz kinematične preslikave iz baze robota do začetka segmenta na katerem se točka nahaja in dodane preslikave do izbrane točke (od začetka segmenta do dela kjer se nahaja točka).
-\(T_{\text{base} \rightarrow \text{POI}} = T_{\text{base} \rightarrow 1} \cdot T_{1 \rightarrow 2} \cdot \ldots \cdot T_{(j-1) \rightarrow j} \cdot T_{j \rightarrow \text{POI}}\) Za vsako od točk dobimo tri komponente kartezične hitrosti.
-
-\begin{equation}
-	T_{\text{base} \rightarrow \text{POI}} = T_{\text{base} \rightarrow 1} \cdot T_{1 \rightarrow 2} \cdot \ldots \cdot T_{(j-1) \rightarrow j} \cdot T_{j \rightarrow \text{POI}}	
+\(T_{0 \rightarrow \text{POI}} = T_{0 \rightarrow 1} \cdot T_{1 \rightarrow 2} \cdot \ldots \cdot T_{(j-1) \rightarrow j} \cdot T_{j \rightarrow \text{POI}}\) Za vsako od točk dobimo tri komponente kartezične hitrosti.
+
+\begin{equation}
+	T_{0 \rightarrow \text{POI}} = T_{0 \rightarrow 1} \cdot T_{1 \rightarrow 2} \cdot \ldots \cdot T_{(j-1) \rightarrow j} \cdot T_{j \rightarrow \text{POI}}	
 \end{equation}

### 2024-03-14 Code/figure_vel_dist_join_jvel_vrep_stopexec.m
started creating figure of velocities, errors, ....

@@ -1,0 +1,62 @@
+figure()
+
+%% distance,velocity, angular_vel, exec_Stop, avoidance_total
+subplot(2,2,1)
+
+% distance
+
+% velocity
+
+
+% angular_vel
+
+
+% exec_Stop
+
+% avoidance_total
+
+
+%% joint velocities
+subplot(2,2,2)
+
+plot((output.q_velocities)')
+
+
+%% joint positions
+
+subplot(2,2,3)
+
+plot((output.joints_positions)')
+
+
+
+%% repulsive velocities
+
+subplot(2,2,4)
+
+% Assuming the cell array is stored in a variable called 'POI_values'
+numTimesteps = 50; % Number of timesteps to process
+numVectors = 7; % Number of velocity vectors per timestep
+
+% Preallocate a matrix to store the norms for efficiency
+velocityNorms = zeros(numTimesteps, numVectors);
+
+for t = 1:numTimesteps
+    % Extract the 3x7 matrix for the current timestep
+    currentMatrix = output.POI_values{t};
+    
+    for v = 1:numVectors
+        % Extract the velocity vector for the current point/object
+        velocityVector = currentMatrix(:, v);
+        
+        % Calculate the norm (magnitude) of the velocity vector
+        velocityNorms(t, v) = norm(velocityVector);
+    end
+end
+
+% Now, velocityNorms contains the norms of the velocity vectors for each object at each timestep
+disp(velocityNorms);
+
+
+plot(velocityNorms)
+

### 2024-03-14 Code/Full_RTConvolution_Multiple_Points.m
created 4 subplots with values

@@ -120,9 +120,10 @@ output.repulsive_field = [];
 output.POI_locations = [];
 output.POI_values = [];
 output.EE_orientation = [];
-output.goal_orientation = [];
+output.goal_orientation = [current_orient];
 manipulability_primary = 0;
 manipulability_avoidance = [];
+output.exec_slowdown = [];
 
 
 %% GET KERNELS
@@ -457,6 +458,9 @@ while (current_dist + current_orient) > goal_dist && Niter <= Nmax
     % save manipulability measurements of jacobians
     output.manipulability_primary = [output.manipulability_primary manipulability_primary];
     output.manipulability_avoidance = [output.manipulability_avoidance manipulability_avoidance];
+
+    % save exec slowdown value
+    output.exec_slowdown = [output.exec_slowdown exec_slowdown];
 
 
     % ITER COUNTER

### 2024-03-14 Code/figure_vel_dist_join_jvel_vrep_stopexec.m

@@ -1,39 +1,61 @@
+% Set the maximum step for the x-axis
+Nmax = 50; % You can adjust this value as needed
+
 figure()
 
-%% distance,velocity, angular_vel, exec_Stop, avoidance_total
+%% Subplot 1: distance, velocity, angular_vel, exec_Stop
 subplot(2,2,1)
 
 % distance
+plot(output.goal_distances, 'DisplayName', 'Position error')
+title('Primary Task')
+hold on
 
-% velocity
-
-
-% angular_vel
-
+% orientation
+plot(output.goal_orientation, 'DisplayName', 'Orientation error')
 
 % exec_Stop
+plot(output.exec_slowdown*50, 'DisplayName', 'Primary Task Slowdown')
+hold off
 
-% avoidance_total
+% Add legend
+legend('show')
 
+% Set x-axis limits
+xlim([0 Nmax])
 
-%% joint velocities
+%% Subplot 2: joint velocities (theta dot with index)
 subplot(2,2,2)
+hold on
+for i = 1:size(output.q_velocities, 1)
+    plot(output.q_velocities(i, :)', 'DisplayName', ['$\dot{\theta}_{' num2str(i) '}$'])
+end
+title('Joint Velocities')
+hold off
 
-plot((output.q_velocities)')
+% Add legend
+legend('Interpreter', 'latex', 'Location', 'east')
 
+% Set x-axis limits
+xlim([0 Nmax])
 
-%% joint positions
-
+%% Subplot 3: joint positions (theta with index)
 subplot(2,2,3)
+hold on
+for i = 1:size(output.joints_positions, 1)
+    plot(output.joints_positions(i, :)', 'DisplayName', ['$\theta_{' num2str(i) '}$'])
+end
+title('Joint Positions')
+hold off
 
-plot((output.joints_positions)')
+% Add legend
+legend('Interpreter', 'latex', 'Location', 'east')
 
+% Set x-axis limits
+xlim([0 Nmax])
 
-
-%% repulsive velocities
-
+%% Subplot 4: repulsive velocities (v_rep with index)
 subplot(2,2,4)
-
 % Assuming the cell array is stored in a variable called 'POI_values'
 numTimesteps = 50; % Number of timesteps to process
 numVectors = 7; % Number of velocity vectors per timestep
@@ -54,9 +76,16 @@ for t = 1:numTimesteps
     end
 end
 
[…]

### 2024-03-14 Articles/IROS/root.tex
.
fixed, removed duplicated hadamard equation

@@
 
-The operating environment is modelled by discrete voxels.
+The robots environment is modelled by discrete voxels.
 As the robots environment can dynamically change, we propose a method that looks at the surrounding space of the robot and calculates these direction away from all the surrounding obstacles in real time.
@@
 
-By employing the Hadamard (element-wise) product (eq.~\ref{eq:hadamard}) between the cutout segment of the obstacle grid \(A_d\) and the corresponding 3D convolutional kernel \(K_d\) for direction \(d\), we derive the resultant matrix \(C_d\), which can be expressed as:
+By employing the Hadamard (element-wise) product (eq.~\ref{eq:hadamard}) between the cutout segment of the obstacle grid \(G_d\) and the corresponding 3D convolutional kernel \(W_d\) and than summoning all of the matrix values for each of the directions $d \in \{x, y, z\}$, we derive the resultant repulsive velocities vector \( \vec{v}_i = \left[ v_{x_i} v_{y_i} v_{z_i}\right] \).
 
@@
 \begin{equation}
-	S = \sum_{i,j,k} (W \odot A)_{ijk} = \sum_{i}\sum_{j}\sum_{k} w_{\Delta i \Delta j \Delta k} ~  a_{\Delta i \Delta j \Delta k}
-\end{equation}
-
-Subsequently, the avoidance velocity in the Cartesian coordinate system is obtained (eq.~\ref{eq:summation}) by summing all the values in the resulting matrix $C_d$, which can be represented as:
-
-\begin{equation}
-	\dot{x}_{poi} = \sum_{i}\sum_{j}\sum_{k} c_{dijk} 
-	\label{eq:summation}
-\end{equation}
-
-\alert{remove the above equation and add a vector of calculated velocities equation instead}
-
-\comment{}{TODO: add the indexes for x,y,z directions}
+	v_d = \sum_{i,j,k} (G \odot W)_{ijk} = \sum_{i}\sum_{j}\sum_{k} g_{\Delta i \Delta j \Delta k} ~  w_{\Delta i \Delta j \Delta k}
+\end{equation}
+
+%\begin{equation}
+%	\dot{x}_{poi} = \sum_{i}\sum_{j}\sum_{k} c_{dijk} 
+%	\label{eq:summation}
+%\end{equation}
+%
+%\alert{remove the above equation and add a vector of calculated velocities equation instead}
+%
+%\comment{}{TODO: add the indexes for x,y,z directions}
 
@@
 \comment{}{PLOT: kernel 2D images}
-\comment{}{EQUATION: kernel generation / values equations}
 

### 2024-03-14 Code/figure_vel_dist_join_jvel_vrep_stopexec.m
added labels and titles to the figures

@@ -7,19 +7,23 @@ figure()
 subplot(2,2,1)
 
 % distance
-plot(output.goal_distances, 'DisplayName', 'Position error')
-title('Primary Task')
+plot(output.goal_distances*10, 'DisplayName', 'Position error')
+title('Primary Task', 'Interpreter', 'latex')
 hold on
 
 % orientation
 plot(output.goal_orientation, 'DisplayName', 'Orientation error')
 
 % exec_Stop
-plot(output.exec_slowdown*50, 'DisplayName', 'Primary Task Slowdown')
+plot((output.exec_slowdown)*100, 'DisplayName', 'Primary Task Slowdown')
 hold off
 
 % Add legend
-legend('show')
+legend('show', 'Interpreter', 'latex')
+
+% Add labels
+xlabel('Step', 'Interpreter', 'latex')
+ylabel('Pos. [cm], Orient. [1], Slowd. [%]', 'Interpreter', 'latex')
 
 % Set x-axis limits
 xlim([0 Nmax])
@@ -30,11 +34,15 @@ hold on
 for i = 1:size(output.q_velocities, 1)
     plot(output.q_velocities(i, :)', 'DisplayName', ['$\dot{\theta}_{' num2str(i) '}$'])
 end
-title('Joint Velocities')
+title('$\dot{\theta}$ (Joint Velocities)', 'Interpreter', 'latex')
 hold off
 
 % Add legend
 legend('Interpreter', 'latex', 'Location', 'east')
+
+% Add labels
+xlabel('Step', 'Interpreter', 'latex')
+ylabel('$\dot{\theta}$ [rad/s]', 'Interpreter', 'latex')
 
 % Set x-axis limits
 xlim([0 Nmax])
@@ -45,47 +53,36 @@ hold on
 for i = 1:size(output.joints_positions, 1)
     plot(output.joints_positions(i, :)', 'DisplayName', ['$\theta_{' num2str(i) '}$'])
 end
-title('Joint Positions')
+title('$\theta$ (Joint Positions)', 'Interpreter', 'latex')
 hold off
 
 % Add legend
-legend('Interpreter', 'latex', 'Location', 'east')
+legend('Interpreter', 'latex', 'Location', 'east', 'Interpreter', 'latex')
+
+% Add labels
+xlabel('Step', 'Interpreter', 'latex')
+ylabel('$\theta$ [rad]', 'Interpreter', 'latex')
 
[…]

### 2024-03-14 Code/figure_vel_dist_join_jvel_vrep_stopexec.m
fixed figure scale

@@ -23,7 +23,7 @@ legend('show', 'Interpreter', 'latex')
 
 % Add labels
 xlabel('Step', 'Interpreter', 'latex')
-ylabel('Pos. [cm], Orient. [1], Slowd. [%]', 'Interpreter', 'latex')
+ylabel('Pos. [dm] Orient. [1] Slowd. [\%]', 'Interpreter', 'latex')
 
 % Set x-axis limits
 xlim([0 Nmax])

### 2024-03-14 Code/figure_vel_dist_join_jvel_vrep_stopexec.m
fixed disappeared code

@@ -68,6 +68,25 @@ xlim([0 Nmax])
 
 %% Subplot 4: repulsive velocities (v_rep with index)
 subplot(2,2,4)
+% Assuming the cell array is stored in a variable called 'POI_values'
+numTimesteps = 50; % Number of timesteps to process
+numVectors = 7; % Number of velocity vectors per timestep
+
+% Preallocate a matrix to store the norms for efficiency
+velocityNorms = zeros(numTimesteps, numVectors);
+
+for t = 1:numTimesteps
+    % Extract the 3x7 matrix for the current timestep
+    currentMatrix = output.POI_values{t};
+    
+    for v = 1:numVectors
+        % Extract the velocity vector for the current point/object
+        velocityVector = currentMatrix(:, v);
+        
+        % Calculate the norm (magnitude) of the velocity vector
+        velocityNorms(t, v) = norm(velocityVector);
+    end
+end
 
 % Plotting the repulsive velocities norms with index
 hold on

### 2024-03-14 Code/Full_RTConvolution_Moving_Ball.m
added 4 plots

@@ -34,7 +34,7 @@ weights_avoidance = weights_avoidance / norm(weights_avoidance,1) / 10;
 % -----------------------------------------------------------
 
 Tstep = 0.1 % time step
-Nmax = 100 % max number of iterations
+Nmax = 75 % max number of iterations
 space_resolution = grid.resolution; % resolution of the obstacles grid
 
 % weights for different tasks
@@ -119,7 +119,8 @@ output.repulsive_field = [];
 output.POI_locations = [];
 output.POI_values = [];
 output.EE_orientation = [];
-output.goal_orientation = [];
+output.goal_orientation = [current_orient];
+output.exec_slowdown = [];
 output.grids = {};
 manipulability_primary = 0;
 manipulability_avoidance = [];
@@ -142,6 +143,7 @@ Niter = 1;
 
 % trajectory calculation loop
 while Niter <= Nmax  
+    Niter
 
     %% UPDATE BALL POSITION
     x_ball = 1.25;
@@ -482,6 +484,9 @@ while Niter <= Nmax
     % save manipulability measurements of jacobians
     output.manipulability_primary = [output.manipulability_primary manipulability_primary];
     output.manipulability_avoidance = [output.manipulability_avoidance manipulability_avoidance];
+
+    % save exec slowdown value
+    output.exec_slowdown = [output.exec_slowdown exec_slowdown];
 
     % save grid values
     output.grids{Niter} = grid.grid;

### 2024-03-14 Code/figure_vel_dist_join_jvel_vrep_stopexec.m

@@ -1,0 +1,108 @@
+% Set the maximum step for the x-axis
+Nmax = 100; % You can adjust this value as needed
+
+figure()
+
+%% Subplot 1: distance, velocity, angular_vel, exec_Stop
+subplot(2,2,1)
+
+% distance
+plot(output.goal_distances, 'DisplayName', 'Position error')
+title('Primary Task', 'Interpreter', 'latex')
+hold on
+
+% orientation
+% plot(output.goal_orientation, 'DisplayName', 'Orientation error')
+
+% exec_Stop
+% plot((output.exec_slowdown)*100, 'DisplayName', 'Primary Task Slowdown')
+hold off
+
+% Add legend
+legend('show', 'Interpreter', 'latex')
+
+% Add labels
+xlabel('Step', 'Interpreter', 'latex')
+ylabel('Pos. [m]', 'Interpreter', 'latex')
+
+% Set x-axis limits
+xlim([0 Nmax])
+
+%% Subplot 2: joint velocities (theta dot with index)
+subplot(2,2,2)
+hold on
+for i = 1:size(output.q_velocities, 1)
+    plot(output.q_velocities(i, :)', 'DisplayName', ['$\dot{\theta}_{' num2str(i) '}$'])
+end
+title('$\dot{\theta}$ (Joint Velocities)', 'Interpreter', 'latex')
+hold off
+
+% Add legend
+legend('Interpreter', 'latex', 'Location', 'east')
+
+% Add labels
+xlabel('Step', 'Interpreter', 'latex')
+ylabel('$\dot{\theta}$ [rad/s]', 'Interpreter', 'latex')
+
+% Set x-axis limits
+xlim([0 Nmax])
+
+%% Subplot 3: joint positions (theta with index)
+subplot(2,2,3)
+hold on
+for i = 1:size(output.joints_positions, 1)
+    plot(output.joints_positions(i, :)', 'DisplayName', ['$\theta_{' num2str(i) '}$'])
+end
+title('$\theta$ (Joint Positions)', 'Interpreter', 'latex')
+hold off
+
+% Add legend
+legend('Interpreter', 'latex', 'Location', 'east', 'Interpreter', 'latex')
+
+% Add labels
+xlabel('Step', 'Interpreter', 'latex')
+ylabel('$\theta$ [rad]', 'Interpreter', 'latex')
+
+% Set x-axis limits
+xlim([0 Nmax])
+
+%% Subplot 4: repulsive velocities (v_rep with index)
+subplot(2,2,4)
+
[…]

### 2024-03-15 Articles/IROS/root.tex
experiment draft description experiment 0 and some for the exp 1

@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{non_and_interp_4_plots.eps} % Replace 'example-image' with the filename of your image
+	\caption{Repulsive field generation via convolutional kernels.
+Left: Original repulsive fields in X (top) and Y (bottom) directions.
+Right: Corresponding fields using interpolation, showcasing enhanced smoothness.}
+	\label{fig:interp-experiment}
+\end{figure}
+
 \section{MANIPULATOR KINEMATICS}
@@
 	\begin{bmatrix}
-		k_p * \vec{v}   \\
+		k_pv* \vec{v}   \\
 		k_{\omega} * \vec{\omega}
@@
 
-\add{describe inverse kinematics integration of velocities}
-
 \subsection{Repulsive Field Visualization}
 
-\add{Visualization of the repulsive field around more or less complicated obstacles.}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.95\linewidth]{demo-field1c.eps} % Replace 'example-image' with the filename of your image
-%	\caption{Example Image}
-%	\label{fig:example-field}
-%\end{figure}
+Prvo predstavimo rezultirajoče se odbojno polje za 2D - dvo dimenzijski zemljevid.
+Načeloma je delovanje na treh dimenzijah enako, seveda pa je to težje vizualizirati.
+Pri tem smo uporabili našo metodo izračuna odbojnega potencialnega polja hitrosti.
+Namesto uporabe v točkah robotskem manipulatorju smo točke vzorčili preko celotnega zemljevida ovir, kot je vidno na figuri \ref{fig:example-field}.
+V vsaki od vzorčenih točk smo izvedli v članku opisani postopek pridobivanja odbojnega polja v X in Y smeri.
+Na spodnjem prikazu figure smo nato vizualizirali odbojno poljo kot vektorsko polje sestavljene iz X in Y komponente.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{demo-field1c.eps} % Replace 'example-image' with the filename of your image
+	\caption{Visualization of the potential field calculated for the entire map using the proposed approach.
+Top: Obstacle distribution via a 2D heatmap.
+Center: Induced repulsive fields along X (left) and Y (right) axes.
+Bottom: Composite vector field showcasing the resultant repulsive velocities for obstacle avoidance.}
+	\label{fig:example-field}
+\end{figure}
 %
@@
 %
-%\subsection{Repulsive Field Interpolation}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=1\linewidth]{non_and_interp_4_plots.eps} % Replace 'example-image' with the filename of your image
-%	\caption{Example Image}
-%	\label{fig:interp-experiment}
-%\end{figure}
-%
-%\subsection{Manipulator Examples}
-%
-%\add{EXAMPLE: MANIPULATOR AND COLUMN}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=1\linewidth]{keyframes_3D.eps} % Replace 'example-image' with the filename of your image
-%	\caption{Example Image}
-%	\label{fig:keyframes-3d-column}
-%\end{figure}
-%
+\subsection{Manipulator Examples}
+
+Delovanje potencialnega polja smo prikazali na dveh različnih primerih manipulatorja.
+Pri tem je ločljivost naše mreže voxlov ovir $R=10cm$.
+Uporaba interpolacije nam omogoča, da imamo relativno grobo mrežo voxlov, kar zniža spominsko zahtevnost mreže prostora, ter pospeši izračun.
+POI so na grafih prikazani s pikami na manipulatorju.
+V bližini ovir iz njih segajo vektorji, ki prikazujejo izračunane odbojno hitrost v točkah.
+
+Med izvajanjem simulacije izvajamo eulerjevo integracijo izračunanih sklepnih hitrosti, s korakom $T_{step}=0.1s$.
+
+V prvem primeru \ref{fig:keyframes-3d-column} manipulator z uporabo potencialnega polja izračunanega na predlagani način, se manipulator varno "zvije" oziroma izbere pot do točke, ki se nahaja na drugi strani stebra.
+V primeru smo izbrali $K = 7$ POI točk interesa, v katerih gledamo oddaljenost od ovir, ki so enakomerno razporejene po segmentih in sklepih od drugega sklepa robota do vrha manipulatorja, torej od prve točke naprej, kjer se je manipulator sposoben izogniti oviri.
+Izbrane konstante primarne naloge so $k_v=5$ in $k_w=20$.
+Izbrane konstante obteženosti posameznih POI so $\alpha=\frac{\left[ \frac{3}{9}, \frac{2}{9}, \frac{1}{9}, \frac{1}{9}, \frac{1}{9}, \frac{1}{9}
+ \right]}{10}$.
+Najkrajša pot vrha v točko pelje direktno skozi steber.
+Primarna naloga je, da vrh preide v želeno pozicijo in orientacijo.
+Ker je naloga vrha primarna, moramo uvesti znižanje primarne hitrosti v bližini ovire $\xi_{p}=1$, da ima sekundarna naloga dovolj prostosti in časa, da lahko izvede rekonfiguracijo za obhod stebra.
+Simulacijo izvajamo $50$ korakov.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{keyframes_3D.eps} % Replace 'example-image' with the filename of your image
+	\caption{Example Image}
+	\label{fig:keyframes-3d-column}
+\end{figure}
+
 %\begin{figure}

### 2024-03-15 Articles/IROS/root.tex
first manipulator example

@@
 Pri tem smo uporabili našo metodo izračuna odbojnega potencialnega polja hitrosti.
-Namesto uporabe v točkah robotskem manipulatorju smo točke vzorčili preko celotnega zemljevida ovir, kot je vidno na figuri \ref{fig:example-field}.
+Namesto uporabe v točkah robotskem manipulatorju smo točke vzorčili preko celotnega zemljevida ovir, kot je vidno na figur \ref{fig:example-field}.
 V vsaki od vzorčenih točk smo izvedli v članku opisani postopek pridobivanja odbojnega polja v X in Y smeri.
@@
 
-Delovanje potencialnega polja smo prikazali na dveh različnih primerih manipulatorja.
+%Delovanje potencialnega polja smo prikazali na dveh različnih primerih manipulatorja.
 Pri tem je ločljivost naše mreže voxlov ovir $R=10cm$.
 Uporaba interpolacije nam omogoča, da imamo relativno grobo mrežo voxlov, kar zniža spominsko zahtevnost mreže prostora, ter pospeši izračun.
+V primerih smo izbrali $K = 7$ POI točk interesa, v katerih gledamo oddaljenost od ovir, ki so enakomerno razporejene po segmentih in sklepih od drugega sklepa robota do vrha manipulatorja, torej od prve točke naprej, kjer se je manipulator sposoben izogniti oviri.
 POI so na grafih prikazani s pikami na manipulatorju.
@@
 
-Med izvajanjem simulacije izvajamo eulerjevo integracijo izračunanih sklepnih hitrosti, s korakom $T_{step}=0.1s$.
-
-V prvem primeru \ref{fig:keyframes-3d-column} manipulator z uporabo potencialnega polja izračunanega na predlagani način, se manipulator varno "zvije" oziroma izbere pot do točke, ki se nahaja na drugi strani stebra.
-V primeru smo izbrali $K = 7$ POI točk interesa, v katerih gledamo oddaljenost od ovir, ki so enakomerno razporejene po segmentih in sklepih od drugega sklepa robota do vrha manipulatorja, torej od prve točke naprej, kjer se je manipulator sposoben izogniti oviri.
+%Med izvajanjem simulacije izvajamo eulerjevo integracijo izračunanih sklepnih hitrosti, s korakom $T_{step}=0.1s$.
+
+The operation of the potential field is demonstrated on two different manipulator cases.
+The resolution of our voxel grid for obstacles is \( R=10cm \).
+The use of interpolation allows us to have a relatively coarse voxel grid, which reduces the memory demand of the space grid and accelerates the computation.
+We selected \( K = 7 \) Points of Interest (POI) to observe the distance from obstacles and are uniformly distributed them along the segments and joints, starting from the second joint of the robot to the tip of the manipulator.
+This distribution begins from the second joint because it is from this point onwards that the manipulator has the capability to avoid obstacles.
+Points of Interest (POIs) are indicated by dots on the manipulator in the graphs.
+Near obstacles, vectors emanate from these points, depicting the calculated repulsive velocities at the locations.
+Throughout the simulation, Euler integration of the calculated joint velocities is performed with a step size of \( T_{step} = 0.1 \) s.
+
+%V prvem primeru \ref{fig:keyframes-3d-column} manipulator z uporabo potencialnega polja izračunanega na predlagani način, se manipulator varno "zvije" oziroma izbere pot do točke, ki se nahaja na drugi strani stebra.
 Izbrane konstante primarne naloge so $k_v=5$ in $k_w=20$.
 Izbrane konstante obteženosti posameznih POI so $\alpha=\frac{\left[ \frac{3}{9}, \frac{2}{9}, \frac{1}{9}, \frac{1}{9}, \frac{1}{9}, \frac{1}{9}
- \right]}{10}$.
+% \right]}{10}$.
 Najkrajša pot vrha v točko pelje direktno skozi steber.
@@
 
+In the first case (Fig.
+\ref{fig:keyframes-3d-column}, \ref{fig:plots-3d-column}), the manipulator safely 'curls' or selects a path to a point located on the other side of a column, avoiding the obstacle with the potential field calculated by the proposed method.
+The constants chosen for the primary task are \( k_v = 5 \) and \( k_w = 20 \).
+The weighting constants for the individual POIs are \( \alpha = \frac{[ \frac{3}{9}, \frac{2}{9}, \frac{1}{9}, \frac{1}{9}, \frac{1}{9}, \frac{1}{9} ]}{10} \), where the biggest weight belongs to the point on the manipulator which is closest to the obstacle and so on.
+As the most direct path for the end-effector to the target passes straight through the column, an approach that would result in a collision, we implement a reduction of the primary speed in the vicinity of the obstacle, setting \( \xi_{p} = 1 \).
+The simulation is performed for 50 steps.
+
 \begin{figure}
@@
 	\includegraphics[width=1\linewidth]{keyframes_3D.eps} % Replace 'example-image' with the filename of your image
-	\caption{Example Image}
+	\caption{Sequential keyframes demonstrating the manipulator's path planning and column obstacle avoidance strategy.}
 	\label{fig:keyframes-3d-column}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{4plots_column.eps} % Replace 'example-image' with the filename of your image
+	\caption{Visualization of the manipulator's path around the column obstacle: (a) Primary Task errors and percentage of the primary speed after applying slowdown, (b) Joint Velocities $\dot{\theta}$, (c) Joint Positions $\theta$, and (d) Norms of Repulsive Velocities $v_{rep}$.}
+	\label{fig:plots-3d-column}
 \end{figure}

### 2024-03-15 Articles/IROS/root.tex
text rephrasing

@@
 
-Prvo predstavimo rezultirajoče se odbojno polje za 2D - dvo dimenzijski zemljevid.
+%Prvo predstavimo rezultirajoče se odbojno polje za 2D - dvo dimenzijski zemljevid.
 Načeloma je delovanje na treh dimenzijah enako, seveda pa je to težje vizualizirati.
@@
 Na spodnjem prikazu figure smo nato vizualizirali odbojno poljo kot vektorsko polje sestavljene iz X in Y komponente.
+
+We introduce the resultant repulsive field on a two-dimensional map.
+While analogous principles apply to three-dimensional spaces, visualization of more dimensions is complex.
+Utilizing our repulsive potential field velocity calculation method, points were sampled across the entire obstacle map as depicted in Figure \ref{fig:example-field}.
+The prescribed algorithm was executed at each sampled point to derive the repulsive fields along the X and Y axes, which were subsequently visualized as a composite vector field.
 

### 2024-03-15 Articles/IROS/root.tex
second draft results

@@
 \begin{equation}
-	\dot{q}_{rep} = \sum_{i=1}^{K} \alpha_i J_{d_i}^{+} \left(v_i - J_{d_i} J^{+} \xi_{p} \vec{v}_{att}\right)
+	\dot{q}_{rep} = k_r \sum_{i=1}^{K} \alpha_i J_{d_i}^{+} \left(v_i - J_{d_i} J^{+} \xi_{p} \vec{v}_{att}\right)
 \end{equation}
@@
 \ref{fig:keyframes-3d-column}, \ref{fig:plots-3d-column}), the manipulator safely 'curls' or selects a path to a point located on the other side of a column, avoiding the obstacle with the potential field calculated by the proposed method.
-The constants chosen for the primary task are \( k_v = 5 \) and \( k_w = 20 \).
+The constants chosen for the primary task are \( k_v = 5 \) and \( k_w = 1.5 \), for the avoidance task \(k_r=20\).
 The weighting constants for the individual POIs are \( \alpha = \frac{[ \frac{3}{9}, \frac{2}{9}, \frac{1}{9}, \frac{1}{9}, \frac{1}{9}, \frac{1}{9} ]}{10} \), where the biggest weight belongs to the point on the manipulator which is closest to the obstacle and so on.
@@
 
+%V drugem primeru (Fig.
+\ref{fig:keyframes-3d-ball}, \ref{fig:plots-ball}) se naš robot giblje v dinamične okolju.
+Predlagan način izračuna odbojnih hitrosti je dobra izbira v primeru, ko potrebujemo lokalno optimizacijsko metodo za upoštevanje dinamičnih spremeb.
+V izbranem primeru imamo žogo, ki se med potekom optimizacije enakomerno premika iz leve strani grafa proti desni strani (torej v smeri y osi od 0.4 v trenutku 0 do 1.4 v zadnjem koraku simulacije).
+Primarna naloga ima sedaj nalogo vzdrževati vrh robota v eni točki, pri čemer smo za večjo sposobnost izogibanja oviri zanemarili izbiro orientacije vrha EE.
+To nam poveča sposobnost gibanja, posledično ne potrebujemo primary speed task slowdown  \( \xi_{p} = 0 \).
+The constants chosen for the primary task are \( k_v = 2 \) and \( k_w = 0 \), for the avoidance task \(k_r=3\).
+The weighting constants for the individual POIs are \( \alpha = \frac{[ \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5} ]}{10} \).
+
+In the second scenario (Fig.
+\ref{fig:keyframes-3d-ball}, \ref{fig:plots-ball}), our robot operates within a dynamic environment.
+The proposed method for calculating repulsive velocities proves to be a suitable choice when a local optimization approach is needed to accommodate dynamic changes.
+In this scenario, a ball moves consistently from the left to the right side of the graph (i.e., along the y-axis from 0.4 at the initial moment to 1.4 in the final step of the simulation), as governed by the equation \( y_{\text{ball}} = 0.4 + (1.4 - 0.4) \times \frac{N_{\text{step}}}{75} \),  with \( x_{\text{ball}} = 1.25 \) and \( z_{\text{ball}} = 0.7 \).
+The primary task is now to maintain the robot's end-effector (EE) at a fixed point, disregarding EE orientation to enhance obstacle avoidance capability.
+This increases maneuverability, thereby eliminating the need for primary speed task slowdown (\( \xi_{p} = 0 \)).
+Constants selected for the primary task are \( k_v = 2 \) and \( k_w = 0 \), and for the avoidance task \( k_r = 3 \).
+Weighting constants for the individual POIs are \( \alpha = \frac{[ \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5} ]}{10} \).
+The simulation is performed for 75 steps.
+
 %\begin{figure}
@@
 %
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=1\linewidth]{keyframes_3D_ball.eps} % Replace 'example-image' with the filename of your image
-%	\caption{Example Image}
-%	\label{fig:keyframes-3d-ball}
-%\end{figure}
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{keyframes_2D_ball.eps} % Replace 'example-image' with the filename of your image
+	\caption{Dynamic obstacle avoidance scenario: Sequential keyframes illustrating the robot’s maneuvering in response to a progressively moving ball from left to right across the workspace.}
+	\label{fig:keyframes-3d-ball}
+\end{figure}
 %
@@
 %
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.75\linewidth]{keyframes_2_columns_ball.eps} % Replace 'example-image' with the filename of your image
-%	\caption{Example Image}
-%	\label{fig:keyframes-2-column-ball}
-%\end{figure}
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{4plots_ball.eps} % Replace 'example-image' with the filename of your image
+	\caption{Performance in a dynamic obstacle avoidance scenario: (a) Primary Task error (b) Joint Velocities $\dot{\theta}$, (c) Joint Positions $\theta$, and (d) Norms of Repulsive Velocities $v_{rep}$.}
+	\label{fig:plots-ball}
+\end{figure}
 %
@@
 \section{CONCLUSION}
+
+\add{TODO}
 

### 2024-03-15 Articles/IROS/root.tex
rewritten introduction

@@
 
-\section{RELATED WORKS}
-
-\section{REPULSIVE FIELD CALCULATION}
-
-The robots environment is modelled by discrete voxels.
+%The robots environment is modelled by discrete voxels.
 As the robots environment can dynamically change, we propose a method that looks at the surrounding space of the robot and calculates these direction away from all the surrounding obstacles in real time.
 We only look in a predefined area / perimeter arround the robot.
-
-In 3D computer graphics, a voxel represents a value on a regular grid in three-dimensional space.
+%
+%In 3D computer graphics, a voxel represents a value on a regular grid in three-dimensional space.
 Each of the voxels holds the probability value of its occupation.
@@
 
+Robots need to quickly adapt to changes in dynamic environment, especially when obstacles appear.
+Our method helps a robot calculate safe directions away from obstacles by looking at the space around it in real-time.
+The focus is on a set area around the robot.
+
+In 3D graphics, a voxel is like a tiny cube in a grid that makes up the space.
+We use voxels to map the workspace of the robot, each voxel has a value that tells us if it’s probably empty or occupied.
+A zero means empty, and a one means definitely something is there, with values in between showing varying levels of certainty.
+
 Mreža voxels je lahko predefinirana, glede na model / 3d zemljevid prostora.
@@
 
-In our method, we compute repulsive velocities within the task space using a novel matrix kernel multiplication approach.
+The grid of voxels can be set in advance to match the layout of the space.
+We can also change the values in the voxels if we know where other objects or agents are moving.
+Besides pre-setting, we can use sensors to detect what's around.
+
+Our method calculates repulsive velocities, which are vectors that guide the robot on where not to go by pointing away from potential collisions.
+These vectors gain strength as the robot approaches an obstacle, signaling an urgent need to maneuver to safety.
+Conversely, when the robot is positioned at an equal distance from obstacles in the vicinity—safely away from all potential collisions—the repulsive velocities weaken because there’s no immediate need to move.
+
+%In our method, we compute repulsive velocities within the task space using a novel matrix kernel multiplication approach.
 Concentrating on the task space is advantageous as it provides a more direct and realistic representation of the environment.
 
-\replace{Naša metoda je posebno primerna za uporabo z senzorskimi sistemi kot so LIDAR ali globinske kamere, saj zaradi upoštevanja celotne okolice točke in ne le razdalje do najbližje točke v okolici efektivno filtriramo senzorski šum.}{prestavi to v diskusijo}
+%\replace{Naša metoda je posebno primerna za uporabo z senzorskimi sistemi kot so LIDAR ali globinske kamere, saj zaradi upoštevanja celotne okolice točke in ne le razdalje do najbližje točke v okolici efektivno filtriramo senzorski šum.}{prestavi to v diskusijo}
 
@@
 That is, it is in the best local minima away from all the obstacles.
+
+\section{RELATED WORKS}
+
+\add{ADD}
+
+\section{REPULSIVE FIELD CALCULATION}
 

### 2024-03-15 Articles/IROS/root.tex
improved introduction

@@
 
-Robots need to quickly adapt to changes in dynamic environment, especially when obstacles appear.
-Our method helps a robot calculate safe directions away from obstacles by looking at the space around it in real-time.
-The focus is on a set area around the robot.
-
-In 3D graphics, a voxel is like a tiny cube in a grid that makes up the space.
-We use voxels to map the workspace of the robot, each voxel has a value that tells us if it’s probably empty or occupied.
+In dynamic environments, a robot manipulator must swiftly respond to emerging obstacles.
+Our approach enables the robot to determine safe movement directions and velocities by real-time assessment of nearby space for obstacles.
+
+In 3D graphics, a voxel is like a tiny cube in a grid that makes up the 3D space.
+We use voxels to represent the workspace of the robot, each voxel has a value that tells us if it’s probabilistically empty or occupied.
 A zero means empty, and a one means definitely something is there, with values in between showing varying levels of certainty.
 
-Mreža voxels je lahko predefinirana, glede na model / 3d zemljevid prostora.
+%Mreža voxels je lahko predefinirana, glede na model / 3d zemljevid prostora.
 Zasedenost voxlov lahko spreminjamo glede na poznane pozicije in trajektorije preostalih agentov v prostoru.
@@
 
-The grid of voxels can be set in advance to match the layout of the space.
+The occupancy grid of voxels can be set in advance to match the layout of the space.
 We can also change the values in the voxels if we know where other objects or agents are moving.
-Besides pre-setting, we can use sensors to detect what's around.
-
-Our method calculates repulsive velocities, which are vectors that guide the robot on where not to go by pointing away from potential collisions.
+Besides pre-setting, we can use depth sensors to detect what's around.
+
+Our method calculates repulsive velocities from existing occupancy voxel grid, which are vectors that guide the robot on where to go by pointing away from potential collisions.
 These vectors gain strength as the robot approaches an obstacle, signaling an urgent need to maneuver to safety.
@@
 As the repulsive field calculation is locally based, it will also go to zero when the agent is surrounded by all directions, equally spaced from all sides.
-That is, it is in the best local minima away from all the obstacles.
+That is, it is in the best local minima away from all the obstacles in its proximity.
 

### 2024-03-15 Articles/IROS/root.tex
work on repulsive field calculation description

@@
 
-\subsection{MAPPING}
-
-Since the obstacle space is discrete (has finite resolution), while the Cartesian space is continuous, we propose two methods for mapping from Cartesian space to the occupancy grid space.
-The simpler approach involves mapping the point directly to the center of the nearest occupancy grid voxel, based on Euclidean distance.\add{is this clear, do i need equation?} However, this discretization can sometimes lead to discontinuities.
+This section introduces our novel approach for calculating repulsive fields, a critical component in enabling robots to detect and avoid obstacles in real-time.
+By leveraging a voxel-based representation of the surrounding space, our method dynamically assesses potential collision threats and calculates directional repulsive velocities.
+These velocities are essential for guiding the robot away from obstacles, ensuring smooth and safe navigation through complex environments.
+Our innovative use of kernel convolution methods allows for the precise calculation of avoidance velocities, providing a robust solution for obstacle avoidance that can be integrated seamlessly into robotic control systems.
+The following subsections delve into the technical details of our approach, including the mapping of Cartesian space to a discrete occupancy grid, kernel selection for optimized field calculation, and the application of tri-linear interpolation for smooth velocity transitions.
+
+%\subsection{MAPPING}
+
+%The occupancy map representation of the robots environment is discrete (has finite resolution), while the Cartesian space is continuous, as rounding the point directly to the center of the nearest occupancy grid voxel, based on Euclidean distance.\add{is this clear, do i need equation?} However, this discretization can sometimes lead to discontinuities.
 Therefore, we propose a second approach: tri-linear interpolation of the calculated repulsive field to achieve a continuous repulsive field value.
+
+The occupancy map representation of the robots environment is discrete (has finite resolution), while the Cartesian space is continuous.
+In the transition from Cartesian space coordinates to discrete occupancy grid representations, approximating or rounding can induce discontinuities in the resulting repulsive field.
+To mitigate this effect, we employ tri-linear interpolation, enabling a seamless transition to a continuous repulsive field value.
 
@@
 
-\subsection{KERNELS SELECTION}
+\subsection{KERNELS AND THEIR MATRIX APPLICATIONS}
+%\subsection{Selecting Kernels and Matrix Operations}
 
 The fundamental concept of our directional kernels lies in computing the repulsive field individually for each direction within the Cartesian coordinate system.
-\remove{Our filters structure was inspired by the Sobel  operator, a 2D convolutional filter frequently utilized in computer vision for calculating image gradients at specific points.}
+
+%\remove{Our filters structure was inspired by the Sobel  operator, a 2D convolutional filter frequently utilized in computer vision for calculating image gradients at specific points.}
 
@@
 
-We propose two different primary axis weights distributions, of course there is no reason why any other distribution of weights could not be used.
+%We propose two different primary axis weights distributions, of course there is no reason why any other distribution of weights could not be used.
 The choice of the weights distribution should depend on the profile of the repulsive velocities we want to archieve for the APF.
-
-The first of the proposed functions is a mirrored normal / gaussian distribution. By changing the sigma we can control how fast or slow does the field value grow when we approach obstacles.
-
-$\Delta i = c_i - i$, $\Delta j = c_j - j$ in $\Delta k = c_k - k$ predstavljajo število celic odmika od centralnega polja matrike, v katerem se nahaja naša točka na agentu v posamezno koordinatno smer.
+%
+%The first of the proposed functions is a mirrored normal / gaussian distribution. By changing the sigma we can control how fast or slow does the field value grow when we approach obstacles.
+%
+%$\Delta i = c_i - i$, $\Delta j = c_j - j$ in $\Delta k = c_k - k$ predstavljajo število celic odmika od centralnega polja matrike, v katerem se nahaja naša točka na agentu v posamezno koordinatno smer.
+
+We introduce two distinct distributions for the primary axis weights, although other distributions could also be applicable.
+The selection of weight distribution is contingent on the desired profile of repulsive velocities within the Artificial Potential Field (APF).
+
+The inaugural proposed function adheres to a mirrored normal or Gaussian distribution, wherein adjusting the standard deviation (\(\sigma\)) modulates the rate at which field values escalate as obstacles are approached.
+
+The terms \(\Delta i = c_i - i\), \(\Delta j = c_j - j\), and \(\Delta k = c_k - k\) denote the displacement count from the matrix's central field, where the field components are calculated, in their respective coordinate directions.
 
@@
 
-Another distribution we used is mirrored linear, where the $l=\lfloor width / 2 \rfloor$ is the rounded down half length of the primary axis kernel.
+Another distribution we used is mirrored linear, where the $\Delta i_{max} = \lfloor \frac{\mathrm{range}}{\Delta \mathrm{R}} \rfloor$ is the rounded down half length of the primary axis kernel.
 
@@
 	\begin{cases} 
-	 	\frac{l - \Delta i}{l} & \text{if } \Delta i > 0 \\
+	 	\frac{\Delta i_{max} - \Delta i}{\Delta i_{max}} & \text{if } \Delta i > 0 \\
 		0 & \text{if } \Delta i = 0 \\
-		\frac{l + \Delta i}{l}& \text{if } \Delta i < 0 
+		\frac{\Delta i_{max} + \Delta i}{\Delta i_{max}}& \text{if } \Delta i < 0 
 	\end{cases}
 \end{equation}
+
+The length $i_{max}$ of the primary axis is critical, as it dictates the detection range for obstacles.
+Longer kernels can detect obstacles further away from the robot, essentially extending the ’safety zone’ around the robot.
+If the primary axis is too long, it can lead to extra calculations and may cause the robot to unnecessarily avoid obstacles that aren't in its immediate path, making its movement and path planning less efficient.
 
@@
 However as we want bigger repulsive field when the obstacle is head on in the direction than when the obstacle is off the cardinal direction, we propose the following multiplicator, to account for the off direction obstacles.
-
-The length of the primary axis is critical, as it dictates the detection range for obstacles. Longer kernels can detect obstacles further away
-from the robot, essentially extending the ’safety zone’ around the robot.
-If the primary axis is too long, it can lead to extra calculations and may cause the robot to unnecessarily avoid obstacles that aren't in its immediate path, making its movement and path planning less efficient.
-A kernel with a primary axis that is too short might restrict the robot's ability to maneuver, detecting obstacles potentially too late, compromising the robots capacity to avoid obstacles effectively (eq.
-\ref{eq: detect range}).
-
-\begin{equation}
-	\label{eq: detect range}
-	num_{primary} = \frac{2 \times \mathrm{range}}{\Delta \mathrm{grid}}\\
-\end{equation}
-
-\add{mogoče dodaj še kak stavek o izbiri dimenzij matrik}
 
@@
 
-The length of the orthogonal axes influences the peripheral detection range for obstacles.
+The length $\Delta j_{max} = \lfloor \frac{\mathrm{width}}{2 \Delta \mathrm{R}} \rfloor$ of the orthogonal axes influences the peripheral detection range for obstacles.
 Excessively wide kernels may generate repulsive velocities for objects that are not in the path of the robot, whereas too narrow kernels might only detect obstacles aligned directly with the Cartesian direction in the point of interest.
-When selecting the width and height of the kernel, we must consider the density of the neighboring points of interest on the agent, ensuring that the \replace{collective fields}{combination of kernels} adequately cover the entire agent's surrounding area.
+When selecting the width and height of the kernel, we must consider the density of the neighboring points of interest on the agent, ensuring that the combination of kernels adequately cover the entire agent's surrounding area.
+
+For point-based agents, such as quadcopters, it is practical to align the dimensions of each matrix—length, width, and height—to be identical.
+This configuration ensures uniform coverage of the spherical area surrounding the agent and reduces the risk of blind spots that might occur when obstacles are close to the agent but situated between the discrete kernels.
 
 For smooth transitions when approaching the obstacles we propose the following sinus based function for the orthagonal axis distributions.
-The proposed equation is the same for both of the orthagonal matrix directions / axis.
-That is of course while operating with axis the $r=\lfloor width / 2 \rfloor$ is the rounded down half length of the selected orthagonal axis kernel.
+The proposed equation is the same for both of the orthagonal matrix axis.
 
@@
 	\begin{cases} 
-		sin(\left| \Delta j \right| \pi / (2 r)) & \text{if } \Delta j < 0 \text{~or }  \Delta j > 0 \\
+		sin(\left| \Delta j \right| \pi / (2 \Delta j_{max})) & \text{if } \Delta j < 0 \text{~or }  \Delta j > 0 \\
 		1 & \text{if } \Delta j = 0 \\
@@
 	\begin{cases} 
-		\frac{l - \Delta j }{l} & \text{if } \Delta j < 0 \text{~or }  \Delta j > 0 \\
+		\frac{\Delta j_{max} - \Delta j }{\Delta j_{max}} & \text{if } \Delta j < 0 \text{~or }  \Delta j > 0 \\
 		1 & \text{if } \Delta j = 0 \\
@@
 
-Če imamo opravka s točkastim agentom, kot je recimo štirikopet - dron, potem je dobra oblika posameznih matrik enaka dimenzija dolžine, širine in višine.
-Tako dobimo enakomerno pokritost krogle prostora v okolici agenta in preprečimo mrtve kote, ki se sicer lahko pojavijo, ko se ovira nahaja v bližini agenta, vendar med posameznimi jedri.
 \add{ADD:primer,slika}
@@
 The simulation is performed for 75 steps.
+
+\add{describe weights and distributions used in the kernels}
 

### 2024-03-15 Articles/IROS/root.tex
rewritten KERNELS AND THEIR MATRIX APPLICATIONS

@@
 
-The occupancy map representation of the robots environment is discrete (has finite resolution), while the Cartesian space is continuous.
-In the transition from Cartesian space coordinates to discrete occupancy grid representations, approximating or rounding can induce discontinuities in the resulting repulsive field.
-To mitigate this effect, we employ tri-linear interpolation, enabling a seamless transition to a continuous repulsive field value.
-
-Once the mapping of these points to the voxel grid is completed, we proceed to employ a specialized kernel convolution method.
-This method is tailored to calculate the components of the avoidance velocity vector in the Cartesian space, we generate the corresponding kernel and extract a segment of the obstacle grid of matching size, centered at agent or the point of interest (POI), resulting in two same size 3D matrices—one is a "window" from the obstacle grid $A_d$ and the other representing the kernel $K_d$.
-
-If the agent is located near the edge of our known voxel grid we can set the eleemnts of the window would be located in the space beyond the matrix as empty in which case the robot might want to move towards this space or as occupied, which will prevent the robot from moving out of the known grid space.
-
-By employing the Hadamard (element-wise) product (eq.~\ref{eq:hadamard}) between the cutout segment of the obstacle grid \(G_d\) and the corresponding 3D convolutional kernel \(W_d\) and than summoning all of the matrix values for each of the directions $d \in \{x, y, z\}$, we derive the resultant repulsive velocities vector \( \vec{v}_i = \left[ v_{x_i} v_{y_i} v_{z_i}\right] \).
-
-%\begin{equation}
-%	C_d = A_d \odot K_d
-%	\label{eq:hadamard}
-%\end{equation}
-
-\begin{equation}
-	v_d = \sum_{i,j,k} (G \odot W)_{ijk} = \sum_{i}\sum_{j}\sum_{k} g_{\Delta i \Delta j \Delta k} ~  w_{\Delta i \Delta j \Delta k}
-\end{equation}
-
-%\begin{equation}
-%	\dot{x}_{poi} = \sum_{i}\sum_{j}\sum_{k} c_{dijk} 
-%	\label{eq:summation}
-%\end{equation}
-%
-%\alert{remove the above equation and add a vector of calculated velocities equation instead}
-%
-%\comment{}{TODO: add the indexes for x,y,z directions}
-
 \subsection{KERNELS AND THEIR MATRIX APPLICATIONS}
@@
 
-\add{ADD:primer,slika}
+%Once the weights of the repulsive kernels have been generated, we can use kernel multiplication to get the repulsive velocity components.
+This method is tailored to calculate the components of the avoidance velocity vector in the Cartesian space, we generate the corresponding kernel and extract a segment of the obstacle grid of matching size, centered at agent or the point of interest (POI), resulting in two same size 3D matrices—one is a "window" from the obstacle grid $A_d$ and the other representing the kernel $K_d$.
+
+Following the establishment of repulsive kernel weights, kernel multiplication is employed to ascertain the repulsive velocity components within Cartesian space.
+We select a corresponding segment from the obstacle grid, aligning in size and centered around the point of interest (POI) on the robot, where we want to calculate the repulsive velocities.
+This results in two identically sized 3D matrices: one serving as a "window" into the obstacle grid (\(A_d\)) and the other as the representative kernel (\(K_d\)).
+
+%If the agent is located near the edge of our known voxel grid we can set the elements of the window would be located in the space beyond the matrix as empty in which case the robot might want to move towards this space or as occupied, which will prevent the robot from moving out of the known grid space.
+
+Positioning of the agent at the edge of the known voxel grid necessitates a strategic decision regarding the window's boundary elements.
+These elements, located beyond the matrix's spatial domain, can either be considered vacant—encouraging the robot to explore these new areas—or occupied—thereby deterring the robot from exiting the mapped grid space.
+
+%By employing the Hadamard (element-wise) product (eq.~\ref{eq:matrix-product}) between the cutout segment of the obstacle grid \(G_d\) and the corresponding 3D convolutional kernel \(W_d\) and than summoning all of the matrix values for each of the directions $d \in \{x, y, z\}$, we derive the resultant repulsive velocities vector \( \vec{v}_i = \left[ v_{x_i} v_{y_i} v_{z_i}\right] \).
+
+Applying the Hadamard (element-wise) product to the segment of the obstacle grid \(G_d\) and the kernel \(W_d\) results in a new matrix for each Cartesian direction (\(d \in \{x, y, z\}\)).
+By summing over all values in this resultant matrix, we calculate the repulsive velocity component for that specific direction.
+This process is independently conducted for each Cartesian coordinate, ultimately yielding the comprehensive repulsive velocity vector \(\vec{v}_i = [ v_{x_i}, v_{y_i}, v_{z_i}]\).
+
+\begin{equation}
+	v_d = \sum_{i,j,k} (G \odot W)_{ijk} = \sum_{i}\sum_{j}\sum_{k} g_{\Delta i \Delta j \Delta k} ~  w_{\Delta i \Delta j \Delta k}
+	\label{eq:matrix-product}
+\end{equation}
+
+%\begin{equation}
+%	\dot{x}_{poi} = \sum_{i}\sum_{j}\sum_{k} c_{dijk} 
+%	\label{eq:summation}
+%\end{equation}
+%
 
@@
 \subsection{3D INTERPOLATION}
+
+The occupancy map representation of the robots environment is discrete (has finite resolution), while the Cartesian space is continuous.
+In the transition from Cartesian space coordinates to discrete occupancy grid representations, approximating or rounding can induce discontinuities in the resulting repulsive field.
+To mitigate this effect, we employ tri-linear interpolation, enabling a seamless transition to a continuous repulsive field value.
+
+\remove{---}
 

### 2024-03-17 Articles/IROS/root.tex
edited interpolation description

@@
 
-The occupancy map representation of the robots environment is discrete (has finite resolution), while the Cartesian space is continuous.
+%The occupancy map representation of the robots environment is discrete (has finite resolution), while the Cartesian space is continuous.
 In the transition from Cartesian space coordinates to discrete occupancy grid representations, approximating or rounding can induce discontinuities in the resulting repulsive field.
-To mitigate this effect, we employ tri-linear interpolation, enabling a seamless transition to a continuous repulsive field value.
-
-\remove{---}
-
-It is essential that the velocity contributions affecting the robot change smoothly.
+To mitigate this effect, we employ tri-linear interpolation, enabling a seamless transition to a continuous repulsive field value (Fig.
+\ref{fig:interp-experiment}).
+%
+%It is essential that the velocity contributions affecting the robot change smoothly.
 However, since our obstacle grid is discretely defined, achieving perfect continuity can be challenging.
-Increasing the resolution of the obstacle field can theoretically bring us closer to continuous behavior, but in practice, we are constrained by finite resolution.
+Increasing the resolution of the obstacle field can theoretically bring us closer to continuous behavior, but in practice, we are constrained by finite resolution of our sensors and of the grid.
+Poleg tega bo višja resolucija posledično povzročila povečanje spomina potrebnega za shranjevanje zemljevida / mreže okolja in potrebne procesorske moči.
 To ensure that the velocity remains continuous when transitioning from one cell of the obstacle grid to another at a point of interest (POI), we employ trilinear interpolation.
-\remove{This technique allows for a smooth and continuous linear approximation of velocities in all three Cartesian directions (x, y, and z) as the POI moves between cells.}
+
+The occupancy grid's representation of the robot's environment is inherently discrete, possessing finite resolution, in contrast to the continuous nature of Cartesian space.
+The approximation or rounding necessary to transition from Cartesian coordinates to discrete grid representations may introduce discontinuities in the calculated repulsive field.
+To counteract this, tri-linear interpolation is utilized, facilitating a smooth transition to a continuous repulsive field value, as illustrated in Fig.
+\ref{fig:interp-experiment}.
+
+Maintaining a smooth change in velocity contributions to the robot is crucial.
+However, the discrete nature of our obstacle grid presents a challenge in achieving complete continuity.
+While enhancing the obstacle grid's resolution could theoretically yield a more continuous behavior, practical limitations arise due to the finite resolution of sensors and the grid representation itself.
+Moreover, an increase in resolution would inevitably require more memory for storing the occupancy grid and additional computational resources.
+Thus, to ensure velocity continuity across adjacent cells within the obstacle grid for a point of interest (POI), trilinear interpolation is employed.
 
@@
 	\label{eq:scale grid}
-	\vec{P} = \vec{p}_{\mathrm{POI}} \times\Delta \mathrm{grid}
+	\vec{P} = \vec{p}_{\mathrm{POI}} \times\Delta \mathrm{R}
 \end{equation}
@@
 Before we can start with the interpolation we need to calculate the distance between POI and smaller coordinates of the cells where we calculated the repulsive velocities (eq.~\ref{eq: deltas interp}).
-\replace{Since the repulsive values we calculate for the cells are alligned with the centers of the cells, we need to move before the interpolation the positions of known grid points by half of the cell width.}{The calculated repulsive velocity values are located at the centers of the cells.
+The calculated repulsive velocity values are located at the centers of the cells.
 Therefore, before interpolation, we shift the values of the cells coordinates by half the resolution of the obstacle grid for each direction.
-}
+That is, the cell coordinates are same as x,y,z, cell index and a half multiplied by grid resolution $\Delta \mathrm{R}$.
 
@@
 	\begin{bmatrix}
-		\frac{\left( P_x - \left( X(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( X(2) - X(1) \right)} \\
-		\frac{\left( P_y - \left( Y(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Y(2) - Y(1) \right)} \\
-		\frac{\left( P_z - \left( Z(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Z(2) - Z(1) \right)} \\
+		\frac{\left( P_x - \left( X(1) + \frac{1}{2} \Delta \mathrm{R} \right)  \right)}{\left( X(2) - X(1) \right)} \\
+		\frac{\left( P_y - \left( Y(1) + \frac{1}{2} \Delta \mathrm{R} \right)  \right)}{\left( Y(2) - Y(1) \right)} \\
+		\frac{\left( P_z - \left( Z(1) + \frac{1}{2} \Delta \mathrm{R} \right)  \right)}{\left( Z(2) - Z(1) \right)} \\
 	\end{bmatrix}
@@
 
-The result of the interpolation is independent of the order of the operations. We first interpolate along the x-axis, followed by along the y-axis and finally along z-axis.
+The result of the interpolation is agnostic of the order of the operations. We first interpolate along the x-axis, followed by along the y-axis and finally along z-axis.
 
@@
 
-The final result is a repulsive velocity vector that transitions smoothly between the discrete values calculated at distinct points in the obstacle grid.
+%The final result is a smooth repulsive velocity field that transitions smoothly between the discrete values calculated at distinct points in the obstacle grid.
+
+The end product is a uniformly smooth repulsive velocity field, facilitating continuous transitions across discrete points determined within the obstacle grid.
 

### 2024-03-17 Articles/IROS/root.tex
edited translational velocity

@@
 
+\alert{equations break column}
+
 %\begin{equation}
@@
 
-\begin{equation}
-	\dot{q} = J^+ \xi_{p} \mathrm{\vec{v}_{att}} + \dot{q}_{rep}
-\end{equation}
-
-The damped Moore-Penrose pseudo-inverse, \(J^+~=~J'(JJ' + \sigma_{ee} I)^{-1}\), is utilized to mitigate singularity issues and improve numerical stability in inverse kinematics computations.
+%\begin{equation}
+%	\dot{\vec{q}} = J^+ \xi_{p} \mathrm{\vec{v}_{att}} + \dot{\vec{q}}_{rep}
+%\end{equation}
+
+\begin{equation}
+	\dot{\vec{q}} = \mathbf{J}^+ \xi_{p} \mathrm{\vec{v}{att}} + \dot{\vec{q}}_{rep}
+\end{equation}
+
+The damped Moore-Penrose pseudo-inverse, \(\mathbf{J}^+~=~\mathbf{J}'(\mathbf{J}\mathbf{J}' + \sigma_{ee} \mathbf{I})^{-1}\), is utilized to mitigate singularity issues and improve numerical stability in inverse kinematics computations.
 $\xi_{p}$ is the primary task execution slowdown constant.
-Finally $\dot{q}_{REP}$ are the weighted sum of avoidance joint velocities, each transformed into the null space of primary velocities, as described in the chapter below.
-\remove{For redundant robots, it optimizes solution selection by minimizing the joint velocity norm, while damping is applied to limit excessive joint velocities near singular configurations, thus improving solution robustness.}
-
-\add{maybe add the middle joints component in the equation (it is not necessary, doesn't make much change in the chosen experiments)}
+Finally $\dot{\vec{q}}_{rep}$ are the weighted sum of avoidance joint velocities, each transformed into the null space of primary velocities, as described in the chapter below.
+
+%\remove{For redundant robots, it optimizes solution selection by minimizing the joint velocity norm, while damping is applied to limit excessive joint velocities near singular configurations, thus improving solution robustness.}
+
+%\add{maybe add the middle joints component in the equation (it is not necessary, doesn't make much change in the chosen experiments)}
 
@@
 
-Vodenje vrha manipulatorja (end-effector) je naloga z najvišjo prioriteto.
-
-\remove{Our method employs inverse kinematics approach (IK) to guide the end effector (EE) towards its target, marking a departure from Khatib's joint coordinates approach in favor of a Cartesian coordinates framework.
-This is particularly beneficial in scenarios involving redundant manipulators, where determining an optimal goal joint configuration in advance is challenging.}
-
-When calculating translational velocity, we avoid the conventional gradient of the squared distance approach, which leads to high initial velocities and subsequently slow speeds near the target.
-Our aim is a consistent velocity throughout the trajectory, with controlled deceleration near the goal.
-This is achieved by first calculating the unit vector towards the target for direction, then modulating its magnitude using a sigmoid function, specifically the arctangent function, to prevent overshooting and ensure stable approaching motion.
+%Vodenje vrha manipulatorja (end-effector) je naloga z najvišjo prioriteto.
+
+Prioritizing control over the end-effector's (EE) translational velocity is essential for a consistent and stable target approach.
+The commonly used method, where translational velocity is proportional to squared distance, leads to impractically high initial velocities, which can prevent obstacle avoidance and real-world execution, followed by disproportionately slow velocities when nearing the goal position.
+Our aim is a consistent velocity profile throughout the trajectory, with controlled deceleration near the goal.
+This is achieved by first calculating the unit vector towards the target for direction, then modulating its magnitude using a sigmoid function, specifically the arctangent function (Eq.
+\ref{eq: v_att}).
 
@@
 
-In the above equation (eq.~\ref{eq: v_att}), $\vec{v}$  represents the end effector's translational velocity towards the target, combining direction and magnitude.
+%In the above equation (eq.~\ref{eq: v_att}), $\vec{v}$  represents the end effector's translational velocity towards the target, combining direction and magnitude.
 The terms $\vec{x}_{EE}$ and $\vec{x}_g$ denote the current and goal positions of the EE, respectively, in Cartesian coordinates.
@@
 Finally, the sigmoid function, particularly the arctangent component, modulates this velocity to avoid overshooting, balancing speed and precision.
-The constant $k_{sigm}$ allows us to set how close to the goal does the robot EE start slowing down.
+The EE velocity will be transformed using the sigmoid curve when near the goal.
+When the distance from goal point is high, the sigmoid will cause the velocity to get into saturation, that is upper limit.
+The constant $k_{sigm}$ allows us to set how close to the goal does the robot EE start slowing down and what is the velocity profile when approaching the goal point.
+
+In Eq.~\ref{eq: v_att}, $\vec{v}$ specifies the end-effector's (EE) translational velocity directed towards the target, integrating both its direction and magnitude.
+Here, $\vec{x}_{EE}$ represents the EE's current position, and $\vec{x}g$ is the target position, both in Cartesian coordinates.
+The calculation $\frac{\vec{x}{EE}-\vec{x}g}{||\vec{x}{EE}-\vec{x}g||}$ generates a unit vector pointing towards the target, ensuring targeted movement.
+The velocity's modulation by the arctangent sigmoid function curtails overshooting by moderating speed in proximity to the target and saturating velocity at a predefined upper limit when distant.
+The parameter \(k_{sigm}\) in the arctangent function adjusts the curve's steepness, affecting how quickly the end-effector decelerates near the target.
+A higher \(k_{sigm}\) maintains speed until closer to the target for a sharp deceleration, while a lower value starts slowing down earlier for a gradual approach.
 
@@
 
+\alert{add the limits of joint positions}
+
 %Delovanje potencialnega polja smo prikazali na dveh različnih primerih manipulatorja.
@@
 \add{describe weights and distributions used in the kernels}
+
+\add{sigmoid weights}
 

### 2024-03-17 Articles/IROS/root.tex
edited rotational velocities

@@
 \end{abstract}
+
+\add{ADD}
 
@@
 Our aim is a consistent velocity profile throughout the trajectory, with controlled deceleration near the goal.
-This is achieved by first calculating the unit vector towards the target for direction, then modulating its magnitude using a sigmoid function, specifically the arctangent function (Eq.
+This is achieved by first calculating the unit vector towards the target for direction, then modulating its magnitude using a sigmoid function, specifically the arctangent function (eq.
 \ref{eq: v_att}).
@@
 
-The rotational velocity error of the EE is needed for ensuring goal orientation of the EE.
+%The rotational error of the EE is needed for ensuring goal orientation of the EE.
 In our approach, orientations are depicted using rotation matrices.
@@
 
-\begin{equation}
-	dR = \frac{gR \cdot cR^{T}}{||gR \cdot cR^{T}||}
+%Determination of the requisite rotation adjustment utilizes the relative rotation matrix \(dR\), calculated from the multiplication of \(gR\) by the \(cR\) transpose, and subsequent normalization to ensure a pure rotational representation.
+This method guarantees that \(dR\) accurately captures the orientation discrepancy, devoid of any scaling factor.
+
+%The current orientation of the end-effector (EE) and its goal orientation are encoded through rotation matrices, \(cR\) for the current state and \(gR\) for the goal state.
+The necessary rotational adjustment is identified through the relative rotation matrix \(dR\), generated by multiplying \(gR\) with the transpose of \(cR\) and then normalizing the outcome (eq.
+\ref{eq: rot_diff_mat}).
+This process ensures \(dR\) accurately represents rotational differences, without any scaling factors.
+%
+%To enhance computational efficiency and prevent gimbal lock, the approach transitions from matrix to quaternion representations for precise and efficient rotational velocity adjustments, essential for achieving the target orientation of the EE.
+
+The current orientation of the end-effector (EE) and its goal orientation are encoded through rotation matrices, \(\mathbf{cR}\) for the current state and \(\mathbf{gR}\) for the goal state.
+The necessary rotational adjustment is identified through the relative rotation matrix \(\mathbf{dR}\), generated by multiplying \(\mathbf{gR}\) with the transpose of \(\mathbf{cR}\) and then normalizing the outcome (eq.
+\ref{eq: rot_diff_mat}).
+This process ensures \(\mathbf{dR}\) accurately represents rotational differences, without the effect of any scaling factors.
+
+%\begin{equation}
+%	dR = \frac{gR \cdot cR^{T}}{||gR \cdot cR^{T}||}
+%	\label{eq: rot_diff_mat}
+%\end{equation}
+
+%\begin{equation}
+%	\boldsymbol{dR} = \frac{\boldsymbol{gR} \cdot \boldsymbol{cR}^{T}}{\left\lVert\boldsymbol{gR} \cdot \boldsymbol{cR}^{T}\right\rVert}
+%	\label{eq: rot_diff_mat}
+%\end{equation}
+
+\begin{equation}
+	\mathbf{dR} = \frac{\mathbf{gR} \cdot \mathbf{cR}^{T}}{||\mathbf{gR} \cdot \mathbf{cR}^{T}||}
 	\label{eq: rot_diff_mat}
@@
 
-The relative rotation matrix value is converted into a quaternion, which is then logarithmically transformed \remove{to represent the rotational error vectorially}.
-The components of this quaternion, excluding the real part, than form the rotational error vector \( \omega \).
-
-\begin{equation}
-	dR \mapsto dQ = a + b \, i + c \, j + d \, k
+To enhance computational efficiency and prevent gimbal lock, the approach transitions from matrix to quaternion representations for precise and efficient rotational velocity adjustments, essential for achieving the target orientation of the EE.
+
+The relative rotation matrix is converted into a quaternion, which is subsequently logarithmically transformed.
+The resulting quaternion's components, excluding the real part, constitute the rotational error vector \( \vec{\omega} \).
+%
+%\begin{equation}
+%	dR \mapsto dQ = a + b \, i + c \, j + d \, k
+%	\label{eq: quat_mapsto}
+%\end{equation}
+%
+%\begin{equation}
+%	dQl = 2 \cdot \log(dQ) = al + bl \, i + cl \, j + dl \, k
+%	\label{eq:quat_log}
+%\end{equation}
+
+\begin{equation}
+	\boldsymbol{dR} \mapsto \boldsymbol{dQ} = a + b \, \vec{i} + c \, \vec{j} + d \, \vec{k}
 	\label{eq: quat_mapsto}
@@
 \begin{equation}
-	dQl = 2 \cdot \log(dQ) = al + bl \, i + cl \, j + dl \, k
+	\boldsymbol{dQl} = 2 \cdot \log(\boldsymbol{dQ}) = a_{log} + b_{log} \, \vec{i} + c_{log} \, \vec{j} + d_{log} \, \vec{k}
 	\label{eq:quat_log}
@@
 
-\alert{There needs to be an explanation why log, where is this from. Maybe a reference. cite: DMP Quaternions article Petrič, Žlajpah, Ude}
+\alert{There needs to be an explanation why log, cite: DMP Quaternions article Žlajpah, Ude}
 
@@
 	\begin{bmatrix}
-		bl \\
-		cl \\
-		dl
+		b_{log} \\
+		c_{log} \\
+		d_{log}
 	\end{bmatrix}
@@
 
-To get the full velocity of the end effector (EE), \( \mathrm{\vec{v}_{ATT}} \), we combines translational and rotational velocities, which we scale using proportional gains \( k_p \) and \( k_r \).
+To get the full velocity of the end effector (EE), \( \mathrm{\vec{v}_{att}} \), we combines translational and rotational velocities, which we scale using proportional gains \( k_v \) and \( k_\omega \).
 
@@
 	\begin{bmatrix}
-		k_pv* \vec{v}   \\
+		k_{v} * \vec{v}   \\
 		k_{\omega} * \vec{\omega}
@@
 
-\add{TODO}
+\add{ADD}
 

### 2024-03-18 Articles/IROS/root.tex
edited primary slowdown

@@
 
-Ker uporabljamo preslikavo v ničelni prostor primarne hitrosti (null space), lahko v primeru velikih hitrosti primarne naloge preostane premalo prostostnih stopenj, da bi se lahko varno izognili oviram.
+%Ker uporabljamo preslikavo v ničelni prostor primarne hitrosti (null space), lahko v primeru velikih hitrosti primarne naloge preostane premalo prostostnih stopenj, da bi se lahko varno izognili oviram.
+Torej sekundarna naloga izogibanja nima dovolj sposobnosti, prostosti za delovanje.
+%
+%To ensure the primary task doesn't overpower the secondary task, we've integrated a primary task execution slowdown.
+This mechanism can reduce the manipulator's velocity towards its primary goal, leaving more maneuverability space for the secondary tasks.
+%
+%\begin{equation}
+%	\label{eq:slowdown}
+%	\xi_{p}=
+%	\frac{1}{1 + \kappa_{\text{sec}} v_{\text{Rmax}}}
+%\end{equation}
+%
+%The slowdown factor (eq.~\ref{eq:slowdown}) is influenced by the constant \( \kappa_{\text{sec}} \) and the maximum repulsive velocity vectors magnitude, which depends on robot's minimum distance from an obstacle \( v_{\text{Rmax}} \).
+We calculate the minimal distance in the repulsive velocities phase of the algorithm, as explained in section~\ref{chap:repulsive velocity}.
+As per the equation, a small \( v_{\text{Rmax}} \) minimizes the slowdown effect, allowing uninterrupted primary task execution.
+Conversely, a large \( v_{\text{Rmax}} \) increases the slowdown, by making the $\xi_{p}$ factor smaller, giving the secondary task more time for corrective actions.
+
+Utilizing the mapping into the null space of the primary velocity, it becomes evident that at high velocities of the primary task, there may be an insufficient number of degrees of freedom remaining to safely navigate around obstacles.
+Consequently, the secondary task, which is aimed at obstacle avoidance, lacks the requisite capability and freedom to operate effectively.
 
@@
 
-To ensure the primary task doesn't overpower the secondary task, we've integrated a primary task execution slowdown.
-This mechanism can reduce the manipulator's velocity towards its primary goal, leaving more maneuverability space for the secondary tasks.
+To mitigate the dominance of the primary task over the secondary task, a mechanism for the deceleration of the primary task's execution has been implemented.
+This strategy effectively decreases the velocity of the manipulator towards its primary objective, thereby allocating additional maneuverability for secondary tasks.
 
@@
 	\xi_{p}=
-	\frac{1}{1 + \kappa_{\text{sec}} \delta_{min}}
-\end{equation}
-
-The slowdown factor (eq.~\ref{eq:slowdown}) is influenced by the constant \( \kappa_{\text{sec}} \) and the robot's minimum distance from an obstacle \( \delta_{\text{min}} \).
-We calculate the minimal distance in the repulsive velocities phase of the algorithm, as explained in section~\ref{chap:repulsive velocity}.
-As per the equation, a large \( \delta_{\text{min}} \) minimizes the slowdown effect, allowing uninterrupted primary task execution.
-Conversely, a small \( \delta_{\text{min}} \) increases the slowdown, by making the $\xi_{p}$ factor smaller, giving the secondary task more time for corrective actions.
+	\frac{1}{1 + \kappa_{\text{sec}} v_{\text{Rmax}}}
+\end{equation}
+
+The deceleration factor (eq.~\ref{eq:slowdown}) is determined by the constant \( \kappa_{\text{sec}} \) and the magnitude of the maximum repulsive velocity vectors, which is contingent upon the robot's minimal distance from an obstacle \( v_{\text{Rmax}} \).
+According to the equation, a small \( v_{\text{Rmax}} \) diminishes the deceleration effect, permitting the uninterrupted execution of the primary task.
+Conversely, large \( v_{\text{Rmax}} \) enhances the deceleration by reducing the $\xi_{p}$ factor, thereby affording the secondary task increased opportunity for obstacles avoidance motion.
 

### 2024-03-18 Articles/IROS/root.tex
edited POI

@@
 \section{REPULSIVE FIELD CALCULATION}
+\label{section:repulsive_vel}
 
@@
 	\xi_{p}=
-	\frac{1}{1 + \kappa_{\text{sec}} v_{\text{Rmax}}}
+	\frac{1}{1 + \kappa_{\text{sec}} ||v_{\text{Rmax}}||}
 \end{equation}
@@
 
-To move the manipulator away from the obstacles in its environment, we cover the manipulator evenly with virtual points.
+%To move the manipulator away from the obstacles in its environment, we cover the manipulator evenly with virtual detection points (POI).
 Gostota virtualnih točk je odvisna od velikosti matrik.
@@
 
-\begin{equation}
-	T_{0 \rightarrow \text{POI}} = T_{0 \rightarrow 1} \cdot T_{1 \rightarrow 2} \cdot \ldots \cdot T_{(j-1) \rightarrow j} \cdot T_{j \rightarrow \text{POI}}	
-\end{equation}
-
-Ko dobimo za vsako od točk odbojno hitrost (repulsive / avoidance velocity), izračunamo drugo normo vsake od hitrosti.
+To enable the avoidance of obstacles by the manipulator within its operating environment, the manipulator is uniformly coated with virtual detection points (POIs).
+The density of these virtual points is dependent upon the dimensions of matrices.
+Given a matrix width $w$ perpendicular to the primary direction, it is advantageous for these points to be spaced at intervals of $\Delta = \frac{w}{3}$.
+This spacing ensures partial overlap between matrices, thereby achieving a uniform distribution of POIs across the entirety of the robot.
+
+During kinematic optimization, repulsive velocities for each point are computed following the methodology described in the previous section (sect.
+\ref{section:repulsive_vel}).
+This is achieved by mapping each point into the task space, where the calculation of repulsive velocity is performed.
+The cartesian mapping for each point comprises a kinematic transformation from the robot's base to the beginning of the segment where the point is located, followed by an additional mapping to the specific point on the segment (eq.
+\ref{eq:transformations}).
+For each point, three components of cartesian velocity are obtained.
+
+%\begin{equation}
+%	T_{0 \rightarrow \text{POI}} = T_{0 \rightarrow 1} \cdot T_{1 \rightarrow 2} \cdot \ldots \cdot T_{(j-1) \rightarrow j} \cdot T_{j \rightarrow \text{POI}}	
+%	\label{eq:transformations}
+%\end{equation}
+
+\begin{equation}
+	\mathbf{T}_{0 \rightarrow \text{POI}} = \mathbf{T}_{0 \rightarrow 1} \cdot \mathbf{T}_{1 \rightarrow 2} \cdot \ldots \cdot \mathbf{T}_{(j-1) \rightarrow j} \cdot \mathbf{T}_{j \rightarrow \text{POI}}
+	\label{eq:transformations}
+\end{equation}
+
+%Ko dobimo za vsako od točk odbojno hitrost (repulsive / avoidance velocity), izračunamo drugo normo vsake od hitrosti.
 Za $K$ točk, v katerih so izogibne hitrosti največje in so posledično najbližje oviri nato izračunamo sklepne hitrosti, ki povzročijo izogibanje oviri.
 
+Upon obtaining the repulsive velocity for each point, the second norm of each velocity is calculated.
+For a number of $K$ points where the avoidance velocities are greatest, indicative of closest proximity to an obstacle, the resulting joint velocities that facilitate obstacle avoidance are computed using inverse kinematic equations for the selected points on the manipulator and transformed into the null-space velocities of the primary task.
+
 %\begin{equation}
@@
 
-\begin{equation}
-	J_{d_i}^{+} = N J_{d_i}' (J_{d_i} N J_{d_i}' + \sigma_{rep})^{-1}, \quad \text{for } i = 1, 2, \ldots, K;
-\end{equation}
-
-Za vsako od točk na robotu izračunamo Jacobijevo transformacijo hitrosti, ki preslika translacijske hitrosti izbrane točke na robotu iz kartezičnega prostora, v prostor sklepov na robotu, ki povezujejo bazo z izbrano točko.
-Transforamcije kotnih hitrosti v točki nas ne zanimajo.
+%\begin{equation}
+%	J_{d_i}^{+} = N J_{d_i}' (J_{d_i} N J_{d_i}' + \sigma_{rep} I)^{-1}, \quad \text{for } i = 1, 2, \ldots, K;
+%\end{equation}
+
+\begin{equation}
+	\mathbf{J}_{d_i}^{+} = \mathbf{N} \mathbf{J}_{d_i}' (\mathbf{J}_{d_i} \mathbf{N} \mathbf{J}_{d_i}' + \sigma_{rep} \mathbf{I})^{-1}, \quad \text{for } i = 1, 2, \ldots, K;
+\end{equation}
+
+%Za vsako od izbranih točk na robotu izračunamo Jacobijevo transformacijo hitrosti, ki preslika translacijske hitrosti izbrane točke na robotu iz kartezičnega prostora, v prostor sklepov na robotu, ki povezujejo bazo z izbrano točko.
+V izbranih točkah nas ne zanimajo transformacije napak orientacij.
 V resnici nas zanima samo komponenta translacijske hitrosti v smeri normale, ki kaže od najbližje točke na objektu proti izbrani točki interesa (POI) na manipulatorju.
-Naš prostor operacij se posledično skrči na eno dimenzijo in the Jacobian, which relates the joint space velocities q and the velocity in the direction of do, can be calculated as  $J_{d_i} = \vec{n}_i^T J_i$.
+
+%We calculate translational velocity Jacobian for every of the $K$ selected points on the robot.
+Since we are only interested in the part of velocity that leads us directly away from the obstacles, we can
+
+%
+%Naš prostor operacij se posledično skrči na eno dimenzijo in the Jacobian, which relates the joint space velocities q and the velocity in the direction of do, can be calculated as  $J_{d_i} = \vec{n}_i^T J_i$.
 \add{citat:žlajpah} Pri čemer je vektor normale želene avoidance velocity kar enotski vektor hitrosti, ki ga dobimo s pomočjo naših matrik $\vec{n}_{i} = \frac{\vec{v}_{poi i}}{||\vec{v}_{poi i}||}$.
@@
 
+We calculate the translational velocity Jacobian for each of the $K$ selected points on the robot.
+Since our interest lies exclusively in the velocity component that directs us away from obstacles, we consequently narrow our operational space to a single dimension.
+Consequently, the Jacobian, correlating joint space velocities \(q\) with the directional velocity \(d_i\), is represented as \(J_{d_i} = \vec{n}_i^T \mathbf{J}_i\).
+The normal vector, signifying the desired avoidance velocity, is essentially the unit velocity vector, derived from our matrices as \(\vec{n}_{i} = \frac{\vec{v}_{poi i}}{||\vec{v}_{poi i}||}\).
+This approach streamlines computational processes by avoiding the necessity for complex matrix inversions, transforming the Jacobian from a matrix to a vector of dimensions \(1 \times n\), where \(n\) signifies the count of joints preceding the POI, thus facilitating more straightforward calculations of repulsive velocities.
+
 \begin{equation}
@@
 
-Celotne izogibne sklepne hitrosti dobimo z obteženo vsoto izogibnih sklepnih hitrosti izbranih $K$ najbližjih točk, pri čemer upoštevamo vpliv primarne naloge gibanja vrha manipulatorja na sklepne hitrosti.
+%Celotne izogibne sklepne hitrosti dobimo z obteženo vsoto izogibnih sklepnih hitrosti izbranih $K$ najbližjih točk, pri čemer upoštevamo vpliv primarne naloge gibanja vrha manipulatorja na sklepne hitrosti.
+
+The overall avoidance joint velocities are obtained through a weighted sum of the avoidance joint velocities for the selected \(K\) nearest points, taking into account the influence of the primary task of manipulator EE movement on the joint velocities.
 

### 2024-03-18 Articles/IROS/root.tex
added related works

@@
 \title{\LARGE \bf
-Matrix Multiplication-Driven Repulsive Fields for 3D Voxel-Based TSDF Calculation
+Matrix Multiplication-Driven Repulsive Fields for Manipulator Kinematic Obstacle Avoidance
 }
@@
 If it is definitly occupied it holds the value of 1.
+
+\alert{---}
+
+\subsection{Motion Planning}
+
+% PREDSTAVITEV REDUNDANCE
+
+Kinematic redundancy~\cite{siciliano1990kinematic, siciliano2016springer} enables a manipulator to follow a predefined task space trajectory using the endeffector (EE), while simultaneously, optimising for an additional task with the remaining movement capacity without impacting the trajectory adherence.
+This is possible because the robot's degrees of freedom (DOF) go beyond what is required to perform the primary task.
+Consequently, the robot can adopt different joint configurations optimised according to the secondary task while performing the primary task.
+Common secondary tasks~\cite{siciliano2010robot} include avoiding singularities, optimising the manipulability measure, minimising joint torques and avoiding obstacles in the operating space.
+
+% PREDSTAVITEV PRISTOPOV PLANIRANJA POTI V ROBOTIKI
+The task of finding the joint trajectories of a manipulator is called motion planning~\cite{IDEASLab2023}.
+It consists of finding a sequence of joint configurations for a robot so that the robot can move along this path from its initial configuration to the goal configuration without colliding with itself, static obstacles or other agents in the environment.
+In addition to collision avoidance, motion planning for manipulators can optionally take into account various constraints, such as position, velocity, acceleration or jerk constraints for the joint angle or end effector, precision of the end effector with respect to position and orientation, stability of the manipulator, avoidance of singularities, or any number of other criteria.
+
+There are numerous methods for planning manipulator movements~\cite{c51, c52}, they can rougly be seperated into global and local approaches.
+Global, sampling-based, methods such as PRM~\cite{vsvestka1997motion}, RRT*~\cite{lavalle1998rapidly, karaman2010incremental}, RRT-Connect~\cite{kuffner2000rrt}, Informerd RRT*~\cite{gammell2014informed}, BIT*~\cite{gammell2015batch} and others offer a globally optimal solution based on a global search in configuration space.
+However, the generated trajectories are not always smooth or optimal, and the performance of the methods may be insufficient for real-time operation.
+
+%Global planning methods are typically more comprehensive, aiming to find a globally optimal trajectory for the manipulator.
+These methods often involve mapping out the entire path before movement begins, taking into account the known environment, which usually includes static obstacles.
+However, the downside of global planning methods is their computational intensity, which often renders them too slow for real-time applications.
+This limitation is particularly pronounced in dynamic or unpredictable environments where conditions change rapidly and require quick responses.
+Examples of global planning methods include graph-based algorithms like A* and Dijkstra's algorithm, which are known for their effectiveness in pathfinding but also for their computational demands [(Lavalle, 2006)].
+
+%Sampling-based methods such as PRM, RRT, RRT*, RRT-Connect, Informerd RRT*, BIT* and others offer a solution based on a global search in configuration space.
+However, the generated trajectories are not always smooth or optimal, and the performance of the methods may be insufficient for real-time operation.
+
+%\remove{Recently, a number of learning-based methods using data-driven techniques have been proposed to improve or accelerate the functionality of sampling-based methods. }
+%
+%\remove{Trajectory optimisation methods such as CHOMP, STOMP and TrajOpt, on the other hand, use optimization to improve an initial seed trajectory.
+Consequently, the optimality of the solution is highly dependent on this initial trajectory.
+Nevertheless, these methods are capable of generating smooth trajectories, and although they can be too computationally intensive for high DOF dynamic real-time environments, they are generally effective in finding constrained motion plans.
+}
+
+Local motion planning approaches employ optimization techniques, two common ones are inverse kinematics~\cite{c29,c38}, that is based on finding a least squares solution of the manipulator joint velocities, and quadratic programming (QP)~\cite{c21,c22, haviland2021neo, c23}.
+Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
+However, since they do not plan further than one step ahead, they tend to get stuck in local minima.
+Therefore, they are often combined with a higher-level planner, for global static environment based \replace{motion}{path} planning, while local optimisation takes dynamic environment changes into account.
+In the following text we will focus on inverse kinematics based approaches.
+
+\subsection{Kinematic Obstacle Avoidance}
+
+%\remove{Different control schemes have been proposed.
+Acceleration-based control excels in precise handling of motion changes, velocity-based control offers consistent and smooth movement, while force and torque-based control provides direct control over joint forces for robust physical interactions.
+(CITATIONS in COMMENT BELLOW)}
+
+Various researchers have adopted different approaches to kinematic avoidance, as detailed in~\cite{c30, c39}.
+
+Colbaugh and Glass (1989)~\cite{c31, c32} tackled this problem in a two-step process.
+Initially, they calculated the robot's end-effector trajectory.
+Subsequently, they used optimization to enhance the robot's dynamic response during trajectory execution and for tasks like obstacle avoidance.
+
+Sciavicco and Siciliano~\cite{c34, c35}, as well as Egeland~\cite{c36}, independently introduced the concept of task-space augmentation, later revisited by Searji~\cite{c37}.
+This method involves extending the primary task Jacobian with the secondary task Jacobian into a square matrix, yielding a singular solution.
+
+Maciejewski and Klein (1985)~\cite{c29}, and Nakamura and Hanafusa (1987)~\cite{c38}, presented the concept of task prioritization and a null space for avoidance.
+In this framework, the secondary task is restricted to velocities that do not directly affect the primary task's movement.
+Their method calculates avoidance joint velocities based on the minimum distance between a point on the manipulator and an avoidance point, object, or obstacle.
+Žlajpah~\cite{c41} improved the concept by proposing a reduced operational space formulation, that reshapes the Jacobain of the avoidance task from three cartesian axis to only the direction of obstacle avoidance.
+Petrič~\cite{c42} suggested a way to smoothly transition between avoidance and trajectory following tasks.
+
+\subsection{Artificial Potential Field}
+
+These approaches typically do not focus on the environmental context, often representing it as geometric primitives and then calculating the distance between these primitives and the lines describing the manipulator.
+
+Khatib~\cite{c33} proposes the concept of an Artificial Potential Field.
+This field consists of a repulsive component that deflects the end effector away from obstacles, depicted as geometric primitives, and an attractive component that draws the manipulator towards its target.
+In the following years many different modifications and improvement of the original APF idea have been proposed, often focused on removal of local minimas in the potential field.
+Khim and Khosla~\cite{c40} suggested the use of harmonic functions to solve the problem of local minima.
+Pinto et al.
+\cite{c43} proposes to vary the field based on the distance of robot from obstacles to fill the local minima.
+Many researchers tried using APF as a heuristic to better guide sampling based approaches~\cite{c45, c46, c47}.
+
+Many of the recent works focus on use of a variation of artificial potential field to plan motion of the manipulator.
+Xia et al.~\cite{c49} uses a variation of APF for manipulator motion~\cite{c49}.
+Park et al.~\cite{park2020trajectory} used a numerical Jacobian in combination with APF for motion planning.
+Zhang et al.~\cite{zhang2021obstacle} proposes dynamic repulsive field based on direction and speed between point on robot and obstacle, it also suggests decision making force that moves the robot away from certain local minima.
+Chen et al.
+proposes\remove{ an application }using APF in joint space and a variable kinematic optimization step~\cite{c50}.
+Long~\cite{c44} suggests creating motion plan of the manipulator using APF, he extends it using RRT to calculate virtual attractive point for the robot to move towards in case of local minima.
+Zhu et al.~\cite{c48} proposes use of APF in combination with MPC, to plan in environment with dynamic obstacles.
+
+\alert{---}
 

### 2024-03-18 Articles/IROS/root.tex
removed redundant parts

@@
 
-% PREDSTAVITEV REDUNDANCE
-
 Kinematic redundancy~\cite{siciliano1990kinematic, siciliano2016springer} enables a manipulator to follow a predefined task space trajectory using the endeffector (EE), while simultaneously, optimising for an additional task with the remaining movement capacity without impacting the trajectory adherence.
@@
 
-% PREDSTAVITEV PRISTOPOV PLANIRANJA POTI V ROBOTIKI
 The task of finding the joint trajectories of a manipulator is called motion planning~\cite{IDEASLab2023}.
 It consists of finding a sequence of joint configurations for a robot so that the robot can move along this path from its initial configuration to the goal configuration without colliding with itself, static obstacles or other agents in the environment.
-In addition to collision avoidance, motion planning for manipulators can optionally take into account various constraints, such as position, velocity, acceleration or jerk constraints for the joint angle or end effector, precision of the end effector with respect to position and orientation, stability of the manipulator, avoidance of singularities, or any number of other criteria.
+
+%In addition to collision avoidance, motion planning for manipulators can optionally take into account various constraints, such as position, velocity, acceleration or jerk constraints for the joint angle or end effector, precision of the end effector with respect to position and orientation, stability of the manipulator, avoidance of singularities, or any number of other criteria.
 
@@
 
-Various researchers have adopted different approaches to kinematic avoidance, as detailed in~\cite{c30, c39}.
-
-Colbaugh and Glass (1989)~\cite{c31, c32} tackled this problem in a two-step process.
+%Various researchers have adopted different approaches to kinematic avoidance, as detailed in~\cite{c30, c39}.
+%
+%Colbaugh and Glass (1989)~\cite{c31, c32} tackled this problem in a two-step process.
 Initially, they calculated the robot's end-effector trajectory.
@@
 Zhang et al.~\cite{zhang2021obstacle} proposes dynamic repulsive field based on direction and speed between point on robot and obstacle, it also suggests decision making force that moves the robot away from certain local minima.
-Chen et al.
+
+%Chen et al.
 proposes\remove{ an application }using APF in joint space and a variable kinematic optimization step~\cite{c50}.
@@
 Zhu et al.~\cite{c48} proposes use of APF in combination with MPC, to plan in environment with dynamic obstacles.
+%
 

### 2024-03-19 Articles/IROS/root.tex
removed more parts

--- 2024-03-18 Articles/IROS/root.tex
@@
 
+\alert{---}
+
+\subsection{Motion Planning}
+
+Motion planning is a process of finding a sequence of joint configurations that lead the manipulator from initial joints configuration to the goal configuration, while adhering to the constraints and optimization criteria.
+This is possible because the manipulator has a redundant number of joints~\cite{siciliano1990kinematic}.
+That means that it has more degrees of freedom (DOF) - joints, than it needs to execute the primary task, that is reaching the End Effector (EE) goal.
+Consequently, the robot can adopt different constrained joint configurations optimised according to the secondary task while performing the primary task.
+Common secondary tasks~\cite{siciliano2010robot} include avoiding singularities, optimising the manipulability measure, minimising joint torques and avoiding obstacles in the operating space.
+
+%Kinematic redundancy~\cite{siciliano1990kinematic, siciliano2016springer} enables a manipulator to follow a predefined task space trajectory using the endeffector (EE), while simultaneously, optimising for an additional task with the remaining movement capacity without impacting the trajectory adherence.
+This is possible because the robot's degrees of freedom (DOF) go beyond what is required to perform the primary task.
+Consequently, the robot can adopt different joint configurations optimised according to the secondary task while performing the primary task.
+Common secondary tasks~\cite{siciliano2010robot} include avoiding singularities, optimising the manipulability measure, minimising joint torques and avoiding obstacles in the operating space.
+%
+%The task of finding the joint trajectories of a manipulator is called motion planning~\cite{IDEASLab2023}.
+It consists of finding a sequence of joint configurations for a robot so that the robot can move along this path from its initial configuration to the goal configuration without colliding with itself, static obstacles or other agents in the environment.
+
+%In addition to collision avoidance, motion planning for manipulators can optionally take into account various constraints, such as position, velocity, acceleration or jerk constraints for the joint angle or end effector, precision of the end effector with respect to position and orientation, stability of the manipulator, avoidance of singularities, or any number of other criteria.
+
+There are global and local motion planning methods.
+Global, sampling-based, methods~\cite{vsvestka1997motion,lavalle1998rapidly, karaman2010incremental,kuffner2000rrt,gammell2015batch} offer a globally optimal solution based on a global search in configuration space.
+However, the generated trajectories are not smooth and optimal, and the performance of methods is insufficient for real-time operation.
+
+%There are global and local motion planning methods.
+Global, sampling-based, methods such as PRM~\cite{vsvestka1997motion}, RRT*~\cite{lavalle1998rapidly, karaman2010incremental} RRT-Connect~\cite{kuffner2000rrt}, Informerd RRT*~\cite{gammell2014informed}, BIT*~\cite{gammell2015batch} offer a globally optimal solution based on a global search in configuration space.
+However, the generated trajectories are not smooth and optimal, and the performance of methods is insufficient for real-time operation.
+
+Local motion planning methods employ optimization techniques, two common ones are inverse kinematics~\cite{c29,c38}, that is based on finding a least squares solution of the manipulator joint velocities, and quadratic programming (QP)~\cite{c21,c23}.
+Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
+However, since they do not plan further than one step ahead, they tend to get stuck in local minima.
+
+Often best solution to motion planning is a combination of a higher-level planner, for global static environment based path planning and local planner, that takes dynamic environment changes into account.
+
+%\subsection{Kinematic Obstacle Avoidance}
+%
+%When using inverse kinematics control for motion planning, one method is augmentation of task space to encompass supplementary parameters, in addition to the position and orientation of the end effector~\cite{c34, c35, c36}.
+This method returns a singular solution of a weighted combination of both tasks.
+Alternatively, our paper utilizes task prioritization~\cite{c29, c38, c41}, which involves calculating the null space for the primary task velocities and applying only those secondary velocities that do not directly influence the primary task velocities.
+
+\subsection{Artificial Potential Field}
+
+Khatib~\cite{c33} proposes the concept of an Artificial Potential Field.
+This field consists of a repulsive component that deflects the end effector away from obstacles, depicted as geometric primitives, and an attractive component that draws the manipulator towards its target.
+In the following years many different modifications and improvement of the original APF idea have been proposed, often focused on removal of local minimas in the potential field.
+Khim and Khosla~\cite{c40} suggested the use of harmonic functions to solve the problem of local minima.
+Pinto et al.
+\cite{c43} proposes to vary the field based on the distance of robot from obstacles to fill the local minima.
+Many researchers tried using APF as a heuristic to better guide sampling based approaches~\cite{c45, c46, c47}.
+
+Many of the recent works focus on use of a variation of artificial potential field to plan motion of the manipulator.
+Xia et al.~\cite{c49} uses a variation of APF for manipulator motion~\cite{c49}.
+Park et al.~\cite{park2020trajectory} used a numerical Jacobian in combination with APF for motion planning.
+
+%Zhang et al.~\cite{zhang2021obstacle} proposes dynamic repulsive field based on direction and speed between point on robot and obstacle, it also suggests decision making force that moves the robot away from certain local minima.
+
+%Chen et al.
+proposes\remove{ an application }using APF in joint space and a variable kinematic optimization step~\cite{c50}.
+Long~\cite{c44} suggests creating motion plan of the manipulator using APF, he extends it using RRT to calculate virtual attractive point for the robot to move towards in case of local minima.
+Zhu et al.~\cite{c48} proposes use of APF in combination with MPC, to plan in environment with dynamic obstacles.
+%
+
+\subsection{ESDF creation}
+
+ESDF grids can be generated directly using sensor measurements.
+Oleynikova et al.~\cite{oleynikova2017voxblox} proposed a method for calculating the ESDF from TSDF.
+Han et al.~\cite{han2019fiesta} proposed a way to integrate point-cloud data into ESDF using ray-casting.
+
+Another way is to generate ESDF from occupancy grids, that can be previously generated using sensors or based on predefined maps.
+When generating ESDF from occupancy grid, common approach is the Brushfire method~\cite{lau2010improved}, that spreads from obstacles until it calculates the distance for every field on a grid.
+Jump Flooding Algorithm (JFA)~\cite{rong2006jump} is a similar method, that can be implemented on a GPU for faster parallelized distance calculation.
+
+Methods that allow for the dynamical distance field generation are rare.
+Zhou et al.~\cite{zhou2021egoplanner} propose use of pairs of points on trajectory and obstacles in their quadcopter trajectory optimization algorithm.
+
+%\remove{The field of image processing invented many different morphological algorithms, that work on binary images and calculate a value for every pixel in image, that represents the distance from nearest occupied binary pixel~\cite{szeliski2021computer, felzenszwalb2012distance, rosenfeld1968distance}.
+Approaches can be seperated based on the distance metric that they use.
+}
+
+%Euclidean Signed Distance Fields (ESDFs) are valuable in robotic navigation and 3D reconstruction, providing a measure of the shortest distance to the nearest obstacle for each point in space.
+The creation of ESDF grids can leverage sensor measurements directly or be derived from other data structures like Truncated Signed Distance Fields (TSDFs) or occupancy grids.
+%
+%One prominent method for generating ESDFs is converting TSDFs into ESDFs.
+Oleynikova et al.
+[(2017)] introduced a technique that efficiently computes the ESDF from TSDFs, utilizing incremental updates for dynamic environments.
+This approach is significant for real-time applications in robotics, where rapid updates to the environment model are crucial.
+Similarly, Han et al.
+[(2019)] developed a method that integrates point-cloud data into ESDF using ray-casting, offering a way to construct ESDFs directly from sensor data.
+%
+%Alternatively, ESDFs can be generated from occupancy grids, which themselves are often derived from sensor measurements.
+The Brushfire algorithm, as improved by Lau et al.
+[(2010)], is a commonly used technique in this context.
+It operates by spreading from obstacles within the grid, calculating the minimum distance to an obstacle for each grid cell.
+This approach is particularly effective for static environments where the layout of obstacles does not change frequently.
+%
+%Furthermore, the Jump Flooding Algorithm (JFA) offers another method for generating ESDFs from occupancy grids.
+JFA is notable for its suitability for implementation on GPUs, enabling faster parallel distance calculations.
+This advantage is critical in applications requiring rapid processing of large datasets, such as in high-resolution 3D mapping and complex robotic navigation tasks.
+%
+%In conclusion, the generation of ESDFs from sensor measurements or derived data structures like TSDFs and occupancy grids is a crucial aspect of robotic navigation and 3D environment modeling.
+Techniques like those proposed by Oleynikova et al.
+[(2017)], Han et al.
+[(2019)], and the improvements to the Brushfire algorithm by Lau et al.
+[(2010)], as well as the application of JFA, highlight the ongoing advancements in efficient and accurate ESDF creation methodologies.
+
+\alert{---}
+
 In dynamic environments, a robot manipulator must swiftly respond to emerging obstacles.
@@
 That is, it is in the best local minima away from all the obstacles in its proximity.
-
-\section{RELATED WORKS}
-
-\add{ADD}
 

### 2024-03-19 Articles/IROS/root.tex
added introduction

--- 2024-03-18 Articles/IROS/root.tex
@@
 
-\alert{---}
-
-\subsection{Motion Planning}
-
-Kinematic redundancy~\cite{siciliano1990kinematic, siciliano2016springer} enables a manipulator to follow a predefined task space trajectory using the endeffector (EE), while simultaneously, optimising for an additional task with the remaining movement capacity without impacting the trajectory adherence.
+%Motion planning involves finding paths for manipulators through joint configurations from an initial to a goal position, respecting constraints and optimization criteria due to redundant degrees of freedom (DOF)~\cite{siciliano1990kinematic}.
+This redundancy allows for diverse configurations, optimizing secondary tasks like avoiding singularities and obstacles while maintaining the primary goal of reaching the End Effector (EE) target~\cite{siciliano2010robot}.
+%
+%Global motion planning methods search the entire configuration space for optimal paths but lack real-time applicability and smoothness~\cite{vsvestka1997motion,lavalle1998rapidly, karaman2010incremental,kuffner2000rrt,gammell2015batch}.
+Local methods, notably inverse kinematics and quadratic programming, offer real-time, smooth solutions but can get stuck in local minima due to their short-sighted planning approach~\cite{c29,c38,c21,c23}.
+%
+%The Artificial Potential Field (APF) concept by Khatib~\cite{c33}, which integrates repulsive forces for obstacle avoidance and attractive forces towards goals, has seen various improvements to mitigate issues like local minima~\cite{c40,c43,c45,c46,c47}.
+Recent works have adapted APF variations for manipulator motion planning, combining them with techniques like numerical Jacobians for enhanced path planning~\cite{c49,park2020trajectory}.
+%
+%For dynamic environments, strategies like ESDF grids, calculated directly from sensor data or occupancy grids, assist in real-time obstacle detection and avoidance~\cite{oleynikova2017voxblox,han2019fiesta,lau2010improved,rong2006jump,zhou2021egoplanner}.
+Our novel approach utilizes occupancy voxel grids to generate repulsive velocities, dynamically directing the robot away from obstacles based on proximity, thus enabling effective navigation in changing environments.
+This method ensures the robot maintains a safe distance from obstacles, with repulsive velocities adjusting in real-time to navigate the manipulator through the safest path possible.
+
+Motion planning is a technique in robotics, that is used to calculate joint changes leading manipulators from initial to a target configuration, while addressing constraints and optimization criteria.
+This task is facilitated by the manipulators' redundant degrees of freedom (DOF), enabling it to undertake various configurations to not only reach their End Effector (EE) target but also optimize for secondary objectives such as obstacle avoidance and minimizing joint torques~\cite{siciliano1990kinematic, siciliano2010robot}.
+
+While global motion planning techniques provide comprehensive solutions by examining the entire configuration space, they fall short in terms of smoothness and real-time application, making them less ideal for dynamic settings~\cite{vsvestka1997motion,lavalle1998rapidly, karaman2010incremental,kuffner2000rrt,gammell2015batch}.
+Conversely, local planning methods, particularly inverse kinematics and quadratic programming, offer prompt, smooth trajectories suitable for real-time operations.
+However, these methods are prone to fall into local minima due to their incremental planning approach~\cite{c29,c38,c21,c23}.
+
+Often used approach is a combination of global and local planning strategies~\cite{c44}, leveraging global planners for static environment navigation and local planners for adjusting to dynamic changes.
+This implementation enhances the manipulator's ability to navigate complex environments by combining the comprehensive pathfinding capabilities of global methods with the adaptability and efficiency of local techniques.
+
+The Artificial Potential Field (APF) concept introduced by Khatib~\cite{c33}, employs repulsive forces for obstacle deflection and attractive forces to guide towards targets, many modifications of the method have been proposed over the year, often trying to mitigate local minima~\cite{c40,c43,c45,c46,c47}.
+Some methods have further adapted APF for manipulator motion planning, incorporating elements such as numerical Jacobians to refine trajectory planning~\cite{c49,park2020trajectory}.
+
+In addressing dynamic environments, technologies such as ESDF grids, derived from sensor data or pre-established occupancy grids, play a crucial role in real-time obstacle avoidance~\cite{oleynikova2017voxblox,han2019fiesta,lau2010improved,rong2006jump,zhou2021egoplanner}.
+
+%Our approach advances this field by utilizing occupancy voxel grids to generate repulsive velocities that guide the robot away from obstacles based on their proximity, thereby ensuring safe navigation.
+This technique dynamically adjusts repulsive forces to steer the manipulator along the safest path, maintaining an optimal distance from potential hazards.
+
+Our approach utilizes occupancy voxel grids to enable safe and dynamic navigation for robot manipulators in changing environments.
+By representing the workspace with voxels and calculating repulsive velocities based on proximity to obstacles, our method allows the robot to determine safe movement directions and velocities in real-time.
+Repulsive velocities gain strength as the robot approaches an obstacle, guiding it to move away from potential collisions, and weaken when the robot is at a safe distance, maintaining an optimal distance from hazards.
+This locally-calculated repulsive field ensures the robot moves safely and efficiently in dynamic environments.
+
+%Our methodology employs occupancy voxel grids to create repulsive velocities.
+These velocities dynamically guide robots away from obstacles based on proximity, facilitating safe navigation.
+By adjusting repulsive forces in real time, our approach ensures robots maintain a safe distance from hazards, effectively integrating global and local planning strategies to navigate dynamic environments efficiently.
+This pragmatic solution leverages voxel-based spatial assessments to avoid potential collisions, ensuring pathfinding and obstacle avoidance.
+%
+%\alert{---}
+%
+%\subsection{Motion Planning}
+%
+%Motion planning is a process of finding a sequence of joint configurations that lead the manipulator from initial joints configuration to the goal configuration, while adhering to the constraints and optimization criteria.
+This is possible because the manipulator has a redundant number of joints~\cite{siciliano1990kinematic}.
+That means that it has more degrees of freedom (DOF) - joints, than it needs to execute the primary task, that is reaching the End Effector (EE) goal.
+Consequently, the robot can adopt different constrained joint configurations optimised according to the secondary task while performing the primary task.
+Common secondary tasks~\cite{siciliano2010robot} include avoiding singularities, optimising the manipulability measure, minimising joint torques and avoiding obstacles in the operating space.
+%
+%%Kinematic redundancy~\cite{siciliano1990kinematic, siciliano2016springer} enables a manipulator to follow a predefined task space trajectory using the endeffector (EE), while simultaneously, optimising for an additional task with the remaining movement capacity without impacting the trajectory adherence.
 This is possible because the robot's degrees of freedom (DOF) go beyond what is required to perform the primary task.
@@
 Common secondary tasks~\cite{siciliano2010robot} include avoiding singularities, optimising the manipulability measure, minimising joint torques and avoiding obstacles in the operating space.
-
-The task of finding the joint trajectories of a manipulator is called motion planning~\cite{IDEASLab2023}.
+%%
+%%The task of finding the joint trajectories of a manipulator is called motion planning~\cite{IDEASLab2023}.
 It consists of finding a sequence of joint configurations for a robot so that the robot can move along this path from its initial configuration to the goal configuration without colliding with itself, static obstacles or other agents in the environment.
-
-%In addition to collision avoidance, motion planning for manipulators can optionally take into account various constraints, such as position, velocity, acceleration or jerk constraints for the joint angle or end effector, precision of the end effector with respect to position and orientation, stability of the manipulator, avoidance of singularities, or any number of other criteria.
-
-There are numerous methods for planning manipulator movements~\cite{c51, c52}, they can rougly be seperated into global and local approaches.
-Global, sampling-based, methods such as PRM~\cite{vsvestka1997motion}, RRT*~\cite{lavalle1998rapidly, karaman2010incremental}, RRT-Connect~\cite{kuffner2000rrt}, Informerd RRT*~\cite{gammell2014informed}, BIT*~\cite{gammell2015batch} and others offer a globally optimal solution based on a global search in configuration space.
-However, the generated trajectories are not always smooth or optimal, and the performance of the methods may be insufficient for real-time operation.
-
-%Global planning methods are typically more comprehensive, aiming to find a globally optimal trajectory for the manipulator.
-These methods often involve mapping out the entire path before movement begins, taking into account the known environment, which usually includes static obstacles.
-However, the downside of global planning methods is their computational intensity, which often renders them too slow for real-time applications.
-This limitation is particularly pronounced in dynamic or unpredictable environments where conditions change rapidly and require quick responses.
-Examples of global planning methods include graph-based algorithms like A* and Dijkstra's algorithm, which are known for their effectiveness in pathfinding but also for their computational demands [(Lavalle, 2006)].
-
-%Sampling-based methods such as PRM, RRT, RRT*, RRT-Connect, Informerd RRT*, BIT* and others offer a solution based on a global search in configuration space.
-However, the generated trajectories are not always smooth or optimal, and the performance of the methods may be insufficient for real-time operation.
-
-%\remove{Recently, a number of learning-based methods using data-driven techniques have been proposed to improve or accelerate the functionality of sampling-based methods. }
-%
-%\remove{Trajectory optimisation methods such as CHOMP, STOMP and TrajOpt, on the other hand, use optimization to improve an initial seed trajectory.
-Consequently, the optimality of the solution is highly dependent on this initial trajectory.
-Nevertheless, these methods are capable of generating smooth trajectories, and although they can be too computationally intensive for high DOF dynamic real-time environments, they are generally effective in finding constrained motion plans.
-}
-
-Local motion planning approaches employ optimization techniques, two common ones are inverse kinematics~\cite{c29,c38}, that is based on finding a least squares solution of the manipulator joint velocities, and quadratic programming (QP)~\cite{c21,c22, haviland2021neo, c23}.
+%
+%%In addition to collision avoidance, motion planning for manipulators can optionally take into account various constraints, such as position, velocity, acceleration or jerk constraints for the joint angle or end effector, precision of the end effector with respect to position and orientation, stability of the manipulator, avoidance of singularities, or any number of other criteria.
+%
+%There are global and local motion planning methods.
+Global, sampling-based, methods~\cite{vsvestka1997motion,lavalle1998rapidly, karaman2010incremental,kuffner2000rrt,gammell2015batch} offer a globally optimal solution based on a global search in configuration space.
+However, the generated trajectories are not smooth and optimal, and the performance of methods is insufficient for real-time operation.
+%
+%%There are global and local motion planning methods.
+Global, sampling-based, methods such as PRM~\cite{vsvestka1997motion}, RRT*~\cite{lavalle1998rapidly, karaman2010incremental} RRT-Connect~\cite{kuffner2000rrt}, Informerd RRT*~\cite{gammell2014informed}, BIT*~\cite{gammell2015batch} offer a globally optimal solution based on a global search in configuration space.
+However, the generated trajectories are not smooth and optimal, and the performance of methods is insufficient for real-time operation.
+%
+%Local motion planning methods employ optimization techniques, two common ones are inverse kinematics~\cite{c29,c38}, that is based on finding a least squares solution of the manipulator joint velocities, and quadratic programming (QP)~\cite{c21,c23}.
 Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
 However, since they do not plan further than one step ahead, they tend to get stuck in local minima.
-Therefore, they are often combined with a higher-level planner, for global static environment based \replace{motion}{path} planning, while local optimisation takes dynamic environment changes into account.
-In the following text we will focus on inverse kinematics based approaches.
-
-\subsection{Kinematic Obstacle Avoidance}
-
-%\remove{Different control schemes have been proposed.
-Acceleration-based control excels in precise handling of motion changes, velocity-based control offers consistent and smooth movement, while force and torque-based control provides direct control over joint forces for robust physical interactions.
-(CITATIONS in COMMENT BELLOW)}
-
-%Various researchers have adopted different approaches to kinematic avoidance, as detailed in~\cite{c30, c39}.
-%
-%Colbaugh and Glass (1989)~\cite{c31, c32} tackled this problem in a two-step process.
-Initially, they calculated the robot's end-effector trajectory.
-Subsequently, they used optimization to enhance the robot's dynamic response during trajectory execution and for tasks like obstacle avoidance.
-
-Sciavicco and Siciliano~\cite{c34, c35}, as well as Egeland~\cite{c36}, independently introduced the concept of task-space augmentation, later revisited by Searji~\cite{c37}.
-This method involves extending the primary task Jacobian with the secondary task Jacobian into a square matrix, yielding a singular solution.
-
-Maciejewski and Klein (1985)~\cite{c29}, and Nakamura and Hanafusa (1987)~\cite{c38}, presented the concept of task prioritization and a null space for avoidance.
-In this framework, the secondary task is restricted to velocities that do not directly affect the primary task's movement.
-Their method calculates avoidance joint velocities based on the minimum distance between a point on the manipulator and an avoidance point, object, or obstacle.
-Žlajpah~\cite{c41} improved the concept by proposing a reduced operational space formulation, that reshapes the Jacobain of the avoidance task from three cartesian axis to only the direction of obstacle avoidance.
-Petrič~\cite{c42} suggested a way to smoothly transition between avoidance and trajectory following tasks.
-
-\subsection{Artificial Potential Field}
-
-These approaches typically do not focus on the environmental context, often representing it as geometric primitives and then calculating the distance between these primitives and the lines describing the manipulator.
-
-Khatib~\cite{c33} proposes the concept of an Artificial Potential Field.
+%
+%Often best solution to motion planning is a combination of a higher-level planner, for global static environment based path planning and local planner, that takes dynamic environment changes into account.
+%
+%%\subsection{Kinematic Obstacle Avoidance}
+%%
+%%When using inverse kinematics control for motion planning, one method is augmentation of task space to encompass supplementary parameters, in addition to the position and orientation of the end effector~\cite{c34, c35, c36}.
+This method returns a singular solution of a weighted combination of both tasks.
+Alternatively, our paper utilizes task prioritization~\cite{c29, c38, c41}, which involves calculating the null space for the primary task velocities and applying only those secondary velocities that do not directly influence the primary task velocities.
+%
+%\subsection{Artificial Potential Field}
+%
+%Khatib~\cite{c33} proposes the concept of an Artificial Potential Field.
 This field consists of a repulsive component that deflects the end effector away from obstacles, depicted as geometric primitives, and an attractive component that draws the manipulator towards its target.
@@
 Many researchers tried using APF as a heuristic to better guide sampling based approaches~\cite{c45, c46, c47}.
-
-Many of the recent works focus on use of a variation of artificial potential field to plan motion of the manipulator.
+%
+%Many of the recent works focus on use of a variation of artificial potential field to plan motion of the manipulator.
 Xia et al.~\cite{c49} uses a variation of APF for manipulator motion~\cite{c49}.
 Park et al.~\cite{park2020trajectory} used a numerical Jacobian in combination with APF for motion planning.
-Zhang et al.~\cite{zhang2021obstacle} proposes dynamic repulsive field based on direction and speed between point on robot and obstacle, it also suggests decision making force that moves the robot away from certain local minima.
-
-%Chen et al.
+%
+%%Zhang et al.~\cite{zhang2021obstacle} proposes dynamic repulsive field based on direction and speed between point on robot and obstacle, it also suggests decision making force that moves the robot away from certain local minima.
+%
+%%Chen et al.
 proposes\remove{ an application }using APF in joint space and a variable kinematic optimization step~\cite{c50}.
@@
 Zhu et al.~\cite{c48} proposes use of APF in combination with MPC, to plan in environment with dynamic obstacles.
-%
-
-\alert{---}
-
-In dynamic environments, a robot manipulator must swiftly respond to emerging obstacles.
+%%
+%
+%\subsection{ESDF creation}
+%
+%ESDF grids can be generated directly using sensor measurements.
+Oleynikova et al.~\cite{oleynikova2017voxblox} proposed a method for calculating the ESDF from TSDF.
+Han et al.~\cite{han2019fiesta} proposed a way to integrate point-cloud data into ESDF using ray-casting.
+%
+%Another way is to generate ESDF from occupancy grids, that can be previously generated using sensors or based on predefined maps.
+When generating ESDF from occupancy grid, common approach is the Brushfire method~\cite{lau2010improved}, that spreads from obstacles until it calculates the distance for every field on a grid.
+Jump Flooding Algorithm (JFA)~\cite{rong2006jump} is a similar method, that can be implemented on a GPU for faster parallelized distance calculation.
+%
+%Methods that allow for the dynamical distance field generation are rare.
+Zhou et al.~\cite{zhou2021egoplanner} propose use of pairs of points on trajectory and obstacles in their quadcopter trajectory optimization algorithm.
+%
+%%\remove{The field of image processing invented many different morphological algorithms, that work on binary images and calculate a value for every pixel in image, that represents the distance from nearest occupied binary pixel~\cite{szeliski2021computer, felzenszwalb2012distance, rosenfeld1968distance}.
+Approaches can be seperated based on the distance metric that they use.
+}
+%
+%%Euclidean Signed Distance Fields (ESDFs) are valuable in robotic navigation and 3D reconstruction, providing a measure of the shortest distance to the nearest obstacle for each point in space.
+The creation of ESDF grids can leverage sensor measurements directly or be derived from other data structures like Truncated Signed Distance Fields (TSDFs) or occupancy grids.
+%%
+%%One prominent method for generating ESDFs is converting TSDFs into ESDFs.
+Oleynikova et al.
+[(2017)] introduced a technique that efficiently computes the ESDF from TSDFs, utilizing incremental updates for dynamic environments.
+This approach is significant for real-time applications in robotics, where rapid updates to the environment model are crucial.
+Similarly, Han et al.
+[(2019)] developed a method that integrates point-cloud data into ESDF using ray-casting, offering a way to construct ESDFs directly from sensor data.
+%%
+%%Alternatively, ESDFs can be generated from occupancy grids, which themselves are often derived from sensor measurements.
+The Brushfire algorithm, as improved by Lau et al.
+[(2010)], is a commonly used technique in this context.
+It operates by spreading from obstacles within the grid, calculating the minimum distance to an obstacle for each grid cell.
+This approach is particularly effective for static environments where the layout of obstacles does not change frequently.
+%%
+%%Furthermore, the Jump Flooding Algorithm (JFA) offers another method for generating ESDFs from occupancy grids.
+JFA is notable for its suitability for implementation on GPUs, enabling faster parallel distance calculations.
+This advantage is critical in applications requiring rapid processing of large datasets, such as in high-resolution 3D mapping and complex robotic navigation tasks.
+%%
+%%In conclusion, the generation of ESDFs from sensor measurements or derived data structures like TSDFs and occupancy grids is a crucial aspect of robotic navigation and 3D environment modeling.
+Techniques like those proposed by Oleynikova et al.
+[(2017)], Han et al.
+[(2019)], and the improvements to the Brushfire algorithm by Lau et al.
+[(2010)], as well as the application of JFA, highlight the ongoing advancements in efficient and accurate ESDF creation methodologies.
+%
+%
+%
+%\alert{---}
+%
+%In dynamic environments, a robot manipulator must swiftly respond to emerging obstacles.
 Our approach enables the robot to determine safe movement directions and velocities by real-time assessment of nearby space for obstacles.
-
-In 3D graphics, a voxel is like a tiny cube in a grid that makes up the 3D space.
+%
+%In 3D graphics, a voxel is like a tiny cube in a grid that makes up the 3D space.
 We use voxels to represent the workspace of the robot, each voxel has a value that tells us if it’s probabilistically empty or occupied.
 A zero means empty, and a one means definitely something is there, with values in between showing varying levels of certainty.
-
-%Mreža voxels je lahko predefinirana, glede na model / 3d zemljevid prostora.
+%
+%%Mreža voxels je lahko predefinirana, glede na model / 3d zemljevid prostora.
 Zasedenost voxlov lahko spreminjamo glede na poznane pozicije in trajektorije preostalih agentov v prostoru.
 Kot omenjeno v uvodu, pa lahko zasedenost voxlov pridobimo tudi z senzorskimi sistemi.
-
-The occupancy grid of voxels can be set in advance to match the layout of the space.
+%
+%The occupancy grid of voxels can be set in advance to match the layout of the space.
 We can also change the values in the voxels if we know where other objects or agents are moving.
 Besides pre-setting, we can use depth sensors to detect what's around.
-
-Our method calculates repulsive velocities from existing occupancy voxel grid, which are vectors that guide the robot on where to go by pointing away from potential collisions.
+%
+%Our method calculates repulsive velocities from existing occupancy voxel grid, which are vectors that guide the robot on where to go by pointing away from potential collisions.
 These vectors gain strength as the robot approaches an obstacle, signaling an urgent need to maneuver to safety.
 Conversely, when the robot is positioned at an equal distance from obstacles in the vicinity—safely away from all potential collisions—the repulsive velocities weaken because there’s no immediate need to move.
-
-%In our method, we compute repulsive velocities within the task space using a novel matrix kernel multiplication approach.
+%
+%%In our method, we compute repulsive velocities within the task space using a novel matrix kernel multiplication approach.
 Concentrating on the task space is advantageous as it provides a more direct and realistic representation of the environment.
-
-%\replace{Naša metoda je posebno primerna za uporabo z senzorskimi sistemi kot so LIDAR ali globinske kamere, saj zaradi upoštevanja celotne okolice točke in ne le razdalje do najbližje točke v okolici efektivno filtriramo senzorski šum.}{prestavi to v diskusijo}
-
-Repulsive velocities tell the agent in which direction to move, so that it avoids nearby obstacles.
+%
+%%\replace{Naša metoda je posebno primerna za uporabo z senzorskimi sistemi kot so LIDAR ali globinske kamere, saj zaradi upoštevanja celotne okolice točke in ne le razdalje do najbližje točke v okolici efektivno filtriramo senzorski šum.}{prestavi to v diskusijo}
+%
+%Repulsive velocities tell the agent in which direction to move, so that it avoids nearby obstacles.
 These velocities drop to zero when the agent maintains a minimum safe distance from obstacles, and rise to their highest when it nears an obstacle, facilitating immediate evasive action.
@@
 \subsection{Manipulator Examples}
-
-\alert{add the limits of joint positions}
 

### 2024-03-19 Articles/IROS/root.tex
managing references - part 1

@@
 
-While global motion planning techniques provide comprehensive solutions by examining the entire configuration space, they fall short in terms of smoothness and real-time application, making them less ideal for dynamic settings~\cite{vsvestka1997motion,lavalle1998rapidly, karaman2010incremental,kuffner2000rrt,gammell2015batch}.
-Conversely, local planning methods, particularly inverse kinematics and quadratic programming, offer prompt, smooth trajectories suitable for real-time operations.
-However, these methods are prone to fall into local minima due to their incremental planning approach~\cite{c29,c38,c21,c23}.
+While global motion planning techniques provide comprehensive solutions by examining the entire configuration space, they fall short in terms of smoothness and real-time application, making them less ideal for dynamic settings~\cite{lavalle1998rapidly, gammell2015batch, karaman2010incremental,kuffner2000rrt}.
+Conversely, local planning methods, particularly inverse kinematics~\cite{c29,c38} and quadratic programming~\cite{c21,c23}, offer prompt, smooth trajectories suitable for real-time operations.
+However, these methods are prone to fall into local minima due to their incremental planning approach.
 
@@
 
-The Artificial Potential Field (APF) concept introduced by Khatib~\cite{c33}, employs repulsive forces for obstacle deflection and attractive forces to guide towards targets, many modifications of the method have been proposed over the year, often trying to mitigate local minima~\cite{c40,c43,c45,c46,c47}.
-Some methods have further adapted APF for manipulator motion planning, incorporating elements such as numerical Jacobians to refine trajectory planning~\cite{c49,park2020trajectory}.
+The Artificial Potential Field (APF) concept introduced by Khatib~\cite{c33}, employs repulsive forces for obstacle deflection and attractive forces to guide towards targets, many modifications of the method have been proposed over the year, often trying to mitigate local minima~\cite{c43,c45,c47,klancar2022robot}.
+Some methods have further adapted APF for manipulator motion planning, incorporating elements such as numerical Jacobians to refine trajectory planning~\cite{c49,park2020trajectory,baumgartner2023potential}.
 
@@
 
-\bibitem{IDEASLab2023}
-IDEAS Lab, "Motion and Path Planning," presented at Purdue University, 2023. [Online]. Available: \url{https://ideas.cs.purdue.edu/research/robotics/planning/}. Accessed on: Jan. 9, 2024.
+\bibitem{siciliano1990kinematic} B. Siciliano, “Kinematic control of redundant robot manipulators: A tutorial,” J. Intell. Robot. Syst., vol. 3, no. 3, Art. no. 3, 1990, doi: 10.1007/BF00126069.
+
+\bibitem{siciliano2010robot} B.
+Siciliano, L.
+Sciavicco, L.
+Villani, and G.
+Oriolo, Robot.
+Model.
+Plan.
+Control, 2010, pp.
+161–189.
+[Online].
+Available: http://link.springer.com/10.1007/978-1-84628-642-1\_4
+
+%\bibitem{vsvestka1997motion} P.
+{\v{S}}vestka and M.
+H.
+Overmars, “Motion planning for carlike robots using a probabilistic learning approach,” The International Journal of Robotics Research, vol.
+16, no.
+2, pp.
+119–143, 1997.
+
+\bibitem{lavalle1998rapidly} S. LaValle, “Rapidly-exploring random trees: A new tool for path planning,” Research Report 9811, 1998.
+
+\bibitem{gammell2015batch} J.
+D.
+Gammell, S.
+S.
+Srinivasa, and T.
+D.
+Barfoot, “Batch Informed Trees (BIT*): Sampling-based Optimal Planning via the Heuristically Guided Search of Implicit Random Geometric Graphs,” in Proc.
+of the 2015 IEEE International Conference on Robotics and Automation (ICRA), May 2015, pp.
+3067–3074.
+doi: 10.1109/ICRA.2015.7139620.
+
+\bibitem{karaman2010incremental} S. Karaman and E. Frazzoli, "Incremental sampling-based algorithms for optimal motion planning," in Proc. Robotics: Science and Systems (RSS), 2010.
+
+\bibitem{kuffner2000rrt} J.
+J.
+Kuffner and S.
+M.
+LaValle, “RRT-connect: An efficient approach to single-query path planning,” in Proceedings of the 2000 ICRA.
+Millennium Conference.
+IEEE International Conference on Robotics and Automation.
+Symposia Proceedings (Cat.
+No.00CH37065), Apr.
+2000, pp.
+995–1001 vol.2.
+doi: 10.1109/ROBOT.2000.844730.
+
+\bibitem{c29} A.
+A.
+Maciejewski and C.
+A.
+Klein, “Obstacle Avoidance for Kinematically Redundant Manipulators in Dynamically Varying Environments,” The International Journal of Robotics Research, vol.
+4, no.
+3, pp.
+109–117, Sep.
+1985.
+doi: 10.1177/027836498500400308.
+
+\bibitem{c38} Y.
+Nakamura, H.
+Hanafusa, and T.
+Yoshikawa, “Task-Priority Based Redundancy Control of Robot Manipulators,” The International Journal of Robotics Research, vol.
+6, no.
+2, pp.
+3–15, Jun.
+1987.
+doi: 10.1177/027836498700600201.
 
@@
 
-\bibitem{c22} H.
+\bibitem{c23} H.
 Toshani and M.
@@
 
-\bibitem{haviland2021neo} J.
-Haviland and P.
-Corke, “NEO: A Novel Expeditious Optimisation Algorithm for Reactive Motion Control of Manipulators,” IEEE Robot.
-Autom.
-Lett., vol.
-6, no.
-2, Art.
-no.
-2, Apr.
-2021, doi: 10.1109/LRA.2021.3056060.
-
-\bibitem{c23} Y.
+%\bibitem{c22} Y.
 Zhang, S.
@@
 
-\bibitem{c24} J.
-Nakanishi, R.
-Cory, M.
-Mistry, J.
-Peters, and S.
-Schaal, “Comparative experiments on task space control with redundancy resolution,” in Proc.
-2005 IEEE/RSJ Int.
-Conf.
-on Intelligent Robots and Systems, Edmonton, Alta., Canada, 2005, pp.
-3901–3908.
-doi: 10.1109/IROS.2005.1545203.
-
-\bibitem{c25} M.
-H.
-Raibert and J.
-J.
-Craig, “Hybrid Position/Force Control of Manipulators,” Journal of Dynamic Systems, Measurement, and Control, vol.
-103, no.
-2, pp.
-126–133, Jun.
-1981.
-doi: 10.1115/1.3139652.
-
-\bibitem{c26} T.
-Yoshikawa, “Dynamic hybrid position/force control of robot manipulators--Description of hand constraints and calculation of joint driving force,” IEEE J.
-Robot.
-Automat., vol.
-3, no.
-5, pp.
-386–392, Oct.
-1987.
-doi: 10.1109/JRA.1987.1087120.
-
-\bibitem{c27} O.
-Khatib, “A unified approach for motion and force control of robot manipulators: The operational space formulation,” IEEE J.
-Robot.
-Automat., vol.
-3, no.
-1, pp.
-43–53, Feb.
-1987.
-doi: 10.1109/JRA.1987.1087068.
-
-\bibitem{c28} N. Hogan, “Impedance Control: An Approach to Manipulation,” in Proc. 1984 American Control Conf., San Diego, CA, USA, Jul. 1984, pp. 304–313. doi: 10.23919/ACC.1984.4788393.
-
-\bibitem{c29} A.
-A.
-Maciejewski and C.
-A.
-Klein, “Obstacle Avoidance for Kinematically Redundant Manipulators in Dynamically Varying Environments,” The International Journal of Robotics Research, vol.
-4, no.
-3, pp.
-109–117, Sep.
-1985.
-doi: 10.1177/027836498500400308.
-
-\bibitem{c30} L.
-Lajpah and T.
-Petri, “Obstacle Avoidance for Redundant Manipulators as Control Problem,” in Serial and Parallel Robot Manipulators - Kinematics, Dynamics, Control and Optimization, S.
-Kucuk, Ed.
-InTech, 2012.
-doi: 10.5772/32651.
-
-\bibitem{c31} R. Colbaugh and K. Glass, “Cartesian control of redundant robots,” J. Robotic Syst., vol. 6, no. 4, pp. 427–459, Aug. 1989. doi: 10.1002/rob.4620060409.
-
-\bibitem{c32} K.
-Glass, R.
-Colbaugh, D.
-Lim, and H.
-Seraji, “Real-time collision avoidance for redundant manipulators,” IEEE Trans.
-Robot.
-Automat., vol.
-11, no.
-3, pp.
-448–457, Jun.
-1995.
-doi: 10.1109/70.388789.
+\bibitem{c44} Z.
+Long, “Virtual target point-based obstacle-avoidance method for manipulator systems in a cluttered environment,” Engineering Optimization, vol.
+52, no.
+11, Art.
+no.
+11, Nov.
+2020.
+doi: 10.1080/0305215X.2019.1681986.
 
@@
 
-\bibitem{c34} L.
-Sciavicco and B.
-Siciliano, “A solution algorithm to the inverse kinematic problem for redundant manipulators,” IEEE J.
-Robot.
-Automat., vol.
-4, no.
-4, pp.
-403–410, Aug.
-1988.
-doi: 10.1109/56.804.
-
-\bibitem{c35} L.
-Sciavicco and B.
-Siciliano, “Solving the Inverse Kinematic Problem for Robotic Manipulators,” in RoManSy 6, A.
-Morecki, G.
-Bianchi, and K.
-Kedzior, Eds., Boston, MA: Springer US, 1987, pp.
-107–114.
-doi: 10.1007/978-1-4684-6915-8\_9.
-
-\bibitem{c36} O. Egeland, “Task-space tracking with redundant manipulators,” IEEE J. Robot. Automat., vol. 3, no. 5, pp. 471–475, Oct. 1987. doi: 10.1109/JRA.1987.1087118.
-
-\bibitem{c37} H. Seraji, “Configuration control of redundant manipulators: theory and implementation,” IEEE Trans. Robot. Automat., vol. 5, no. 4, pp. 472–490, Aug. 1989. doi: 10.1109/70.88062.
-
-\bibitem{c38} Y.
-Nakamura, H.
-Hanafusa, and T.
-Yoshikawa, “Task-Priority Based Redundancy Control of Robot Manipulators,” The International Journal of Robotics Research, vol.
-6, no.
-2, pp.
-3–15, Jun.
-1987.
-doi: 10.1177/027836498700600201.
-
-\bibitem{c39} B. Siciliano and O. Khatib, Eds., Springer Handbook of Robotics. in Springer Handbooks. Cham: Springer International Publishing, 2016. doi: 10.1007/978-3-319-32552-1.
-
-\bibitem{c40} J.-O. Kim and P. Khosla, “Real-Time Obstacle Avoidance Using Harmonic Potential Functions,” 1992. doi: 10.1109/70.143352.
-
-\bibitem{c41} L.
-Zlajpah and B.
-Nemec, “Kinematic control algorithms for on-line obstacle avoidance for redundant manipulators,” in Proc.
-IEEE/RSJ International Conference on Intelligent Robots and Systems, Lausanne, Switzerland, 2002, pp.
-1898–1903.
-doi: 10.1109/IRDS.2002.1044033.
-
-\bibitem{c42} T.
-Petrič and L.
-Žlajpah, “Smooth continuous transition between tasks on a kinematic control level: Obstacle avoidance as a control problem,” Robotics and Autonomous Systems, vol.
-61, no.
-9, Art.
-no.
-9, Sep.
-2013.
-doi: 10.1016/j.robot.2013.04.019.
+%\bibitem{c40} J.-O. Kim and P. Khosla, “Real-Time Obstacle Avoidance Using Harmonic Potential Functions,” 1992. doi: 10.1109/70.143352.
 
@@
 
-\bibitem{c44} Z.
-Long, “Virtual target point-based obstacle-avoidance method for manipulator systems in a cluttered environment,” Engineering Optimization, vol.
-52, no.
-11, Art.
-no.
-11, Nov.
-2020.
-doi: 10.1080/0305215X.2019.1681986.
-
 \bibitem{c45} A. H. Qureshi and Y. Ayaz, “Potential Functions based Sampling Heuristic For Optimal Path Planning,” Auton Robot, vol. 40, no. 6, Art. no. 6, Aug. 2016. doi: 10.1007/s10514-015-9518-0.
 
-\bibitem{c46} A.
+%\bibitem{c46} A.
 H.
@@
 
-\bibitem{c48} T.
-Zhu, J.
-Mao, L.
-Han, C.
-Zhang, and J.
-Yang, “Real-Time Dynamic Obstacle Avoidance for Robot Manipulators Based on Cascaded Nonlinear MPC With Artificial Potential Field,” IEEE Trans.
-Ind.
-Electron., pp.
-1–11, 2023.
-doi: 10.1109/TIE.2023.3306405.
+\bibitem{klancar2022robot}
+G.
+Klančar, A.
+Zdešar, and M.
+Krishnan, “Robot Navigation Based on Potential Field and Gradient Obtained by Bilinear Interpolation and a Grid-Based Search,” Sensors, vol.
+22, no.
+9, art.
+no.
+3295, pp.
+1-24, 2022.
+doi: 10.3390/s22093295.
 
 \bibitem{c49} X. Xia et al., “Path Planning for Obstacle Avoidance of Robot Arm Based on Improved Potential Field Method,” Sensors, vol. 23, no. 7, Art. no. 7, Apr. 2023. doi: 10.3390/s23073754.
-
-\bibitem{c50} Y.
-Chen, L.
-Chen, J.
-Ding, and Y.
-Liu, “Research on Real-Time Obstacle Avoidance Motion Planning of Industrial Robotic Arm Based on Artificial Potential Field Method in Joint Space,” Applied Sciences, vol.
-13, no.
-12, p.
-6973, Jan.
-2023.
-doi: 10.3390/app13126973.
-
-\bibitem{c51} S. M. LaValle, Planning Algorithms. Cambridge: Cambridge University Press, 2006.
-
-\bibitem{c52} M.
-G.
-Tamizi, M.
-Yaghoubi, and H.
-Najjaran, “A review of recent trend in motion planning of industrial robots,” Int J Intell Robot Appl, vol.
-7, no.
-2, Art.
-no.
-2, Jun.
-2023.
-doi: 10.1007/s41315-023-00274-2.
-
-\bibitem{vsvestka1997motion} P.
-{\v{S}}vestka and M.
-H.
-Overmars, “Motion planning for carlike robots using a probabilistic learning approach,” The International Journal of Robotics Research, vol.
-16, no.
-2, pp.
-119–143, 1997.
-
-\bibitem{lavalle1998rapidly} S. LaValle, “Rapidly-exploring random trees: A new tool for path planning,” Research Report 9811, 1998.
-
-\bibitem{gammell2015batch} J.
-D.
-Gammell, S.
-S.
-Srinivasa, and T.
-D.
-Barfoot, “Batch Informed Trees (BIT*): Sampling-based Optimal Planning via the Heuristically Guided Search of Implicit Random Geometric Graphs,” in Proc.
-of the 2015 IEEE International Conference on Robotics and Automation (ICRA), May 2015, pp.
-3067–3074.
-doi: 10.1109/ICRA.2015.7139620.
-
-\bibitem{karaman2010incremental} S. Karaman and E. Frazzoli, "Incremental sampling-based algorithms for optimal motion planning," in Proc. Robotics: Science and Systems (RSS), 2010.
-
-\bibitem{gammell2014informed} J.
-D.
-Gammell, S.
-S.
-Srinivasa, and T.
-D.
-Barfoot, “Informed RRT*: Optimal sampling-based path planning focused via direct sampling of an admissible ellipsoidal heuristic,” in Proc.
-of the 2014 IEEE/RSJ International Conference on Intelligent Robots and Systems, Chicago, IL, USA, Sep.
-2014, pp.
-2997–3004.
-doi: 10.1109/IROS.2014.6942976.
-
-\bibitem{kuffner2000rrt} J.
-J.
-Kuffner and S.
-M.
-LaValle, “RRT-connect: An efficient approach to single-query path planning,” in Proceedings of the 2000 ICRA.
-Millennium Conference.
-IEEE International Conference on Robotics and Automation.
-Symposia Proceedings (Cat.
-No.00CH37065), Apr.
-2000, pp.
-995–1001 vol.2.
-doi: 10.1109/ROBOT.2000.844730.
-
-\bibitem{siciliano1990kinematic} B. Siciliano, “Kinematic control of redundant robot manipulators: A tutorial,” J. Intell. Robot. Syst., vol. 3, no. 3, Art. no. 3, 1990, doi: 10.1007/BF00126069.
-
-\bibitem{siciliano2016springer} B. Siciliano and O. Khatib, Eds., Springer Handbook of Robotics, in Springer Handbooks. Cham: Springer International Publishing, 2016. doi: 10.1007/978-3-319-32552-1.
-
-\bibitem{siciliano2010robot} B.
-Siciliano, L.
-Sciavicco, L.
-Villani, and G.
-Oriolo, Robot.
-Model.
-Plan.
-Control, 2010, pp.
-161–189.
-[Online].
-Available: http://link.springer.com/10.1007/978-1-84628-642-1\_4
-
-\bibitem{dai2022review} Y.
-Dai, C.
-Xiang, Y.
-Zhang, Y.
-Jiang, W.
-Qu, and Q.
-Zhang, “A Review of Spatial Robotic Arm Trajectory Planning,” Aerospace, vol.
-9, p.
-361, Jul.
-2022, doi: 10.3390/aerospace9070361.
-
-\bibitem{gottschalk1996obbtree} S.
-Gottschalk, M.
-C.
-Lin, and D.
-Manocha, “OBBTree: a hierarchical structure for rapid interference detection,” in Proceedings of the 23rd Annual Conference on Computer Graphics and Interactive Techniques, ACM, Aug.
-1996, pp.
-171–180.
-doi: 10.1145/237170.237244.
-
-\bibitem{vandenbergen1997efficient} G.
-van den Bergen, “Efficient Collision Detection of Complex Deformable Models using AABB Trees,” Journal of Graphics Tools, vol.
-2, no.
-4, pp.
-1–13, 1997.
-doi: 10.1080/10867651.1997.10487480.
-
-\bibitem{chen2018path} G.
-Chen, D.
-Liu, Y.
-Wang, Q.
-Jia, and X.
-Zhang, “Path planning method with obstacle avoidance for manipulators in dynamic environment,” International Journal of Advanced Robotic Systems, vol.
-15, no.
-6, Art.
-no.
-1729881418820223, Nov.
-2018, doi: 10.1177/1729881418820223.
-
-\bibitem{puiu2011realtime} D.
-Puiu and F.
-Moldoveanu, “Real-time collision avoidance for redundant manipulators,” in Proc.
-of the 2011 6th IEEE International Symposium on Applied Computational Intelligence and Informatics (SACI), Timisoara, Romania, 2011, pp.
-403–408, doi: 10.1109/SACI.2011.5873037.
-
-\bibitem{oleynikova2017voxblox} H.
-Oleynikova, Z.
-Taylor, M.
-Fehr, R.
-Siegwart, and J.
-Nieto, “Voxblox: Incremental 3D Euclidean Signed Distance Fields for on-board MAV planning,” in Proc.
-of the 2017 IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS), Vancouver, BC, Sep.
-2017, pp.
-1366–1373.
-doi: 10.1109/IROS.2017.8202315.
-
-\bibitem{wurmOctoMap} K.
-M.
-Wurm, A.
-Hornung, M.
-Bennewitz, C.
-Stachniss, and W.
-Burgard, “OctoMap: A Probabilistic, Flexible, and Compact 3D Map Representation for Robotic Systems,” [Details of publication, e.g., in Proc.
-of the Conference/Journal Name, Year, pp.
-Page numbers].
-[DOI or URL if available].
-
-\bibitem{gao2019flying} F.
-Gao, W.
-Wu, W.
-Gao, and S.
-Shen, “Flying on point clouds: Online trajectory generation and autonomous navigation for quadrotors in cluttered environments,” Journal of Field Robotics, vol.
-36, no.
-4, pp.
-710–733, 2019, doi: 10.1002/rob.21842.
-
-\bibitem{elfes1989using} A. Elfes, “Using occupancy grids for mobile robot perception and navigation,” Computer, vol. 22, no. 6, pp. 46–57, Jun. 1989, doi: 10.1109/2.30720.
-
-\bibitem{han2019fiesta} L.
-Han, F.
-Gao, B.
-Zhou, and S.
-Shen, “FIESTA: Fast Incremental Euclidean Distance Fields for Online Motion Planning of Aerial Robots,” arXiv, Jul.
-26, 2019.
-Accessed: Jan.
-11, 2024.
-[Online].
-Available: http://arxiv.org/abs/1903.02144
-
-\bibitem{xu2021voxel} Y.
-Xu, X.
-Tong, and U.
-Stilla, “Voxel-based representation of 3D point clouds: Methods, applications, and its potential use in the construction industry,” Automation in Construction, vol.
-126, p.
-103675, Jun.
-2021, doi: 10.1016/j.autcon.2021.103675.
-
-\bibitem{niessner2013realtime} M.
-Nießner, M.
-Zollhöfer, S.
-Izadi, and M.
-Stamminger, “Real-time 3D reconstruction at scale using voxel hashing,” ACM Trans.
-Graph., vol.
-32, no.
-6, pp.
-1–11, Nov.
-2013, doi: 10.1145/2508363.2508374.
-
-\bibitem{dryanovski2010multivolume} I.
-Dryanovski, W.
-Morris, and J.
-Xiao, “Multi-volume occupancy grids: An efficient probabilistic 3D mapping model for micro aerial vehicles,” in Proc.
-of the 2010 IEEE/RSJ International Conference on Intelligent Robots and Systems, Taipei, Oct.
-2010, pp.
-1553–1559.
-doi: 10.1109/IROS.2010.5652494.
-
-\bibitem{thrun2002probabilistic} S. Thrun, Probabilistic robotics, vol. 45, 2002. Accessed: Jun. 14, 2023. [Online]. Available: https://dl.acm.org/doi/10.1145/504729.504754
-
-\bibitem{lau2010improved} B.
-Lau, C.
-Sprunk, and W.
-Burgard, “Improved updating of Euclidean distance maps and Voronoi diagrams,” in Proc.
-of the 2010 IEEE/RSJ International Conference on Intelligent Robots and Systems, Taipei, Oct.
-2010, pp.
-281–286.
-doi: 10.1109/IROS.2010.5650794.
-
-\bibitem{luo2018collisionfree} L.
-Luo et al., “Collision-Free Path-Planning for Six-DOF Serial Harvesting Robot Based on Energy Optimal and Artificial Potential Field,” Complexity, vol.
-2018, pp.
-1–12, Nov.
-2018, doi: 10.1155/2018/3563846.
-
-\bibitem{zhang2021obstacle} W.
-Zhang, H.
-Cheng, L.
-Hao, X.
-Li, M.
-Liu, and X.
-Gao, “An obstacle avoidance algorithm for robot manipulators based on decision-making force,” Robotics and Computer-Integrated Manufacturing, vol.
-71, p.
-102114, Oct.
-2021, doi: 10.1016/j.rcim.2020.102114.
 
@@
 
-\bibitem{han2018dynamic} D.
-Han, H.
-Nie, J.
-Chen, and M.
-Chen, “Dynamic obstacle avoidance for manipulators using distance calculation and discrete detection,” Robotics and Computer-Integrated Manufacturing, vol.
-49, pp.
-98–104, Feb.
-2018, doi: 10.1016/j.rcim.2017.05.013.
+\bibitem{baumgartner2023potential}
+J.
+Baumgartner, T.
+Petrič, and G.
+Klančar, “Potential Field Control of a Redundant Nonholonomic Mobile Manipulator with Corridor-Constrained Base Motion,” Machines, vol.
+11, no.
+2, art.
+no.
+293, pp.
+1-28, 2023.
+doi: 10.3390/machines11020293.
+
+\bibitem{oleynikova2017voxblox} H.
+Oleynikova, Z.
+Taylor, M.
+Fehr, R.
+Siegwart, and J.
+Nieto, “Voxblox: Incremental 3D Euclidean Signed Distance Fields for on-board MAV planning,” in Proc.
+of the 2017 IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS), Vancouver, BC, Sep.
+2017, pp.
+1366–1373.
+doi: 10.1109/IROS.2017.8202315.
+
+\bibitem{han2019fiesta} L.
+Han, F.
+Gao, B.
+Zhou, and S.
+Shen, “FIESTA: Fast Incremental Euclidean Distance Fields for Online Motion Planning of Aerial Robots,” arXiv, Jul.
+26, 2019.
+Accessed: Jan.
+11, 2024.
+[Online].
+Available: http://arxiv.org/abs/1903.02144
+
+\bibitem{lau2010improved} B.
+Lau, C.
+Sprunk, and W.
+Burgard, “Improved updating of Euclidean distance maps and Voronoi diagrams,” in Proc.
+of the 2010 IEEE/RSJ International Conference on Intelligent Robots and Systems, Taipei, Oct.
+2010, pp.
+281–286.
+doi: 10.1109/IROS.2010.5650794.
 
@@
 doi: 10.1145/1111411.1111431.
-
-\bibitem{szeliski2021computer} R. Szeliski, “Computer Vision: Algorithms and Applications, 2nd Edition,” Springer, 2021.
-
-\bibitem{rosenfeld1968distance} A. Rosenfeld and J. L. Pfaltz, “Distance functions on digital pictures,” Pattern Recognition, vol. 1, no. 1, pp. 33–61, Jul. 1968, doi: 10.1016/0031-3203(68)90013-7.
-
-\bibitem{felzenszwalb2012distance} P.
-F.
-Felzenszwalb and D.
-P.
-Huttenlocher, “Distance Transforms of Sampled Functions,” Theory of Comput., vol.
-8, no.
-1, pp.
-415–428, 2012, doi: 10.4086/toc.2012.v008a019.
 
@@
 
+%\bibitem{gammell2014informed} J.
+D.
+Gammell, S.
+S.
+Srinivasa, and T.
+D.
+Barfoot, “Informed RRT*: Optimal sampling-based path planning focused via direct sampling of an admissible ellipsoidal heuristic,” in Proc.
+of the 2014 IEEE/RSJ International Conference on Intelligent Robots and Systems, Chicago, IL, USA, Sep.
+2014, pp.
+2997–3004.
+doi: 10.1109/IROS.2014.6942976.
+%
+%	
+%\bibitem{IDEASLab2023}
+%IDEAS Lab, "Motion and Path Planning," presented at Purdue University, 2023. [Online]. Available: \url{https://ideas.cs.purdue.edu/research/robotics/planning/}. Accessed on: Jan. 9, 2024.
+%
+
+\bibitem{haviland2021neo} J.
+Haviland and P.
+Corke, “NEO: A Novel Expeditious Optimisation Algorithm for Reactive Motion Control of Manipulators,” IEEE Robot.
+Autom.
+Lett., vol.
+6, no.
+2, Art.
+no.
+2, Apr.
+2021, doi: 10.1109/LRA.2021.3056060.
+
+\bibitem{c24} J.
+Nakanishi, R.
+Cory, M.
+Mistry, J.
+Peters, and S.
+Schaal, “Comparative experiments on task space control with redundancy resolution,” in Proc.
+2005 IEEE/RSJ Int.
+Conf.
+on Intelligent Robots and Systems, Edmonton, Alta., Canada, 2005, pp.
+3901–3908.
+doi: 10.1109/IROS.2005.1545203.
+
+\bibitem{c25} M.
+H.
+Raibert and J.
+J.
+Craig, “Hybrid Position/Force Control of Manipulators,” Journal of Dynamic Systems, Measurement, and Control, vol.
+103, no.
+2, pp.
+126–133, Jun.
+1981.
+doi: 10.1115/1.3139652.
+
+\bibitem{c26} T.
+Yoshikawa, “Dynamic hybrid position/force control of robot manipulators--Description of hand constraints and calculation of joint driving force,” IEEE J.
+Robot.
+Automat., vol.
+3, no.
+5, pp.
+386–392, Oct.
+1987.
+doi: 10.1109/JRA.1987.1087120.
+
+\bibitem{c27} O.
+Khatib, “A unified approach for motion and force control of robot manipulators: The operational space formulation,” IEEE J.
+Robot.
+Automat., vol.
+3, no.
+1, pp.
+43–53, Feb.
+1987.
+doi: 10.1109/JRA.1987.1087068.
+
+\bibitem{c28} N. Hogan, “Impedance Control: An Approach to Manipulation,” in Proc. 1984 American Control Conf., San Diego, CA, USA, Jul. 1984, pp. 304–313. doi: 10.23919/ACC.1984.4788393.
+
+\bibitem{c30} L.
+Lajpah and T.
+Petri, “Obstacle Avoidance for Redundant Manipulators as Control Problem,” in Serial and Parallel Robot Manipulators - Kinematics, Dynamics, Control and Optimization, S.
+Kucuk, Ed.
+InTech, 2012.
+doi: 10.5772/32651.
+
+\bibitem{c31} R. Colbaugh and K. Glass, “Cartesian control of redundant robots,” J. Robotic Syst., vol. 6, no. 4, pp. 427–459, Aug. 1989. doi: 10.1002/rob.4620060409.
+
+\bibitem{c32} K.
+Glass, R.
+Colbaugh, D.
+Lim, and H.
+Seraji, “Real-time collision avoidance for redundant manipulators,” IEEE Trans.
+Robot.
+Automat., vol.
+11, no.
+3, pp.
+448–457, Jun.
+1995.
+doi: 10.1109/70.388789.
+
+\bibitem{c34} L.
+Sciavicco and B.
+Siciliano, “A solution algorithm to the inverse kinematic problem for redundant manipulators,” IEEE J.
+Robot.
+Automat., vol.
+4, no.
+4, pp.
+403–410, Aug.
+1988.
+doi: 10.1109/56.804.
+
+\bibitem{c35} L.
+Sciavicco and B.
+Siciliano, “Solving the Inverse Kinematic Problem for Robotic Manipulators,” in RoManSy 6, A.
+Morecki, G.
+Bianchi, and K.
+Kedzior, Eds., Boston, MA: Springer US, 1987, pp.
+107–114.
+doi: 10.1007/978-1-4684-6915-8\_9.
+
+\bibitem{c36} O. Egeland, “Task-space tracking with redundant manipulators,” IEEE J. Robot. Automat., vol. 3, no. 5, pp. 471–475, Oct. 1987. doi: 10.1109/JRA.1987.1087118.
+
+\bibitem{c37} H. Seraji, “Configuration control of redundant manipulators: theory and implementation,” IEEE Trans. Robot. Automat., vol. 5, no. 4, pp. 472–490, Aug. 1989. doi: 10.1109/70.88062.
+
+\bibitem{c39} B. Siciliano and O. Khatib, Eds., Springer Handbook of Robotics. in Springer Handbooks. Cham: Springer International Publishing, 2016. doi: 10.1007/978-3-319-32552-1.
+
+\bibitem{c41} L.
+Zlajpah and B.
+Nemec, “Kinematic control algorithms for on-line obstacle avoidance for redundant manipulators,” in Proc.
+IEEE/RSJ International Conference on Intelligent Robots and Systems, Lausanne, Switzerland, 2002, pp.
+1898–1903.
+doi: 10.1109/IRDS.2002.1044033.
+
+\bibitem{c42} T.
+Petrič and L.
+Žlajpah, “Smooth continuous transition between tasks on a kinematic control level: Obstacle avoidance as a control problem,” Robotics and Autonomous Systems, vol.
+61, no.
+9, Art.
+no.
+9, Sep.
+2013.
+doi: 10.1016/j.robot.2013.04.019.
+
+\bibitem{c48} T.
+Zhu, J.
+Mao, L.
+Han, C.
+Zhang, and J.
+Yang, “Real-Time Dynamic Obstacle Avoidance for Robot Manipulators Based on Cascaded Nonlinear MPC With Artificial Potential Field,” IEEE Trans.
+Ind.
+Electron., pp.
+1–11, 2023.
+doi: 10.1109/TIE.2023.3306405.
+
+\bibitem{c50} Y.
+Chen, L.
+Chen, J.
+Ding, and Y.
+Liu, “Research on Real-Time Obstacle Avoidance Motion Planning of Industrial Robotic Arm Based on Artificial Potential Field Method in Joint Space,” Applied Sciences, vol.
+13, no.
+12, p.
+6973, Jan.
+2023.
+doi: 10.3390/app13126973.
+
+\bibitem{c51} S. M. LaValle, Planning Algorithms. Cambridge: Cambridge University Press, 2006.
+
+\bibitem{c52} M.
+G.
+Tamizi, M.
+Yaghoubi, and H.
+Najjaran, “A review of recent trend in motion planning of industrial robots,” Int J Intell Robot Appl, vol.
+7, no.
+2, Art.
+no.
+2, Jun.
+2023.
+doi: 10.1007/s41315-023-00274-2.
+
+\bibitem{siciliano2016springer} B. Siciliano and O. Khatib, Eds., Springer Handbook of Robotics, in Springer Handbooks. Cham: Springer International Publishing, 2016. doi: 10.1007/978-3-319-32552-1.
+
+\bibitem{dai2022review} Y.
+Dai, C.
+Xiang, Y.
+Zhang, Y.
+Jiang, W.
+Qu, and Q.
+Zhang, “A Review of Spatial Robotic Arm Trajectory Planning,” Aerospace, vol.
+9, p.
+361, Jul.
+2022, doi: 10.3390/aerospace9070361.
+
+\bibitem{gottschalk1996obbtree} S.
+Gottschalk, M.
+C.
+Lin, and D.
+Manocha, “OBBTree: a hierarchical structure for rapid interference detection,” in Proceedings of the 23rd Annual Conference on Computer Graphics and Interactive Techniques, ACM, Aug.
+1996, pp.
+171–180.
+doi: 10.1145/237170.237244.
+
+\bibitem{vandenbergen1997efficient} G.
+van den Bergen, “Efficient Collision Detection of Complex Deformable Models using AABB Trees,” Journal of Graphics Tools, vol.
+2, no.
+4, pp.
+1–13, 1997.
+doi: 10.1080/10867651.1997.10487480.
+
+\bibitem{chen2018path} G.
+Chen, D.
+Liu, Y.
+Wang, Q.
+Jia, and X.
+Zhang, “Path planning method with obstacle avoidance for manipulators in dynamic environment,” International Journal of Advanced Robotic Systems, vol.
+15, no.
+6, Art.
+no.
+1729881418820223, Nov.
+2018, doi: 10.1177/1729881418820223.
+
+\bibitem{puiu2011realtime} D.
+Puiu and F.
+Moldoveanu, “Real-time collision avoidance for redundant manipulators,” in Proc.
+of the 2011 6th IEEE International Symposium on Applied Computational Intelligence and Informatics (SACI), Timisoara, Romania, 2011, pp.
+403–408, doi: 10.1109/SACI.2011.5873037.
+
+\bibitem{wurmOctoMap} K.
+M.
+Wurm, A.
+Hornung, M.
+Bennewitz, C.
+Stachniss, and W.
+Burgard, “OctoMap: A Probabilistic, Flexible, and Compact 3D Map Representation for Robotic Systems,” [Details of publication, e.g., in Proc.
+of the Conference/Journal Name, Year, pp.
+Page numbers].
+[DOI or URL if available].
+
+\bibitem{gao2019flying} F.
+Gao, W.
+Wu, W.
+Gao, and S.
+Shen, “Flying on point clouds: Online trajectory generation and autonomous navigation for quadrotors in cluttered environments,” Journal of Field Robotics, vol.
+36, no.
+4, pp.
+710–733, 2019, doi: 10.1002/rob.21842.
+
+\bibitem{elfes1989using} A. Elfes, “Using occupancy grids for mobile robot perception and navigation,” Computer, vol. 22, no. 6, pp. 46–57, Jun. 1989, doi: 10.1109/2.30720.
+
+\bibitem{xu2021voxel} Y.
+Xu, X.
+Tong, and U.
+Stilla, “Voxel-based representation of 3D point clouds: Methods, applications, and its potential use in the construction industry,” Automation in Construction, vol.
+126, p.
+103675, Jun.
+2021, doi: 10.1016/j.autcon.2021.103675.
+
+\bibitem{niessner2013realtime} M.
+Nießner, M.
+Zollhöfer, S.
+Izadi, and M.
+Stamminger, “Real-time 3D reconstruction at scale using voxel hashing,” ACM Trans.
+Graph., vol.
+32, no.
+6, pp.
+1–11, Nov.
+2013, doi: 10.1145/2508363.2508374.
+
+\bibitem{dryanovski2010multivolume} I.
+Dryanovski, W.
+Morris, and J.
+Xiao, “Multi-volume occupancy grids: An efficient probabilistic 3D mapping model for micro aerial vehicles,” in Proc.
+of the 2010 IEEE/RSJ International Conference on Intelligent Robots and Systems, Taipei, Oct.
+2010, pp.
+1553–1559.
+doi: 10.1109/IROS.2010.5652494.
+
+\bibitem{thrun2002probabilistic} S. Thrun, Probabilistic robotics, vol. 45, 2002. Accessed: Jun. 14, 2023. [Online]. Available: https://dl.acm.org/doi/10.1145/504729.504754
+
+\bibitem{luo2018collisionfree} L.
+Luo et al., “Collision-Free Path-Planning for Six-DOF Serial Harvesting Robot Based on Energy Optimal and Artificial Potential Field,” Complexity, vol.
+2018, pp.
+1–12, Nov.
+2018, doi: 10.1155/2018/3563846.
+
+\bibitem{zhang2021obstacle} W.
+Zhang, H.
+Cheng, L.
+Hao, X.
+Li, M.
+Liu, and X.
+Gao, “An obstacle avoidance algorithm for robot manipulators based on decision-making force,” Robotics and Computer-Integrated Manufacturing, vol.
+71, p.
+102114, Oct.
+2021, doi: 10.1016/j.rcim.2020.102114.
+
+\bibitem{han2018dynamic} D.
+Han, H.
+Nie, J.
+Chen, and M.
+Chen, “Dynamic obstacle avoidance for manipulators using distance calculation and discrete detection,” Robotics and Computer-Integrated Manufacturing, vol.
+49, pp.
+98–104, Feb.
+2018, doi: 10.1016/j.rcim.2017.05.013.
+
+\bibitem{szeliski2021computer} R. Szeliski, “Computer Vision: Algorithms and Applications, 2nd Edition,” Springer, 2021.
+
+\bibitem{rosenfeld1968distance} A. Rosenfeld and J. L. Pfaltz, “Distance functions on digital pictures,” Pattern Recognition, vol. 1, no. 1, pp. 33–61, Jul. 1968, doi: 10.1016/0031-3203(68)90013-7.
+
+\bibitem{felzenszwalb2012distance} P.
+F.
+Felzenszwalb and D.
+P.
+Huttenlocher, “Distance Transforms of Sampled Functions,” Theory of Comput., vol.
+8, no.
+1, pp.
+415–428, 2012, doi: 10.4086/toc.2012.v008a019.
+
 \bibitem{dryanovski2010multivolume} I.

### 2024-03-19 Articles/IROS/root.tex
added sections description to introduction

@@
 This locally-calculated repulsive field ensures the robot moves safely and efficiently in dynamic environments.
+
+This article introduces a method for obstacle avoidance in robotic manipulators via matrix multiplication-driven repulsive fields.
+Section II outlines the mathematical framework for generating repulsive velocities.
+Section III integrates this with inverse kinematic control, focusing on obstacle navigation.
+Section IV presents simulations demonstrating the method's effectiveness in various scenarios, highlighting its potential in robotic motion planning and obstacle avoidance.
 

### 2024-03-19 Articles/IROS/root.tex
removed more references

@@
 
-Motion planning is a technique in robotics, that is used to calculate joint changes leading manipulators from initial to a target configuration, while addressing constraints and optimization criteria.
-This task is facilitated by the manipulators' redundant degrees of freedom (DOF), enabling it to undertake various configurations to not only reach their End Effector (EE) target but also optimize for secondary objectives such as obstacle avoidance and minimizing joint torques~\cite{siciliano1990kinematic, siciliano2010robot}.
+Motion planning~\cite{c52} is a technique in robotics, that is used to calculate joint changes leading manipulators from initial to a target configuration, while addressing constraints and optimization criteria.
+This task is facilitated by the manipulators' redundant degrees of freedom (DOF), enabling it to undertake various configurations to not only reach their End Effector (EE) target but also optimize for secondary objectives such as obstacle avoidance and minimizing joint torques~\cite{siciliano1990kinematic}.
 
 While global motion planning techniques provide comprehensive solutions by examining the entire configuration space, they fall short in terms of smoothness and real-time application, making them less ideal for dynamic settings~\cite{lavalle1998rapidly, gammell2015batch, karaman2010incremental,kuffner2000rrt}.
-Conversely, local planning methods, particularly inverse kinematics~\cite{c29,c38} and quadratic programming~\cite{c21,c23}, offer prompt, smooth trajectories suitable for real-time operations.
+Conversely, local planning methods, particularly inverse kinematics~\cite{c29,c38} and quadratic programming~\cite{c21,c23, haviland2021neo}, offer prompt, smooth trajectories suitable for real-time operations.
 However, these methods are prone to fall into local minima due to their incremental planning approach.
@@
 
-Our approach utilizes occupancy voxel grids to enable safe and dynamic navigation for robot manipulators in changing environments.
+Our approach utilizes occupancy voxel grids~\cite{han2018dynamic} to enable safe and dynamic navigation for robot manipulators in changing environments.
 By representing the workspace with voxels and calculating repulsive velocities based on proximity to obstacles, our method allows the robot to determine safe movement directions and velocities in real-time.
@@
 
+\bibitem{c52} M.
+G.
+Tamizi, M.
+Yaghoubi, and H.
+Najjaran, “A review of recent trend in motion planning of industrial robots,” Int J Intell Robot Appl, vol.
+7, no.
+2, Art.
+no.
+2, Jun.
+2023.
+doi: 10.1007/s41315-023-00274-2.
+
 \bibitem{siciliano1990kinematic} B. Siciliano, “Kinematic control of redundant robot manipulators: A tutorial,” J. Intell. Robot. Syst., vol. 3, no. 3, Art. no. 3, 1990, doi: 10.1007/BF00126069.
 
-\bibitem{siciliano2010robot} B.
+%\bibitem{siciliano2010robot} B.
 Siciliano, L.
@@
 Available: http://link.springer.com/10.1007/978-1-84628-642-1\_4
-
-%\bibitem{vsvestka1997motion} P.
-{\v{S}}vestka and M.
-H.
-Overmars, “Motion planning for carlike robots using a probabilistic learning approach,” The International Journal of Robotics Research, vol.
-16, no.
-2, pp.
-119–143, 1997.
 
@@
 
-%\bibitem{c22} Y.
-Zhang, S.
-S.
-Ge, and T.
-H.
-Lee, “A Unified Quadratic-Programming-Based Dynamical System Approach to Joint Torque Optimization of Physically Constrained Redundant Manipulators,” IEEE Trans.
-Syst., Man, Cybern.
-B, vol.
-34, no.
-5, pp.
-2126–2132, Oct.
-2004.
-doi: 10.1109/TSMCB.2004.830347.
+\bibitem{haviland2021neo} J.
+Haviland and P.
+Corke, “NEO: A Novel Expeditious Optimisation Algorithm for Reactive Motion Control of Manipulators,” IEEE Robot.
+Autom.
+Lett., vol.
+6, no.
+2, Art.
+no.
+2, Apr.
+2021, doi: 10.1109/LRA.2021.3056060.
 
@@
 doi: 10.1109/ROBOT.1985.1087247.
-
-%\bibitem{c40} J.-O. Kim and P. Khosla, “Real-Time Obstacle Avoidance Using Harmonic Potential Functions,” 1992. doi: 10.1109/70.143352.
 
@@
 
-%\bibitem{c46} A.
-H.
-Qureshi et al., “Adaptive Potential guided directional-RRT*,” in Proc.
-2013 IEEE International Conference on Robotics and Biomimetics (ROBIO), Shenzhen, China, Dec.
-2013, pp.
-1887–1892.
-doi: 10.1109/ROBIO.2013.6739744.
-
 \bibitem{c47} J.
@@
 
-%\bibitem{gammell2014informed} J.
-D.
-Gammell, S.
-S.
-Srinivasa, and T.
-D.
-Barfoot, “Informed RRT*: Optimal sampling-based path planning focused via direct sampling of an admissible ellipsoidal heuristic,” in Proc.
-of the 2014 IEEE/RSJ International Conference on Intelligent Robots and Systems, Chicago, IL, USA, Sep.
-2014, pp.
-2997–3004.
-doi: 10.1109/IROS.2014.6942976.
-%
-%	
-%\bibitem{IDEASLab2023}
-%IDEAS Lab, "Motion and Path Planning," presented at Purdue University, 2023. [Online]. Available: \url{https://ideas.cs.purdue.edu/research/robotics/planning/}. Accessed on: Jan. 9, 2024.
-%
-
-\bibitem{haviland2021neo} J.
-Haviland and P.
-Corke, “NEO: A Novel Expeditious Optimisation Algorithm for Reactive Motion Control of Manipulators,” IEEE Robot.
-Autom.
-Lett., vol.
-6, no.
-2, Art.
-no.
-2, Apr.
-2021, doi: 10.1109/LRA.2021.3056060.
-
-\bibitem{c24} J.
-Nakanishi, R.
-Cory, M.
-Mistry, J.
-Peters, and S.
-Schaal, “Comparative experiments on task space control with redundancy resolution,” in Proc.
-2005 IEEE/RSJ Int.
-Conf.
-on Intelligent Robots and Systems, Edmonton, Alta., Canada, 2005, pp.
-3901–3908.
-doi: 10.1109/IROS.2005.1545203.
-
-\bibitem{c25} M.
-H.
-Raibert and J.
-J.
-Craig, “Hybrid Position/Force Control of Manipulators,” Journal of Dynamic Systems, Measurement, and Control, vol.
-103, no.
-2, pp.
-126–133, Jun.
-1981.
-doi: 10.1115/1.3139652.
-
-\bibitem{c26} T.
-Yoshikawa, “Dynamic hybrid position/force control of robot manipulators--Description of hand constraints and calculation of joint driving force,” IEEE J.
-Robot.
-Automat., vol.
-3, no.
-5, pp.
-386–392, Oct.
-1987.
-doi: 10.1109/JRA.1987.1087120.
-
-\bibitem{c27} O.
-Khatib, “A unified approach for motion and force control of robot manipulators: The operational space formulation,” IEEE J.
-Robot.
-Automat., vol.
-3, no.
-1, pp.
-43–53, Feb.
-1987.
-doi: 10.1109/JRA.1987.1087068.
-
-\bibitem{c28} N. Hogan, “Impedance Control: An Approach to Manipulation,” in Proc. 1984 American Control Conf., San Diego, CA, USA, Jul. 1984, pp. 304–313. doi: 10.23919/ACC.1984.4788393.
+\bibitem{han2018dynamic} D.
+Han, H.
+Nie, J.
+Chen, and M.
+Chen, “Dynamic obstacle avoidance for manipulators using distance calculation and discrete detection,” Robotics and Computer-Integrated Manufacturing, vol.
+49, pp.
+98–104, Feb.
+2018, doi: 10.1016/j.rcim.2017.05.013.
 
 \bibitem{c30} L.
-Lajpah and T.
-Petri, “Obstacle Avoidance for Redundant Manipulators as Control Problem,” in Serial and Parallel Robot Manipulators - Kinematics, Dynamics, Control and Optimization, S.
+Žlajpah and T.
+Petrič, “Obstacle Avoidance for Redundant Manipulators as Control Problem,” in Serial and Parallel Robot Manipulators - Kinematics, Dynamics, Control and Optimization, S.
 Kucuk, Ed.
@@
 doi: 10.5772/32651.
-
-\bibitem{c31} R. Colbaugh and K. Glass, “Cartesian control of redundant robots,” J. Robotic Syst., vol. 6, no. 4, pp. 427–459, Aug. 1989. doi: 10.1002/rob.4620060409.
 
@@
 
-\bibitem{c34} L.
-Sciavicco and B.
-Siciliano, “A solution algorithm to the inverse kinematic problem for redundant manipulators,” IEEE J.
-Robot.
-Automat., vol.
-4, no.
-4, pp.
-403–410, Aug.
-1988.
-doi: 10.1109/56.804.
-
-\bibitem{c35} L.
-Sciavicco and B.
-Siciliano, “Solving the Inverse Kinematic Problem for Robotic Manipulators,” in RoManSy 6, A.
-Morecki, G.
-Bianchi, and K.
-Kedzior, Eds., Boston, MA: Springer US, 1987, pp.
-107–114.
-doi: 10.1007/978-1-4684-6915-8\_9.
-
-\bibitem{c36} O. Egeland, “Task-space tracking with redundant manipulators,” IEEE J. Robot. Automat., vol. 3, no. 5, pp. 471–475, Oct. 1987. doi: 10.1109/JRA.1987.1087118.
-
-\bibitem{c37} H. Seraji, “Configuration control of redundant manipulators: theory and implementation,” IEEE Trans. Robot. Automat., vol. 5, no. 4, pp. 472–490, Aug. 1989. doi: 10.1109/70.88062.
-
-\bibitem{c39} B. Siciliano and O. Khatib, Eds., Springer Handbook of Robotics. in Springer Handbooks. Cham: Springer International Publishing, 2016. doi: 10.1007/978-3-319-32552-1.
-
 \bibitem{c41} L.
-Zlajpah and B.
+Žlajpah and B.
 Nemec, “Kinematic control algorithms for on-line obstacle avoidance for redundant manipulators,” in Proc.
@@
 
-\bibitem{c42} T.
-Petrič and L.
-Žlajpah, “Smooth continuous transition between tasks on a kinematic control level: Obstacle avoidance as a control problem,” Robotics and Autonomous Systems, vol.
-61, no.
-9, Art.
-no.
-9, Sep.
-2013.
-doi: 10.1016/j.robot.2013.04.019.
-
-\bibitem{c48} T.
-Zhu, J.
-Mao, L.
-Han, C.
-Zhang, and J.
-Yang, “Real-Time Dynamic Obstacle Avoidance for Robot Manipulators Based on Cascaded Nonlinear MPC With Artificial Potential Field,” IEEE Trans.
-Ind.
-Electron., pp.
-1–11, 2023.
-doi: 10.1109/TIE.2023.3306405.
-
-\bibitem{c50} Y.
-Chen, L.
-Chen, J.
-Ding, and Y.
-Liu, “Research on Real-Time Obstacle Avoidance Motion Planning of Industrial Robotic Arm Based on Artificial Potential Field Method in Joint Space,” Applied Sciences, vol.
-13, no.
-12, p.
-6973, Jan.
-2023.
-doi: 10.3390/app13126973.
-
-\bibitem{c51} S. M. LaValle, Planning Algorithms. Cambridge: Cambridge University Press, 2006.
-
-\bibitem{c52} M.
-G.
-Tamizi, M.
-Yaghoubi, and H.
-Najjaran, “A review of recent trend in motion planning of industrial robots,” Int J Intell Robot Appl, vol.
-7, no.
-2, Art.
-no.
-2, Jun.
-2023.
-doi: 10.1007/s41315-023-00274-2.
-
-\bibitem{siciliano2016springer} B. Siciliano and O. Khatib, Eds., Springer Handbook of Robotics, in Springer Handbooks. Cham: Springer International Publishing, 2016. doi: 10.1007/978-3-319-32552-1.
-
-\bibitem{dai2022review} Y.
-Dai, C.
-Xiang, Y.
-Zhang, Y.
-Jiang, W.
-Qu, and Q.
-Zhang, “A Review of Spatial Robotic Arm Trajectory Planning,” Aerospace, vol.
-9, p.
-361, Jul.
-2022, doi: 10.3390/aerospace9070361.
-
-\bibitem{gottschalk1996obbtree} S.
-Gottschalk, M.
-C.
-Lin, and D.
-Manocha, “OBBTree: a hierarchical structure for rapid interference detection,” in Proceedings of the 23rd Annual Conference on Computer Graphics and Interactive Techniques, ACM, Aug.
-1996, pp.
-171–180.
-doi: 10.1145/237170.237244.
-
-\bibitem{vandenbergen1997efficient} G.
-van den Bergen, “Efficient Collision Detection of Complex Deformable Models using AABB Trees,” Journal of Graphics Tools, vol.
-2, no.
-4, pp.
-1–13, 1997.
-doi: 10.1080/10867651.1997.10487480.
-
-\bibitem{chen2018path} G.
-Chen, D.
-Liu, Y.
-Wang, Q.
-Jia, and X.
-Zhang, “Path planning method with obstacle avoidance for manipulators in dynamic environment,” International Journal of Advanced Robotic Systems, vol.
-15, no.
-6, Art.
-no.
-1729881418820223, Nov.
-2018, doi: 10.1177/1729881418820223.
-
-\bibitem{puiu2011realtime} D.
-Puiu and F.
-Moldoveanu, “Real-time collision avoidance for redundant manipulators,” in Proc.
-of the 2011 6th IEEE International Symposium on Applied Computational Intelligence and Informatics (SACI), Timisoara, Romania, 2011, pp.
-403–408, doi: 10.1109/SACI.2011.5873037.
-
-\bibitem{wurmOctoMap} K.
-M.
-Wurm, A.
-Hornung, M.
-Bennewitz, C.
-Stachniss, and W.
-Burgard, “OctoMap: A Probabilistic, Flexible, and Compact 3D Map Representation for Robotic Systems,” [Details of publication, e.g., in Proc.
-of the Conference/Journal Name, Year, pp.
-Page numbers].
-[DOI or URL if available].
-
-\bibitem{gao2019flying} F.
-Gao, W.
-Wu, W.
-Gao, and S.
-Shen, “Flying on point clouds: Online trajectory generation and autonomous navigation for quadrotors in cluttered environments,” Journal of Field Robotics, vol.
-36, no.
-4, pp.
-710–733, 2019, doi: 10.1002/rob.21842.
-
-\bibitem{elfes1989using} A. Elfes, “Using occupancy grids for mobile robot perception and navigation,” Computer, vol. 22, no. 6, pp. 46–57, Jun. 1989, doi: 10.1109/2.30720.
-
-\bibitem{xu2021voxel} Y.
-Xu, X.
-Tong, and U.
-Stilla, “Voxel-based representation of 3D point clouds: Methods, applications, and its potential use in the construction industry,” Automation in Construction, vol.
-126, p.
-103675, Jun.
-2021, doi: 10.1016/j.autcon.2021.103675.
-
-\bibitem{niessner2013realtime} M.
-Nießner, M.
-Zollhöfer, S.
-Izadi, and M.
-Stamminger, “Real-time 3D reconstruction at scale using voxel hashing,” ACM Trans.
-Graph., vol.
-32, no.
-6, pp.
-1–11, Nov.
-2013, doi: 10.1145/2508363.2508374.
-
-\bibitem{dryanovski2010multivolume} I.
-Dryanovski, W.
-Morris, and J.
-Xiao, “Multi-volume occupancy grids: An efficient probabilistic 3D mapping model for micro aerial vehicles,” in Proc.
-of the 2010 IEEE/RSJ International Conference on Intelligent Robots and Systems, Taipei, Oct.
-2010, pp.
-1553–1559.
-doi: 10.1109/IROS.2010.5652494.
-
-\bibitem{thrun2002probabilistic} S. Thrun, Probabilistic robotics, vol. 45, 2002. Accessed: Jun. 14, 2023. [Online]. Available: https://dl.acm.org/doi/10.1145/504729.504754
-
-\bibitem{luo2018collisionfree} L.
-Luo et al., “Collision-Free Path-Planning for Six-DOF Serial Harvesting Robot Based on Energy Optimal and Artificial Potential Field,” Complexity, vol.
-2018, pp.
-1–12, Nov.
-2018, doi: 10.1155/2018/3563846.
-
-\bibitem{zhang2021obstacle} W.
-Zhang, H.
-Cheng, L.
-Hao, X.
-Li, M.
-Liu, and X.
-Gao, “An obstacle avoidance algorithm for robot manipulators based on decision-making force,” Robotics and Computer-Integrated Manufacturing, vol.
-71, p.
-102114, Oct.
-2021, doi: 10.1016/j.rcim.2020.102114.
-
-\bibitem{han2018dynamic} D.
-Han, H.
-Nie, J.
-Chen, and M.
-Chen, “Dynamic obstacle avoidance for manipulators using distance calculation and discrete detection,” Robotics and Computer-Integrated Manufacturing, vol.
-49, pp.
-98–104, Feb.
-2018, doi: 10.1016/j.rcim.2017.05.013.
-
-\bibitem{szeliski2021computer} R. Szeliski, “Computer Vision: Algorithms and Applications, 2nd Edition,” Springer, 2021.
-
-\bibitem{rosenfeld1968distance} A. Rosenfeld and J. L. Pfaltz, “Distance functions on digital pictures,” Pattern Recognition, vol. 1, no. 1, pp. 33–61, Jul. 1968, doi: 10.1016/0031-3203(68)90013-7.
-
-\bibitem{felzenszwalb2012distance} P.
-F.
-Felzenszwalb and D.
-P.
-Huttenlocher, “Distance Transforms of Sampled Functions,” Theory of Comput., vol.
-8, no.
-1, pp.
-415–428, 2012, doi: 10.4086/toc.2012.v008a019.
-
-\bibitem{dryanovski2010multivolume} I.
-Dryanovski, W.
-Morris, and J.
-Xiao, “Multi-volume occupancy grids: An efficient probabilistic 3D mapping model for micro aerial vehicles,” in 2010 IEEE/RSJ International Conference on Intelligent Robots and Systems, Taipei: IEEE, Oct.
-2010, pp.
-1553–1559, doi: 10.1109/IROS.2010.5652494.
-
 \end{thebibliography}

### 2024-03-19 Articles/IROS/root.tex
removed historic versions
Mentor Corrections: PAGE 1

@@
 
-%Our methodology employs occupancy voxel grids to create repulsive velocities.
-These velocities dynamically guide robots away from obstacles based on proximity, facilitating safe navigation.
-By adjusting repulsive forces in real time, our approach ensures robots maintain a safe distance from hazards, effectively integrating global and local planning strategies to navigate dynamic environments efficiently.
-This pragmatic solution leverages voxel-based spatial assessments to avoid potential collisions, ensuring pathfinding and obstacle avoidance.
-%
-%\alert{---}
-%
-%\subsection{Motion Planning}
-%
-%Motion planning is a process of finding a sequence of joint configurations that lead the manipulator from initial joints configuration to the goal configuration, while adhering to the constraints and optimization criteria.
-This is possible because the manipulator has a redundant number of joints~\cite{siciliano1990kinematic}.
-That means that it has more degrees of freedom (DOF) - joints, than it needs to execute the primary task, that is reaching the End Effector (EE) goal.
-Consequently, the robot can adopt different constrained joint configurations optimised according to the secondary task while performing the primary task.
-Common secondary tasks~\cite{siciliano2010robot} include avoiding singularities, optimising the manipulability measure, minimising joint torques and avoiding obstacles in the operating space.
-%
-%%Kinematic redundancy~\cite{siciliano1990kinematic, siciliano2016springer} enables a manipulator to follow a predefined task space trajectory using the endeffector (EE), while simultaneously, optimising for an additional task with the remaining movement capacity without impacting the trajectory adherence.
-This is possible because the robot's degrees of freedom (DOF) go beyond what is required to perform the primary task.
-Consequently, the robot can adopt different joint configurations optimised according to the secondary task while performing the primary task.
-Common secondary tasks~\cite{siciliano2010robot} include avoiding singularities, optimising the manipulability measure, minimising joint torques and avoiding obstacles in the operating space.
-%%
-%%The task of finding the joint trajectories of a manipulator is called motion planning~\cite{IDEASLab2023}.
-It consists of finding a sequence of joint configurations for a robot so that the robot can move along this path from its initial configuration to the goal configuration without colliding with itself, static obstacles or other agents in the environment.
-%
-%%In addition to collision avoidance, motion planning for manipulators can optionally take into account various constraints, such as position, velocity, acceleration or jerk constraints for the joint angle or end effector, precision of the end effector with respect to position and orientation, stability of the manipulator, avoidance of singularities, or any number of other criteria.
-%
-%There are global and local motion planning methods.
-Global, sampling-based, methods~\cite{vsvestka1997motion,lavalle1998rapidly, karaman2010incremental,kuffner2000rrt,gammell2015batch} offer a globally optimal solution based on a global search in configuration space.
-However, the generated trajectories are not smooth and optimal, and the performance of methods is insufficient for real-time operation.
-%
-%%There are global and local motion planning methods.
-Global, sampling-based, methods such as PRM~\cite{vsvestka1997motion}, RRT*~\cite{lavalle1998rapidly, karaman2010incremental} RRT-Connect~\cite{kuffner2000rrt}, Informerd RRT*~\cite{gammell2014informed}, BIT*~\cite{gammell2015batch} offer a globally optimal solution based on a global search in configuration space.
-However, the generated trajectories are not smooth and optimal, and the performance of methods is insufficient for real-time operation.
-%
-%Local motion planning methods employ optimization techniques, two common ones are inverse kinematics~\cite{c29,c38}, that is based on finding a least squares solution of the manipulator joint velocities, and quadratic programming (QP)~\cite{c21,c23}.
-Both methods are fast, suitable for real-time applications in dynamic environments and provide smooth solutions.
-However, since they do not plan further than one step ahead, they tend to get stuck in local minima.
-%
-%Often best solution to motion planning is a combination of a higher-level planner, for global static environment based path planning and local planner, that takes dynamic environment changes into account.
-%
-%%\subsection{Kinematic Obstacle Avoidance}
-%%
-%%When using inverse kinematics control for motion planning, one method is augmentation of task space to encompass supplementary parameters, in addition to the position and orientation of the end effector~\cite{c34, c35, c36}.
-This method returns a singular solution of a weighted combination of both tasks.
-Alternatively, our paper utilizes task prioritization~\cite{c29, c38, c41}, which involves calculating the null space for the primary task velocities and applying only those secondary velocities that do not directly influence the primary task velocities.
-%
-%\subsection{Artificial Potential Field}
-%
-%Khatib~\cite{c33} proposes the concept of an Artificial Potential Field.
-This field consists of a repulsive component that deflects the end effector away from obstacles, depicted as geometric primitives, and an attractive component that draws the manipulator towards its target.
-In the following years many different modifications and improvement of the original APF idea have been proposed, often focused on removal of local minimas in the potential field.
-Khim and Khosla~\cite{c40} suggested the use of harmonic functions to solve the problem of local minima.
-Pinto et al.
-\cite{c43} proposes to vary the field based on the distance of robot from obstacles to fill the local minima.
-Many researchers tried using APF as a heuristic to better guide sampling based approaches~\cite{c45, c46, c47}.
-%
-%Many of the recent works focus on use of a variation of artificial potential field to plan motion of the manipulator.
-Xia et al.~\cite{c49} uses a variation of APF for manipulator motion~\cite{c49}.
-Park et al.~\cite{park2020trajectory} used a numerical Jacobian in combination with APF for motion planning.
-%
-%%Zhang et al.~\cite{zhang2021obstacle} proposes dynamic repulsive field based on direction and speed between point on robot and obstacle, it also suggests decision making force that moves the robot away from certain local minima.
-%
-%%Chen et al.
-proposes\remove{ an application }using APF in joint space and a variable kinematic optimization step~\cite{c50}.
-Long~\cite{c44} suggests creating motion plan of the manipulator using APF, he extends it using RRT to calculate virtual attractive point for the robot to move towards in case of local minima.
-Zhu et al.~\cite{c48} proposes use of APF in combination with MPC, to plan in environment with dynamic obstacles.
-%%
-%
-%\subsection{ESDF creation}
-%
-%ESDF grids can be generated directly using sensor measurements.
-Oleynikova et al.~\cite{oleynikova2017voxblox} proposed a method for calculating the ESDF from TSDF.
-Han et al.~\cite{han2019fiesta} proposed a way to integrate point-cloud data into ESDF using ray-casting.
-%
-%Another way is to generate ESDF from occupancy grids, that can be previously generated using sensors or based on predefined maps.
-When generating ESDF from occupancy grid, common approach is the Brushfire method~\cite{lau2010improved}, that spreads from obstacles until it calculates the distance for every field on a grid.
-Jump Flooding Algorithm (JFA)~\cite{rong2006jump} is a similar method, that can be implemented on a GPU for faster parallelized distance calculation.
-%
-%Methods that allow for the dynamical distance field generation are rare.
-Zhou et al.~\cite{zhou2021egoplanner} propose use of pairs of points on trajectory and obstacles in their quadcopter trajectory optimization algorithm.
-%
-%%\remove{The field of image processing invented many different morphological algorithms, that work on binary images and calculate a value for every pixel in image, that represents the distance from nearest occupied binary pixel~\cite{szeliski2021computer, felzenszwalb2012distance, rosenfeld1968distance}.
-Approaches can be seperated based on the distance metric that they use.
-}
-%
-%%Euclidean Signed Distance Fields (ESDFs) are valuable in robotic navigation and 3D reconstruction, providing a measure of the shortest distance to the nearest obstacle for each point in space.
-The creation of ESDF grids can leverage sensor measurements directly or be derived from other data structures like Truncated Signed Distance Fields (TSDFs) or occupancy grids.
-%%
-%%One prominent method for generating ESDFs is converting TSDFs into ESDFs.
-Oleynikova et al.
-[(2017)] introduced a technique that efficiently computes the ESDF from TSDFs, utilizing incremental updates for dynamic environments.
-This approach is significant for real-time applications in robotics, where rapid updates to the environment model are crucial.
-Similarly, Han et al.
-[(2019)] developed a method that integrates point-cloud data into ESDF using ray-casting, offering a way to construct ESDFs directly from sensor data.
-%%
-%%Alternatively, ESDFs can be generated from occupancy grids, which themselves are often derived from sensor measurements.
-The Brushfire algorithm, as improved by Lau et al.
-[(2010)], is a commonly used technique in this context.
-It operates by spreading from obstacles within the grid, calculating the minimum distance to an obstacle for each grid cell.
-This approach is particularly effective for static environments where the layout of obstacles does not change frequently.
-%%
-%%Furthermore, the Jump Flooding Algorithm (JFA) offers another method for generating ESDFs from occupancy grids.
-JFA is notable for its suitability for implementation on GPUs, enabling faster parallel distance calculations.
-This advantage is critical in applications requiring rapid processing of large datasets, such as in high-resolution 3D mapping and complex robotic navigation tasks.
-%%
-%%In conclusion, the generation of ESDFs from sensor measurements or derived data structures like TSDFs and occupancy grids is a crucial aspect of robotic navigation and 3D environment modeling.
-Techniques like those proposed by Oleynikova et al.
-[(2017)], Han et al.
-[(2019)], and the improvements to the Brushfire algorithm by Lau et al.
-[(2010)], as well as the application of JFA, highlight the ongoing advancements in efficient and accurate ESDF creation methodologies.
-%
-%
-%
-%\alert{---}
-%
-%In dynamic environments, a robot manipulator must swiftly respond to emerging obstacles.
-Our approach enables the robot to determine safe movement directions and velocities by real-time assessment of nearby space for obstacles.
-%
-%In 3D graphics, a voxel is like a tiny cube in a grid that makes up the 3D space.
-We use voxels to represent the workspace of the robot, each voxel has a value that tells us if it’s probabilistically empty or occupied.
-A zero means empty, and a one means definitely something is there, with values in between showing varying levels of certainty.
-%
-%%Mreža voxels je lahko predefinirana, glede na model / 3d zemljevid prostora.
-Zasedenost voxlov lahko spreminjamo glede na poznane pozicije in trajektorije preostalih agentov v prostoru.
-Kot omenjeno v uvodu, pa lahko zasedenost voxlov pridobimo tudi z senzorskimi sistemi.
-%
-%The occupancy grid of voxels can be set in advance to match the layout of the space.
-We can also change the values in the voxels if we know where other objects or agents are moving.
-Besides pre-setting, we can use depth sensors to detect what's around.
-%
-%Our method calculates repulsive velocities from existing occupancy voxel grid, which are vectors that guide the robot on where to go by pointing away from potential collisions.
-These vectors gain strength as the robot approaches an obstacle, signaling an urgent need to maneuver to safety.
-Conversely, when the robot is positioned at an equal distance from obstacles in the vicinity—safely away from all potential collisions—the repulsive velocities weaken because there’s no immediate need to move.
-%
-%%In our method, we compute repulsive velocities within the task space using a novel matrix kernel multiplication approach.
-Concentrating on the task space is advantageous as it provides a more direct and realistic representation of the environment.
-%
-%%\replace{Naša metoda je posebno primerna za uporabo z senzorskimi sistemi kot so LIDAR ali globinske kamere, saj zaradi upoštevanja celotne okolice točke in ne le razdalje do najbližje točke v okolici efektivno filtriramo senzorski šum.}{prestavi to v diskusijo}
-%
-%Repulsive velocities tell the agent in which direction to move, so that it avoids nearby obstacles.
-These velocities drop to zero when the agent maintains a minimum safe distance from obstacles, and rise to their highest when it nears an obstacle, facilitating immediate evasive action.
-As the repulsive field calculation is locally based, it will also go to zero when the agent is surrounded by all directions, equally spaced from all sides.
-That is, it is in the best local minima away from all the obstacles in its proximity.
-
 \section{REPULSIVE FIELD CALCULATION}
@@
 
-This section introduces our novel approach for calculating repulsive fields, a critical component in enabling robots to detect and avoid obstacles in real-time.
-By leveraging a voxel-based representation of the surrounding space, our method dynamically assesses potential collision threats and calculates directional repulsive velocities.
-These velocities are essential for guiding the robot away from obstacles, ensuring smooth and safe navigation through complex environments.
-Our innovative use of kernel convolution methods allows for the precise calculation of avoidance velocities, providing a robust solution for obstacle avoidance that can be integrated seamlessly into robotic control systems.
-The following subsections delve into the technical details of our approach, including the mapping of Cartesian space to a discrete occupancy grid, kernel selection for optimized field calculation, and the application of tri-linear interpolation for smooth velocity transitions.
-
-%\subsection{MAPPING}
-
-%The occupancy map representation of the robots environment is discrete (has finite resolution), while the Cartesian space is continuous, as rounding the point directly to the center of the nearest occupancy grid voxel, based on Euclidean distance.\add{is this clear, do i need equation?} However, this discretization can sometimes lead to discontinuities.
-Therefore, we propose a second approach: tri-linear interpolation of the calculated repulsive field to achieve a continuous repulsive field value.
+We leverage a voxel-based representation of the surrounding space, to dynamically assesses potential collision threats and calculates directional repulsive velocities for smooth and safe navigation through complex environments.
 
 \subsection{KERNELS AND THEIR MATRIX APPLICATIONS}
-%\subsection{Selecting Kernels and Matrix Operations}
-
-The fundamental concept of our directional kernels lies in computing the repulsive field individually for each direction within the Cartesian coordinate system.
-
-%\remove{Our filters structure was inspired by the Sobel  operator, a 2D convolutional filter frequently utilized in computer vision for calculating image gradients at specific points.}
-
-Our kernels are designed as three-dimensional matrixes with a primary kernel axis aligned along a specific Cartesian direction, corresponding to the calculated repulsive velocity.
-The two secondary kernel axes are orthogonal to this primary axis.
-The distribution of values along the primary axis is inversely symmetric, exhibiting positive values on one side and negative values on the other, with the zero valued cell in the center of the kernel, where jump between max positive and max negative magnitude is.
+
+The fundamental concept of directional kernels lies in computing the repulsive field individually for each direction within the Cartesian coordinate system.
+
+Kernels are designed as three-dimensional matrixes with a primary kernel axis aligned along a specific Cartesian direction, corresponding to the calculated repulsive velocity.
+The two secondary kernel axes are orthogonal to the primary axis.
+The distribution of values along the primary axis is inversely symmetric, exhibiting positive values on one side and negative values on the other, with the zero valued cell in the center of the kernel, where the values jump from max negative to max positive magnitude (Fig.~\cite{fig:kernels}).
 The function of the increase in magnitude along the primary axis of the kernel defines the shape of the repulsive velocity field, determining how the repulsive velocity changes as the agent approaches an obstacle.
@@
 
-% KERNEL WEIGHTS
-
-% PRIMARY AXIS
-
-%We propose two different primary axis weights distributions, of course there is no reason why any other distribution of weights could not be used.
-The choice of the weights distribution should depend on the profile of the repulsive velocities we want to archieve for the APF.
-%
-%The first of the proposed functions is a mirrored normal / gaussian distribution. By changing the sigma we can control how fast or slow does the field value grow when we approach obstacles.
-%
-%$\Delta i = c_i - i$, $\Delta j = c_j - j$ in $\Delta k = c_k - k$ predstavljajo število celic odmika od centralnega polja matrike, v katerem se nahaja naša točka na agentu v posamezno koordinatno smer.
-
-We introduce two distinct distributions for the primary axis weights, although other distributions could also be applicable.
+We introduce two distributions for the primary axis weights.
 The selection of weight distribution is contingent on the desired profile of repulsive velocities within the Artificial Potential Field (APF).
 
-The inaugural proposed function adheres to a mirrored normal or Gaussian distribution, wherein adjusting the standard deviation (\(\sigma\)) modulates the rate at which field values escalate as obstacles are approached.
-
-The terms \(\Delta i = c_i - i\), \(\Delta j = c_j - j\), and \(\Delta k = c_k - k\) denote the displacement count from the matrix's central field, where the field components are calculated, in their respective coordinate directions.
+First proposed function is mirrored normal or Gaussian distribution, the standard deviation (\(\sigma\)) modulates the rate at which field values escalate as obstacles are approached.
+
+Terms \(\Delta i = c_i - i\), \(\Delta j = c_j - j\), and \(\Delta k = c_k - k\) denote the displacement index from the matrix's central field, where fields' components are calculated.
 
@@
 
-%\begin{equation}
-%w_i = 
-%\begin{cases} 
-%	\frac{e^{-\frac{\Delta i^2}{2\sigma^2}}}{\sigma \sqrt{2\pi}} & \text{if } \Delta i > 0 \\
-%	0 & \text{if } \Delta i = 0 \\
-%	-\frac{e^{-\frac{\Delta i^2}{2\sigma^2}}}{\sigma \sqrt{2\pi}} & \text{if } \Delta i < 0 
-%\end{cases}
-%\end{equation}
-
 Another distribution we used is mirrored linear, where the $\Delta i_{max} = \lfloor \frac{\mathrm{range}}{\Delta \mathrm{R}} \rfloor$ is the rounded down half length of the primary axis kernel.
@@
 
-% ORTHAGONAL AXIS
-
 If the matrix would be only 1 field width and height the field would work kind of as ray tracing in each of the main cartesian coordinate directions.
@@
 However as we want bigger repulsive field when the obstacle is head on in the direction than when the obstacle is off the cardinal direction, we propose the following multiplicator, to account for the off direction obstacles.
-
-%\remove{In the orthogonal axes, the magnitude of values diminishes with increasing distance from the core of the kernel, nearing zero at the edges.
-This gradation is vital for a smooth velocity transition when obstacles emerge in the kernels range.
-The peak magnitude is located at the kernel's core, along the primary axis.
-The polarity of the values is determined by their position relative to the primary axis.
-}
 
@@
 
-%Once the weights of the repulsive kernels have been generated, we can use kernel multiplication to get the repulsive velocity components.
-This method is tailored to calculate the components of the avoidance velocity vector in the Cartesian space, we generate the corresponding kernel and extract a segment of the obstacle grid of matching size, centered at agent or the point of interest (POI), resulting in two same size 3D matrices—one is a "window" from the obstacle grid $A_d$ and the other representing the kernel $K_d$.
-
 Following the establishment of repulsive kernel weights, kernel multiplication is employed to ascertain the repulsive velocity components within Cartesian space.
@@
 
-%If the agent is located near the edge of our known voxel grid we can set the elements of the window would be located in the space beyond the matrix as empty in which case the robot might want to move towards this space or as occupied, which will prevent the robot from moving out of the known grid space.
-
 Positioning of the agent at the edge of the known voxel grid necessitates a strategic decision regarding the window's boundary elements.
 These elements, located beyond the matrix's spatial domain, can either be considered vacant—encouraging the robot to explore these new areas—or occupied—thereby deterring the robot from exiting the mapped grid space.
-
-%By employing the Hadamard (element-wise) product (eq.~\ref{eq:matrix-product}) between the cutout segment of the obstacle grid \(G_d\) and the corresponding 3D convolutional kernel \(W_d\) and than summoning all of the matrix values for each of the directions $d \in \{x, y, z\}$, we derive the resultant repulsive velocities vector \( \vec{v}_i = \left[ v_{x_i} v_{y_i} v_{z_i}\right] \).
 
@@
 
-%\begin{equation}
-%	\dot{x}_{poi} = \sum_{i}\sum_{j}\sum_{k} c_{dijk} 
-%	\label{eq:summation}
-%\end{equation}
-%
-
 \comment{}{PLOT: kernel 2D images}
@@
 \subsection{3D INTERPOLATION}
-
-%The occupancy map representation of the robots environment is discrete (has finite resolution), while the Cartesian space is continuous.
-In the transition from Cartesian space coordinates to discrete occupancy grid representations, approximating or rounding can induce discontinuities in the resulting repulsive field.
-To mitigate this effect, we employ tri-linear interpolation, enabling a seamless transition to a continuous repulsive field value (Fig.
-\ref{fig:interp-experiment}).
-%
-%It is essential that the velocity contributions affecting the robot change smoothly.
-However, since our obstacle grid is discretely defined, achieving perfect continuity can be challenging.
-Increasing the resolution of the obstacle field can theoretically bring us closer to continuous behavior, but in practice, we are constrained by finite resolution of our sensors and of the grid.
-Poleg tega bo višja resolucija posledično povzročila povečanje spomina potrebnega za shranjevanje zemljevida / mreže okolja in potrebne procesorske moči.
-To ensure that the velocity remains continuous when transitioning from one cell of the obstacle grid to another at a point of interest (POI), we employ trilinear interpolation.
 
@@
 We get the indexes of the surrounding cells by first scaling the POI position by grid resolution and than rounding the position to the nearest lower and upper integer positions (eq.~\ref{eq:floor and ceil}).
-
-%\begin{equation}
-% X = \left[ \lfloor p_{x} \rfloor, \lceil p_{x} \rceil \right] 
-%\end{equation}
-%
-%\begin{equation}
-% Y = \left[ \lfloor p_{y} \rfloor, \lceil p_{y} \rceil \right] 
-%\end{equation}
-%
-% \begin{equation}
-% Z = \left[ \lfloor p_{z} \rfloor, \lceil p_{z} \rceil \right] 
-%\end{equation}
 
@@
 
-%\begin{equation}
-%	\Delta X = \frac{\left( P_x - \left( X(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( X(2) - X(1) \right)}
-%\end{equation}
-%
-%\begin{equation}
-%	\Delta Y = \frac{\left( P_y - \left( Y(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Y(2) - Y(1) \right)}
-%\end{equation}
-%
-%\begin{equation}
-%	\Delta Z = \frac{\left( P_z - \left( Z(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Z(2) - Z(1) \right)}
-%\end{equation}
-
 \begin{equation}
@@
 
-%\begin{equation}
-%	\vec{V}_{rep,00} = \vec{V}_{rep,000}(1 - x_d) + \vec{V}_{rep,100}x_d
-%\end{equation}
-%\begin{equation}
-%	\vec{V}_{rep,01} = \vec{V}_{rep,001}(1 - x_d) + \vec{V}_{rep,101}x_d
-%\end{equation}
-%\begin{equation}
-%	\vec{V}_{rep,10} = \vec{V}_{rep,010}(1 - x_d) + \vec{V}_{rep,110}x_d
-%\end{equation}
-%\begin{equation}
-%	\vec{V}_{rep,11} = \vec{V}_{rep,011}(1 - x_d) + \vec{V}_{rep,111}x_d
-%\end{equation}
-
 \begin{equation}
@@
 
-%\begin{equation}
-%	\vec{V}rep_{xyz,0} = \vec{V}rep_{xyz,00}(1 - \Delta y) + \vec{V}rep_{xyz,10} \, \Delta y
-%\end{equation}
-%
-%\begin{equation}
-%	\vec{V}rep_{xyz,1} = \vec{V}rep_{xyz,01}(1 - \Delta y) + \vec{V}rep_{xyz,11} \, \Delta y
-%\end{equation}
-
 \begin{equation}
@@
 
-%The final result is a smooth repulsive velocity field that transitions smoothly between the discrete values calculated at distinct points in the obstacle grid.
-
 The end product is a uniformly smooth repulsive velocity field, facilitating continuous transitions across discrete points determined within the obstacle grid.
@@
 	\centering
-	\includegraphics[width=1\linewidth]{non_and_interp_4_plots.eps} % Replace 'example-image' with the filename of your image
+	\includegraphics[width=1\linewidth]{non_and_interp_4_plots.eps} 
 	\caption{Repulsive field generation via convolutional kernels.
@@
 
-%\begin{equation}
-%	\dot{\vec{q}} = J^+ \xi_{p} \mathrm{\vec{v}_{att}} + \dot{\vec{q}}_{rep}
-%\end{equation}
-
 \begin{equation}
@@
 
-%\remove{For redundant robots, it optimizes solution selection by minimizing the joint velocity norm, while damping is applied to limit excessive joint velocities near singular configurations, thus improving solution robustness.}
-
-%\add{maybe add the middle joints component in the equation (it is not necessary, doesn't make much change in the chosen experiments)}
-
 We run the inverse kinematics algorithm once for every time step, until we reach the desired cost function or reach the selected time limit.
@@
 \subsection{END-EFFECTOR VELOCITY}
-
-%Vodenje vrha manipulatorja (end-effector) je naloga z najvišjo prioriteto.
 
@@
 \end{equation}
-
-%In the above equation (eq.~\ref{eq: v_att}), $\vec{v}$  represents the end effector's translational velocity towards the target, combining direction and magnitude.
-The terms $\vec{x}_{EE}$ and $\vec{x}_g$ denote the current and goal positions of the EE, respectively, in Cartesian coordinates.
-The unit vector calculation, $\frac{\vec{x}_{EE}-\vec{x}_g}{||\vec{x}_{EE}-\vec{x}_g||}$, ensures motion directed towards the target.
-Finally, the sigmoid function, particularly the arctangent component, modulates this velocity to avoid overshooting, balancing speed and precision.
-The EE velocity will be transformed using the sigmoid curve when near the goal.
-When the distance from goal point is high, the sigmoid will cause the velocity to get into saturation, that is upper limit.
-The constant $k_{sigm}$ allows us to set how close to the goal does the robot EE start slowing down and what is the velocity profile when approaching the goal point.
 
@@
 
-%The rotational error of the EE is needed for ensuring goal orientation of the EE.
-In our approach, orientations are depicted using rotation matrices.
-Specifically, \( R \) represents the current EE orientation, while \( gR \) signifies the goal EE orientation.
-The disparity between these orientations is encapsulated by the relative rotation matrix \( dR \).
-This matrix is formulated by multiplying the goal orientation matrix \( gR \) with the transpose of the current orientation matrix \( cR^{T} \).
-To ensure that it represents a pure rotation without any scaling we than normalize the so gotten matrix .
-
-%Determination of the requisite rotation adjustment utilizes the relative rotation matrix \(dR\), calculated from the multiplication of \(gR\) by the \(cR\) transpose, and subsequent normalization to ensure a pure rotational representation.
-This method guarantees that \(dR\) accurately captures the orientation discrepancy, devoid of any scaling factor.
-
-%The current orientation of the end-effector (EE) and its goal orientation are encoded through rotation matrices, \(cR\) for the current state and \(gR\) for the goal state.
-The necessary rotational adjustment is identified through the relative rotation matrix \(dR\), generated by multiplying \(gR\) with the transpose of \(cR\) and then normalizing the outcome (eq.
-\ref{eq: rot_diff_mat}).
-This process ensures \(dR\) accurately represents rotational differences, without any scaling factors.
-%
-%To enhance computational efficiency and prevent gimbal lock, the approach transitions from matrix to quaternion representations for precise and efficient rotational velocity adjustments, essential for achieving the target orientation of the EE.
-
 The current orientation of the end-effector (EE) and its goal orientation are encoded through rotation matrices, \(\mathbf{cR}\) for the current state and \(\mathbf{gR}\) for the goal state.
@@
 
-%\begin{equation}
-%	dR = \frac{gR \cdot cR^{T}}{||gR \cdot cR^{T}||}
-%	\label{eq: rot_diff_mat}
-%\end{equation}
-
-%\begin{equation}
-%	\boldsymbol{dR} = \frac{\boldsymbol{gR} \cdot \boldsymbol{cR}^{T}}{\left\lVert\boldsymbol{gR} \cdot \boldsymbol{cR}^{T}\right\rVert}
-%	\label{eq: rot_diff_mat}
-%\end{equation}
-
 \begin{equation}
@@
 The resulting quaternion's components, excluding the real part, constitute the rotational error vector \( \vec{\omega} \).
-%
-%\begin{equation}
-%	dR \mapsto dQ = a + b \, i + c \, j + d \, k
-%	\label{eq: quat_mapsto}
-%\end{equation}
-%
-%\begin{equation}
-%	dQl = 2 \cdot \log(dQ) = al + bl \, i + cl \, j + dl \, k
-%	\label{eq:quat_log}
-%\end{equation}
 
@@
 
-%Ker uporabljamo preslikavo v ničelni prostor primarne hitrosti (null space), lahko v primeru velikih hitrosti primarne naloge preostane premalo prostostnih stopenj, da bi se lahko varno izognili oviram.
-Torej sekundarna naloga izogibanja nima dovolj sposobnosti, prostosti za delovanje.
-%
-%To ensure the primary task doesn't overpower the secondary task, we've integrated a primary task execution slowdown.
-This mechanism can reduce the manipulator's velocity towards its primary goal, leaving more maneuverability space for the secondary tasks.
-%
-%\begin{equation}
-%	\label{eq:slowdown}
-%	\xi_{p}=
-%	\frac{1}{1 + \kappa_{\text{sec}} v_{\text{Rmax}}}
-%\end{equation}
-%
-%The slowdown factor (eq.~\ref{eq:slowdown}) is influenced by the constant \( \kappa_{\text{sec}} \) and the maximum repulsive velocity vectors magnitude, which depends on robot's minimum distance from an obstacle \( v_{\text{Rmax}} \).
-We calculate the minimal distance in the repulsive velocities phase of the algorithm, as explained in section~\ref{chap:repulsive velocity}.
-As per the equation, a small \( v_{\text{Rmax}} \) minimizes the slowdown effect, allowing uninterrupted primary task execution.
-Conversely, a large \( v_{\text{Rmax}} \) increases the slowdown, by making the $\xi_{p}$ factor smaller, giving the secondary task more time for corrective actions.
-
 Utilizing the mapping into the null space of the primary velocity, it becomes evident that at high velocities of the primary task, there may be an insufficient number of degrees of freedom remaining to safely navigate around obstacles.
@@
 \subsection{AVOIDANCE VELOCITIES}
-
-%To move the manipulator away from the obstacles in its environment, we cover the manipulator evenly with virtual detection points (POI).
-Gostota virtualnih točk je odvisna od velikosti matrik.
-Če je $w$ širina matrike v smeri pravokotno na primarno smer, je dobro, če so točke medsebojno razmaknjene glede na potek segmentov za $\Delta = \frac{w}{3}$.
-Tedaj se bodo posamezne matrike deloma pokrivale in dosežemo enakomerno porazdelitev točk (POI) po celotnem robotu.
-Med izvajanjem kinematične optimizacije izračunavamo izogibne / odbojne hitrosti za vsako od točk po postopku opisanem v zgornjem poglavju.
-\add{referenca poglavja} To dobimo tako, da izvedemo preslikavo vsake od točk v task space, kjer izvajamo izračun odbojne hitrosti.
-Pri tem je posamezna kartezična preslikava sestavljena iz kinematične preslikave iz baze robota do začetka segmenta na katerem se točka nahaja in dodane preslikave do izbrane točke (od začetka segmenta do dela kjer se nahaja točka).
-\(T_{0 \rightarrow \text{POI}} = T_{0 \rightarrow 1} \cdot T_{1 \rightarrow 2} \cdot \ldots \cdot T_{(j-1) \rightarrow j} \cdot T_{j \rightarrow \text{POI}}\) Za vsako od točk dobimo tri komponente kartezične hitrosti.
 
@@
 
-%\begin{equation}
-%	T_{0 \rightarrow \text{POI}} = T_{0 \rightarrow 1} \cdot T_{1 \rightarrow 2} \cdot \ldots \cdot T_{(j-1) \rightarrow j} \cdot T_{j \rightarrow \text{POI}}	
-%	\label{eq:transformations}
-%\end{equation}
-
 \begin{equation}
@@
 
-%Ko dobimo za vsako od točk odbojno hitrost (repulsive / avoidance velocity), izračunamo drugo normo vsake od hitrosti.
-Za $K$ točk, v katerih so izogibne hitrosti največje in so posledično najbližje oviri nato izračunamo sklepne hitrosti, ki povzročijo izogibanje oviri.
-
 Upon obtaining the repulsive velocity for each point, the second norm of each velocity is calculated.
@@
 
-%\begin{equation}
-%J_{d_0}^+ = N J_{d_0}' (J_{d_0} N J_{d_0}' + \sigma_{rep})^{-1};
-%\end{equation}
-
-%\begin{equation}
-%	J_{d_i}^{+} = N J_{d_i}' (J_{d_i} N J_{d_i}' + \sigma_{rep} I)^{-1}, \quad \text{for } i = 1, 2, \ldots, K;
-%\end{equation}
-
 \begin{equation}
@@
 \end{equation}
-
-%Za vsako od izbranih točk na robotu izračunamo Jacobijevo transformacijo hitrosti, ki preslika translacijske hitrosti izbrane točke na robotu iz kartezičnega prostora, v prostor sklepov na robotu, ki povezujejo bazo z izbrano točko.
-V izbranih točkah nas ne zanimajo transformacije napak orientacij.
-V resnici nas zanima samo komponenta translacijske hitrosti v smeri normale, ki kaže od najbližje točke na objektu proti izbrani točki interesa (POI) na manipulatorju.
-
-%We calculate translational velocity Jacobian for every of the $K$ selected points on the robot.
-Since we are only interested in the part of velocity that leads us directly away from the obstacles, we can
-
-%
-%Naš prostor operacij se posledično skrči na eno dimenzijo in the Jacobian, which relates the joint space velocities q and the velocity in the direction of do, can be calculated as  $J_{d_i} = \vec{n}_i^T J_i$.
-\add{citat:žlajpah} Pri čemer je vektor normale želene avoidance velocity kar enotski vektor hitrosti, ki ga dobimo s pomočjo naših matrik $\vec{n}_{i} = \frac{\vec{v}_{poi i}}{||\vec{v}_{poi i}||}$.
-This method offers computational efficiency by eliminating complex matrix inversions, as the Jacobian changes from matrix into a vector dimensions $1 \times n$, where n is number of joints that are located from before the POI and simplifying repulsive velocity calculations.
 
@@
 
-%Celotne izogibne sklepne hitrosti dobimo z obteženo vsoto izogibnih sklepnih hitrosti izbranih $K$ najbližjih točk, pri čemer upoštevamo vpliv primarne naloge gibanja vrha manipulatorja na sklepne hitrosti.
-
 The overall avoidance joint velocities are obtained through a weighted sum of the avoidance joint velocities for the selected \(K\) nearest points, taking into account the influence of the primary task of manipulator EE movement on the joint velocities.
@@
 \subsection{Repulsive Field Visualization}
-
-%Prvo predstavimo rezultirajoče se odbojno polje za 2D - dvo dimenzijski zemljevid.
-Načeloma je delovanje na treh dimenzijah enako, seveda pa je to težje vizualizirati.
-Pri tem smo uporabili našo metodo izračuna odbojnega potencialnega polja hitrosti.
-Namesto uporabe v točkah robotskem manipulatorju smo točke vzorčili preko celotnega zemljevida ovir, kot je vidno na figur \ref{fig:example-field}.
-V vsaki od vzorčenih točk smo izvedli v članku opisani postopek pridobivanja odbojnega polja v X in Y smeri.
-Na spodnjem prikazu figure smo nato vizualizirali odbojno poljo kot vektorsko polje sestavljene iz X in Y komponente.
 
@@
 \end{figure}
-%
-%\add{add label, caption, every plot has different axis scale}
-%
-%\comment{}{we dont have the close obstacles bottlenecks, we do have weird bottlenecks in the orthagonal directions - maybe}
-%
+
 \subsection{Manipulator Examples}
-
-%Delovanje potencialnega polja smo prikazali na dveh različnih primerih manipulatorja.
-Pri tem je ločljivost naše mreže voxlov ovir $R=10cm$.
-Uporaba interpolacije nam omogoča, da imamo relativno grobo mrežo voxlov, kar zniža spominsko zahtevnost mreže prostora, ter pospeši izračun.
-V primerih smo izbrali $K = 7$ POI točk interesa, v katerih gledamo oddaljenost od ovir, ki so enakomerno razporejene po segmentih in sklepih od drugega sklepa robota do vrha manipulatorja, torej od prve točke naprej, kjer se je manipulator sposoben izogniti oviri.
-POI so na grafih prikazani s pikami na manipulatorju.
-V bližini ovir iz njih segajo vektorji, ki prikazujejo izračunane odbojno hitrost v točkah.
-
-%Med izvajanjem simulacije izvajamo eulerjevo integracijo izračunanih sklepnih hitrosti, s korakom $T_{step}=0.1s$.
 
@@
 
-%V prvem primeru \ref{fig:keyframes-3d-column} manipulator z uporabo potencialnega polja izračunanega na predlagani način, se manipulator varno "zvije" oziroma izbere pot do točke, ki se nahaja na drugi strani stebra.
-Izbrane konstante primarne naloge so $k_v=5$ in $k_w=20$.
-Izbrane konstante obteženosti posameznih POI so $\alpha=\frac{\left[ \frac{3}{9}, \frac{2}{9}, \frac{1}{9}, \frac{1}{9}, \frac{1}{9}, \frac{1}{9}
-% \right]}{10}$.
-Najkrajša pot vrha v točko pelje direktno skozi steber.
-Primarna naloga je, da vrh preide v želeno pozicijo in orientacijo.
-Ker je naloga vrha primarna, moramo uvesti znižanje primarne hitrosti v bližini ovire $\xi_{p}=1$, da ima sekundarna naloga dovolj prostosti in časa, da lahko izvede rekonfiguracijo za obhod stebra.
-Simulacijo izvajamo $50$ korakov.
-
 In the first case (Fig.
@@
 \end{figure}
-
-%V drugem primeru (Fig.
-\ref{fig:keyframes-3d-ball}, \ref{fig:plots-ball}) se naš robot giblje v dinamične okolju.
-Predlagan način izračuna odbojnih hitrosti je dobra izbira v primeru, ko potrebujemo lokalno optimizacijsko metodo za upoštevanje dinamičnih spremeb.
-V izbranem primeru imamo žogo, ki se med potekom optimizacije enakomerno premika iz leve strani grafa proti desni strani (torej v smeri y osi od 0.4 v trenutku 0 do 1.4 v zadnjem koraku simulacije).
-Primarna naloga ima sedaj nalogo vzdrževati vrh robota v eni točki, pri čemer smo za večjo sposobnost izogibanja oviri zanemarili izbiro orientacije vrha EE.
-To nam poveča sposobnost gibanja, posledično ne potrebujemo primary speed task slowdown  \( \xi_{p} = 0 \).
-The constants chosen for the primary task are \( k_v = 2 \) and \( k_w = 0 \), for the avoidance task \(k_r=3\).
-The weighting constants for the individual POIs are \( \alpha = \frac{[ \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5} ]}{10} \).
 
@@
 
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=1\linewidth]{keyframes_2D_a.eps} % Replace 'example-image' with the filename of your image
-%	\caption{Example Image}
-%	\label{fig:keyframes-2d-column}
-%\end{figure}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=1\linewidth]{keyframes_2_columns.eps} % Replace 'example-image' with the filename of your image
-%	\caption{Example Image}
-%	\label{fig:keyframes-2-column}
-%\end{figure}
-%
-%
-%\add{EXAMPLE: MANIPULATOR AND MOVING NOISY BALL}
-%
 \begin{figure}
@@
 \end{figure}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=1\linewidth]{keyframes_2D_ball.eps} % Replace 'example-image' with the filename of your image
-%	\caption{Example Image}
-%	\label{fig:keyframes-2d-ball}
-%\end{figure}
-%
+
 \begin{figure}
@@
 \bibitem{siciliano1990kinematic} B. Siciliano, “Kinematic control of redundant robot manipulators: A tutorial,” J. Intell. Robot. Syst., vol. 3, no. 3, Art. no. 3, 1990, doi: 10.1007/BF00126069.
-
-%\bibitem{siciliano2010robot} B.
-Siciliano, L.
-Sciavicco, L.
-Villani, and G.
-Oriolo, Robot.
-Model.
-Plan.
-Control, 2010, pp.
-161–189.
-[Online].
-Available: http://link.springer.com/10.1007/978-1-84628-642-1\_4
 

### 2024-03-19 Articles/IROS/root.tex
Mentor Corrections: PAGE 2

@@
 
+Kernel multiplication is employed to ascertain the repulsive velocity components within Cartesian space.
+We select a segment from the occupation grid, that is centered in the point of interest (POI) on the robot, where we want to calculate the repulsive velocities.
+The segment needs to be of the same size as the directional matrix, for the cordinate axis in which we are calulating the repulsive velocity.
+This results in two identically sized 3D matrices: one serving as a "window" into the obstacle grid (\(G_d\)) and the other as the directional kernel (\(W_d\)).
+
+Applying the Hadamard (element-wise) product to the segment of the occupation grid \(G_d\) and the kernel \(W_d\) results in a new matrix for each Cartesian direction (\(d \in \{x, y, z\}\)).
+By summing over all values in this resultant matrix, we calculate the repulsive velocity component for that specific direction.
+This process is independently conducted for each Cartesian coordinate, ultimately yielding the comprehensive repulsive velocity vector \(\vec{v}_i = [ v_{x_i}, v_{y_i}, v_{z_i}]\).
+
+\begin{equation}
+	v_d = \sum_{i,j,k} (G \odot W)_{ijk} = \sum_{i}\sum_{j}\sum_{k} g_{\Delta i \Delta j \Delta k} ~  w_{\Delta i \Delta j \Delta k}
+	\label{eq:matrix-product}
+\end{equation}
+
+If the agent is positioned at the edge of the known voxel grid, elements beyond the window's boundary must be filled.
+These elements, located outside the matrix's spatial domain, can either be considered vacant—encouraging the robot to explore these new areas—or occupied—thereby deterring the robot from exiting the mapped grid space.
+
 \subsection{KERNELS AND THEIR MATRIX APPLICATIONS}
@@
 
-Kernels are designed as three-dimensional matrixes with a primary kernel axis aligned along a specific Cartesian direction, corresponding to the calculated repulsive velocity.
-The two secondary kernel axes are orthogonal to the primary axis.
-The distribution of values along the primary axis is inversely symmetric, exhibiting positive values on one side and negative values on the other, with the zero valued cell in the center of the kernel, where the values jump from max negative to max positive magnitude (Fig.~\cite{fig:kernels}).
+Kernels are designed as three-dimensional matrixes with a primary kernel axis $i$ aligned along a specific Cartesian direction, corresponding to the calculated repulsive velocity.
+The two secondary kernel axes $j,k$ are orthogonal to the primary axis.
+The distribution of values along the primary axis is inversely symmetric, exhibiting positive values on one side and negative values on the other, with the zero valued cell in the center of the kernel, where the values jump from max negative to max positive magnitude (Fig.~\cite{fig:kernels} \alert{TBA}).
 The function of the increase in magnitude along the primary axis of the kernel defines the shape of the repulsive velocity field, determining how the repulsive velocity changes as the agent approaches an obstacle.
@@
 
-We introduce two distributions for the primary axis weights.
+We introduce two distributions for the primary axis weights, namely gaussian and linear.
 The selection of weight distribution is contingent on the desired profile of repulsive velocities within the Artificial Potential Field (APF).
@@
 First proposed function is mirrored normal or Gaussian distribution, the standard deviation (\(\sigma\)) modulates the rate at which field values escalate as obstacles are approached.
-
-Terms \(\Delta i = c_i - i\), \(\Delta j = c_j - j\), and \(\Delta k = c_k - k\) denote the displacement index from the matrix's central field, where fields' components are calculated.
 
@@
 
-Another distribution we used is mirrored linear, where the $\Delta i_{max} = \lfloor \frac{\mathrm{range}}{\Delta \mathrm{R}} \rfloor$ is the rounded down half length of the primary axis kernel.
+Terms \(\Delta i = c_i - i\), \(\Delta j = c_j - j\), and \(\Delta k = c_k - k\) denote the displacement index from the matrix's central field, where fields' components are calculated.
+
+Second proposed distribution is mirrored linear, where the $\Delta i_{max} = \lfloor \frac{\mathrm{range}}{\Delta \mathrm{R}} \rfloor$ is the rounded down half length of the primary axis kernel.
 
@@
 If the matrix would be only 1 field width and height the field would work kind of as ray tracing in each of the main cartesian coordinate directions.
-Since our need is to detect also obstacles that dont align perfectly along the cartesian direction, it is important that our matrixes have width and height.
+Since our need is to detect also obstacles that dont align perfectly along the cartesian direction, it is important that our matrixes have width ($i$ axis) and height ($j$ axis).
 However as we want bigger repulsive field when the obstacle is head on in the direction than when the obstacle is off the cardinal direction, we propose the following multiplicator, to account for the off direction obstacles.
 
-The length $\Delta j_{max} = \lfloor \frac{\mathrm{width}}{2 \Delta \mathrm{R}} \rfloor$ of the orthogonal axes influences the peripheral detection range for obstacles.
+The length $\Delta j_{max} = \lfloor \frac{\mathrm{width}}{2 \Delta \mathrm{R}} \rfloor$ and height $\Delta k_{max} = \lfloor \frac{\mathrm{height}}{2 \Delta \mathrm{R}} \rfloor$ of the orthogonal axes influences the peripheral detection range for obstacles.
 Excessively wide kernels may generate repulsive velocities for objects that are not in the path of the robot, whereas too narrow kernels might only detect obstacles aligned directly with the Cartesian direction in the point of interest.
@@
 
-Finally we get three matrix kernels, one for each of the three coordinate axis directions by multiplieing weights components for primary and orthagonal axis.
+By multiplieing weights components for primary and both orthagonal axis we get matrix kernels. We get one kernel for every cartesian direction.
 
@@
 	w_{\Delta i \Delta j \Delta k} = w_{\Delta i} * w_{\Delta j} * w_{\Delta k}
-\end{equation}
-
-Following the establishment of repulsive kernel weights, kernel multiplication is employed to ascertain the repulsive velocity components within Cartesian space.
-We select a corresponding segment from the obstacle grid, aligning in size and centered around the point of interest (POI) on the robot, where we want to calculate the repulsive velocities.
-This results in two identically sized 3D matrices: one serving as a "window" into the obstacle grid (\(A_d\)) and the other as the representative kernel (\(K_d\)).
-
-Positioning of the agent at the edge of the known voxel grid necessitates a strategic decision regarding the window's boundary elements.
-These elements, located beyond the matrix's spatial domain, can either be considered vacant—encouraging the robot to explore these new areas—or occupied—thereby deterring the robot from exiting the mapped grid space.
-
-Applying the Hadamard (element-wise) product to the segment of the obstacle grid \(G_d\) and the kernel \(W_d\) results in a new matrix for each Cartesian direction (\(d \in \{x, y, z\}\)).
-By summing over all values in this resultant matrix, we calculate the repulsive velocity component for that specific direction.
-This process is independently conducted for each Cartesian coordinate, ultimately yielding the comprehensive repulsive velocity vector \(\vec{v}_i = [ v_{x_i}, v_{y_i}, v_{z_i}]\).
-
-\begin{equation}
-	v_d = \sum_{i,j,k} (G \odot W)_{ijk} = \sum_{i}\sum_{j}\sum_{k} g_{\Delta i \Delta j \Delta k} ~  w_{\Delta i \Delta j \Delta k}
-	\label{eq:matrix-product}
 \end{equation}
@@
 
-We start by scalling the coordinates of POI into the grid koordinate system, by multiplying it by grid resolution (eq.~\ref{eq:scale grid}).
-
-\begin{equation}
-	\label{eq:scale grid}
-	\vec{P} = \vec{p}_{\mathrm{POI}} \times\Delta \mathrm{R}
-\end{equation}
-
-We get the indexes of the surrounding cells by first scaling the POI position by grid resolution and than rounding the position to the nearest lower and upper integer positions (eq.~\ref{eq:floor and ceil}).
+The indexes of the eight surrounding cells are calculated by scaling the POI $x,y,z$ by grid resolution and than rounding the position to the nearest lower and upper integer positions (eq.~\ref{eq:floor and ceil}).
 
@@
 	\begin{bmatrix}
-		\; \lfloor \vec{p}_{\mathrm{POI}}(1) \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(1) \rceil \;  \\
-		\; \lfloor \vec{p}_{\mathrm{POI}}(2) \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(2) \rceil \; \\
-		\; \lfloor \vec{p}_{\mathrm{POI}}(3) \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(3) \rceil \; 
+		\; \lfloor \vec{p}_{\mathrm{POI}}(1) \times\Delta \mathrm{R} \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(1) \times\Delta \mathrm{R} \rceil \;  \\
+		\; \lfloor \vec{p}_{\mathrm{POI}}(2) \times\Delta \mathrm{R} \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(2) \times\Delta \mathrm{R} \rceil \; \\
+		\; \lfloor \vec{p}_{\mathrm{POI}}(3) \times\Delta \mathrm{R} \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(3) \times\Delta \mathrm{R} \rceil \; 
 	\end{bmatrix}
@@
 
-Once we got the indexes of the eight surrounding cells of our POI, we use our kernel matrix multiplication method, to calculate the 3x1 repulsive velocity vectors for all the cells (eq.~\ref{eq: calc rep vel}).
+Once we got the indexes of the eight surrounding cells of our POI, we calculate the repulsive velocity vectors in each point using kernel matrix multiplication method (seperately for each cartesian direction) (eq.~\ref{eq: calc rep vel}).
 
@@
 The calculated repulsive velocity values are located at the centers of the cells.
-Therefore, before interpolation, we shift the values of the cells coordinates by half the resolution of the obstacle grid for each direction.
-That is, the cell coordinates are same as x,y,z, cell index and a half multiplied by grid resolution $\Delta \mathrm{R}$.
+Therefore, before interpolation, we shift the values of the cells coordinates by half the resolution of the obstacle grid in each direction.
+That is, the cell coordinates are same as $x,y,z$, cell index and a half multiplied by grid resolution $\Delta \mathrm{R}$.
 
@@
 	\label{eq: deltas interp}
-	\Delta \vec{P} =
 	\begin{bmatrix}
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{manipulator-kernels-visualization-small.eps} % Replace 'example-image' with the filename of your image
+	\caption{Visualization of 3 repulsive Y-direction kernels in POI on the manipulator.}
+	\label{fig:kernels}
+\end{figure}
+
 To enable the avoidance of obstacles by the manipulator within its operating environment, the manipulator is uniformly coated with virtual detection points (POIs).
@@
 
-\add{describe weights and distributions used in the kernels}
+\add{describe weights and distributions used in experiments the kernels}
 

### 2024-03-20 Articles/IROS/root.tex
changed interpolation euations
removed the rotational calculation section

@@
 	\label{eq: calc rep vel}
-	\vec{V}rep_{xyz,ijk} = \mathrm{calc\_rep\_vel}(X[i], Y[j], Z[k]) \quad \forall i, j, k \in \{1, 2\}
+	\vec{V}rep_{xyz,ijk} = \mathrm{rep}(X[i], Y[j], Z[k]) \quad \forall i, j, k \in \{1, 2\}
 \end{equation}
@@
 \begin{equation}
-	\dot{\vec{q}} = \mathbf{J}^+ \xi_{p} \mathrm{\vec{v}{att}} + \dot{\vec{q}}_{rep}
+	\dot{\vec{q}} = \mathbf{J}^+ \xi_{p} \mathrm{\vec{v}_{att}} + \dot{\vec{q}}_{rep}
 \end{equation}
@@
 \begin{equation} 
-	\vec{v} = \frac{\vec{x}_{EE} - \vec{x}_g}{||\vec{x}_{EE} - \vec{x}_g||} \times \frac{\arctan(k_{sigm} \; ||\vec{x}_{EE} - \vec{x}_g||) }{pi/2}
+	\vec{v} = \frac{\vec{x}_{EE} - \vec{x}_g}{||\vec{x}_{EE} - \vec{x}_g||} \times \frac{\arctan(k_{sigm} \; ||\vec{x}_{EE} - \vec{x}_g||) }{\pi/2}
 	\label{eq: v_att}
@@
 In Eq.~\ref{eq: v_att}, $\vec{v}$ specifies the end-effector's (EE) translational velocity directed towards the target, integrating both its direction and magnitude.
-Here, $\vec{x}_{EE}$ represents the EE's current position, and $\vec{x}g$ is the target position, both in Cartesian coordinates.
-The calculation $\frac{\vec{x}{EE}-\vec{x}g}{||\vec{x}{EE}-\vec{x}g||}$ generates a unit vector pointing towards the target, ensuring targeted movement.
+Here, $\vec{x}_{EE}$ represents the EE's current position, and $\vec{x}_g$ is the target position, both in Cartesian coordinates.
+The calculation $\frac{\vec{x}_{EE}-\vec{x}_g}{||\vec{x}_{EE}-\vec{x}_g||}$ generates a unit vector pointing towards the target, ensuring targeted movement.
 The velocity's modulation by the arctangent sigmoid function curtails overshooting by moderating speed in proximity to the target and saturating velocity at a predefined upper limit when distant.
@@
 The current orientation of the end-effector (EE) and its goal orientation are encoded through rotation matrices, \(\mathbf{cR}\) for the current state and \(\mathbf{gR}\) for the goal state.
-The necessary rotational adjustment is identified through the relative rotation matrix \(\mathbf{dR}\), generated by multiplying \(\mathbf{gR}\) with the transpose of \(\mathbf{cR}\) and then normalizing the outcome (eq.
-\ref{eq: rot_diff_mat}).
-This process ensures \(\mathbf{dR}\) accurately represents rotational differences, without the effect of any scaling factors.
-
-\begin{equation}
-	\mathbf{dR} = \frac{\mathbf{gR} \cdot \mathbf{cR}^{T}}{||\mathbf{gR} \cdot \mathbf{cR}^{T}||}
-	\label{eq: rot_diff_mat}
-\end{equation}
-
-To enhance computational efficiency and prevent gimbal lock, the approach transitions from matrix to quaternion representations for precise and efficient rotational velocity adjustments, essential for achieving the target orientation of the EE.
-
-The relative rotation matrix is converted into a quaternion, which is subsequently logarithmically transformed.
+The necessary rotational adjustment is identified through the relative rotation matrix \(\mathbf{dR}\), which is transformed into a quaternion representation, where we than calculate the rotational error using method presented in the article \cite{Zlajpah2017}.
+%
+%\begin{equation}
+%	\mathbf{dR} = \frac{\mathbf{gR} \cdot \mathbf{cR}^{T}}{||\mathbf{gR} \cdot \mathbf{cR}^{T}||}
+%	\label{eq: rot_diff_mat}
+%\end{equation}
+%
+%To enhance computational efficiency and prevent gimbal lock, the approach transitions from matrix to quaternion representations for precise and efficient rotational velocity adjustments, essential for achieving the target orientation of the EE.
+%
+%The relative rotation matrix is converted into a quaternion, which is subsequently logarithmically transformed.
 The resulting quaternion's components, excluding the real part, constitute the rotational error vector \( \vec{\omega} \).
-
-\begin{equation}
-	\boldsymbol{dR} \mapsto \boldsymbol{dQ} = a + b \, \vec{i} + c \, \vec{j} + d \, \vec{k}
-	\label{eq: quat_mapsto}
-\end{equation}
-
-\begin{equation}
-	\boldsymbol{dQl} = 2 \cdot \log(\boldsymbol{dQ}) = a_{log} + b_{log} \, \vec{i} + c_{log} \, \vec{j} + d_{log} \, \vec{k}
-	\label{eq:quat_log}
-\end{equation}
-
-\alert{There needs to be an explanation why log, cite: DMP Quaternions article Žlajpah, Ude}
-
-\begin{equation}
-	\vec{\omega} =
-	\begin{bmatrix}
-		b_{log} \\
-		c_{log} \\
-		d_{log}
-	\end{bmatrix}
-	\label{eq:rot_error_vector}
-\end{equation}
+%
+%\begin{equation}
+%	\boldsymbol{dR} \mapsto \boldsymbol{dQ} = a + b \, \vec{i} + c \, \vec{j} + d \, \vec{k}
+%	\label{eq: quat_mapsto}
+%\end{equation}
+%
+%\begin{equation}
+%	\boldsymbol{dQl} = 2 \cdot \log(\boldsymbol{dQ}) = a_{log} + b_{log} \, \vec{i} + c_{log} \, \vec{j} + d_{log} \, \vec{k}
+%	\label{eq:quat_log}
+%\end{equation}
+%
+%\alert{There needs to be an explanation why log, cite: DMP Quaternions article Žlajpah, Ude}
+%
+%
+%
+%\begin{equation}
+%	\vec{\omega} =
+%	\begin{bmatrix}
+%		b_{log} \\
+%		c_{log} \\
+%		d_{log}
+%	\end{bmatrix}
+%	\label{eq:rot_error_vector}
+%\end{equation}
 
@@
 \end{figure}
+
+\comment{}{fix units in  plots}
 
@@
 
+\bibitem{Zlajpah2017}
+L.
+Zlajpah, “On orientation control of functional redundant robots,” in Proc.
+of the 2017 IEEE International Conference on Robotics and Automation (ICRA), Singapore, May 2017, pp.
+2475–2482.
+doi: 10.1109/ICRA.2017.7989288.
+
 \end{thebibliography}

### 2024-03-20 Articles/IROS/root.tex
added abstract

@@
 
-\add{ADD}
+The article presents a new method for obstacle avoidance in robotic manipulators.
+It uses occupancy voxel grids to calculate repulsive velocities based on the proximity of obstacles.
+This approach allows for safe and agile navigation in changing environments.
+Repulsive velocities generated through matrix multiplication are processed using tri-linear interpolation, resulting in smooth and efficient movement away from potential hazards.
+The integration of this method with inverse kinematic control improves manipulator motion planning and obstacle avoidance capabilities.
+Simulation results indicate the effectiveness of this method in various scenarios, showcasing its potential for local motion planning and obstacle avoidance.
 
@@
 \section{INTRODUCTION}
-
-%The robots environment is modelled by discrete voxels.
-As the robots environment can dynamically change, we propose a method that looks at the surrounding space of the robot and calculates these direction away from all the surrounding obstacles in real time.
-We only look in a predefined area / perimeter arround the robot.
-%
-%In 3D computer graphics, a voxel represents a value on a regular grid in three-dimensional space.
-Each of the voxels holds the probability value of its occupation.
-In case of voxel being empty it holds the value of 0, if the voxel is occupied it holds the value of non-zero, depending on our assurance of it being occupied.
-If it is definitly occupied it holds the value of 1.
-
-%Motion planning involves finding paths for manipulators through joint configurations from an initial to a goal position, respecting constraints and optimization criteria due to redundant degrees of freedom (DOF)~\cite{siciliano1990kinematic}.
-This redundancy allows for diverse configurations, optimizing secondary tasks like avoiding singularities and obstacles while maintaining the primary goal of reaching the End Effector (EE) target~\cite{siciliano2010robot}.
-%
-%Global motion planning methods search the entire configuration space for optimal paths but lack real-time applicability and smoothness~\cite{vsvestka1997motion,lavalle1998rapidly, karaman2010incremental,kuffner2000rrt,gammell2015batch}.
-Local methods, notably inverse kinematics and quadratic programming, offer real-time, smooth solutions but can get stuck in local minima due to their short-sighted planning approach~\cite{c29,c38,c21,c23}.
-%
-%The Artificial Potential Field (APF) concept by Khatib~\cite{c33}, which integrates repulsive forces for obstacle avoidance and attractive forces towards goals, has seen various improvements to mitigate issues like local minima~\cite{c40,c43,c45,c46,c47}.
-Recent works have adapted APF variations for manipulator motion planning, combining them with techniques like numerical Jacobians for enhanced path planning~\cite{c49,park2020trajectory}.
-%
-%For dynamic environments, strategies like ESDF grids, calculated directly from sensor data or occupancy grids, assist in real-time obstacle detection and avoidance~\cite{oleynikova2017voxblox,han2019fiesta,lau2010improved,rong2006jump,zhou2021egoplanner}.
-Our novel approach utilizes occupancy voxel grids to generate repulsive velocities, dynamically directing the robot away from obstacles based on proximity, thus enabling effective navigation in changing environments.
-This method ensures the robot maintains a safe distance from obstacles, with repulsive velocities adjusting in real-time to navigate the manipulator through the safest path possible.
 

### 2024-03-20 Articles/IROS/root.tex
added conclusion

@@
 
-\end{abstract}
-
 The article presents a new method for obstacle avoidance in robotic manipulators.
@@
 Simulation results indicate the effectiveness of this method in various scenarios, showcasing its potential for local motion planning and obstacle avoidance.
+
+\end{abstract}
 
@@
 	\centering
-	\includegraphics[width=1\linewidth]{manipulator-kernels-visualization-small.eps} % Replace 'example-image' with the filename of your image
+	\includegraphics[width=1.0\linewidth]{manipulator-kernels-visualization-small.eps} % Replace 'example-image' with the filename of your image
 	\caption{Visualization of 3 repulsive Y-direction kernels in POI on the manipulator.}
@@
 
-\add{ADD}
+%The article proposes a new method for calculating repulsive velocities for obstacle avoidance when path planning using for robotic manipulators.
+The environment is presented using discretized voxel grids, that represent obstacles.
+The presented method uses matrix multiplication to generate values for the repuslive field in each of the centers of this voxels.
+Tri-linear interpolation is than used to generate a smooth and continues velocity field for the points located between the centers of voxels.
+This method was than combined with an modified inverse-kinematics optimization approach, that combines EE goal task execution and obstacle avoidance using these fields.
+Method is highely efficient, as the voxels can be quite big, which speeds up the calculation of the avoidance velocities and makes lower grid representation size.
+The method was tested in the simulated environemnt, where it was shown that the mathod is applicable for local path planning and for dynamic enviornemt path corrections.
+%
+%In the future we aim to modify this repulsive field, to be one kernel shaped arround the manipulator instead of the seperate directional kernels located on the POI.
+As the balance of multiple repulsive velocities can be complicated, we plan to use learning based approach, to create a combined controller for primary and avoidance task.
+Potential path of research is also to use Model Predictive Controller (MPC), to calculate movement of the arm in accordance with more than just one step ahead.
+
+This study introduces an innovative methodology for computing repulsive velocities to facilitate obstacle avoidance in the motion planning of robotic manipulators.
+It employs discretized voxel grids to model environmental obstacles and utilizes matrix multiplication for calculating repulsive field values at voxel centers.
+Tri-linear interpolation is applied to achieve a continuous and smooth velocity field between voxels.
+The approach is further integrated with a modified inverse kinematics optimization strategy, which concurrently manages the end-effector's goal attainment and obstacle circumvention.
+The method's efficiency is heightened by the permissible largeness of voxel dimensions, which accelerates the computation of avoidance velocities and allows for a coarser grid resolution.
+The technique's efficacy and applicability for local path planning and dynamic environment adjustments were validated through simulation.
+
+Future work will focus on refining the repulsive field generation by encapsulating the manipulator within a singular kernel, rather than relying on separate directional kernels at points of interest.
+Given the complexity of balancing multiple repulsive velocities, a learning-based approach is proposed for developing a unified controller for primary tasks and obstacle avoidance.
+Additionally, the exploration of a Model Predictive Control (MPC) strategy is anticipated, aiming to extend the manipulator's motion planning capabilities beyond immediate step considerations.
 

### 2024-03-20 Articles/IROS/root.tex
DRAFT 1 - LONG VERSION

@@
 The two secondary kernel axes $j,k$ are orthogonal to the primary axis.
-The distribution of values along the primary axis is inversely symmetric, exhibiting positive values on one side and negative values on the other, with the zero valued cell in the center of the kernel, where the values jump from max negative to max positive magnitude (Fig.~\cite{fig:kernels} \alert{TBA}).
+The distribution of values along the primary axis is inversely symmetric, exhibiting positive values on one side and negative values on the other, with the zero valued cell in the center of the kernel, where the values jump from max negative to max positive magnitude.
 The function of the increase in magnitude along the primary axis of the kernel defines the shape of the repulsive velocity field, determining how the repulsive velocity changes as the agent approaches an obstacle.
@@
 \end{equation}
-
-\comment{}{PLOT: kernel 2D images}
 

### 2024-03-20 Articles/IROS/root.tex
marked parts to cut off

@@
 
-If the agent is positioned at the edge of the known voxel grid, elements beyond the window's boundary must be filled.
-These elements, located outside the matrix's spatial domain, can either be considered vacant—encouraging the robot to explore these new areas—or occupied—thereby deterring the robot from exiting the mapped grid space.
+\remove{If the agent is positioned at the edge of the known voxel grid, elements beyond the window's boundary must be filled.
+These elements, located outside the matrix's spatial domain, can either be considered vacant—encouraging the robot to explore these new areas—or occupied—thereby deterring the robot from exiting the mapped grid space.}
 
@@
 The length $i_{max}$ of the primary axis is critical, as it dictates the detection range for obstacles.
-Longer kernels can detect obstacles further away from the robot, essentially extending the ’safety zone’ around the robot.
+\remove{Longer kernels can detect obstacles further away from the robot, essentially extending the ’safety zone’ around the robot.
 If the primary axis is too long, it can lead to extra calculations and may cause the robot to unnecessarily avoid obstacles that aren't in its immediate path, making its movement and path planning less efficient.
+}
 
-If the matrix would be only 1 field width and height the field would work kind of as ray tracing in each of the main cartesian coordinate directions.
-Since our need is to detect also obstacles that dont align perfectly along the cartesian direction, it is important that our matrixes have width ($i$ axis) and height ($j$ axis).
+\remove{If the matrix would be only 1 field width and height the field would work kind of as ray tracing in each of the main cartesian coordinate directions.} Since our need is to detect also obstacles that dont align perfectly along the cartesian direction, it is important that our matrixes have width ($i$ axis) and height ($j$ axis).
 However as we want bigger repulsive field when the obstacle is head on in the direction than when the obstacle is off the cardinal direction, we propose the following multiplicator, to account for the off direction obstacles.
@@
 The length $\Delta j_{max} = \lfloor \frac{\mathrm{width}}{2 \Delta \mathrm{R}} \rfloor$ and height $\Delta k_{max} = \lfloor \frac{\mathrm{height}}{2 \Delta \mathrm{R}} \rfloor$ of the orthogonal axes influences the peripheral detection range for obstacles.
-Excessively wide kernels may generate repulsive velocities for objects that are not in the path of the robot, whereas too narrow kernels might only detect obstacles aligned directly with the Cartesian direction in the point of interest.
-When selecting the width and height of the kernel, we must consider the density of the neighboring points of interest on the agent, ensuring that the combination of kernels adequately cover the entire agent's surrounding area.
+\remove{Excessively wide kernels may generate repulsive velocities for objects that are not in the path of the robot, whereas too narrow kernels might only detect obstacles aligned directly with the Cartesian direction in the point of interest.} When selecting the width and height of the kernel, we must consider the density of the neighboring points of interest on the manipulator, ensuring that the combination of kernels adequately cover the entire manipulator's surrounding area.
 
-For point-based agents, such as quadcopters, it is practical to align the dimensions of each matrix—length, width, and height—to be identical.
-This configuration ensures uniform coverage of the spherical area surrounding the agent and reduces the risk of blind spots that might occur when obstacles are close to the agent but situated between the discrete kernels.
+\remove{For point-based agents, such as quadcopters, it is practical to align the dimensions of each matrix—length, width, and height—to be identical.
+This configuration ensures uniform coverage of the spherical area surrounding the agent and reduces the risk of blind spots that might occur when obstacles are close to the agent but situated between the discrete kernels.}
 
-For smooth transitions when approaching the obstacles we propose the following sinus based function for the orthagonal axis distributions.
-The proposed equation is the same for both of the orthagonal matrix axis.
+\replace{For smooth transitions when approaching the obstacles we propose the following sinus based function for the orthagonal axis distributions.
+The proposed equation is the same for both of the orthagonal matrix axis.}{We propose a sinusoidal function for smooth orthogonal axis transitions when approaching obstacles.}
 
@@
 
-Maintaining a smooth change in velocity contributions to the robot is crucial.
+\remove{Maintaining a smooth change in velocity contributions to the robot is crucial.
 However, the discrete nature of our obstacle grid presents a challenge in achieving complete continuity.
@@
 Moreover, an increase in resolution would inevitably require more memory for storing the occupancy grid and additional computational resources.
-Thus, to ensure velocity continuity across adjacent cells within the obstacle grid for a point of interest (POI), trilinear interpolation is employed.
+Thus, to ensure velocity continuity across adjacent cells within the obstacle grid for a point of interest (POI), trilinear interpolation is employed.}
 
@@
 
-The end product is a uniformly smooth repulsive velocity field, facilitating continuous transitions across discrete points determined within the obstacle grid.
+The end product is a uniformly smooth repulsive velocity field vector. \remove{, facilitating continuous transitions across discrete points determined within the obstacle grid.}
 
@@
 
-We run the inverse kinematics algorithm once for every time step, until we reach the desired cost function or reach the selected time limit.
+\remove{We run the inverse kinematics algorithm once for every time step, until we reach the desired cost function or reach the selected time limit. }
 
@@
 The commonly used method, where translational velocity is proportional to squared distance, leads to impractically high initial velocities, which can prevent obstacle avoidance and real-world execution, followed by disproportionately slow velocities when nearing the goal position.
-Our aim is a consistent velocity profile throughout the trajectory, with controlled deceleration near the goal.
-This is achieved by first calculating the unit vector towards the target for direction, then modulating its magnitude using a sigmoid function, specifically the arctangent function (eq.
+\remove{Our aim is a consistent velocity profile throughout the trajectory, with controlled deceleration near the goal.
+This is achieved by first calculating the unit vector towards the target for direction, then modulating its magnitude using a sigmoid function, specifically the arctangent function }%(eq.
 \ref{eq: v_att}).
@@
 
-Utilizing the mapping into the null space of the primary velocity, it becomes evident that at high velocities of the primary task, there may be an insufficient number of degrees of freedom remaining to safely navigate around obstacles.
-Consequently, the secondary task, which is aimed at obstacle avoidance, lacks the requisite capability and freedom to operate effectively.
+Utilizing the mapping into the null space of the primary velocity, it becomes evident that at high velocities of the primary task, there may be an insufficient number of degrees of freedom remaining for the secondary task to safely navigate around obstacles.
+\remove{Consequently, the secondary task, which is aimed at obstacle avoidance, lacks the requisite capability and freedom to operate effectively.}
 
@@
 
-We calculate the translational velocity Jacobian for each of the $K$ selected points on the robot.
-Since our interest lies exclusively in the velocity component that directs us away from obstacles, we consequently narrow our operational space to a single dimension.
+The translational velocity Jacobian is calculated for each of the $K$ selected points on the robot.
+Velocity space can be restricted to just one dimension, on the velocity component that directs movement away from obstacles.
 Consequently, the Jacobian, correlating joint space velocities \(q\) with the directional velocity \(d_i\), is represented as \(J_{d_i} = \vec{n}_i^T \mathbf{J}_i\).

### 2024-03-20 Articles/IROS/root.tex
cut off parts

@@
 
-\remove{If the agent is positioned at the edge of the known voxel grid, elements beyond the window's boundary must be filled.
+%\remove{If the agent is positioned at the edge of the known voxel grid, elements beyond the window's boundary must be filled.
 These elements, located outside the matrix's spatial domain, can either be considered vacant—encouraging the robot to explore these new areas—or occupied—thereby deterring the robot from exiting the mapped grid space.}
@@
 The length $i_{max}$ of the primary axis is critical, as it dictates the detection range for obstacles.
-\remove{Longer kernels can detect obstacles further away from the robot, essentially extending the ’safety zone’ around the robot.
+
+%\remove{Longer kernels can detect obstacles further away from the robot, essentially extending the ’safety zone’ around the robot.
 If the primary axis is too long, it can lead to extra calculations and may cause the robot to unnecessarily avoid obstacles that aren't in its immediate path, making its movement and path planning less efficient.
@@
 
-\remove{If the matrix would be only 1 field width and height the field would work kind of as ray tracing in each of the main cartesian coordinate directions.} Since our need is to detect also obstacles that dont align perfectly along the cartesian direction, it is important that our matrixes have width ($i$ axis) and height ($j$ axis).
+%\remove{If the matrix would be only 1 field width and height the field would work kind of as ray tracing in each of the main cartesian coordinate directions.}
+
+Since our need is to detect also obstacles that dont align perfectly along the cartesian direction, it is important that our matrixes have width ($i$ axis) and height ($j$ axis).
 However as we want bigger repulsive field when the obstacle is head on in the direction than when the obstacle is off the cardinal direction, we propose the following multiplicator, to account for the off direction obstacles.
@@
 The length $\Delta j_{max} = \lfloor \frac{\mathrm{width}}{2 \Delta \mathrm{R}} \rfloor$ and height $\Delta k_{max} = \lfloor \frac{\mathrm{height}}{2 \Delta \mathrm{R}} \rfloor$ of the orthogonal axes influences the peripheral detection range for obstacles.
-\remove{Excessively wide kernels may generate repulsive velocities for objects that are not in the path of the robot, whereas too narrow kernels might only detect obstacles aligned directly with the Cartesian direction in the point of interest.} When selecting the width and height of the kernel, we must consider the density of the neighboring points of interest on the manipulator, ensuring that the combination of kernels adequately cover the entire manipulator's surrounding area.
 
-\remove{For point-based agents, such as quadcopters, it is practical to align the dimensions of each matrix—length, width, and height—to be identical.
+%\remove{Excessively wide kernels may generate repulsive velocities for objects that are not in the path of the robot, whereas too narrow kernels might only detect obstacles aligned directly with the Cartesian direction in the point of interest.}
+
+When selecting the width and height of the kernel, we must consider the density of the neighboring points of interest on the manipulator, ensuring that the combination of kernels adequately cover the entire manipulator's surrounding area.
+
+%\remove{For point-based agents, such as quadcopters, it is practical to align the dimensions of each matrix—length, width, and height—to be identical.
 This configuration ensures uniform coverage of the spherical area surrounding the agent and reduces the risk of blind spots that might occur when obstacles are close to the agent but situated between the discrete kernels.}
+%
+%\replace{For smooth transitions when approaching the obstacles we propose the following sinus based function for the orthagonal axis distributions.
+The proposed equation is the same for both of the orthagonal matrix axis.}{We propose a sinusoidal function for smooth orthogonal axis transitions when approaching obstacles.}
 
-\replace{For smooth transitions when approaching the obstacles we propose the following sinus based function for the orthagonal axis distributions.
-The proposed equation is the same for both of the orthagonal matrix axis.}{We propose a sinusoidal function for smooth orthogonal axis transitions when approaching obstacles.}
+We propose a sinusoidal function for smooth orthogonal axis transitions when approaching obstacles.
 
@@
 
-\remove{Maintaining a smooth change in velocity contributions to the robot is crucial.
+%\remove{Maintaining a smooth change in velocity contributions to the robot is crucial.
 However, the discrete nature of our obstacle grid presents a challenge in achieving complete continuity.
@@
 
-The end product is a uniformly smooth repulsive velocity field vector. \remove{, facilitating continuous transitions across discrete points determined within the obstacle grid.}
+The end product is a uniformly smooth repulsive velocity field vector.
+
+%\remove{, facilitating continuous transitions across discrete points determined within the obstacle grid.}
 
@@
 
-\remove{We run the inverse kinematics algorithm once for every time step, until we reach the desired cost function or reach the selected time limit. }
+%\remove{We run the inverse kinematics algorithm once for every time step, until we reach the desired cost function or reach the selected time limit. }
 
@@
 The commonly used method, where translational velocity is proportional to squared distance, leads to impractically high initial velocities, which can prevent obstacle avoidance and real-world execution, followed by disproportionately slow velocities when nearing the goal position.
-\remove{Our aim is a consistent velocity profile throughout the trajectory, with controlled deceleration near the goal.
+
+%\remove{Our aim is a consistent velocity profile throughout the trajectory, with controlled deceleration near the goal.
 This is achieved by first calculating the unit vector towards the target for direction, then modulating its magnitude using a sigmoid function, specifically the arctangent function }%(eq.
@@
 Utilizing the mapping into the null space of the primary velocity, it becomes evident that at high velocities of the primary task, there may be an insufficient number of degrees of freedom remaining for the secondary task to safely navigate around obstacles.
-\remove{Consequently, the secondary task, which is aimed at obstacle avoidance, lacks the requisite capability and freedom to operate effectively.}
+
+%\remove{Consequently, the secondary task, which is aimed at obstacle avoidance, lacks the requisite capability and freedom to operate effectively.}
 

### 2024-03-20 Articles/IROS/root.tex
fixed equations that broke line

@@
 
+%\begin{equation}
+%	\label{eq: interp x}
+%	\vec{V}rep_{xyz,jk} = \vec{V}rep_{xyz,0jk}(1 - \Delta x) + \vec{V}rep_{xyz,1jk} \, \Delta x \quad \forall \, j, k \in \{1, 2\}
+%\end{equation}
+
 \begin{equation}
-	\label{eq: interp x}
-	\vec{V}rep_{xyz,jk} = \vec{V}rep_{xyz,0jk}(1 - \Delta x) + \vec{V}rep_{xyz,1jk} \, \Delta x \quad \forall \, j, k \in \{1, 2\}
+\begin{aligned}
+	\vec{V}rep_{xyz,jk} = &\vec{V}rep_{xyz,0jk}(1 - \Delta x) \\
+	&+ \vec{V}rep_{xyz,1jk} \, \Delta x \quad \forall \, j, k \in \{1, 2\}
+\end{aligned}
 \end{equation}
 
-\alert{equations break column}
+%\begin{equation}
+%	\label{eq: interp y}
+%	\vec{V}rep_{xyz,k} = \vec{V}rep_{xyz,0k}(1 - \Delta y) + \vec{V}rep_{xyz,1k} \, \Delta y \quad \forall \, k \in \{1, 2\}
+%\end{equation}
 
 \begin{equation}
-	\label{eq: interp y}
-	\vec{V}rep_{xyz,k} = \vec{V}rep_{xyz,0k}(1 - \Delta y) + \vec{V}rep_{xyz,1k} \, \Delta y \quad \forall \, k \in \{1, 2\}
+	\begin{aligned}
+		\vec{V}rep_{xyz,k} = &\vec{V}rep_{xyz,0k}(1 - \Delta y) \\
+		&+ \vec{V}rep_{xyz,1k} \, \Delta y \quad \forall \, k \in \{1, 2\}
+	\end{aligned}
 \end{equation}
@@
 	\centering
-	\includegraphics[width=1\linewidth]{demo-field1c.eps} % Replace 'example-image' with the filename of your image
+	\includegraphics[width=0.8\linewidth]{demo-field1c.eps} % Replace 'example-image' with the filename of your image
 	\caption{Visualization of the potential field calculated for the entire map using the proposed approach.
@@
 \end{figure}
-
-\comment{}{fix units in  plots}
 
@@
 The simulation is performed for 75 steps.
-
-\add{describe weights and distributions used in experiments the kernels}
-
-\add{sigmoid weights}
 

### 2024-03-20 Articles/IROS/root.tex
removed a bunch of references

@@
 Motion planning~\cite{c52} is a technique in robotics, that is used to calculate joint changes leading manipulators from initial to a target configuration, while addressing constraints and optimization criteria.
-This task is facilitated by the manipulators' redundant degrees of freedom (DOF), enabling it to undertake various configurations to not only reach their End Effector (EE) target but also optimize for secondary objectives such as obstacle avoidance and minimizing joint torques~\cite{siciliano1990kinematic}.
+This task is facilitated by the manipulators' redundant degrees of freedom (DOF), enabling it to undertake various configurations to not only reach their End Effector (EE) target but also optimize for secondary objectives such as obstacle avoidance and minimizing joint torques~\cite{c30}.
 
-While global motion planning techniques provide comprehensive solutions by examining the entire configuration space, they fall short in terms of smoothness and real-time application, making them less ideal for dynamic settings~\cite{lavalle1998rapidly, gammell2015batch, karaman2010incremental,kuffner2000rrt}.
-Conversely, local planning methods, particularly inverse kinematics~\cite{c29,c38} and quadratic programming~\cite{c21,c23, haviland2021neo}, offer prompt, smooth trajectories suitable for real-time operations.
+While global motion planning techniques provide comprehensive solutions by examining the entire configuration space, they fall short in terms of smoothness and real-time application, making them less ideal for dynamic settings~\cite{c52}.
+Conversely, local planning methods, particularly inverse kinematics~\cite{c38} and quadratic programming~\cite{c21,c23, haviland2021neo}, offer prompt, smooth trajectories suitable for real-time operations.
 However, these methods are prone to fall into local minima due to their incremental planning approach.
 
-Often used approach is a combination of global and local planning strategies~\cite{c44}, leveraging global planners for static environment navigation and local planners for adjusting to dynamic changes.
+Often used approach is a combination of global and local planning strategies, leveraging global planners for static environment navigation and local planners for adjusting to dynamic changes.
 This implementation enhances the manipulator's ability to navigate complex environments by combining the comprehensive pathfinding capabilities of global methods with the adaptability and efficiency of local techniques.
 
-The Artificial Potential Field (APF) concept introduced by Khatib~\cite{c33}, employs repulsive forces for obstacle deflection and attractive forces to guide towards targets, many modifications of the method have been proposed over the year, often trying to mitigate local minima~\cite{c43,c45,c47,klancar2022robot}.
-Some methods have further adapted APF for manipulator motion planning, incorporating elements such as numerical Jacobians to refine trajectory planning~\cite{c49,park2020trajectory,baumgartner2023potential}.
+The Artificial Potential Field (APF) concept introduced by Khatib~\cite{c33}, employs repulsive forces for obstacle deflection and attractive forces to guide the robot towards target.
+Different variations of the method have been proposed adapting APF for manipulator motion planning~\cite{klancar2022robot, c49,park2020trajectory,baumgartner2023potential, c30}.
 
-In addressing dynamic environments, technologies such as ESDF grids, derived from sensor data or pre-established occupancy grids, play a crucial role in real-time obstacle avoidance~\cite{oleynikova2017voxblox,han2019fiesta,lau2010improved,rong2006jump,zhou2021egoplanner}.
+In addressing dynamic environments, technologies such as ESDF grids, derived from sensor data or pre-established occupancy grids, play a crucial role in real-time obstacle avoidance~\cite{oleynikova2017voxblox,han2019fiesta,zhou2021egoplanner}.
 
@@
 
-\bibitem{siciliano1990kinematic} B. Siciliano, “Kinematic control of redundant robot manipulators: A tutorial,” J. Intell. Robot. Syst., vol. 3, no. 3, Art. no. 3, 1990, doi: 10.1007/BF00126069.
+%\bibitem{siciliano1990kinematic} B. Siciliano, “Kinematic control of redundant robot manipulators: A tutorial,” J. Intell. Robot. Syst., vol. 3, no. 3, Art. no. 3, 1990, doi: 10.1007/BF00126069.
 
-\bibitem{lavalle1998rapidly} S. LaValle, “Rapidly-exploring random trees: A new tool for path planning,” Research Report 9811, 1998.
+%\bibitem{lavalle1998rapidly} S. LaValle, “Rapidly-exploring random trees: A new tool for path planning,” Research Report 9811, 1998.
 
-\bibitem{gammell2015batch} J.
+%\bibitem{gammell2015batch} J.
 D.
@@
 
-\bibitem{karaman2010incremental} S. Karaman and E. Frazzoli, "Incremental sampling-based algorithms for optimal motion planning," in Proc. Robotics: Science and Systems (RSS), 2010.
+%\bibitem{karaman2010incremental} S. Karaman and E. Frazzoli, "Incremental sampling-based algorithms for optimal motion planning," in Proc. Robotics: Science and Systems (RSS), 2010.
 
-\bibitem{kuffner2000rrt} J.
+%\bibitem{kuffner2000rrt} J.
 J.
@@
 
-\bibitem{c29} A.
+%\bibitem{c29} A.
 A.
@@
 
-\bibitem{c44} Z.
+%\bibitem{c44} Z.
 Long, “Virtual target point-based obstacle-avoidance method for manipulator systems in a cluttered environment,” Engineering Optimization, vol.
@@
 
-\bibitem{c43} M.
+%\bibitem{c43} M.
 F.
@@
 
-\bibitem{c45} A. H. Qureshi and Y. Ayaz, “Potential Functions based Sampling Heuristic For Optimal Path Planning,” Auton Robot, vol. 40, no. 6, Art. no. 6, Aug. 2016. doi: 10.1007/s10514-015-9518-0.
+%\bibitem{c45} A. H. Qureshi and Y. Ayaz, “Potential Functions based Sampling Heuristic For Optimal Path Planning,” Auton Robot, vol. 40, no. 6, Art. no. 6, Aug. 2016. doi: 10.1007/s10514-015-9518-0.
 
-\bibitem{c47} J.
+%\bibitem{c47} J.
 Yi, Q.
@@
 
+\bibitem{c30} L.
+Žlajpah and T.
+Petrič, “Obstacle Avoidance for Redundant Manipulators as Control Problem,” in Serial and Parallel Robot Manipulators - Kinematics, Dynamics, Control and Optimization, S.
+Kucuk, Ed.
+InTech, 2012.
+doi: 10.5772/32651.
+
 \bibitem{oleynikova2017voxblox} H.
@@
 
-\bibitem{lau2010improved} B.
+%\bibitem{lau2010improved} B.
 Lau, C.
@@
 
-\bibitem{rong2006jump} G.
+%\bibitem{rong2006jump} G.
 Rong and T.-S.
@@
 
-\bibitem{c30} L.
-Žlajpah and T.
-Petrič, “Obstacle Avoidance for Redundant Manipulators as Control Problem,” in Serial and Parallel Robot Manipulators - Kinematics, Dynamics, Control and Optimization, S.
-Kucuk, Ed.
-InTech, 2012.
-doi: 10.5772/32651.
-
-\bibitem{c32} K.
+%\bibitem{c32} K.
 Glass, R.
@@
 
-\bibitem{c41} L.
+%\bibitem{c41} L.
 Žlajpah and B.

### 2024-03-20 Articles/IROS/root.tex
references cleared
Mentor Corrections: Experiments

@@
 The use of interpolation allows us to have a relatively coarse voxel grid, which reduces the memory demand of the space grid and accelerates the computation.
-We selected \( K = 7 \) Points of Interest (POI) to observe the distance from obstacles and are uniformly distributed them along the segments and joints, starting from the second joint of the robot to the tip of the manipulator.
+We selected \( K = 7 \) POI to observe the distance from obstacles and are uniformly distributed them along the segments and joints, starting from the second joint of the robot to the tip of the manipulator.
 This distribution begins from the second joint because it is from this point onwards that the manipulator has the capability to avoid obstacles.
-Points of Interest (POIs) are indicated by dots on the manipulator in the graphs.
+POIs are indicated by dots on the manipulator in the figures.
 Near obstacles, vectors emanate from these points, depicting the calculated repulsive velocities at the locations.
@@
 As the most direct path for the end-effector to the target passes straight through the column, an approach that would result in a collision, we implement a reduction of the primary speed in the vicinity of the obstacle, setting \( \xi_{p} = 1 \).
-The simulation is performed for 50 steps.
 
@@
 The proposed method for calculating repulsive velocities proves to be a suitable choice when a local optimization approach is needed to accommodate dynamic changes.
-In this scenario, a ball moves consistently from the left to the right side of the graph (i.e., along the y-axis from 0.4 at the initial moment to 1.4 in the final step of the simulation), as governed by the equation \( y_{\text{ball}} = 0.4 + (1.4 - 0.4) \times \frac{N_{\text{step}}}{75} \),  with \( x_{\text{ball}} = 1.25 \) and \( z_{\text{ball}} = 0.7 \).
+In this scenario, a ball moves as governed by the equation \( y_{\text{ball}} = 0.4 + (1.4 - 0.4) \frac{N_{\text{step}}}{75} \),  with \( x_{\text{ball}} = 1.25 \) and \( z_{\text{ball}} = 0.7 \).
 The primary task is now to maintain the robot's end-effector (EE) at a fixed point, disregarding EE orientation to enhance obstacle avoidance capability.
@@
 Weighting constants for the individual POIs are \( \alpha = \frac{[ \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{1}{5} ]}{10} \).
-The simulation is performed for 75 steps.
 
@@
 
-%\bibitem{siciliano1990kinematic} B. Siciliano, “Kinematic control of redundant robot manipulators: A tutorial,” J. Intell. Robot. Syst., vol. 3, no. 3, Art. no. 3, 1990, doi: 10.1007/BF00126069.
-
-%\bibitem{lavalle1998rapidly} S. LaValle, “Rapidly-exploring random trees: A new tool for path planning,” Research Report 9811, 1998.
-
-%\bibitem{gammell2015batch} J.
-D.
-Gammell, S.
-S.
-Srinivasa, and T.
-D.
-Barfoot, “Batch Informed Trees (BIT*): Sampling-based Optimal Planning via the Heuristically Guided Search of Implicit Random Geometric Graphs,” in Proc.
-of the 2015 IEEE International Conference on Robotics and Automation (ICRA), May 2015, pp.
-3067–3074.
-doi: 10.1109/ICRA.2015.7139620.
-
-%\bibitem{karaman2010incremental} S. Karaman and E. Frazzoli, "Incremental sampling-based algorithms for optimal motion planning," in Proc. Robotics: Science and Systems (RSS), 2010.
-
-%\bibitem{kuffner2000rrt} J.
-J.
-Kuffner and S.
-M.
-LaValle, “RRT-connect: An efficient approach to single-query path planning,” in Proceedings of the 2000 ICRA.
-Millennium Conference.
-IEEE International Conference on Robotics and Automation.
-Symposia Proceedings (Cat.
-No.00CH37065), Apr.
-2000, pp.
-995–1001 vol.2.
-doi: 10.1109/ROBOT.2000.844730.
-
-%\bibitem{c29} A.
-A.
-Maciejewski and C.
-A.
-Klein, “Obstacle Avoidance for Kinematically Redundant Manipulators in Dynamically Varying Environments,” The International Journal of Robotics Research, vol.
-4, no.
-3, pp.
-109–117, Sep.
-1985.
-doi: 10.1177/027836498500400308.
+\bibitem{c30} L.
+Žlajpah and T.
+Petrič, “Obstacle Avoidance for Redundant Manipulators as Control Problem,” in Serial and Parallel Robot Manipulators - Kinematics, Dynamics, Control and Optimization, S.
+Kucuk, Ed.
+InTech, 2012.
+doi: 10.5772/32651.
 
@@
 
-%\bibitem{c44} Z.
-Long, “Virtual target point-based obstacle-avoidance method for manipulator systems in a cluttered environment,” Engineering Optimization, vol.
-52, no.
-11, Art.
-no.
-11, Nov.
-2020.
-doi: 10.1080/0305215X.2019.1681986.
-
 \bibitem{c33} O.
@@
 doi: 10.1109/ROBOT.1985.1087247.
-
-%\bibitem{c43} M.
-F.
-Pinto, T.
-R.
-F.
-Mendonça, L.
-R.
-Olivi, E.
-B.
-Costa, and A.
-L.
-M.
-Marcato, “Modified approach using variable charges to solve inherent limitations of potential fields method,” in Proc.
-2014 11th IEEE/IAS International Conference on Industry Applications, Dec.
-2014, pp.
-1–6.
-doi: 10.1109/INDUSCON.2014.7059414.
-
-%\bibitem{c45} A. H. Qureshi and Y. Ayaz, “Potential Functions based Sampling Heuristic For Optimal Path Planning,” Auton Robot, vol. 40, no. 6, Art. no. 6, Aug. 2016. doi: 10.1007/s10514-015-9518-0.
-
-%\bibitem{c47} J.
-Yi, Q.
-Yuan, R.
-Sun, and H.
-Bai, “Path planning of a manipulator based on an improved P\_RRT* algorithm,” Complex Intell.
-Syst., vol.
-8, no.
-3, pp.
-2227–2245, Jun.
-2022.
-doi: 10.1007/s40747-021-00628-y.
 
@@
 
-\bibitem{c30} L.
-Žlajpah and T.
-Petrič, “Obstacle Avoidance for Redundant Manipulators as Control Problem,” in Serial and Parallel Robot Manipulators - Kinematics, Dynamics, Control and Optimization, S.
-Kucuk, Ed.
-InTech, 2012.
-doi: 10.5772/32651.
-
 \bibitem{oleynikova2017voxblox} H.
@@
 Available: http://arxiv.org/abs/1903.02144
-
-%\bibitem{lau2010improved} B.
-Lau, C.
-Sprunk, and W.
-Burgard, “Improved updating of Euclidean distance maps and Voronoi diagrams,” in Proc.
-of the 2010 IEEE/RSJ International Conference on Intelligent Robots and Systems, Taipei, Oct.
-2010, pp.
-281–286.
-doi: 10.1109/IROS.2010.5650794.
-
-%\bibitem{rong2006jump} G.
-Rong and T.-S.
-Tan, “Jump flooding in GPU with applications to Voronoi diagram and distance transform,” in Proceedings of the 2006 Symposium on Interactive 3D Graphics and Games - SI3D ’06, Redwood City, California, 2006, p.
-109.
-doi: 10.1145/1111411.1111431.
 
@@
 
-%\bibitem{c32} K.
-Glass, R.
-Colbaugh, D.
-Lim, and H.
-Seraji, “Real-time collision avoidance for redundant manipulators,” IEEE Trans.
-Robot.
-Automat., vol.
-11, no.
-3, pp.
-448–457, Jun.
-1995.
-doi: 10.1109/70.388789.
-
-%\bibitem{c41} L.
-Žlajpah and B.
-Nemec, “Kinematic control algorithms for on-line obstacle avoidance for redundant manipulators,” in Proc.
-IEEE/RSJ International Conference on Intelligent Robots and Systems, Lausanne, Switzerland, 2002, pp.
-1898–1903.
-doi: 10.1109/IRDS.2002.1044033.
-
 \bibitem{Zlajpah2017}
 L.
-Zlajpah, “On orientation control of functional redundant robots,” in Proc.
+Žlajpah, “On orientation control of functional redundant robots,” in Proc.
 of the 2017 IEEE International Conference on Robotics and Automation (ICRA), Singapore, May 2017, pp.

### 2024-03-20 Articles/IROS/root.tex
fixes page 1

@@
 
-The article presents a new method for obstacle avoidance in robotic manipulators.
+The article presents a new method for obstacle avoidance for robotic manipulators.
 It uses occupancy voxel grids to calculate repulsive velocities based on the proximity of obstacles.
@@
 Repulsive velocities generated through matrix multiplication are processed using tri-linear interpolation, resulting in smooth and efficient movement away from potential hazards.
-The integration of this method with inverse kinematic control improves manipulator motion planning and obstacle avoidance capabilities.
-Simulation results indicate the effectiveness of this method in various scenarios, showcasing its potential for local motion planning and obstacle avoidance.
+The integration of this method with inverse kinematic control enables manipulator motion planning and obstacle avoidance capabilities.
+Simulation results prove the effectiveness of this method in various scenarios, showcasing its potential for local motion planning and obstacle avoidance.
 
@@
 
-Motion planning~\cite{c52} is a technique in robotics, that is used to calculate joint changes leading manipulators from initial to a target configuration, while addressing constraints and optimization criteria.
+Motion planning~\cite{c52} is a technique in robotics, that is used to calculate joint changes leading a manipulator from initial to a target configuration, while addressing constraints and optimization criteria.
 This task is facilitated by the manipulators' redundant degrees of freedom (DOF), enabling it to undertake various configurations to not only reach their End Effector (EE) target but also optimize for secondary objectives such as obstacle avoidance and minimizing joint torques~\cite{c30}.
@@
 
-The Artificial Potential Field (APF) concept introduced by Khatib~\cite{c33}, employs repulsive forces for obstacle deflection and attractive forces to guide the robot towards target.
+The Artificial Potential Field (APF) concept~\cite{c33}, employs repulsive forces for obstacle repulsion and attractive forces to guide the robot towards target.
 Different variations of the method have been proposed adapting APF for manipulator motion planning~\cite{klancar2022robot, c49,park2020trajectory,baumgartner2023potential, c30}.
 
-In addressing dynamic environments, technologies such as ESDF grids, derived from sensor data or pre-established occupancy grids, play a crucial role in real-time obstacle avoidance~\cite{oleynikova2017voxblox,han2019fiesta,zhou2021egoplanner}.
+In addressing dynamic environments, technologies such as Euclidian Signed Distance Field (ESDF) grids, derived from sensor data or pre-established occupancy grids, play a crucial role in real-time obstacle avoidance~\cite{oleynikova2017voxblox,han2019fiesta,zhou2021egoplanner}.
 
@@
 
-Our approach utilizes occupancy voxel grids~\cite{han2018dynamic} to enable safe and dynamic navigation for robot manipulators in changing environments.
+The introduced approach utilizes occupancy voxel grids~\cite{han2018dynamic} to enable safe and dynamic navigation for robot manipulators in changing environments.
+A method for obstacle avoidance in robotic manipulators via matrix multiplication-driven repulsive fields is presented.
 By representing the workspace with voxels and calculating repulsive velocities based on proximity to obstacles, our method allows the robot to determine safe movement directions and velocities in real-time.
@@
 
-This article introduces a method for obstacle avoidance in robotic manipulators via matrix multiplication-driven repulsive fields.
 Section II outlines the mathematical framework for generating repulsive velocities.
@@
 
-We leverage a voxel-based representation of the surrounding space, to dynamically assesses potential collision threats and calculates directional repulsive velocities for smooth and safe navigation through complex environments.
+The method leverages a voxel-based representation of the surrounding space, to dynamically assess potential collision threats and calculate directional repulsive velocities for smooth and safe navigation through complex environments.
 
@@
 
-Applying the Hadamard (element-wise) product to the segment of the occupation grid \(G_d\) and the kernel \(W_d\) results in a new matrix for each Cartesian direction (\(d \in \{x, y, z\}\)).
+Applying the Hadamard (element-wise) product to the segment of the occupation grid \(G_d\) and the kernel \(W_d\) results in a new matrix.
 By summing over all values in this resultant matrix, we calculate the repulsive velocity component for that specific direction.

### 2024-03-20 Articles/IROS/root.tex
fixes page 2

@@
 
-By multiplieing weights components for primary and both orthagonal axis we get matrix kernels. We get one kernel for every cartesian direction.
+By multiplieing weights components for primary and both orthagonal axis we get matrix kernels. We calculate a different kernel for every cartesian direction.
 
@@
 The occupancy grid's representation of the robot's environment is inherently discrete, possessing finite resolution, in contrast to the continuous nature of Cartesian space.
-The approximation or rounding necessary to transition from Cartesian coordinates to discrete grid representations may introduce discontinuities in the calculated repulsive field.
+The approximation or rounding necessary to transition from Cartesian coordinates to discrete grid representations introduces discontinuities in the calculated repulsive field.
 To counteract this, tri-linear interpolation is utilized, facilitating a smooth transition to a continuous repulsive field value, as illustrated in Fig.
@@
 Therefore, before interpolation, we shift the values of the cells coordinates by half the resolution of the obstacle grid in each direction.
-That is, the cell coordinates are same as $x,y,z$, cell index and a half multiplied by grid resolution $\Delta \mathrm{R}$.
+That is, the cell coordinates are same as $x,y,z$, cell index and plus a half of grid resolution $\Delta \mathrm{R}$ step.
 

### 2024-03-20 Articles/IROS/root.tex
fixes page 3

@@
 $\xi_{p}$ is the primary task execution slowdown constant.
-Finally $\dot{\vec{q}}_{rep}$ are the weighted sum of avoidance joint velocities, each transformed into the null space of primary velocities, as described in the chapter below.
+Finally $\dot{\vec{q}}_{rep}$ are the weighted sum of avoidance joint velocities, each transformed into the null space of primary velocities.
 
@@
 
-To get the full velocity of the end effector (EE), \( \mathrm{\vec{v}_{att}} \), we combines translational and rotational velocities, which we scale using proportional gains \( k_v \) and \( k_\omega \).
+To get the full velocity of the end effector (EE), \( \mathrm{\vec{v}_{att}} \), translational and rotational velocities are scaled using proportional gains and combined \( k_v \) and \( k_\omega \).
 
@@
 To mitigate the dominance of the primary task over the secondary task, a mechanism for the deceleration of the primary task's execution has been implemented.
-This strategy effectively decreases the velocity of the manipulator towards its primary objective, thereby allocating additional maneuverability for secondary tasks.
+This strategy effectively decreases the velocity of the manipulator towards its primary objective, thereby allocating additional maneuverability for secondary task.
 

### 2024-03-20 Articles/IROS/root.tex
fixes page 4

@@
 The deceleration factor (eq.~\ref{eq:slowdown}) is determined by the constant \( \kappa_{\text{sec}} \) and the magnitude of the maximum repulsive velocity vectors, which is contingent upon the robot's minimal distance from an obstacle \( v_{\text{Rmax}} \).
-According to the equation, a small \( v_{\text{Rmax}} \) diminishes the deceleration effect, permitting the uninterrupted execution of the primary task.
-Conversely, large \( v_{\text{Rmax}} \) enhances the deceleration by reducing the $\xi_{p}$ factor, thereby affording the secondary task increased opportunity for obstacles avoidance motion.
+According to the equation, a small \( v_{\text{Rmax}} \) minimizes the deceleration effect, permitting the uninterrupted execution of the primary task.
+Conversely, large \( v_{\text{Rmax}} \) boosts the deceleration by reducing the $\xi_{p}$ factor, thereby affording the secondary task increased opportunity for obstacles avoidance motion.
 
@@
 Given a matrix width $w$ perpendicular to the primary direction, it is advantageous for these points to be spaced at intervals of $\Delta = \frac{w}{3}$.
-This spacing ensures partial overlap between matrices, thereby achieving a uniform distribution of POIs across the entirety of the robot.
+This spacing ensures partial overlap between matrices, thereby achieving a uniform distribution of repulsive vectors across the entirety of the robot.
 
@@
 Upon obtaining the repulsive velocity for each point, the second norm of each velocity is calculated.
-For a number of $K$ points where the avoidance velocities are greatest, indicative of closest proximity to an obstacle, the resulting joint velocities that facilitate obstacle avoidance are computed using inverse kinematic equations for the selected points on the manipulator and transformed into the null-space velocities of the primary task.
+For a number of $K$ points where the avoidance velocities are greatest, indicative of closest proximity to an obstacle, the resulting joint velocities that facilitate obstacle avoidance are computed using inverse kinematic equations for the selected points on the manipulator and transformed into the null-space velocities of the primary task (eq.~\ref{eq:point_nullspace}).
 
@@
 	\mathbf{J}_{d_i}^{+} = \mathbf{N} \mathbf{J}_{d_i}' (\mathbf{J}_{d_i} \mathbf{N} \mathbf{J}_{d_i}' + \sigma_{rep} \mathbf{I})^{-1}, \quad \text{for } i = 1, 2, \ldots, K;
+	\label{eq:point_nullspace}
 \end{equation}

### 2024-03-21 Articles/IROS/root.tex
Mentor Corrections: Round 2

@@
 The article presents a new method for obstacle avoidance for robotic manipulators.
-It uses occupancy voxel grids to calculate repulsive velocities based on the proximity of obstacles.
+It uses occupancy voxel grids and convolutional matrix kernels to calculate repulsive velocities based on the proximity of obstacles.
 This approach allows for safe and agile navigation in changing environments.
@@
 
-Often used approach is a combination of global and local planning strategies, leveraging global planners for static environment navigation and local planners for adjusting to dynamic changes.
+Often used approach is a combination~\cite{a15} of global and local planning strategies, leveraging global planners for static environment navigation and local planners for adjusting to dynamic changes.
 This implementation enhances the manipulator's ability to navigate complex environments by combining the comprehensive pathfinding capabilities of global methods with the adaptability and efficiency of local techniques.
@@
 The segment needs to be of the same size as the directional matrix, for the cordinate axis in which we are calulating the repulsive velocity.
-This results in two identically sized 3D matrices: one serving as a "window" into the obstacle grid (\(G_d\)) and the other as the directional kernel (\(W_d\)).
+This results in two identically sized 3D matrices: one serving as a "window" into the obstacle grid $\mathbf{(G_d)}$ and the other as the directional kernel $\mathbf{(W_d)}$.
 
-Applying the Hadamard (element-wise) product to the segment of the occupation grid \(G_d\) and the kernel \(W_d\) results in a new matrix.
+Applying the Hadamard (element-wise) product to the segment of the occupation grid $\mathbf{(G_d)}$ and the kernel $\mathbf{(W_d)}$ results in a new matrix.
 By summing over all values in this resultant matrix, we calculate the repulsive velocity component for that specific direction.
-This process is independently conducted for each Cartesian coordinate, ultimately yielding the comprehensive repulsive velocity vector \(\vec{v}_i = [ v_{x_i}, v_{y_i}, v_{z_i}]\).
+This process is independently conducted for each Cartesian coordinate, ultimately yielding the comprehensive repulsive velocity vector \(\mathbf{v}_i = [ v_{x_i}, v_{y_i}, v_{z_i}]\).
 
 \begin{equation}
-	v_d = \sum_{i,j,k} (G \odot W)_{ijk} = \sum_{i}\sum_{j}\sum_{k} g_{\Delta i \Delta j \Delta k} ~  w_{\Delta i \Delta j \Delta k}
+	v_d = \sum_{i,j,k} (\mathbf{G} \odot \mathbf{W})_{ijk} = \sum_{i}\sum_{j}\sum_{k} g_{\Delta i \Delta j \Delta k} ~  w_{\Delta i \Delta j \Delta k}
 	\label{eq:matrix-product}
 \end{equation}
+
+Where $g_{\Delta i \Delta j \Delta k}$ and $w_{\Delta i \Delta j \Delta k}$ denote $i$-th, $j$-th, $k$-th indexed elements of each matrix.
 
@@
 
-The fundamental concept of directional kernels lies in computing the repulsive field individually for each direction within the Cartesian coordinate system.
+The fundamental concept of convolutional directional kernels lies in computing the repulsive field individually for each direction within the Cartesian coordinate system.
 
@@
 
-We introduce two distributions for the primary axis weights, namely gaussian and linear.
+We introduce two distributions for the primary axis weights, namely Gaussian and linear.
 The selection of weight distribution is contingent on the desired profile of repulsive velocities within the Artificial Potential Field (APF).
@@
 	\begin{cases} 
-		sin(\left| \Delta j \right| \pi / (2 \Delta j_{max})) & \text{if } \Delta j < 0 \text{~or }  \Delta j > 0 \\
+		\sin(\left| \Delta j \right| \pi / (2 \Delta j_{max})) & \text{if } \Delta j < 0 \text{~or }  \Delta j > 0 \\
 		1 & \text{if } \Delta j = 0 \\
@@
 \begin{equation}
-	w_{\Delta i \Delta j \Delta k} = w_{\Delta i} * w_{\Delta j} * w_{\Delta k}
+	w_{\Delta i \Delta j \Delta k} = w_{\Delta i} \cdot w_{\Delta j} \cdot w_{\Delta k}
 \end{equation}
@@
 
-The indexes of the eight surrounding cells are calculated by scaling the POI $x,y,z$ by grid resolution and than rounding the position to the nearest lower and upper integer positions (eq.~\ref{eq:floor and ceil}).
+The indexes of the eight surrounding cells $s=1...8$ are calculated by scaling the selected POI coordinates $x,y,z$ by grid resolution and than rounding the position to the nearest lower and upper integer positions (eq.~(\ref{eq:floor and ceil})).
 
@@
 	\begin{bmatrix}
-		X \\
-		Y \\
-		Z \\
+		\mathbf{X} \\
+		\mathbf{Y} \\
+		\mathbf{Z} \\
 	\end{bmatrix}
@@
 	\begin{bmatrix}
-		\; \lfloor \vec{p}_{\mathrm{POI}}(1) \times\Delta \mathrm{R} \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(1) \times\Delta \mathrm{R} \rceil \;  \\
-		\; \lfloor \vec{p}_{\mathrm{POI}}(2) \times\Delta \mathrm{R} \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(2) \times\Delta \mathrm{R} \rceil \; \\
-		\; \lfloor \vec{p}_{\mathrm{POI}}(3) \times\Delta \mathrm{R} \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(3) \times\Delta \mathrm{R} \rceil \; 
+		\; \lfloor \mathbf{p}_{\mathrm{POI}}(1) \times\Delta \mathrm{R} \rfloor \; \lceil \mathbf{p}_{\mathrm{POI}}(1) \times\Delta \mathrm{R} \rceil \;  \\
+		\; \lfloor \mathbf{p}_{\mathrm{POI}}(2) \times\Delta \mathrm{R} \rfloor \; \lceil \mathbf{p}_{\mathrm{POI}}(2) \times\Delta \mathrm{R} \rceil \; \\
+		\; \lfloor \mathbf{p}_{\mathrm{POI}}(3) \times\Delta \mathrm{R} \rfloor \; \lceil \mathbf{p}_{\mathrm{POI}}(3) \times\Delta \mathrm{R} \rceil \; 
 	\end{bmatrix}
@@
 
-Once we got the indexes of the eight surrounding cells of our POI, we calculate the repulsive velocity vectors in each point using kernel matrix multiplication method (seperately for each cartesian direction) (eq.~\ref{eq: calc rep vel}).
+Once we got the indexes of the eight surrounding cells of our POI, we calculate the repulsive velocity vectors in each point using kernel matrix multiplication method (seperately for each cartesian direction) (eq.~(\ref{eq: calc rep vel})).
 
@@
 	\label{eq: calc rep vel}
-	\vec{V}rep_{xyz,ijk} = \mathrm{rep}(X[i], Y[j], Z[k]) \quad \forall i, j, k \in \{1, 2\}
+	\mathbf{V}rep_{xyz,nml} = \mathrm{rep}(\mathbf{X}[n], \mathbf{Y}[m], \mathbf{Z}[l]) \quad \forall n, m, l \in \{1, 2\}
 \end{equation}
@@
 Trilinear interpolation method works on a 3-dimensional regular grid.
-Before we can start with the interpolation we need to calculate the distance between POI and smaller coordinates of the cells where we calculated the repulsive velocities (eq.~\ref{eq: deltas interp}).
+Before we can start with the interpolation we need to calculate the distance between POI and smaller coordinates of the cells where we calculated the repulsive velocities (eq.~(\ref{eq: deltas interp})).
 The calculated repulsive velocity values are located at the centers of the cells.
@@
 	\begin{bmatrix}
-		\frac{\left( P_x - \left( X(1) + \frac{1}{2} \Delta \mathrm{R} \right)  \right)}{\left( X(2) - X(1) \right)} \\
-		\frac{\left( P_y - \left( Y(1) + \frac{1}{2} \Delta \mathrm{R} \right)  \right)}{\left( Y(2) - Y(1) \right)} \\
-		\frac{\left( P_z - \left( Z(1) + \frac{1}{2} \Delta \mathrm{R} \right)  \right)}{\left( Z(2) - Z(1) \right)} \\
+		\frac{\left( P_x - \left( \mathbf{X}(1) + \frac{1}{2} \Delta \mathit{R} \right)  \right)}{\left( \mathbf{X}(2) - \mathbf{X}(1) \right)} \\
+		\frac{\left( P_y - \left( \mathbf{Y}(1) + \frac{1}{2} \Delta \mathit{R} \right)  \right)}{\left( \mathbf{Y}(2) - \mathbf{Y}(1) \right)} \\
+		\frac{\left( P_z - \left( \mathbf{Z}(1) + \frac{1}{2} \Delta \mathit{R} \right)  \right)}{\left( \mathbf{Z}(2) - \mathbf{Z}(1) \right)} \\
 	\end{bmatrix}
@@
 \begin{aligned}
-	\vec{V}rep_{xyz,jk} = &\vec{V}rep_{xyz,0jk}(1 - \Delta x) \\
-	&+ \vec{V}rep_{xyz,1jk} \, \Delta x \quad \forall \, j, k \in \{1, 2\}
+	\mathbf{V}rep_{xyz,jk} = &\mathbf{V}rep_{xyz,0jk}(1 - \Delta x) \\
+	&+ \mathbf{V}rep_{xyz,1jk} \, \Delta x \quad \forall \, j, k \in \{1, 2\}
 \end{aligned}
@@
 	\begin{aligned}
-		\vec{V}rep_{xyz,k} = &\vec{V}rep_{xyz,0k}(1 - \Delta y) \\
-		&+ \vec{V}rep_{xyz,1k} \, \Delta y \quad \forall \, k \in \{1, 2\}
+		\mathbf{V}rep_{xyz,k} = &\mathbf{V}rep_{xyz,0k}(1 - \Delta y) \\
+		&+ \mathbf{V}rep_{xyz,1k} \, \Delta y \quad \forall \, k \in \{1, 2\}
 	\end{aligned}
@@
 	\label{eq: interp z}
-	\vec{V}rep_{xyz} = \vec{V}rep_{xyz,0}(1 - \Delta z) + \vec{V}rep_{xyz,1} \, \Delta z 
+	\mathbf{V}rep_{xyz} = \mathbf{V}rep_{xyz,0}(1 - \Delta z) + \mathbf{V}rep_{xyz,1} \, \Delta z 
 \end{equation}
@@
 	\caption{Repulsive field generation via convolutional kernels.
-Left: Original repulsive fields in X (top) and Y (bottom) directions.
+Left: Original repulsive fields in $\mathit{X}$ (top) and $\mathit{Y}$ (bottom) directions.
 Right: Corresponding fields using interpolation, showcasing enhanced smoothness.}
@@
 \begin{equation}
-	\dot{\vec{q}} = \mathbf{J}^+ \xi_{p} \mathrm{\vec{v}_{att}} + \dot{\vec{q}}_{rep}
+	\dot{\mathbf{q}} = \mathbf{J}^+ \xi_{p} \mathrm{\mathbf{v}_{att}} + \dot{\mathbf{q}}_{rep}
 \end{equation}
@@
 $\xi_{p}$ is the primary task execution slowdown constant.
-Finally $\dot{\vec{q}}_{rep}$ are the weighted sum of avoidance joint velocities, each transformed into the null space of primary velocities.
+Finally $\dot{\mathbf{q}}_{rep}$ are the weighted sum of avoidance joint velocities, each transformed into the null space of primary velocities.
 
@@
 \begin{equation} 
-	\vec{v} = \frac{\vec{x}_{EE} - \vec{x}_g}{||\vec{x}_{EE} - \vec{x}_g||} \times \frac{\arctan(k_{sigm} \; ||\vec{x}_{EE} - \vec{x}_g||) }{\pi/2}
+	\mathbf{v} = \frac{\mathbf{x}_{EE} - \mathbf{x}_g}{||\mathbf{x}_{EE} - \mathbf{x}_g||} \times \frac{\arctan(k_{sigm} \; ||\mathbf{x}_{EE} - \mathbf{x}_g||) }{\pi/2}
 	\label{eq: v_att}
@@
 
-In Eq.~\ref{eq: v_att}, $\vec{v}$ specifies the end-effector's (EE) translational velocity directed towards the target, integrating both its direction and magnitude.
-Here, $\vec{x}_{EE}$ represents the EE's current position, and $\vec{x}_g$ is the target position, both in Cartesian coordinates.
-The calculation $\frac{\vec{x}_{EE}-\vec{x}_g}{||\vec{x}_{EE}-\vec{x}_g||}$ generates a unit vector pointing towards the target, ensuring targeted movement.
+In Eq.~(\ref{eq: v_att}), $\mathbf{v}$ specifies the end-effector's (EE) translational velocity directed towards the target, integrating both its direction and magnitude.
+Here, $\mathbf{x}_{EE}$ represents the EE's current position, and $\mathbf{x}_g$ is the target position, both in Cartesian coordinates.
+The calculation $\frac{\mathbf{x}_{EE}-\mathbf{x}_g}{||\mathbf{x}_{EE}-\mathbf{x}_g||}$ generates a unit vector pointing towards the target, ensuring targeted movement.
 The velocity's modulation by the arctangent sigmoid function curtails overshooting by moderating speed in proximity to the target and saturating velocity at a predefined upper limit when distant.
@@
 
-To get the full velocity of the end effector (EE), \( \mathrm{\vec{v}_{att}} \), translational and rotational velocities are scaled using proportional gains and combined \( k_v \) and \( k_\omega \).
+To get the full velocity of the end effector (EE), \( \mathrm{\mathbf{v}_{att}} \), translational and rotational components are multiplied using proportional gains and combined.
 
 \begin{equation}
-	\mathrm{\vec{v}_{att}} = 
+	\mathrm{\mathbf{v}_{att}} = 
 	\begin{bmatrix}
-		k_{v} * \vec{v}   \\
-		k_{\omega} * \vec{\omega}
+		k_{v} \cdot \mathbf{v}   \\
+		k_{\omega} \cdot \mathbf{\omega}
 	\end{bmatrix}
@@
 
-The deceleration factor (eq.~\ref{eq:slowdown}) is determined by the constant \( \kappa_{\text{sec}} \) and the magnitude of the maximum repulsive velocity vectors, which is contingent upon the robot's minimal distance from an obstacle \( v_{\text{Rmax}} \).
+The deceleration factor (eq.~(\ref{eq:slowdown})) is determined by the constant \( \kappa_{\text{sec}} \) and the magnitude of the maximum repulsive velocity vectors, which is contingent upon the robot's minimal distance from an obstacle \( v_{\text{Rmax}} \).
 According to the equation, a small \( v_{\text{Rmax}} \) minimizes the deceleration effect, permitting the uninterrupted execution of the primary task.
@@
 	\includegraphics[width=1.0\linewidth]{manipulator-kernels-visualization-small.eps} % Replace 'example-image' with the filename of your image
-	\caption{Visualization of 3 repulsive Y-direction kernels in POI on the manipulator.}
+	\caption{Visualization of only $Y$ repulsive direction kernels ($X$,$Z$ not shown), for 3 seperate POI on the manipulator.
+The colours of the kernels display their absolute values.
+We can see, that the magnitude of the kernel values falls when approaching kernel edge.}
 	\label{fig:kernels}
@@
 This is achieved by mapping each point into the task space, where the calculation of repulsive velocity is performed.
-The cartesian mapping for each point comprises a kinematic transformation from the robot's base to the beginning of the segment where the point is located, followed by an additional mapping to the specific point on the segment (eq.
-\ref{eq:transformations}).
+The cartesian mapping for each point comprises a kinematic transformation from the robot's base to the beginning of the segment where the point is located, followed by an additional mapping to the specific point on the segment (eq.~(\ref{eq:transformations})).
 For each point, three components of cartesian velocity are obtained.
@@
 Upon obtaining the repulsive velocity for each point, the second norm of each velocity is calculated.
-For a number of $K$ points where the avoidance velocities are greatest, indicative of closest proximity to an obstacle, the resulting joint velocities that facilitate obstacle avoidance are computed using inverse kinematic equations for the selected points on the manipulator and transformed into the null-space velocities of the primary task (eq.~\ref{eq:point_nullspace}).
+For a number of $K$ points where the avoidance velocities are greatest, indicative of closest proximity to an obstacle, the resulting joint velocities that facilitate obstacle avoidance are computed using inverse kinematic equations for the selected points on the manipulator and transformed into the null-space velocities of the primary task (eq.~(\ref{eq:point_nullspace})).
 
@@
 Velocity space can be restricted to just one dimension, on the velocity component that directs movement away from obstacles.
-Consequently, the Jacobian, correlating joint space velocities \(q\) with the directional velocity \(d_i\), is represented as \(J_{d_i} = \vec{n}_i^T \mathbf{J}_i\).
-The normal vector, signifying the desired avoidance velocity, is essentially the unit velocity vector, derived from our matrices as \(\vec{n}_{i} = \frac{\vec{v}_{poi i}}{||\vec{v}_{poi i}||}\).
+Consequently, the Jacobian, correlating joint space velocities \(q\) with the directional velocity \(d_i\), is represented as \(\mathbf{J}_{d_i} = \mathbf{n}_i^T \mathbf{J}_i\).
+The normal vector, signifying the desired avoidance velocity, is essentially the unit velocity vector, derived from our matrices as \(\mathbf{n}_{i} = \frac{\mathbf{v}_{poi i}}{||\mathbf{v}_{poi i}||}\).
 This approach streamlines computational processes by avoiding the necessity for complex matrix inversions, transforming the Jacobian from a matrix to a vector of dimensions \(1 \times n\), where \(n\) signifies the count of joints preceding the POI, thus facilitating more straightforward calculations of repulsive velocities.
@@
 \begin{equation}
-	\dot{q}_{rep} = k_r \sum_{i=1}^{K} \alpha_i J_{d_i}^{+} \left(v_i - J_{d_i} J^{+} \xi_{p} \vec{v}_{att}\right)
+	\dot{\mathbf{q}}_{rep} = k_r \sum_{i=1}^{K} \alpha_i \mathbf{J}_{d_i}^{+} \left(v_i - \mathbf{J}_{d_i} \mathbf{J}^{+} \xi_{p} \mathbf{v}_{att}\right)
 \end{equation}
@@
 Top: Obstacle distribution via a 2D heatmap.
-Center: Induced repulsive fields along X (left) and Y (right) axes.
+Center: Induced repulsive fields along $\mathit{X}$ (left) and $\mathit{Y}$ (right) axes.
 Bottom: Composite vector field showcasing the resultant repulsive velocities for obstacle avoidance.}
@@
 The operation of the potential field is demonstrated on two different manipulator cases.
-The resolution of our voxel grid for obstacles is \( R=10cm \).
+The resolution of our voxel grid for obstacles is \( R=10\) cm.
 The use of interpolation allows us to have a relatively coarse voxel grid, which reduces the memory demand of the space grid and accelerates the computation.
@@
 The proposed method for calculating repulsive velocities proves to be a suitable choice when a local optimization approach is needed to accommodate dynamic changes.
-In this scenario, a ball moves as governed by the equation \( y_{\text{ball}} = 0.4 + (1.4 - 0.4) \frac{N_{\text{step}}}{75} \),  with \( x_{\text{ball}} = 1.25 \) and \( z_{\text{ball}} = 0.7 \).
+In this scenario, a ball moves as governed by the equation \( y_{{ball}} = 0.4 + (1.4 - 0.4) \frac{N_{{step}}}{75} \),  with \( x_{{ball}} = 1.25 \) and \( z_{{ball}} = 0.7 \).
 The primary task is now to maintain the robot's end-effector (EE) at a fixed point, disregarding EE orientation to enhance obstacle avoidance capability.
@@
 
+\bibitem{a15} J.
+Amiryan and M.
+Jamzad, "Adaptive motion planning with artificial potential fields using a prior path," in 2015 3rd RSI International Conference on Robotics and Mechatronics (ICROM), Tehran, Iran, 2015, pp.
+731-736, doi: 10.1109/ICRoM.2015.7367873.
+Keywords: Planning; Force; Motion segmentation; Vehicles; Path planning; Mobile robots; Artificial Potential Fields; Autonomous vehicle; Evolutionary Algorithms; Random Sampling; Reactive Motion Planning.
+
 \bibitem{c33} O.
